### 2016-07-24 draft/draft_v0.tex
Minor update

% Version 0; preprint format; Written by SH

\documentclass[preprint]{aastex}

% Packages
\usepackage{emulateapj5}
\usepackage{apjfonts}
\usepackage{amssymb, amsmath}
\usepackage{graphicx}
\usepackage{CJK}
\usepackage{natbib}
\usepackage{hyperref}
\usepackage[usenames]{color}

% Package Settings
\bibliographystyle{apj}
\hypersetup{colorlinks=true,
            citecolor=cyan,
            linkcolor=cyan,
            filecolor=magenta,      
            urlcolor=cyan}
\urlstyle{same}

% Figure extention
\DeclareGraphicsExtensions{.pdf,.png,.jpg}
\input psfig.tex

%%%%%%%%%%%%: User Defined Commands %%%%%%%%%%%%

% Song Huang's definition 
\def\arcsec{{\prime\prime}}
\def\arcmin{{\prime}}
\def\degree{{\circ}}
\def\h{\hskip -3 mm}
\def\aa{{A\&A}}
\def\aas{{ A\&AS}}
\def\aj{{AJ}}
\def\al{$\alpha$}
\def\bet{$\beta$}
\def\amin{$^\prime$}
\def\annrev{{ARA\&A}}
\def\apj{{ApJ}}
\def\apjs{{ApJS}}
\def\asec{$^{\prime\prime}$}
\def\deg{$^{\circ}$}
\def\ddeg{{\rlap.}$^{\circ}$}
\def\dsec{{\rlap.}$^{\prime\prime}$}
\def\cc{cm$^{-3}$}
\def\flamb{erg s$^{-1}$ cm$^{-2}$ \AA$^{-1}$}
\def\flux{erg s$^{-1}$ cm$^{-2}$}
\def\fnu{erg s$^{-1}$ cm$^{-2}$ Hz$^{-1}$}
\def\hst{{\textit{HST}}}
\def\kms{km s$^{-1}$}
\def\lamb{$\lambda$}
\def\lax{{$\mathrel{\hbox{\rlap{\hbox{\lower4pt\hbox{$\sim$}}}\hbox{$<$}}}$}}
\def\gax{{$\mathrel{\hbox{\rlap{\hbox{\lower4pt\hbox{$\sim$}}}\hbox{$>$}}}$}}
\def\simlt{\lower.5ex\hbox{$\; \buildrel < \over \sim \;$}}
\def\simgt{\lower.5ex\hbox{$\; \buildrel > \over \sim \;$}}
\def\micron{{$\mu$m}}
\def\mnras{{MNRAS}}
\def\nat{{Nature}}
\def\pasp{{PASP}}
\def\perang{\AA$^{-1}$}
\def\peryr{yr$^{-1}$}
\def\reference{\noindent\pp}
\def\refindent{\par\noindent\parskip=2pt\hangindent=3pc\hangafter=1 }
\def\sb{mag~arcsec$^{-2}$}
\def\solum{$L_\odot$}
\def\solmass{$M_\odot$}
\def\sigs{$\sigma_*$}
\newcommand{\lt}{<}
\newcommand{\gt}{>}

\def\etal{{\ et al.~}}
\def\galfit{{\tt GALFIT}}
\def\ser{{S\'{e}rsic\ }}
\def\redm{\texttt{redMaPPer}~}
\def\redbcg{$\Lambda \ge 30$}
\def\nonbcg{$\Lambda < 30$}
\def\logms{$\log (M_{\star}/M_{\odot})$}
\def\logmh{$\log (M_{\mathrm{Halo}}/M_{\odot})$}

% Commenting:
\newcommand{\todo}[1]{\textcolor{red}{\textbf{TODO:~#1}}}
\newcommand{\plan}[1]{\textcolor{cyan}{#1}}
\newcommand{\addref}{{\textcolor{red}{REF}}}
\newcommand{\note}[2]{\textcolor{blue}{\textbf{[Comment (#1): #2]}}}
\newcommand{\song}[1]{\textcolor{magenta}{\textbf{[Song: #1]}}}
\newcommand{\alexie}[1]{\textcolor{blue}{\textbf{[Alexie: #1]}}}
\newcommand{\kevin}[1]{\textcolor{green}{\textbf{[Kevin: #1]}}}

%%%%%%%%%%%%: Header and Version %%%%%%%%%%%%

\slugcomment{Draft version 0}
\email{song.huang@ipmu.jp}
\shorttitle{ENVIRONMENT DEPENDENCE OF MASS-SIZE RELATION}
\shortauthors{HUANG ET AL.}

\begin{document}

\begin{CJK*}{UTF8}{gbsn}

%%%%%%%%%%%%: Title and Affiliations %%%%%%%%%%%%

\title{The Environmental Dependence of Structures for Massive Galaxies
from the Hyper Suprime-Cam Survey}

\author{Song Huang (黄崧)\altaffilmark{1} 
        Alexie Leauthaud\altaffilmark{1}, 
        Kevin Bundy\altaffilmark{1},
        Michael Strauss\altaffilmark{2},
        Yen-Ting Lin\altaffilmark{3},
        Rachel Mandelbaum\altaffilmark{4}
        }
\date{}

\altaffiltext{1}{Kavli Institute for the Physics and Mathematics of the 
    Universe, The University of Tokyo Institutes for Advanced Study, 
    the University of Tokyo (Kavli IPMU, WPI), Kashiwa 277--8583, Japan}
\altaffiltext{2}{Princeton University Observatory, 
    Peyton Hall, Princeton, NJ 08544, USA}
\altaffiltext{3}{Institute of Astronomy and Astrophysics, Academia Sinica, 
    P.O.~Box 23--141, Taipei 10617, Taiwan}
\altaffiltext{4}{Carnegie-Mellon University \todo{correct affiliation here}}

%%%%%%%%%%%%: Abstract and Keywords %%%%%%%%%%%%

\begin{abstract}
    \todo{Place holder; This is the abstract used for giving a talk.}\\
    Under the most popular formation scenario, the structures of massive central galaxies
    depend on environment at fixed stellar mass due to different assembly histories shaped
    by their host halos. Yet, clear evidence of such effect is still lacking. Using deep,
    multi-band images for a large sample of massive galaxies at 0.2 < z < 0.5 from the
    Hyper Suprime-Cam (HSC) survey, we discover subtle, but systematic structural
    difference for massive galaxies in low and high mass haloes. The differences are
    consistent with richer merger history in more massive halos. We show that the average
    profiles of mass, shape, and color, along with relations among masses within different
    radius (as proxy of mass assembled at different time) can help us gain more insights
    of their assembly history, and the weak lensing analysis enabled by HSC survey further
    helps us connect the differences we find to the average halo properties.
\end{abstract}
\keywords{galaxies: elliptical and lenticular, cD --- galaxies: formation --- 
          galaxies: photometry --- galaxies: structure --- galaxies: surveys}

\maketitle

%%%%%%%%%%%%: Main Text %%%%%%%%%%%%

%% ------------------------------------------------------------------------------------ %% 
\section{Introduction}

    \plan{Scientific Background}
    \begin{itemize}
        \item \plan{Massive galaxies are important cosmic probe and unique labs to study 
            galaxy evolution.} 
        \item \plan{Briefly explain why massive early-type galaxies are important through 
            the difficulties of stellar-halo mass relation and stellar mass function at the 
            high-mass end.}
        \item \plan{Brief summary of the current understanding of their cosmic assembly 
            history.}
        \item \plan{Explain why we should care about the environment, and why we expect to 
            see some environemtnal dependence in the structure and other properties of
            massive galaxies.}
        \item \plan{Brief review of the current observations.  It is still not clear 
            whether there is a clear environmental dependence.}
    \end{itemize}

    \plan{Observational difficulties (a.k.a Why we need HSC)}
    \begin{itemize}
        \item \plan{Explain why it is important to study the mass distribution of 
            massive galaxies out to large physical radius; and, given their unique light 
            profiles, why it is more difficult to study them compared to late-type 
            galaxies.}
        \item \plan{Very brief summary of past observational efforts, and why they are not 
            good enough (Not enough number of really massive galaxy; Shallow images; 
            Background subtraction issue; and stacking analysis can be dangerous as while)}
    \end{itemize}

    \plan{Basica idea of this work}
    \begin{itemize}
        \item \plan{Here we take advantages of the ambitious Hyper-Suprime camera 
            survey\ldots}
    \end{itemize}

    The paper is organized as follows. Section~2 gives a brief overview of the HSC
    observation and data reduction.  We will also summarize the process of sample
    selection.  In Section~3, we will describe the method for deriving the stellar mass 
    surface density profile.  The main results are summarized in Section~4.  Section~5 
    provides discussions of the assumptions used in this work, the potentially interesting
    physical implications, and several future improvements, ending with a summary in
    Section~6.

    All the magnitudes used here are in AB system (\addref), and are corrected 
    for Galactic extinction using calibrations from \citet{Schlafly11}.
    Within this work, we assume $H_0$ = 70~km~s$^{-1}$ Mpc$^{-1}$, ${\Omega}_m=0.3$, 
    and ${\Omega}_{\Lambda}=0.7$.

%% ------------------------------------------------------------------------------------ %%

\section{Observations and Sample Selection}

\subsection{The Hyper Suprime-Cam Survey}

    The Subaru Strategic Program (SSP, \addref) makes use of the new prime-focus camera,
    the Hyper Suprime-Cam (HSC;~\addref: Miyazaki\etal 2012), on the 8.2-m Subaru
    telescope at Mauna Kea.  
    Taking advantage of the large field of view (FoV;~1.5 deg in diameter) of HSC, this
    ambitious multi-layer photometric survey will cover $\sim 1400$ deg$^2$ of sky in 5
    broad bands (\textit{g r i z Y}) to the depth of $r \sim 26$ mag in the \texttt{WIDE}
    part in the next few years.  
    This work is based on the internal data release \texttt{S15B}, which covers $\sim 100$
    square of degree of sky in all 5-band to the required depth of \texttt{WIDE} survey.  
    The regions covered by this release are overlapped with several previous spectroscopic
    survey (e.g. SDSS/BOSS \addref; GAMA \addref).

    The data are processed with \texttt{hscPipe 4.0.1}, a derivative of the Large Synoptic
    Survey Telescope (LSST) pipeline (\addref: Ivezic\etal~2008, Axelrod\etal 2010),
    modified for use with Suprime-Cam and Hyper Suprime-Cam.  
    \texttt{hscPipe} first bias subtract, flat field, model background, and perform object
    detection and measurement on the single exposure data. 
    Then, different exposures are warped onto a common World Coordinate System (WCS) and
    combined into final images with improved signal-to-noise ratio (SNR) after astrometric
    and photometric calibration.  
    The pixel scale of the combined images is 0.168$\arcsec$.  The photometric calibration
    is based on data obtained from the Panoramic Survey Telescope and Rapid Response
    System (Pan-STARRS) 1 imaging survey (\addref: Schlafly\etal 2012, Tonry\etal 2012,
    Magnier\etal 2013). 
    To achieve consistent deblending and photometry across all bands, the \texttt{hscPipe}
    will perform multi-band post-processing on the combined images.  
    The footprints and peaks of detected sources on each band will be merged into a single
    catalog.    
    This consistent set of peaks and footprints is used as starting point for deblend and
    measure objects on the combined images of each band.  
    These measurements are then merged into a reference catalog.  After fixing the
    centroids, shape, and other non-amplitude parameters of every object in this catalog,
    \texttt{hscPipe} will derive forced photometry measurements at each band. 
    Details of \texttt{hscPipe} and the multi-band processing method will be presented in
    \addref (Bosch\etal 2017?).

    In Figure 1, we compare the false color ({\tt gri}) images of three nearby massive
    ETGs from both SDSS and HSC surveys.  
    The images are generated using similar scaling and color schemes, and clearly
    demonstrate HSC's capability to reach to low surface brightness domain.  
    In $i$-band (ignoring the slight difference in filter response curve), the HSC
    \texttt{WIDE} image is $2.5$-$3.0$ magnitude deeper than SDSS on average.  
    This gives us huge advantage in exploring the outskirt of massive ETGs given their
    extended, shallow outer surface brightness profiles.

    Motivated by the requirement of weak lensing analysis, the $i$-band data typically has
    the best seeing in all five bands (the median seeing is around FWHM$\sim 0.8\arcsec$).
    Therefore, we will mostly use the $i$-band images to study the structure of galaxies.

\subsection{Massive Central Galaxies in Different Environments}

    The main scientific goal of this work is to investigate mass assembly history of
    massive, central ETGs, and its dependence on environments (or host dark halo mass)
    through their the spatial distribution of stellar mass.

    To achieve this goal, we will select \logms$ > 11.5$ massive central galaxies within
    host halo mass larger and smaller than \logmh$= 14.0$ at $0.2 < z < 0.5$.  
    As both $M_{\star}$-$M_{\mathrm{Halo}}$ relation and stellar mass function are still
    quite uncertain at \logms$ > 11.5$, it is of great interest to investigate the
    structure of galaxies in this region carefully.
    And, this stellar mass cut can greatly reduce the contamination of satellites (e.g.
    \addref: van~Uitert\etal 2016).  
    Meanwhile, under the adopted cosmology, $1\arcsec$ equals 3.3 and 6.1 kpc at redshit
    0.2 and 0.5.  
    Hence, our redshift bin enables us to reliable measure the total stellar mass within
    the inner 5 to 10 kpc (where the ``in-situ'' component should still dominates) of
    massive galaxies at the high redshift end assuming the typical seeing of data. 
    At the same time, the depth of the data still allows to study the very outskirt ($\sim
    100$ kpc) of these galaxies at $z\sim 0.5$.  Also, such limits on mass and redshift
    ensure us a sizable sample while let us safely ignore any significant structural
    evolution (no star formation, lower merger rate \etal e.g. \addref).

    Based on recent constraints of $M_{\star}$-$M_{\mathrm{Halo}}$ relation (e.g.
    \addref), above \logms$ = 11.5$, there is a large scatter of halo mass at fixed
    stellar mass.  
    At \logmh$<14.0$, these massive galaxies are mostly the centrals of small groups;
    while at \logmh$>14.0$, they start to become the centrals of very massive groups and
    galaxy clusters.   
    Limited by the sample size and the capability of measureing halo mass for individual
    central galaxy, we will simply separate the sample into two broad halo mass bins with
    the help of the \redm cluster catalog\footnote{See:
    http://risa.stanford.edu/redmapper/} ( \texttt{v5.10}, e.g. \addref: Rykoff\etal 2014;
    Rozo\etal 2015b).
    These clusters are selected from the SDSS DR8 photometric data using overdensity of
    red-sequence galaxies. 
    For each cluster, the catalog provides robust estimations of photometric redshift
    $z_{\Lambda}$ and richness $\Lambda$, along with the best candidate of the central
    galaxy (the one with the highest central probability \texttt{P\_CEN}).
    Information about the identified member candidates is also provided separately. 
    Please see Rozo\etal (2014, 2015a, 2015b) for more details about the performance of
    the \redm cluster catalog.      
    Several works (e.g. Li\etal 2015; Saro\etal 2015; Farahi\etal 2016; Simet\etal 2016)
    have tried to calibrate the $M_{200c}-\Lambda$ relation using different methods.
    Despite the slightly different calibrations derived, it is safe to assume that most
    clusters identified by \redm ($\Lambda > 20$) have $\log (M_{200c}/M_{\odot}) \geq
    14.0$.  
    Considering the typical uncertainty of richness, and the fact that \redm catalog
    starts to become incomplete toward low richness ($\Lambda < 40$) end at $z > 0.33$, we
    will focus on the $\Lambda > 30$ clusters in this work.  
    Based on the calibrations from Farahi\etal (2016) or Simet\etal (2016), 
    such richness cut gives us haloes with 
    $M_{200c} \geq 1.56\pm0.35 \times 10^{14} M_{\odot}$ or 
    $M_{200c} \geq 1.60\pm0.11 \times 10^{14} M_{\odot}$.  
    Therefore, \redm can provide us a sample of massive central galaxies in \logmh$\geq
    14.0$ haloes.  
    And, we assume that, in the same footprints and redshift range, the \logms$\geq 11.5$
    galaxies outside \redm clusters are most likely to be the central of \logmh$< 14.0$
    haloes.

    According to Leauthaud\etal (2016), all $0.2 < z < 0.5$ and \logms$\geq 11.5$ galaxies
    should have $i_{\mathrm{SDSS, cModel}} \leq 21.0$ mag.  Ignoring the tiny difference
    in response curves between SDSS and HSC $i$-band filters, we started by selecting all
    galaxies with $i_{\mathrm{HSC, cModel}} \leq 21.5$ in the regions that are covered in
    all five filters, and have already reached the expected depth of \texttt{WIDE} survey.  
    %% The SQL search is saved as "sample.sql" 
    %% The master catalog used here is "dr1_wide_galaxy_icmodel_21.5.fits"
    %%   The original catalogs contain 2275477 objects
    The details of the HSC \texttt{cModel} photometry, and its performence will be
    described in \addref Bosch\etal in.\ prep. 
    It is quite similar to the SDSS \texttt{cModel} in principle as it also fits the total
    flux of an object using a combination of de~Vaucouleur and exponential components
    after the PSF convolution is considered.  
    For this work, we select objects that are classified as extended, have well defined 
    centroids, do not fail during the deblending process, and have \texttt{cModel} 
    magnitudes in all five bands. 
    \song{Should we discuss the impact of failed cModel photometry?} 
    A series of quality control cuts are applied to remove objects that are affected by 
    saturation, cosmic ray, other optical artefacts; and the ones that are too close to 
    the edge of the field in all five bands 
    (each criterion affects less than 8\% of the entire sample).
    After these cuts, a total number of 1760845 galaxies left in the sample.
    This sample will be referred as \texttt{hscPho}. 
    %% Saved as "dr1_wide_galaxy_hscPho.fits"

    We match this sample with the central galaxies of \redm catalogs using a $1.0\arcsec$
    radius, and it results in 704 galaxies.  
    %% Saved as "dr1_redbcg_hsc_sdss_gama_1arcsec.fits"
    A small fraction of \redm centrals in the our footprints do not have matched object in
    our sample due to severe contamination from optical artefacts or bleeding trails from
    saturated objects.  
    And slight change of the mathcing radius has no impact of this sample.  
    For galaxy in this sample, the photometric redshift of the cluster is available for
    each of them; and for 72\% of them, the spectroscopic redshift from SDSS DR8 is also
    provided by \redm catalog when available. 
    This sample of will be referred as \texttt{redBCG} in the rest of the paper.

    For the vast majority of galaxies that are not the centrals of very massive haloes, we
    need other sources for their redshfit estimations. 
    Although there has been many efforts for deriving photometric redshift (photo-$z$)
    using HSC five-band photometry, it is still a working progress now, and its
    performence at $z \leq 0.5$ is expected to be less satisfying due to the lack of
    information at the shorter wavelength side of the 4000$\AA$ break.  
    Therefore, we will only rely on external spectroscopic redshifts (spec-$z$) for this
    work.  
    The HSC database compiles a catalog of external spec-$z$ for this data release by
    matching the detected objects with public data of several spectroscopic surveys (e.g.
    SDSS/BOSS \addref; GAMA \addref; PRIMUS \addref; \etal). 
    Duplicated matches from different sources are merged through interal matching using
    $0.5\arcsec$ radius. 
    For each object, the quality information of the spec-$z$ from different catalogs are 
    homogenized into a single flag that indicates whether the redshfit is secure, and 
    only secure spec-$z$ are used in this work. 
    %% This is catalog: "dr1_specz_use.fits"

    We match the \texttt{hscPho} sample with this external spec-$z$ catalog using a 
    $1.0 \arcsec$ radius, and it leads to 116813 matched objects.  
    %% Saved as "dr1_wide_galaxy_icmodel_21.5.fits"
    Among these matches, there are 42696 objects at $0.2 \leq z \leq 0.5$. 
    Most of these redshifts come from either SDSS or GAMA survey.
    To further exclude contamination from satellite galaxies, we aggressively remove 
    all galaxies that could be associated with the \redm clusters from this catalog.  
    First, the catalog of \texttt{redBCG} central galaxies is merged with the catalog  
    of their member galaxies. 
    Then, through cross matching using $1.0 \arcsec$ radius, all \texttt{hscPho} galaxies 
    that are within 2.5 Mpc of projected distance to any of cluster member are removed 
    from the catalog.  
    After this step, 29973 galaxies are left in the sample. 
    We will refer this sample as \texttt{nonBCG} later.
    %% Saved as "dr1_nonbcg_use_sed5b.fits"

    The \texttt{redBCG} sample is also matched with external spec-$z$ catalog in the same
    way, and we update the redshift of the central galaxies with external spec-$z$ when 
    it is available.  
    For the unmatched ones, only 3 of them have spec-$z$ provided by the \redm catalog 
    (most likely due to different criteria during the selection of spec-$z$ sample).
    For the rest, we will use the photo-$z$ of the cluster from the \redm catalog instead.
    At $0.2 \leq z \leq 0.5$, there are 375 central galaxies left in the \texttt{redBCG} 
    sample. 
    Among them, 67 (17.9\%) of them do not have spec-$z$.
    Using the central galaxis have both spec- and photo-$z$, we estimate the median 
    absolute redshift difference $|z_{\mathrm{Photo}} - z_{\mathrm{Spec}}|$ is 
    $\sim 0.01$ (only 12 have absolute difference larger than 0.05; and the largest 
    difference is $\sim 0.12$).
    Such quality is good enough for the purpose of this work. 
    %% Saved as "dr1_redbcg_use_sed5b.fits"

\subsection{Stellar Masses from SED Fitting}

    To further select galaxies that are truly very massive central galaxies, we use 
    broadband Spectral Engergy Distributions (SEDs) fitting method (see \addref 
    Walcher\etal 2011 for a recent review) to derive stellar mass
    of our samples, and provide $k$-corrected photometry.

\subsection{Galaxies in $\Lambda < 20$ haloes}

    \begin{itemize}
        \item \plan{Remove the galaxy that is close to any \redm cluster}
        \item \plan{Remove the galaxy that is close to any bright star} 
            \todo{Justify why we do not do this to the \redm central sample}
    \end{itemize}

%% ------------------------------------------------------------------------------------ %%

\section{Data Reduction}

\subsection{1-D surface brightness profile}

    \plan{Basica ideas}
    \begin{itemize}
        \item \plan{Briefly explain that 1-D surface brightness profile is a very old 
            fashion, but robust and straightforward way to describe the light or mass 
            distribution of a galaxy.  Although it is not exactly ``model-independent'', 
            it indeed has the advantages of not easily affected by complex substrctures 
            within the galaxy comparing with the more popular 2-D modelling method. }
        \item \plan{Due to the uniquely extended nature of the light distributions of 
            these massive ETGs, it is still unclear which is the most appropriate 2-D 
            model for them (Can not be well described by 1-\ser model). }
    \end{itemize}

    \plan{Step by step description}
    \begin{itemize}
        \item \plan{Generation of cutout image in multiple bands}: \\
            \begin{enumerate}
                \item \plan{Choice of image size: make sure that the cutout covers out 
                    to at least 500 kpc from the center of the galaxy.}
                \item \plan{Also generate the bad pixel masks, and the reconstructed 
                    PSF model for the galaxy center.}
            \end{enumerate}
        \item \plan{Generation of object masks}: \\
            \begin{enumerate}
                \item \plan{Briefly explain why we need to our own photometry instead 
                    of just using the cModel results from the HSC pipeline.}
                \item \plan{We need two kinds of masks: mask of all objects that will 
                    be used to measure the sky value; mask of objects expect for the 
                    central galaxy we want to derive surface brightness profile.  
                    Briefly describe their requirements.}
                \item \plan{Object detection using \texttt{SEP} Python library.}
            \end{enumerate}
        \item \textbf{Re-measurement and correction of sky background}:
    \end{itemize}

\subsection{Average mass-to-light ratio}

    \begin{itemize}
        \item \plan{Briefly explain the method to convert surface brightness profiles 
            into mass density profile, focusing on the key assumptions.}
        \item \plan{Describe the SED fitting process.}    
        \item \plan{Briefly discuss the impact of choices of SSP model, IMF, priors of 
            parameters for SED fitting}
    \end{itemize}

\subsection{Basic properties of the sample}

    \begin{itemize}
        \item \plan{Compare stellar mass from integration of 1-D profile and the 
            \texttt{cModel} photometry}
        \item \plan{Stellar mass, redshift, K-corrected color distributions.}
        \item \plan{Red sequence properties}    
        \item \plan{Potential ``contaminations'' from disc galaxies}
    \end{itemize}

    \todo{Justify why we can assume most very massive galaxies in \nonbcg haloes are
        central galaxies}

%% ------------------------------------------------------------------------------------ %%

\section{Results}

\subsection{Impact of deep photometry on the luminosity and stellar mass function}

\subsection{Comparison of stellar mass density profiles}

    \plan{Brief summary of the process of sample matching}

\subsubsection{Using stellar mass derived from the GAMA survey}

\subsubsection{Using stellar mass within 10 kpc aperture}

\subsubsection{Using stellar mass within 100 kpc aperture}

\subsection{Relations between stellar mass within different physical apertures}

\subsection{$M_{\ast}-R_{\mathrm{50}}$ relation}

%% ------------------------------------------------------------------------------------ %%

\section{Discussion}

\subsection{Implications on assembly history of massive galaxies}

    \plan{Comparison of mass profiles with previous studies at high redsfhit}

\subsection{Ellipticity profiles}

\subsection{The impact of $M/L$ gradient}

\subsection{Massive satellites of the $\Lambda > 20$ haloes}
\song{TBD: optional}

\subsection{Connection and difference with the Intra-Cluster Light}

\subsection{Future improvements}

    \begin{enumerate}
        \item \plan{Improvements of HSC data reduction: better sky modelling; 
            \texttt{redMaPPer} clusters using HSC photometry.}
        \item \plan{Comparison with 2-D image modelling method.}
        \item \plan{``Correct'' the PSF smeearing effect at the center with the help of 
            residual correct 2-D model.}
    \end{enumerate}

%% ------------------------------------------------------------------------------------ %%

\section{Summary}

%% ------------------------------------------------------------------------------------ %%

\acknowledgements

  The Hyper Suprime-Cam (HSC) collaboration includes the astronomical communities of Japan
  and Taiwan, and Princeton University.  The HSC instrumentation and software were
  developed by the National Astronomical Observatory of Japan (NAOJ), the Kavli Institute
  for the Physics and Mathematics of the Universe (Kavli IPMU), the University of Tokyo,
  the High Energy Accelerator Research Organization (KEK), the Academia Sinica Institute
  for Astronomy and Astrophysics in Taiwan (ASIAA), and Princeton University.  Funding was
  contributed by the FIRST program from Japanese Cabinet Office, the Ministry of
  Education, Culture, Sports, Science and Technology (MEXT), the Japan Society for the
  Promotion of Science (JSPS),  Japan Science and Technology Agency (JST),  the Toray
  Science  Foundation, NAOJ, Kavli IPMU, KEK, ASIAA,  and Princeton University.

  Funding for SDSS-III has been provided by the Alfred P. Sloan Foundation, the
  Participating Institutions, the National Science Foundation, and the U.S.  Department of
  Energy. The SDSS-III web site is http://www.sdss3.org.  SDSS-III is managed by the
  Astrophysical Research Consortium for the Participating Institutions of the SDSS-III
  Collaboration including the University of Arizona, the Brazilian Participation Group,
  Brookhaven National Laboratory, University of Cambridge, University of Florida, the
  French Participation Group, the German Participation Group, the Instituto de Astrofisica
  de Canarias, the Michigan State/Notre Dame/JINA Participation Group, Johns Hopkins
  University, Lawrence Berkeley National Laboratory, Max Planck Institute for
  Astrophysics, New Mexico State University, New York University, Ohio State University,
  Pennsylvania State University, University of Portsmouth, Princeton University, the
  Spanish Participation Group, University of Tokyo, University of Utah, Vanderbilt
  University, University of Virginia, University of Washington, and Yale University.

  \todo{Full acknowledgement}\\
  \begin{itemize}
      \item Acknowledgements for Kevin and Alexie's funding 
      \item Acknowledgements for the Python libraries
  \end{itemize}

  This research made use of 
  \href{http://www.astropy.org/}{Astropy}, a community-developed 
  core Python package for Astronomy (Astropy Collaboration, 2013; 
  \href{https://github.com/kbarbary/sep}{sep} 
  Source Extraction and Photometry in Python \citep{PythonSEP}.

%%%%%%%%%%: Bibligraphic Section %%%%%%%%%%

\bibliography{redbcg}

%%%%%%%%%%: Appendix Section %%%%%%%%%%%%

\appendix

\section{A. Extraction of 1-D surface brightness profile}

\section{B. Derive average mass-to-light ratio using \texttt{iSEDFit}}

%%%%%%%%%%%%: Figures Section %%%%%%%%%%%%

% Fig. 1
\clearpage
\figurenum{1}
\begin{figure}
    \centering 
    \includegraphics[width=15.5cm]{fig/fig1.png}
    \caption{Figure.1\todo{Caption}}
    \label{figure:1}
\end{figure}

% Fig. 2
\clearpage
\figurenum{2}
\begin{figure}
    \centering 
    \includegraphics[width=13.5cm]{fig/massive_s15b_g15_sample.png}
    \caption{Figure.2\todo{Caption}}
    \label{figure:2}
\end{figure}

% Fig. 3
\clearpage
\figurenum{3}
\begin{figure}
    \centering 
    \includegraphics[width=13.5cm]{fig/massive_s15b_zuse_hist1.png}
    \caption{Figure.3\todo{Caption}}
    \label{figure:3}
\end{figure}

% Fig. 4
\clearpage
\figurenum{4}
\begin{figure}
    \centering 
    \includegraphics[width=15.5cm]{fig/redbcg_1_HSC-I_full_imgsub_ellip_default_sum.png}
    \caption{Figure.4\todo{Caption}}
    \label{figure:4}
\end{figure}

% Fig. 5
\clearpage
\figurenum{5}
\begin{figure}
    \centering 
    \includegraphics[width=14.0cm]{fig/redbcg_1_HSC-I_full_imgsub_ellip_default_compare.png}
    \caption{Figure.5\todo{Caption}}
    \label{figure:5}
\end{figure}

%%%%%%%%%%%%: End of the File %%%%%%%%%%%%

\end{CJK*}

\clearpage

%%%%%%%%%%%: Possible Tables %%%%%%%%%%%%%
%\input{table1.tex}

\label{lastpage}
\end{document}