### 2016-12-04 draft/ms_v0.tex
version 0

% Version 0; preprint format; Written by SH
% Version 1; emulateapj format; Edited by SH and AL

%\documentclass[twocolumn]{aastex6}
\documentclass{emulateapj}

% Packages
%\usepackage{emulateapj5}
\usepackage{apjfonts}
\usepackage{amssymb, amsmath}
\usepackage{graphicx}
\usepackage{CJK}
\usepackage{natbib}
\usepackage{hyperref}
\usepackage{float}
\usepackage[usenames]{color}

% Package Settings
\bibliographystyle{apj}
\hypersetup{colorlinks=true,
            citecolor=cyan,
            linkcolor=cyan,
            filecolor=magenta,      
            urlcolor=cyan}
\urlstyle{same}

% Figure extention
\DeclareGraphicsExtensions{.pdf,.png,.jpg}
\input psfig.tex

%%%%%%%%%%%%: User Defined Commands %%%%%%%%%%%%

% Song Huang's definition 
\def\arcsec{{\prime\prime}}
\def\arcmin{{\prime}}
\def\degree{{\circ}}
\def\h{\hskip -3 mm}
\def\aa{{A\&A}}
\def\aas{{ A\&AS}}
\def\aj{{AJ}}
\def\al{$\alpha$}
\def\bet{$\beta$}
\def\amin{$^\prime$}
\def\annrev{{ARA\&A}}
\def\apj{{ApJ}}
\def\apjs{{ApJS}}
\def\asec{$^{\prime\prime}$}
\def\deg{$^{\circ}$}
\def\ddeg{{\rlap.}$^{\circ}$}
\def\dsec{{\rlap.}$^{\prime\prime}$}
\def\cc{cm$^{-3}$}
\def\flamb{erg s$^{-1}$ cm$^{-2}$ \AA$^{-1}$}
\def\flux{erg s$^{-1}$ cm$^{-2}$}
\def\fnu{erg s$^{-1}$ cm$^{-2}$ Hz$^{-1}$}
\def\hst{{\textit{HST}}}
\def\kms{km s$^{-1}$}
\def\lamb{$\lambda$}
\def\lax{{$\mathrel{\hbox{\rlap{\hbox{\lower4pt\hbox{$\sim$}}}\hbox{$<$}}}$}}
\def\gax{{$\mathrel{\hbox{\rlap{\hbox{\lower4pt\hbox{$\sim$}}}\hbox{$>$}}}$}}
\def\simlt{\lower.5ex\hbox{$\; \buildrel < \over \sim \;$}}
\def\simgt{\lower.5ex\hbox{$\; \buildrel > \over \sim \;$}}
\def\micron{{$\mu$m}}
\def\mnras{{MNRAS}}
\def\nat{{Nature}}
\def\pasp{{PASP}}
\def\perang{\AA$^{-1}$}
\def\peryr{yr$^{-1}$}
\def\reference{\noindent\pp}
\def\refindent{\par\noindent\parskip=2pt\hangindent=3pc\hangafter=1 }
\def\sb{mag~arcsec$^{-2}$}
\def\lsun{$L_\odot$}
\def\msun{$M_\odot$}
\def\sigs{$\sigma_*$}
\newcommand{\lt}{<}
\newcommand{\gt}{>}

\def\etal{{\ et al.~}}
\def\galfit{{\tt GALFIT}}
\def\ser{{S\'{e}rsic\ }}
\def\redm{\texttt{redMaPPer}}
\def\cmodel{\texttt{cModel}}
% Samples
\def\rbcg{\texttt{cenHighMh}}
\def\nbcg{\texttt{cenLowMh}}
\def\redbcg{{$\lambda \ge 30$}}
\def\nonbcg{{$\lambda < 20$}}
% Mass related 
\def\mstar{{$M_{\star}$}}
\def\mhalo{{$M_{\mathrm{halo}}$}}
\def\logms{{$\log (M_{\star}/M_{\odot})$}}
\def\logmh{{$\log (M_{\mathrm{Halo}}/M_{\odot})$}}

\def\minn{{$M_{\star,10\mathrm{kpc}}$}}
\def\mtot{{$M_{\star,100\mathrm{kpc}}$}}
\def\mout{{$M_{\star,150\mathrm{kpc}}$}}
\def\mgama{{$M_{\star,\mathrm{GAMA}}$}}
\def\mcmodel{{$M_{\star,\mathrm{cModel}}$}}

\def\logminn{{$\log (M_{\star,10\mathrm{kpc}}/M_{\odot})$}}
\def\logmtot{{$\log (M_{\star,100\mathrm{kpc}}/M_{\odot})$}}
\def\logmout{{$\log (M_{\star,150\mathrm{kpc}}/M_{\odot})$}}
\def\logmgama{{$\log (M_{\star,mathrm{GAMA}}/M_{\odot})$}}
\def\logmcmodel{{$\log (M_{\star,mathrm{cModel}}/M_{\odot})$}}

\def\m2l{{$M_{\star}/L_{\star}$}}
\def\mden{{$\mu_{\star}$}}

% Commenting:
\newcommand{\xxx}[1]{\textcolor{red}{\textbf{XXX}}}
\newcommand{\todo}[1]{\textcolor{red}{\textbf{TODO:~#1}}}
\newcommand{\plan}[1]{\textcolor{cyan}{#1}}
\newcommand{\addref}{{\textcolor{red}{REF}}}
\newcommand{\note}[2]{\textcolor{blue}{\textbf{[Comment (#1): #2]}}}
\newcommand{\song}[1]{\textcolor{magenta}{\textbf{[Song: #1]}}}
\newcommand{\alexie}[1]{\textcolor{blue}{\textbf{[Alexie: #1]}}}
\newcommand{\kevin}[1]{\textcolor{green}{\textbf{[Kevin: #1]}}}

%%%%%%%%%%%%: Header and Version %%%%%%%%%%%%

\shorttitle{ENVIRONMENT DEPENDENCE OF MASS-SIZE RELATION}
\shortauthors{HUANG ET AL.}
\slugcomment{Draft version 0}

\begin{document}

\begin{CJK*}{UTF8}{gbsn}

%%%%%%%%%%%%: Title and Affiliations %%%%%%%%%%%%

%\title{The Environment and Structure of Massive Central Galaxies
%       using the Subaru Hyper Suprime-Cam Survey}

\title{First Detection of the Enviromental Dependence of the Masses, Sizes, and Structures of Massive Galaxies }

\author{Song Huang (黄崧)\altaffilmark{1}, 
        HSC collaboration\altaffilmark{1}}

%        \author{Song Huang (黄崧)\altaffilmark{1} 
%        Alexie Leauthaud\altaffilmark{1}, 
%        Kevin Bundy\altaffilmark{1},
%        Michael Strauss\altaffilmark{2},
%        Yen-Ting Lin\altaffilmark{3},
%        Rachel Mandelbaum\altaffilmark{4}
%        }

\email{song.huang@ipmu.jp}

\date{}

\altaffiltext{1}{Kavli Institute for the Physics and Mathematics of the 
    Universe, The University of Tokyo Institutes for Advanced Study, 
    the University of Tokyo (Kavli IPMU, WPI), Kashiwa 277--8583, Japan}

%\altaffiltext{2}{Princeton University Observatory, 
%    Peyton Hall, Princeton, NJ 08544, USA}
%\altaffiltext{3}{Institute of Astronomy and Astrophysics, Academia Sinica, 
%    P.O.~Box 23--141, Taipei 10617, Taiwan}
%\altaffiltext{4}{Carnegie-Mellon University \todo{correct affiliation here}}

%%%%%%%%%%%%: Abstract and Keywords %%%%%%%%%%%%

\begin{abstract}

    Although the environmental dependence of structures for massive central 
    galaxies is predicted by the promising hierarchical assembly model, 
    observations at low redshift seem to find no convincing evidence for that. 
    With the help of deep $i$-band images of a large sample of massive central 
    galaxies at $0.3 < z < 0.5$ from the Subaru Hyper Suprime-Cam (HSC) survey, 
    we map their stellar mass distributions out to radius larger than $100$ kpc, 
    and discover subtle, but systematic and robust structural differences that 
    depend on halo mass. 
    At fixed stellar mass within 100 kpc, the massive central galaxies in more 
    massive ($M_{200m} > 1.6\times 10^{14} M_{\odot}$) halos have a 
    slightly flattened inner profile within $\sim 15$-$20$ kpc, and a more prominent 
    outer envelope compared to ones in less massive 
    ($M_{200m}>8.7\times 10^{13} M_{\odot}$) halos.
    For centrals with $M_{\star} > 10^{11.5} M_{\odot}$, the ones in more massive 
    halos show very significant excess of mass in the outskirt when the two samples 
    are matched using proxies of mass assembled at $z > 1$. 
    Such differences are broadly consistent with richer recent merging history 
    for more massive halos. 
    We suggest that the relation between total stellar mass and mass within inner 5 
    or 10 kpc is potentially interesting for diagnosing the role played by host halo 
    in shaping the structures of massive central galaxies. 
    These results also highlight the importance of deep photometry and the usage of 
    detailed structural information in the study of the assembly history of galaxies.

\end{abstract}
\keywords{galaxies: elliptical and lenticular, cD --- galaxies: formation --- 
          galaxies: photometry --- galaxies: structure --- galaxies: surveys}

\maketitle

%%%%%%%%%%%%: Main Text %%%%%%%%%%%%

%% ------------------------------------------------------------------------------------ %% 
    \begin{figure*}[t!]
        \centering 
        \includegraphics[width=\textwidth]{fig/redbcg_sdss_compare}
        \caption{A comparison between the imaging quality of SDSS and HSC Wide for a sample 
        	of nearby massive elliptical galaxies at $0.2 < z < 0.5$.  
            These images are generated using $gri$ band images with an arcsinh stretch 
            \citep{Lupton2004}. 
            The HSC Wide layer is $3.0$-$4.0$ magnitudes deeper than SDSS which is 
            critical in order to map the outskirts of ETGs out to large radii.}
        \label{fig:sdss_compare}
    \end{figure*}
%% ------------------------------------------------------------------------------------ %%

%% ------------------------------------------------------------------------------------ %% 
\section{Introduction}
    \label{sec:intro}

    \plan{Scientific Background}
    \begin{itemize}
        \item \plan{Massive galaxies are important cosmic probe and unique labs to study 
            galaxy evolution.} 
        \item \plan{Briefly explain why massive early-type galaxies are important through 
            the difficulties of stellar-halo mass relation and stellar mass function at the 
            high-mass end.}
        \item \plan{Brief summary of the current understanding of their cosmic assembly 
            history.}
        \item \plan{Explain why we should care about the environment, and why we expect to 
            see some environmental dependence in the structure and other properties of
            massive galaxies.}
        \item \plan{Brief review of the current observations.  It is still not clear 
            whether there is a clear environmental dependence.}
    \end{itemize}

    \plan{Observational difficulties (a.k.a Why we need HSC)}
    \begin{itemize}
        \item \plan{Explain why it is important to study the mass distribution of 
            massive galaxies out to large physical radius; and, given their unique light 
            profiles, why it is more difficult to study them compared to late-type 
            galaxies.}
        \item \plan{Very brief summary of past observational efforts, and why they are not 
            good enough (Not enough number of really massive galaxy; Shallow images; 
            Background subtraction issue; and stacking analysis can be dangerous as while)}
    \end{itemize}

    \plan{Basic idea of this work}
    \begin{itemize}
        \item \plan{Here we take advantages of the ambitious Hyper-Suprime camera 
            survey\ldots}
    \end{itemize}

\alexie{Because the text is long, at this point in the introduction, let's give reader an overview of our methodology. Let's sim for one or two succint paragraphs. That way they will begin the paper having a better idea of what we want to achieve. We will want to give them some key ideas here. 1) Quality of the HSC data (Reference Fig 1 here) 2) we want to look for enviromental differences, we'll use  a cluster catalog to divide into halo mass bins, 3) define the concepts of M10 and M100 (and reference Fig 10 here). You sort of have the elements in place below, by needs more editing and reorganizing.}

\alexie{Bring Fig 1 and Fig 10 here so they are the two first figures.}

	\song{Ok, will do; I will just say Fig 10 demonstrate the stellar mass and structural 
    	diversity of the massive galaxies}

    To achieve this goal, we select a large sample of massive central galaxies within
    halos that are more or less massive than \logmh{}$\sim 14.0$ at $0.3 < z < 0.5$ 
    from the current $\sim 100$ square degree of HSC data.
    As indicated by the $M_{\star}$-$M_{\mathrm{Halo}}$ relation, at the same \mstar{}, 
    massive central galaxies with similar morphology can live in halos within large range 
    of \mhalo{}, therefore it is very interesting to look into the influence of their halos 
    on the structure of these galaxies. 
    Under the adopted cosmology, $1.0^{\arcsec}$ equals 4.4 and 6.1 kpc at redshift 0.3 
    and 0.5.  
    In this redshift range, we can reliably measure the total stellar mass within the 
    inner 5 to 10 kpc (where the ``in-situ'' stellar component should dominate) given 
    the typical seeing, while still be able to map their stellar distributions out to 
    $\sim 100$ kpc at $z\sim 0.5$ given the imaging depth of these HSC data.
    Also, we can safely ignore significant mass growth and structural evolution 
    (no star formation, lower merger rate \etal~e.g.\ \citealt{Bellstedt2016},
    \citealt{Inagaki2015}; but also see \citealt{Bai2014}) 
    with our redshift and \mstar{} ranges.

    The paper is organized as follows. Section\ref{sec:data} gives a brief overview of 
    the HSC observation and data reduction, along with the summary of the data selection
    processes.  
    Then, we will explain the procedures for extracting 1-D surface brightness profiles 
    (Section \ref{sec:ellipse}) and estimating stellar mass (Section \ref{sec:mstar}) in
    details.   
    We will present the main results in Section \ref{sec:result}, and discuss the several 
    related technical and scientific issues in Section \ref{sec:discussion}, ending with
    summary and future plans in Section \ref{sec:summary}.

    All the magnitudes used here are in AB system (\citealt{Oke1983}), and are corrected 
    for Galactic extinction using calibrations from \citet{Schlafly11}.
    In this work, we assume $H_0$ = 70~km~s$^{-1}$ Mpc$^{-1}$, ${\Omega}_m=0.3$, 
    and ${\Omega}_{\lambda}=0.7$. 
%% ------------------------------------------------------------------------------------ %%

%% ------------------------------------------------------------------------------------ %% 
    \begin{figure*}[bt]
        \centering 
        \includegraphics[width=\textwidth]{fig/redbcg_ellipse_example}
        \caption{Left: example of the 1-D surface brightness and ellipticity profile 
            for a massive galaxy at $z=0.23$ in $i$-band using \texttt{Ellipse}. 
            In this work, we always show the radial profile in $R^{1/4}$ scale as it 
            is the most appropriate one to show the structural details of massive 
            ETGs at both inner and outer regions. 
            The dash-line shows the surface brightness profile after correcting the 
            background. 
            We also plot the brightness profile of the PSF model normalized at the 
            central surface brightness of the galaxy to highlight the region affected 
            by seeing. 
            On the top panel, the dash line shows the ellipticity used for the final 
            isophote.~~~ 
            Right: the masked $i$-band image of this galaxy with the isophotes extracted 
            by \texttt{Ellipse} overlaid. 
            The thick, green isophote highlights the one with $\mu_{i}\sim 28.5$~\sb.}
            \song{Will update completely, takes a little bit longer}
            %\alexie{Can we add physical radii at the top left since $R^{1/4}$ is not 
            %    very intuitive. On the right hand side, why show the background as blue, 
            %    instead of showing the actual image of the galaxy with masks overlaid? 
            %    Add units to x axis of right hand side. 
            %    Red text on left hand side, replace with "1D profile of a $10^{xx}$ 
            %    M$_{\odot}$ galaxy at $z=xx$}.}
        \label{fig:ellipse}
    \end{figure*}
%% ------------------------------------------------------------------------------------ %%

%% ------------------------------------------------------------------------------------ %% 
\section{Data And Sample Selection}
    \label{sec:data}

\subsection{The Hyper Suprime-Cam Survey}
    \label{ssec:hsc}

    The Subaru Strategic Program (SSP, \addref) makes use of the new prime-focus camera,
    the Hyper Suprime-Cam (HSC;~\citealt{Miyazaki2012}), on the 8.2-m Subaru telescope at 
    Mauna Kea. 
    This ambitious multi-layer photometric survey takes advantage of the large field of 
    view (FoV;~1.5 deg in diameter) of HSC and will cover $\sim 1400$ deg$^2$ of sky in 5
    broad bands (\textit{g r i z Y}) to the depth of $r \sim 26$ mag in the \texttt{WIDE}. 
    This work is based on the internal data release \texttt{S15B}, which covers 
    $\sim 100$square of degree of sky in all 5-band to full \texttt{WIDE} depth.  
    The regions covered by this release overlap with a number of spectroscopic surveys 
    (e.g.\ SDSS/BOSS: \citealt{Eisenstein2011}, \citealt{Alam2015}; 
    GAMA: \citealt{Driver2011}, \citealt{Liske2015}).

   The HSC \texttt{WIDE} survey is $3.0$-$4.0$ magnitudes deeper in the $i$-band 
    than SDSS. 
    Combined with the excellent imaging resolution (median $i$-band seeing is 0.6"), 
    and the wide area, this makes the HSC wide layer a tremendous data set to perform a 
    large statistical study of surface brightness profiles of ETGs out to large radii. 
    Fig.\ref{fig:sdss_compare} illustrated the quality of HSC imaging compared to SDSS 
    for a sample of low redshift ETGs. 
    Fig.\ref{fig:sdss_compare} clearly demonstrates that the HSC Wide survey is well 
    suited for mapping the stellar distribution of massive galaxies out to large radii 
    and will be a powerful data set to explore the assembly history of ETGs.

	The HSC $i$-band images typically have the best seeing in all five bands because of 
    strict requirements determined by weak lensing science. 
    We will therefore use the $i$-band images to measure the stellar distributions of 
    massive galaxies.

%% ------------------------------------------------------------------------------------ %% 
\subsection{The HSC galaxy catalog and photometry measurements}
    \label{sec:pipeline}

    The HSC SSP data are processed with \texttt{hscPipe 4.0.1}, a derivative of the Large 
    Synoptic Survey Telescope (LSST) pipeline (e.g.\ \citealt{Juric2015}; 
    \citealt{Axelrod2010}), modified for HSC. 
    \texttt{hscPipe} first performs a number of tasks at the single exposure level 
    (bias subtraction, flat fielding, background modeling, object detection and 
    measurements). 
    Then pipeline perform astrometric and photometric calibration for each single 
    expoures. 
    After that, the \texttt{hscPipe} warp different exposures on to a common
    World Coordinate System (WCS), combine them into coadded images with improved 
    signal-to-noise ratio ($S/N$), and update the images with better astrometric and 
    photometric calibrations at the same time using the common stars among different
    exposures.
    The pixel scale of the combined images is $0.168^{\arcsec}$.  
    The photometric calibration is based on data from the Panoramic Survey Telescope 
    and Rapid Response System (Pan-STARRS) 1 imaging survey 
    (\citealt{Schlafly2012}, \citealt{Tonry2012}, \citealt{Magnier2013}). 
    To achieve consistent deblending and photometry across all bands, \texttt{hscPipe} 
    performs multi-band post-processing at the coadd level. 
    First, the \texttt{hscPipe} performs object detection again on the coadd images
    in each band independently, and identify the above-threshold region (referred as 
    ``footprint'') and the flux peak within it for each source. 
    ``footprints'' and peaks from different bands are then merged together before 
    the \texttt{hscPipe} deblend and measure them in each band. 
    Later, \texttt{hscPipe} selects a ``reference band'' for each object based on the 
    $S/N$ in different bands (for most galaxies in this work, it is the $i$-band). 
    After fixing the centroids, shape, and other non-amplitude parameters of each 
    object in this reference catalog, \texttt{hscPipe} perform forced photometry 
    on the coadd image in each band.
    The PSF and galaxy model fluxes measured in the forced photometry approach are the
    best for color measurements.  
    Please refer to \citet{BoschInPrep} for more details of the \texttt{hscPipe} and 
    the multi-band processing method.

    The HSC \cmodel{} algorithm is similar to the SDSS \cmodel{} one.  
    It fits the flux distribution of an object using a combination of de~Vaucouleur and 
    exponential components after considering the PSF convolution.  
    For more details about the algorithm, please see Bosch\etal~in prep.~.
    We have tested its performance using synthetic objects (Huang\etal~in prep.~), and 
    the results indicate that, generally speaking, the HSC \cmodel{} photometry is 
    accurate down to $i >25.0$ mag.  
    However, for massive ETGs with extended stellar distributions, \cmodel{} currently 
    systematically underestimates their total flux.
    This problem indicates the intrinsic limitation of \cmodel{} as it is incapable of
    modeling profiles that are extremely extended in the outskirt at the depth of 
    HSC survey. 
    At the same time, issues with the deblender also worsen the situation. 
    As the image becomes much deeper, it also significantly increases the fraction 
    objects that are blended with others, and makes reliable deblending process 
    a very challenging problem especially for massive ETGs where satellites and 
    background galaxies often blend with the low surface brightness stellar envelope.
    To make sure the reliable detection of faint objects close to the detection limit, 
    the deblending method implemented by the \texttt{hscPipe} now tends to 
    ``over-deblend'' the surrounding areas of bright galaxies, and further results in 
    under-estimated total flux of massive ETGs (more discussion in 
    Bosch\etal~in prep.~).  
    For these reasons, we will perform customized photometric measurements to derive 
    more accurate total luminosity and stellar mass. 
    We only use the HSC \cmodel{} photometry in the initial selection of parent 
    sample.

%% ------------------------------------------------------------------------------------ %% 
\subsection{Initial Massive Galaxy Sample}
    \label{ssec:initial}

    Because of the caveats that apply to \texttt{hscPipe} measurements for bright 
    galaxies, we will use custom made software to measure the luminosity profiles of
    massive galaxies. 
    Our first step is use the HSC galaxy catalog to create an initial sample of massive
    galaxies that is complete to \xxx{} at $z < 0.5$. 
    We will then use custom made software to re-measure luminosities for all galaxies 
    in this sample. 
    Based on Leauthaud\etal~(2016), most \logms{}$\geq 11.5$ galaxies should have 
    $i_{\mathrm{SDSS, cModel}} \leq 21.0$ mag.
    So we first select all galaxies with $i_{\mathrm{HSC, cModel}} \leq 21.5$ 
    \footnote{We neglect the insignificant differences between the response curves of
    SDSS-$i$ and HSC-$i$ filters}
    in regions that have reached the required depth of \texttt{WIDE} survey in all 
    five bands. 
    %% The SQL search is saved as "sample.sql" 
    %% The master catalog used here is "dr1_wide_galaxy_icmodel_21.5.fits"
    %%   The original catalogs contain 2275477 objects

    We select extended objects that have no error in deblending process, well defined 
    centroids, and \texttt{cModel} magnitudes in all five bands. 
    After removing the objects that have pixel affected by saturation, cosmic-ray, and 
    other optical artefact \footnote{each criterion affects less than 8\% of the entire
    sample}, we select 1760845 galaxies that will be referred as \texttt{hscPho}. 
    %% Saved as "dr1_wide_galaxy_hscPho.fits"

    As reliable photometric redshift using HSC photometry is still a working progress, 
    we only use objects with either available spectroscopic redshift or robust 
    red-sequence photo-$z$ from the \redm{} catalog (see Section \ref{ssec:redmapper}).  
    We first match the \texttt{hscPho} sample with the external spec-$z$ catalog 
    compiled by the HSC database\footnote{It is created by matching HSC objects with 
    public data of several spectroscopic surveys (e.g.\ SDSS/BOSS; GAMA). 
    Duplicated matches from different sources are merged through internal matching 
    using $0.5^{\arcsec}$ radius. 
    For each object, the quality information of the spec-$z$ from different catalogs 
    are homogenized into a single flag that indicates whether the redshift is secure, 
    and only secure spec-$z$ are used in this work.}
    %% This is catalog: "dr1_specz_use.fits"
    using $1.0^{\arcsec}$ radius.
    At $0.2 \leq z \leq 0.5$, most redshifts in the HSC spec-$z$ catalog come from 
    SDSS/BOSS and GAMA surveys.  
    The BOSS survey provides the majority of spec-$z$ in this work.  
    Due to the complex selection criteria for different subsamples within the BOSS
    survey (e.g.\ the \texttt{LOWZ} and \texttt{CMASS}), it is not easy to estimate 
    its \mstar{} completeness.  
    Recently, through comparing with the deeper Stripe 82 Massive Galaxy Catalog
    (\texttt{S82-MGC}; \citealt{Bundy2015}), \citet{Leauthaud2016} suggests that the 
    BOSS spec-$z$ is about 80\% complete at \logms{}$\geq 11.6$ at $0.3 < z < 0.5$. 
    The GAMA survey, which partially overlaps with the HSC footprint, provides 
    additional 14\% unique spec-$z$. 
    Based on \citet{Taylor2011} (e.g.\ their Fig.~6), at $z\sim 0.3$, the GAMA 
    sample is 80\% complete down to $10^{10.8}$\msun; but only 80\% complete to 
    $10^{12.0}$\msun at $z\sim 0.5$. 
    Despite the difference in \mstar{} estimates, we can expect 
    our sample should be quite complete above \logms{}$\geq 11.5$-$11.6$.
    We will further address the issue of \mstar{}-completeness more carefully 
    using the common sample with \texttt{S82-MGC} (see Section\ref{ssec:s82}).

    Among the 116813 matched objects, 42696 are at $0.2 \leq z \leq 0.5$.
    %% Saved as "dr1_wide_galaxy_icmodel_21.5.fits"
    The majority of these redshifts come from either SDSS or BOSS data.  
    GAMA survey contributes another $\sim 14$\%. 
    For objects without external spec-$z$, we match them with the central galaxies 
    from \redm{} catalog using $2.0^{\arcsec}$ radius. 
    The matched objects with useful red-sequence photo-$z$ ($z_{\lambda}$) are also 
    included in the final sample of bright galaxies with reliable redshift between 
    $0.2 \leq z \leq 0.5$
    (will be referred as \texttt{hscZ}).
    As we will show later (see Section \ref{ssec:future}), to ensure reasonable 
    \mstar{}-completeness at the high-\mstar{} end, and make sure the samples we 
    want to compare have well overlapped redshift distributions, we will focus on 
    the galaxies at $0.3 \leq z \leq 0.5$.

%% ------------------------------------------------------------------------------------ %%  
\subsection{\redm{}{} cluster catalog}
    \label{ssec:redmapper}

    In this paper, our study focuses on galaxies which are located at the center of
    their dark matter halos. To limit our sample to central galaxies, we use 
    \texttt{v5.10} of the \redm{}{} cluster 
    catalog\footnote{See: http://risa.stanford.edu/redmapper/} 
    (e.g.\ \citealt{Rykoff2014}; \citealt{Rozo2015b}). 
    These authors have developed a well-tested red-sequence cluster finder that has been 
    run on SDSS DR8 (\citealt{SDSSDR8}) photometric data. 
    For each cluster, the catalog provides a photometric redshift estimate $z_{\lambda}$, 
    a richness $\lambda$, and identified the most likely central
    galaxy (this is the galaxies with the highest value of the central probability 
    $P_{\mathrm{Cen}}$). 
    A list of member galaxies for each cluster, and associated membership probabilities, 
    is also provided. 
    Details about the performance of the \redm{}~cluster catalog can be found in 
    \citet{Rozo2014}, \citet{Rozo2015a}, and \citet{Rozo2015b}. 
    Several studies have published calibrations between the \redm{}~richness estimate, 
    $\lambda$, and halo mass (e.g.\ \citealt{Saro2015}; \citealt{Farahi2016}; 
    \citealt{Simet2016}). 
    The results of these studies are in good agreement and indicate that clusters 
    identified by \redm{}~($\lambda > 20$) have $\log (M_{200,c}/M_{\odot}) \geq 14.0$. 
    Therefore, the \redm{} catalog helps us group the massive central galaxies into 
    samples with different average \mhalo{}.  
    And, to select reliable candidates of central galaxies, we only include the ones 
    with high probability of being the central galaxy ($P_{\mathrm{CEN}} \geq 0.7$).

    Although the \redm{} catalog provides us a good sample of central galaxies in 
    massive halos, the typical uncertainty of $\lambda$ estimate is still at 
    $\sim 5$-10 level.  
    In addition, due to the depth and resolution of SDSS images, the \redm{} catalog 
    becomes slightly incomplete at lower richness ($\lambda < 40$) end at $z > 0.33$.
    To reduce the impacts from uncertain richness and incomplete selection of 
    massive halos, we will focus on a sample of central galaxies in halos with 
    $\lambda > 30$. 
    This will also help up enhance the \mhalo{} contrast in the structural comparison.     
    Based on the calibration in \citet{Simet2016}, the $\lambda \geq 20$ sample should
    have halo mass ($M_{200m}$) more massive than $10^{14.0} M_{\odot}$.  
    For the $\lambda \geq 30$ sample, they should live in halo more massive than
    $10^{14.2} M_{\odot}$ using the same calibration. 
    And, we can confirm that the results presented later do not depend on the choice 
    of $\lambda$ boundary here. 
    For the massive central galaxies that are not in these cluster-level halos, 
    we unfortunately can not estimate their \mhalo{} individually, but it is safe 
    to assume they should have $\log (M_{200m}/M_{\odot}) < 14.0$.     
    %\textcolor{red}{but sometimes we make cuts at $\lambda<20$ and $\lambda>30$. 
    %Is it confusing to have several different thresholds? Need to give a Motivation 
    %for using different numbers here.}

%% ------------------------------------------------------------------------------------ %%

%% ------------------------------------------------------------------------------------ %% 
  \begin{figure*}[bt!]
      \centering 
      \includegraphics[width=\textwidth]{fig/redbcg_smf_new}
      \caption{\textbf{Left:}: We compre the \mstar{} volume-density distributions of 
      	  \rbcg{} (red) and \nbcg{} (black) galaxies using stellar mass based on 
          \mcmodel{} (dash) and \mtot{} (solid) to highlight the increase of \mstar{}
          due to different methods.  
          Here we also show the volume-density distributions of the \texttt{S82-MGC} 
          galaxies (green) to evaluate the \mstar{}-completeness. 
          The vertical line highlight \logms{}$=11.6$ before correcting the \mstar{} using
          total luminosity from 1-D profile.~~
          \textbf{Right:} We plot \mtot{} against the difference between \mtot{} and 
          \mcmodel{} for the \rbcg{} (red squares; color indicates redshift) and \nbcg{}
          (grey dots) galaxies. 
          Two vertical lines highlight the $11.6<$\logmtot{}$<11.9$ mass range where the
          two samples overlap the most.
          We also show the historgrams of the mass differences for galaxies with 
          $11.6<$\logmtot{}$<11.9$ on the right side. 
          \song{The figure is updated, but I think the histogram on the right side is 
          useful; 1) it shows the average mass difference clearly; 2) shows the differences
          between the two samples; Also try to update the caption, let me know what do you 
          think. And, I think the left panel should stay here, because we need to discuss 
          the mass completness here; the discussion section can use another figure where 
          we overplot other people's SMF; we just need some caveats.}
          \alexie{Simplfy the caption here. What is the main message we are trying to convey. 
          I think the point of this figure is to show that cmodel is underestimated. 
          Focus on that message. Some information here is un-necessary. 
          Also looks make some changes to this Figure: 1) right. Remove the right hand 
          histograms. 2) I ont think we learn much by the color coding. Just use a single 
          color (less explaning to do). Label using your names cenLowMh and cenHighMh}}
      \label{fig:smf1}
  \end{figure*}
%% ------------------------------------------------------------------------------------ %%

%% ------------------------------------------------------------------------------------ %% 
\subsection{Massive Central Galaxies from Low and High Mass Halos}
    \label{ssec:mass_central}

    Based on recent constraints of $M_{\star}$-$M_{\mathrm{Halo}}$ relation 
    (e.g.\ \citealt{Leauthaud2012}, \citealt{Behroozi2013}, \citealt{Kravtsov2014}), 
    the \mhalo{} of \logms{}$ > 11.0$ galaxies has a large scatter at fixed \mstar{}.            
    Although we can not yet measure \mhalo{} individually, we can still broadly 
    separate them into galaxies in small groups (\logmh{}$<14.0$) and 
    large groups/clusters (\logmh{}$>14.0$), and investigate the 
    ``environmental''-dependence of their structures.

    % We match this sample with the central galaxies of \redm{}~catalogs using a 
    % $1.0^{\arcsec}$ radius, and it results in 704 galaxies.  
    %% Saved as "dr1_redbcg_hsc_sdss_gama_1arcsec.fits"
    Firstly, we match the \texttt{hscZ} sample with the massive central galaxies of 
    \redm{} clusters with $\lambda \geq 20$ and $P_{\mathrm{0.7}} \geq 0.7$ using 
    $1.0^{\arcsec}$ radius. 
    This step finds 375 matched galaxies at $0.2 \leq z \leq 0.5$ 
    (among them 282 are at $0.3 \leq z \leq 0.5$).  
    We will refer these \textbf{central galaxies in more massive halos} as \rbcg{} 
    sample from now on.
    A small fraction of \redm{} centrals within the HSC footprints are missing  
    due to severe contamination from optical artifacts or bright stars.
    Spec-$z$ is not available for 67 (17.9\%) galaxies in the sample, but their 
    $z_{\lambda}$ should be very accurate (median $|z_{\lambda} - z_{\mathrm{Spec}}|$ is 
    about 0.01 for the ones with spec-$z$).
    The median richness ($\lambda$) of these clusters is $\sim 32$, corresponding to 
    halo mass of $M_{200m}\sim 1.7\times 10^{14} h^{-1} M_{\odot}$. 
    Among these central galaxies, 222 are in clusters with $\lambda \geq 30$. 
    When focusing on this subsample, the median richness is $\sim 41$, corresponding 
    to halos with $M_{200m} \sim 2.2\times 10^{14} h^{-1} M_{\odot}$. 
    Meanwhile, only 15\% of \rbcg{} sample lives in clusters with $\lambda \geq 50$
    ($M_{200m} \sim 3.0\times 10^{14} h^{-1} M_{\odot}$).   
    Therefore, the current \rbcg{} sample is not dominated by very massive clusters 
    with \mhalo{}$>10^{15} M_{\odot}$.  
    %% Saved as "dr1_redbcg_use_sed5b.fits"

    Assuming that the \redm{} catalog correctly identifies all halos with 
    $M_{200m}\sim 1.0\times 10^{14} h^{-1} M_{\odot}$ in the current footprint,
    the unmatched massive galaxies from the above step should be dominated by central 
    galaxies of less massive halos and massive satellites galaxies.  
    To select a purer sample of massive centrals in $M_{200m}<10^{14} h^{-1} M_{\odot}$
    halos, we identify and remove any galaxy that is too close to 
    \emph{any cluster member} in \redm{} catalog in both radius and redshift direction.

    \alexie{This is a rather strange procedure, I think we need to change it (referee and readers will complain!). What we really want to do is to remove any galaxies that have $r<R200$ in projection (from the galaxy with the highest value of Pcen) and remove galaxies within some deltaz of the cluster redshift. The cluster redshift is more well determined than the redshift estiamte of any cluster member. In short, what we are doing here is cutting out a cylindar arond the most likley BCG. The deltaz value can be motivated by the cluster redshift uncertainty from the redmapper papers. The only remaining question is then what radius to use around clusters. To get that number here is what we need to do. 1) Start with Eq 12 and Table 3 in Simet et al. For each cluster, assign a value of M200m (halo mass defined as 200 time background) to each cluster. 3) we then need to estimate the cluster radius R200m (or R200b, same thing). To do this, we need to use a mass-concentration relation. I'm pretty sure there is an easy command to do this in Python using the python package "colossus" by bendikt diemer. Download it and see if youc an get it to work. If it doens't look obvious after 30 min, I can do it on my end using my own code. But good for you to know how to do this. We can simply use the mass-concentration realtion from Benedikts work. 4) Now for every cluster you have R200m. And you define the low halo mass central sample as ones that are not within any cylindar around any cluster. Sorry, I know this might mean re-doing everything.....let's discuss by Skype if this is a major issue.}

    For each \redm{} cluster member, we remove objects within a cylinder region 
    around it from the remaining sample. 
    Such cylinder is defined by 1) physical radius of 3.0 Mpc at the red-sequence 
    redshift of the cluster; 2) $|z - z_{\lambda}| \leq 0.05$ as 
    \nbcg{} later.

    However, we should note that the \nbcg{} sample at this stage spans a large range
    in \mstar{}, extending towards \logms{}$<10^{11}$.  
    Therefore, we should expect contamination from satellites in \logmh{}$< 14.0$ halos
    at certain degree.  
    Later, after we update the total \mstar{} of them, we will mainly focus on the 
    \nbcg{} galaxies at high mass end, especially the ones at \logms{}$ > 11.5$. 
    Based on observational constraints (e.g. \citealt{vanUitert2016}), we can neglect 
    satellite contamination at such high \mstar{} range. 
    %% Saved as "dr1_nonbcg_use_sed5b.fits"

\subsection{Summary of Sample Construction}
    \label{ssec:sample}

    In summary, from the $\sim 100$ deg$^2$ deep HSC images, we select a large sample
    of massive central galaxies with reliable redshift information, and broadly separate 
    them into two groups with different \mhalo{}.  
    To help you go through the results, here are a few key points to keep in mind:

    \begin{itemize}
        \item \texttt{hscPho} sample: this parent sample consists of bright galaxies 
            with $i_{\mathrm{cModel}} \leq 21.0$, good quality images and reliable 
            \texttt{cModel} photometry in all five HSC bands in the \texttt{S15B} 
            data release. 
        \item \texttt{hscZ} sample: through matching the \texttt{hscPho} sample with 
            spectroscopic redshifts from surveys like SDSS/BOSS and GAMA, or accurate 
            red-sequence photometric redshift in the \redm{} catalog, we come up with 
            a large sample of bright galaxies with reliable redshift information. 
        \item \rbcg{} sample: we select a sample of 375 galaxies at $0.2 \leq z \leq 0.5$
            that are identified as central galaxies in $\lambda > 30$ \redm{} clusters. 
            The most reliable ones ($P_{\mathrm{Cen}} >0.7$) among 
            them will represent the central galaxies in very massive halos
            (\logmh{}$\geq 14.0$). 
        \item \nbcg{} sample: after excluding all the galaxies that are close to any 
            \redm{} clusters in both radial and redshift directions from the 
            \texttt{hscZ} sample, we have 29973 bright galaxies at $0.2 \leq z \leq 0.5$.
            With the help of \mstar{} estimates via SED fitting (see Section 
            \ref{ssec:isedfit}), we consider the massive ones among them as the 
            candidates of central galaxies in halos with \logmh{}$<14.0$.  
        \item To compare the \rbcg{} and \nbcg{} samples carefully, we will focus on 
            the redshift range at $0.3 \leq z \leq 0.5$ where their redshift 
            distributions greatly overlap and both samples have acceptable
            \mstar-completeness (see Section \ref{ssec:s82}).  
            Also, during the comparison of structures, we will mainly 
            \song{Will use $>30$ sample as default later; will update the figure first, 
            then correct the number of galaxies when I go through the text again.}
            \alexie{Why mainly use here? We have defined the high mass sample to be above this cut. Lets use $\lambda>30$ everywhere! Making more samples will make readers confused.} use \rbcg{} 
            galaxies with $\lambda > 30$ to ensure that they are from more massive 
            halos than the ones in \nbcg{} sample.
    \end{itemize}

%% ------------------------------------------------------------------------------------ %%

%% ------------------------------------------------------------------------------------ %% 
\section{Measurements of 1-D Surface Brightness Profile}
    \label{sec:ellipse}

    To estimate the total luminosity of massive galaxies, and measure their 
    one-dimensional stellar mass density (\mden{}) profiles, we perform elliptical 
    isophotes fitting using the IRAF task \texttt{Ellipse} (Jedrzejewski 1987).  
    Compared to the popular two-dimensional model fitting method, the isophote fitting
    approach is much less sensitive to the choice of model, number of components, 
    and the initial guesses of free parameters. 
    It is also less affected by the uncertainty of sky background subtraction.  
    This is particularly important for the massive ellipticals in our sample.  
    The much deeper HSC image reveals significantly more extended structures, 
    while also make it inappropriate to fit these galaxies with simple model like 
    the de~Vaucouleurs model or single \ser component model. 
    Such models fail to fit the central region and the stellar halo of massive 
    galaxies simultaneously, and also can not account for the radial variation of 
    ellipticity and position angle. 
    In principle, such massive galaxies can still be described using more complex 
    models (e.g \citealt{Huang2013a}; \citealt{Huang2013b}).  
    However, they are still very sensitive to background subtraction. 
    And it becomes even more difficult to choose initial parameters, and investigate
    the degeneracies among parameters. 
    Although it is certainly worth exploring in the future, we decide that the 1-D 
    method suits the goal of this work better.

    We first prepare large $i$-band cut-out images around these massive galaxies 
    that cover at least 750 kpc in radius, along with the bad pixel masks, and the 
    PSF model reconstructed using the central coordinates. 
    While they include all the visible light of the galaxy, they also leave enough 
    space to evaluate the background subtraction. 
    We choose to use $i$-band images not only considering their excellent seeing
    condition, but also because they trace the stellar distributions of massive 
    galaxies at $0.3 \leq z \leq 0.5$ reasonably well (fall between rest-frame $g$ 
    and $r$ band), and they suffer less from background uncertainty, which is crucial 
    for studying the outskirt of massive ETGs.  
    Although $z$ and Y-band should be better \mden{} tracer, the sky background 
    is much higher and the seeing is considerably worse on average.

    Secondly, using the \texttt{SEP} Python library, we perform \texttt{SExtractor}-like
    background estimation and object detections using different configurations to 
    overcome the ``over-deblending'' challenge met by the \texttt{hscPipe}. 
    Combining detections using different backgrounds and $S/N$ thresholds, 
    we correctly obtain the footprints of all objects, even of the ones that are 
    very close to the center of bright galaxies.
    For the rest objects, we convert their footprints into a mask image for the
    \texttt{Ellipse} procedure after increasing their sizes adaptively based on the 
    brightness and distance to the massive galaxy. 
    We achieve this via convolving the individual footprint with a Gaussian kernel
    so that we can conservatively exclude pixels affected by this object while still 
    making the mask follow the shape of the object. 
    Meanwhile, we create a more aggressive mask of \emph{all} objects in the same 
    approach, median-rebin the remaining pixels using a $6x6$ pixels box, and take 
    the peak value (or mode) of the distribution of these rebinned ``sky'' pixels 
    as the average background value around these bright galaxies.
    As expected, we often found slightly negative value that indicate over-subtraction 
    of background at certain degree. 
    In \texttt{hscPipe}, the background on each CCD is modeled with a 
    Chebyshev-polynomial fit to the smoothed image after excluding pixels with $S/N >5$.
    This algorithm suffer less than the SDSS version (e.g.\ see \citealt{Blanton2011}),
    but still over-subtract background around very bright object. 
    For our massive galaxies, such over-subtracted background creates artifactual
    truncation or steeper surface brightness profile.
    We therefore provide an ad hoc fix using a \texttt{SExtractor}-style background 
    model ($200x200$ pixels background box size, and 6 pixels median filtering size of 
    sky boxes) via running \texttt{SEP} on the image with all objects masked out.
    This model can account for the slightly over-subtracted background at very large 
    scale, shift the distribution of rebinned ``sky'' pixels to 0, and do not affect the 
    intrinsic flux distribution of our target.

    Finally, we run \texttt{Ellipse} on the background-corrected image following similar 
    strategy of \citep{Li2012}.
    In short, we start with free centroid and shape of each isophote, then gradually 
    constrain their behaviors, and eventually extract 1-D surface brightness profile 
    along the major axis using isophotes with fixed centroid and shape.
    The 4th Fourier modes are included to fit the isophote better. 
    In this way, we can extract the radial changes of the centroid, ellipticity, 
    position angle, and average isophotal distortions (more ``disky'' or ``boxy'', 
    e.g.\ \citealt{Kormendy2009}) along with the surface brightness profile.  
    Fig.\ref{fig:ellipse} shows an example of the 1-D surface brightness and 
    ellipticity profile for a massive galaxy at $z\sim0.2$ along with its masked 
    image and a subset of isophotes.  
    Appendix\ref{app:A} includes more details about the \texttt{Ellipse} procedure.

    %At low surface brightness regime, the final profile can be sensitive to several 
    %configuration parameters of \texttt{Ellipse}.  
    %After some tests, we choose to use 0.1 dex in logarithm as the step in semi-major 
    %axis length between successive ellipses, and we use the median pixel value over the
    %elliptical annulus after rejecting outlier pixels twice with $3\sigma$-clipping.
    %These parameters are selected to make the final profile less affected by any nearby
    %objects, and please see Appendix.~B for more details about the extraction of 
    %1-D profile.

    For physical (e.g.\ late-stage major merger) and unphysical (e.g.\ nearby 
    foreground galaxy or bright star) reasons, we can not extract reliable 1-D 
    profiles for small fraction of massive galaxies due to heavily masked out 
    central region. 
    This issue reflects an intrinsic limitation of the 1-D method. 
    It affects $48/375$ \rbcg{} galaxies, and a smaller fraction of \nbcg{} ones. 
    It is worth noting that this excludes most massive galaxies in major merger 
    process from the analysis. 
    We will come back to them in the future using 2-D modeling method.

    We correct these surface brightness profiles for Galactic extinction and 
    cosmological dimming, then integrate them to various radius to get the 
    luminosity within different physical apertures.  
    Although the $i$-band 1-D profile can in principle reach to $\sim 30$ \sb, 
    we notice the profiles at very low surface brightness part are often less 
    reliable by showing truncation or large fluctuation that are due to either 
    background uncertainty or contamination from bright neighbours.
    We therefore only consider the $<28.5$ \sb part of the profile here. 
    This conservative choice already allows us to study the region out to 
    $\sim 120$ kpc for massive galaxies in this sample.

    %We also apply the isophotes from $i$-band images to other bands in 
    %``force-photometry'' mode \texttt{Ellipse} run to get rough estimates of 
    %color profiles.  
    %Without taking the difference in seeing and background into account, 
    %these color profiles can not be used for physical discussion. 
    %But with the help of $K$-correction from the next section, we show that there 
    %is no systematic difference in color gradient between \rbcg{} and 
    %\nbcg{} sample.

%% ------------------------------------------------------------------------------------ %%

\section{Stellar Masses and Stellar Mass Density Profiles}
    \label{sec:mstar}

\subsection{Stellar Masses from SED Fitting}
    \label{ssec:isedfit}

    To convert the luminosity into stellar mass estimates, we assume that these massive 
    galaxies can be well described by an average \m2l{}. 
    This is a reasonable assumption considering that they are ETGs that are dominated by 
    old stellar population and are known to have shallow color gradients. 
    We will further justify this point using the average color profiles in Section
    \ref{sec:discussion}.

    We use the broadband Spectral Energy Distributions (SEDs) fitting 
    (see \citealt{Walcher2011} for a recent review) code 
    \texttt{iSEDFit}\footnote{http://www.sos.siena.edu/~jmoustakas/isedfit/} 
    (\citealt{Moustakas13}) to estimate the average \m2l{} and $k$-corrections using
    5-band HSC \cmodel{} fluxes under forced-photometry mode. 
    Although \cmodel{} tends to underestimate the total flux for bright, extended 
    objects, the forced-photometry results still provides accurate \emph{average} color 
    of the galaxy as it constrains the behaviour of the model using the optimized 
    reference one and taks the PSF convolution into account
    (e.g.\ Huang\etal~in prep.~).

    \texttt{iSEDFit} takes a simplified Bayesian approach. 
    In short, it first generates a large grid of SEDs from synthetic stellar population 
    models by drawing randomly from the prior distributions of relevant parameters 
    (e.g.\ age, metallicity, dust extinction, and star formation history).
    Based on these models, it uses the observed photometry and redshift to compute the 
    statistical likelihood, and generate the posterior probability distribution (PDF) 
    functions of each parameter.  
    To get the best estimate of certain parameter, \texttt{iSEDFit} ``integrates'' the 
    full PDF over all the other ``nuisance'' parameters.
    Then, the median value of the resulting marginalized PDF is used as the best value, 
    while the 1-$\sigma$ uncertainty is derived from the cumulative PDF. 
    Please refer to \citet{Moustakas13} for technical details and performance of 
    \texttt{iSEDFit}.

%% ------------------------------------------------------------------------------------ %% 
    In this work, we derive average \m2l{} using the Flexible Stellar Population 
    Synthesis\footnote{http://scholar.harvard.edu/cconroy/sps-models}
    (FSPS; \texttt{v2.4}; \citealt{FSPS}, \citealt{Conroy2010}) model based on the MILES
    \footnote{http://www.iac.es/proyecto/miles/pages/stellar-libraries/miles-library.php}
    (\citealt{MILES1}, \citealt{MILES2}) stellar library and \citet{Chabrier2003} 
    IMF between 0.1 to 100 \msun. 
    We use a delayed-$\tau$ model with stochastic star burst as the form of star 
    formation history (SFH), and choose a set of reasonable parameters after some tests.  
    Such form of SFH is appropriate for massive galaxies at low redshift 
    (e.g.\ \citealt{Kauffmann2003}). 
    For stellar metallicity, we assume flat distribution between 0.004 to 0.03 (which is 
    the highest metallicity allowed by FSPS models).  
    And, we adopt the \citet{Calzetti2000} extinction law with a order two Gamma 
    distribution of $A_{V}$ between 0 to 2 magnitude.
    Since most galaxies in both samples are red, quiescent galaxies, the results are 
    not very sensitive to parameters related to SFH and internal dust extinction. 
    To achieve reasonable sampling across these parameters, we generate 250000 models.

    We construct five band SED using the forced-photometry \cmodel{} magnitude corrected 
    for the Galactic extinction. 
    As for the photometric error, the current \cmodel{} photometry underestimates 
    the flux errors of bright galaxies as it only includes statistical error, not the 
    systematic uncertainties in the model-fitting process.  
    From tests of synthetic objects, we find that the average accuracy of \cmodel{} 
    photometry for bright galaxy is roughtly at 1\% level. 
    Therefore, we supply \texttt{iSEDFit} with simplified flux errors assuming $S/N = 100$ 
    for $riz$ bands, and $S/N = 80$ for $gY$ bands (on average, images in $gY$ bands are 
    shallower in depth and/or have higher background noise).  
    The typical uncertainty of \logms{} for both samples is around 0.08-0.10 dex at the 
    high-\mstar{} end. 
    Please see Appendix\ref{app:B} for more detailed discussions of the SED fitting
    results.  
    %In Fig.~3, we show an example of SED fitting result. 
    %\song{Not sure if we need to show one example, but put one here for now.}

%% ------------------------------------------------------------------------------------ %%

%% ------------------------------------------------------------------------------------ %% 
  \begin{figure*}[t!]
      \centering 
      \includegraphics[width=\textwidth]{fig/redbcg_sample_stats}
      \caption{\textbf{Left}: The \logms{}-$g-r$ color relation of the \rbcg{} 
          (red circle; detaield format is the same with the right panel of 
          Fig.\ref{fig:smf1}) and \nbcg{} (grey dots) samples.
          We apply the $k$-corrections from \texttt{iSEDFit} fitting to the colors.~~
          \textbf{Middle}: the histograms of \mtot{} for the \rbcg{} (orange-red) and 
          \nbcg{} (grey-black) samples at both $z>0.2$ (step-filled histogram) and 
          $0.3 \leq z \leq 0.5$ (stepped histogram). 
          The vertical lines in both left and middle figures highlight the 
          $11.6<$\logmtot{}$<11.9$ mass range that will be used in the comparison of 
          the \mtot{}-matched samples.~~
          \textbf{Right}: the histograms of the redshift for the \rbcg{} and 
          \nbcg{} galaxies in both \logmtot$>11.2$ and $11.6<$\logmtot{}$<11.9$
          mass bins.
          The vertical lines highlights the $0.3\leq z \leq 0.5$ redshift range 
          \song{Will move to new Appendix later; but the figures are updated here; 
          I will make sure to use consistent symbols, colors, and legends.}
          \alexie{Move this figure to the appendix. It is not clear to be that we need this Figure anyway. We can say that we make matched samples in the text. I think that is enough. If the referee wants to see the Figure, we can show him or her. But I think it just adds clutter.}.
      }
      \label{fig:sample_stats}
  \end{figure*}
%% ------------------------------------------------------------------------------------ %%

\subsection{Comparison with \texttt{S82-MGC}}
    \label{ssec:s82}

    Given the heterogeneous nature of the redshift resources for these samples, we 
    further investigate their \mstar{} completeness via comparing with galaxies that are 
    in common with the \texttt{S82-MGC} sample. 
    The \texttt{S82-MGC} sample matches the deeper SDSS photometric data in the Stripe 82 
    region (\citealt{Annis2014}) with the near infrared data from the United Kingdom 
    Infrared Telescope Infrared Deep Sky Survey (UKIDSS; \citealt{Lawrence2007}). 
    The deeper photometry and better photo-$z$ make the \texttt{S82-MGC} sample complete 
    to \logms{}$\geq 11.2$ at $z<0.7$.

    The match between the \texttt{hscPho} sample and the \texttt{S82-MGC} one results in 
    20453 galaxies at $0.3 \leq z \leq 0.5$ (referred as \texttt{s82Pho} sample).  
    We estimate their \mstar{} using \texttt{cModel} SED in exactly the same way. 
    The \mstar{} using HSC photometry shows tight correlation with ones based on the
    \texttt{S82-MGC} SED (five optical bands from SDSS and NIR data from UKIDSS; and 
    \texttt{S82-MGC} also used \texttt{iSEDFit} under slightly different assumptions).  
    Then we estimate the volume densities distributions of galaxies in \rbcg{}, \nbcg{}, 
    and \texttt{s82Pho} samples (left panel of Fig.\ref{fig:smf1}) using rough estimates 
    of the survey areas they occupied (since we only care about the relative shapes of 
    the density distributions). 
    And, we estimate the uncertanties of these distributions using a 10000 times 
    bootstrap resampling. 
    The distribution of \nbcg{} sample starts to deviates from the \texttt{s82Pho} one 
    below \logms{}$< 11.6$, indicating that the sample starts to become incomplete within
    $0.3 \leq z \leq 0.5$. 
    As for the \rbcg{} sample, their mass completeness should be better thanks to the 
    reliable \redm{} $z_{\lambda}$. 
    Its behaviour is quite consistent with the mass function of galaxies in massive 
    halos with a narrow \mhalo{} distribution (\addref).   
    Later, we will see that the \cmodel{} photometry on average underestimate the 
    \mstar{} of galaxies with \logms{}$>11.5$ by $-0.1$ dex.
    Therefore, we conclude that the \nbcg{} becomes not perfectly complete below 
    \logms{}$< 11.7$.  
    This is most likely due to the incompleteness of the BOSS spec-$z$ sample. 
    Although it is unlikely that the incompleteness will bias the results of our 
    comparisons, we will bear this in mind, and match the two samples carefully in 
    \logms{} and redshift space before comparing them.

%% ------------------------------------------------------------------------------------ %%

%% ------------------------------------------------------------------------------------ %% 
  \begin{figure*}[bt!]
      \centering 
      \includegraphics[width=\textwidth]{fig/average_mass_profiles_fsps1_A}
      \caption{
      	\song{The figures are updated as you suggested; But I want to point out that
        the two samples have different mass completeness (\rbcg{} has photo-$z$ ones), and 
        I simply combine them.  The table at the end will be updated, and I will provide 
        the average profiles from our key figures too.}
      	\alexie{I have a suggestion that I think will make this figure more simple and wil also help with the structure of the text. In the left hand side, don't bin by $\lambda$. Instead, only make mass bins. That way you don't have to worry about getting into the matched samples at this point because you wont be showing the enviromental dependence. And when comparing on the rigth hand side, it is not clear which lines we want to compare with (are previous results for BCgs, for field galaxies,...? readers don't know). I think it will be a more straightforward comparison just to show the stellar mass bins here. For our stellar mass bins, you only need the labels for our data once on the left hand side.}. 
      	Left: Median \mden{} profiles of galaxies in different \mtot{} bins:
      	[11.4, 11.6] (short-dashed line), [11.6, 11.8] (solid line), and 
      	[11.8, 12.0] (long-dashed line). 
      	Here we combine the \rbcg{} and \nbcg{} samples, and we also show a random subset 
      	of individual profiles in the background.
      	The shaded region highlights the region that could be affected by PSF.
        Two vertical lines label the radius of 10 kpc (thin, dotted line) and
        100 kpc (thick, dash line). ~~~ 
        Right: comparison between the average \mden{} profiles with the ones from 
        previous observations and simulations, including: 
        1) average profile of massive elliptical galaxies at $z\sim 0$ from 
        \citet[][Cyan, solid]{Huang2013a}; 
        2) average profile of massive galaxies at $0.25 \leq z < 0.50$ observed by 
        \textit{HST} from \citet[][Red, long-dashed]{Patel2013} ;
        3) average radial stellar distribuitions in massive halos using simulation and 
        particle tagging method (\citealt{Cooper13}; Purple, short-dashed).}
      \label{fig:avg_prof}
  \end{figure*}
%% ------------------------------------------------------------------------------------ %%

%% ------------------------------------------------------------------------------------ %% 
  \begin{figure*}[t!]
      \centering 
      \includegraphics[width=\textwidth]{fig/redbcg_prof_m100A}
      \caption{\alexie{So this is our main figure, right? We really need to make this stand out. Remove all the matched sample plots on the left (appendix?). We only need to say we do the matched samples in the text, we don't need a figure to back this up. Then, you have a seperate figure for $M10$. Why not combine the two into a single four panel figure? I think that would capture the main result pretty well, and show M10 and M100 side by side.} Comparison of the \mden{} profiles for \mtot{}-matched samples of \rbcg{}
      (orange to red) and \nbcg{} (grey to black) galaxies within $11.6<$\logms{}$<11.9$.
      \textbf{Left figure} shows the details of the \mtot{}-matching process.  
      On the top panel, we show the overall distributions of \rbcg{} (light orange boxes) 
      and \nbcg{} (light grey dots) galaxies on the \mtot{}-$z$ plane.  
      And, we match the two sample in the \mtot{}-$z$ space outlined by the shaded region.
      We highlight the \rbcg{} galaxies in this region using bigger boxes in red frames, 
      whose size reflects the $P_{\mathrm{Cen}}$ value.  
      We also color-code them using the richness ($\lambda$) of the host cluster. 
      The matched \nbcg{} galaxies are highlighted using darker color and bigger dots. 
      To further evaluate the matching results, we show the distributions of \mtot{} 
      (middle panel) and redshift (bottom panel) separately. 
      On both panesl, we show the histograms along with their kernel density 
      distributions.  
      And, on the top of each panel, two sets of short vertical lines highlight the median 
      value (solid) and the inter-quartile (dash) of each distribution.~~~
      \textbf{The figure on the right} shows the results of the profile comparison.  
      On its upper panel, we show the \mden{} profile of each galaxy using very light 
      color in the background.  
      We highlight the median profile and its corresponding uncertainties using thick 
      solid line with darker color and shaded region. 
      Other formats are the same with Fig.\ref{fig:avg_prof}.
      Meanwhile, we hightligh the difference between the median profiles and its 
      uncertainty on the bottom panel (brown solid line and shaded region).
      To test the significance of the differential profile, we perform statistical tests 
      by comparing the median profiles of two random groups that are drawn from the 
      mixed sample, and have the same sizes with the matched \rbcg{} and \nbcg{} ones.
      Repeating this process many times, the median random differential profile 
      (black solid line), their 1-$\sigma$ (dark-grey shaded region) and 3-$\sigma$
      (light-grey shaded region) are highlighted.
      \song{Will make a new 4 panels figures here; but I want to make some changes as 
      if we simply put the current ones together, it looks bad...}}  
      \label{fig:prof_m100_A}
  \end{figure*}
%% ------------------------------------------------------------------------------------ %%

\subsection{Stellar Mass Corrected for Total Luminosity}
    \label{ssec:mtotal}

    Using the best-fit stellar mass from \texttt{iSEDFit} (referred as \mcmodel{}), 
    we estimate the average \m2l{} in $i$-band, then use that \m2l{} to convert the 
    1-D luminosity density profiles into stellar mass density profiles; and 
    also convert the luminosity within 10 and 100 kpc apertures into 
    corresponding stellar mass estimates (referred as \minn{} and \mtot{}). 
    From now, we use \mtot{} as the new proxy of ``total'' stellar mass of these 
    galaxies.  
    And we also derive the $k$-corrected luminosity and colors from the results.
    In Appendix\ref{app:B}, we also show the outputs like the average stellar age, 
    metallicity, and internal dust extinction all behave reasonably for both 
    samples, but will not use them for further scientific analysis.

    As expected, the integration of 1-D profile out to very large radius help 
    recover more luminosity (stellar mass) compared to the \cmodel{} results.
    We highlight this in the right panel of Fig.\ref{fig:smf1} as we show the 
    distributions of differences between \mtot{} and \mcmodel{}, and their relations 
    to the \mtot{}.  
    For both samples, we see the ``extra mass'' recovered by 1-D photometry 
    increases with \mtot{}, therefore this improvement impacts the \rbcg{}
    sample more.  
    This is basically consistent with the fact that the structure of massive ETGs 
    depends on its stellar mass in a way that more massive ones tends to be 
    more extended (e.g.\ \addref).  
    Apparently, such mass-dependent difference between \mtot{} and \mcmodel{} will 
    result in difference of the SMF. 
    As the slope and normalization of SMF at high mass end is very sensitive 
    to both cosmology and interesting baryonic processes such as AGN feedback
    (\addref), they are particularly important in constraining the galaxy evolution
    model, and therefore became the center of some intense arguments (\addref).  
    Although the jury is still out on this topic, it has become clear that 
    popular results using SDSS photometry (\addref) underestimate the total 
    luminosity (stellar mass) of massive ETGs.  
    However, past works (e.g.\ \addref) often relied on either stacking of 
    shallow images or the extrapolation of models to investigate this effect. 
    Using the deep HSC images, we have the potential to answer this question 
    much better. 
    Detailed studies of SMF deserves more careful investigations of many 
    systematic issues (e.g.\ \addref), so we just show the impact of the 
    ``extra mass'' on the volume density distributions of both samples 
    (Fig.\ref{fig:smf1}, left panel).  
    Generally speaking, the stellar mass corrected for total luminosity 
    shifts the distribution toward the higher mass end, and slightly modify
    the slope. 
    People who want to use HSC photometry or other \cmodel{}-like photometry to 
    study the properties and evolution of SMF should be aware of such effect.

    As mentioned earlier, our assumption of using an average \m2l{} ignores its
    radial variation.   
    It is well known that massive ETGs have negative optical color gradient which 
    indicates a gradient of \m2l{} (e.g.\ \citealt{LaBarbera2012}; \citealt{DSouza2015}),
    but the gradient is smooth and shallow out to a few times of the effective 
    radius (color gradient at the very outskirt is still not well quantified). 
    And, for the comparison between the \rbcg{} and \nbcg{} samples, 
    Our results should remain intact as long as there is no significant
    differences of color gradient between these two samples.
    We will discuss this more later, but initial results do suggest a smooth color
    gradient out to very large radius, and there is no systematic difference in 
    color gradients of the two samples.  
    Both suggest that the average \m2l{} approach should work reasonably well for
    our goals.

    And, we should mention that our approach is in principle very similar to the 
    one adopted by the GAMA survey (\citealt{Taylor2011}), where they start with 
    \mstar{} estimated using average \m2l{} from SED fitting results of
    PSF-matched aperture photometry, and later correct it with better estimate 
    of total luminosity.  
    However, the GAMA survey relied on multi-band single-\ser model fitting on 
    the SDSS images to get total luminosity (\citealt{Kelvin2012}). 
    It is therefore interesting to compare the \mstar{} (\mgama{} v.s. \mtot{})
    estimates for galaxies in common, given the differences in data and method. 
    Please see the Appendix\ref{app:C} for more details. 
    In short, we notice systematic differences between the two \mstar{} estimates, 
    and it is likely that the GAMA models miss fluxes in the stellar halo that 
    is beyond the depth of SDSS image.

    In the left figure of Fig.\ref{fig:sample_stats}, we show the \mtot{}-color 
    relations using the $k$-corrected $g-r$ color for the \rbcg{} and \nbcg{} 
    samples. 
    Both samples follow the same tight red-sequence with very little contamination 
    from star-forming galaxy in the ``blue cloud''.
    And, at fixed \mtot{}, we see little offset in color distributions of the two 
    samples suggesting that both samples consist of old quiescent galaxies with 
    similar average stellar population properties.  
    This is consistent with previous result that suggests the average stellar 
    population of massive central galaxy does not depend on \mhalo{} 
    (e.g.\ \citealt{Park2007}).  
    To look into the potential environmental dependence of structure, we will 
    focus on the mass range of $11.6 \le$\logmtot{}$\le 11.9$, where both samples 
    have acceptable completeness, and their \mtot{} distributions greatly overlap
    (see the normalized distributions of \mtot{} in the middle panel of 
    Fig.\ref{fig:sample_stats}). 
    Due to the different sources of redshift for the \rbcg{} and \nbcg{} samples, 
    the two samples have quite different redshift distributions 
    (right figure of Fig.\ref{fig:sample_stats}).
    Even within the $11.6 \le$\logmtot{}$\le 11.9$ mass bin, the redshift
    differences still differ from each other a little bit.  
    Apparently, the redshift distribution of the \nbcg{} sample skews toward 
    higher-$z$ end, primarily due to the contribution of BOSS spec-$z$.
    This could lead bias into the comparison of \mden{} profiles and other 
    properties between the two sample (please see Appendix\ref{app:D} for more 
    details).  
    We will address this via matching the two samples in \mtot{} and redshift
    distributions carefully before the comparisons.

%% ------------------------------------------------------------------------------------ %%

\section{Results}
    \label{sec:result}

\subsection{Comparison of Surface Mass Density Profiles}
    \label{ssec:sbp_compare}

%% ------------------------------------------------------------------------------------ %% 
  \begin{figure*}[t!]
      \centering 
      \includegraphics[width=\textwidth]{fig/redbcg_prof_m100B}
      \caption{Comparisons of the \mden{} profiles for \rbcg{} (orange-red) and \nbcg{} 
      	(grey-black) galaxies that are matched using proxies of total \mstar{}. 
        The formats are in consistent with the right figure of Fig.\ref{fig:prof_m100_A}.
        The differences are, here, the samples are matched in slightly differnt ways. 
        From left to right: a) using samples at lower redshift ($0.3 < z < 0.4$); 
        b) using \rbcg{} sample with $\lambda > 20$ instead of 30; 
        c) using \mstar{} within 150 kpc instead of 100 kpc.
        The results are broadly consistent with the one in Fig.\ref{fig:prof_m100_A}.
        \song{Considering to move to a new Appendix.} \alexie{This Figure can be moved to an appendix. It can also be summarized by saying in the main part of the paper that "we have tested that our results are robust to, the choice of the redshift range, the lower $\lambda$ cut, .....". I think it is probably enough to simple say this in the text. Not 1oo percent convinced you need a figure to abck this up.}}
      \label{fig:prof_m100_B}
  \end{figure*}
%% ------------------------------------------------------------------------------------ %%

\subsubsection{Internal Comparison and Comparison with Previous Works}
    \label{sssec:sbp_inter}

    As discussed earlier, although the mass-size scaling relation has been intensely 
    discussed related to the topic of environmental dependence of structure, it may not be 
    the best tool to investigate this problem.    
    Compared to the simple scaling relation, the detailed stellar mass density profiles 
    contain much richer structural details that can help us diagnose the role played by 
    different physical processes, including the environmental effect.

    Before the actual comparisons, we first show the general properties 
    of these 1-D profiles by separating both samples into different bins 
    of \mtot{} and redshift, also by comparing them with results from 
    literature (Fig.\ref{fig:avg_prof}).   
    For profiles in each bin, we derive the median profile along with its 
    uncertainty using bootstrap resampling method (5000 times). 
    The left panel of Fig.\ref{fig:avg_prof} focuses on different mass bins.  
    As expected, for both samples, higher \mtot{} moves the median mass 
    density profile upward a little in the inner region, but makes it 
    more extended in the outer part.
    In the mass range we explored ($\ge 10^{11.2} M_{\odot}$), most of these 
    ETGs are slow-rotating (e.g.\ \citealt{Cappellari13b}) giant ellipticals 
    with boxy inner isophotal shape (e.g.\ \citealt{Kormendy2009}), and 
    flattened density profile at the center (e.g.\ \citealt{Lauer07}), 
    but their structure in the outskirt clearly does \textbf{NOT} response 
    to stellar mass increase in a self-similar way.

    In principe, this is consistent with result from HST observations of 
    BCGs at $0.3 < z <0.9$ (\citealt{Bai2014}) and the claimed positive 
    correlation between luminosity and \ser index
    (e.g.\ \citealt{Savorgnan13})\footnote{But it does not mean such high 
    \ser index model works well for massive ETGs as they fail at the inner-most 
    region dramatically, and can not explain the radial variation of 
    isophotal shape}; it is also consistent with the picture that more 
    massive ETGs experienced more recent (minor) mergers.

\alexie{My comment in Figre \ref{{fig:avg_prof}} wil make this part unnecessary and will improve the flow of the paper.}    At the same time, such behaviour should make us very cautious about 
    any indication of ``environmental dependence of structure'' as it 
    may be degenerate with difference caused by stellar mass when the 
    two samples are not perfectly matched in mass.  
    This is exactly what we see in the median profiles of \rbcg{}
    and \nbcg{} in the same mass bins. 
    Although the comparison shows that \rbcg{} shows much more 
    extended outer envelope than \nbcg{}, the very similar profile
    in the inner $\sim 10$ kpc reveals that this is dominated by the 
    different mass distributions in two samples.

    In the right panel of Fig.\ref{fig:avg_prof}, we compare these median 
    profiles with a few past works.  
    \citep{Huang2013a}, derived the median stellar mass density profile 
    for a small sample of very nearby ellipticals from the 
    Carnegie-Irvine Galaxy Survey (CGS, \citealt{CGS1}).
    Individual profile and total luminosity were derived from 
    three-component 2-D models that describe both very inner and outer 
    luminosity distributions of these galaxies accurately. 
    The average stellar mass of this sample is around $10^{11.3} M_{\odot}$.
    Due to the proximity of this sample ($< 100$ Mpc), the average profile 
    is unaffected by seeing within $\sim 1$ kpc.
    Most galaxies of the CGS sample are not in any cluster.
    The median profile qualitatively agrees with the median profile of 
    \nbcg{} in the low mass bin, especially in the inner region. 
    Outside the inner 15 kpc, the CGS median profile shows a slightly more
    prominent outer envelope.  
    While this could simply reflect the small size of CGS sample ($\sim 30$),
    interestingly, it is also consistent with the expectation if the 
    mild mass growth from $z\sim 0.4$ mostly happened at the outskirt.   
    The CGS images are already slightly deeper than SDSS images in $r$-band,
    however, the median profile can only reach to $\sim 50$ kpc (with much 
    larger uncertainty compared to this work) for ETGs within $\sim 100$ Mpc.
    Such comparison clearly highlights the improvement made by the deep HSC 
    images as individual profile of $z\sim 0.5$ galaxy is reliable out to 
    at least 100 kpc.

%% ------------------------------------------------------------------------------------ %% 
  \begin{figure*}[t]
      \centering 
      \includegraphics[width=14.0cm]{fig/redbcg_prof_m100C}
      \caption{Comparisons of the \mden{} profiles for \rbcg{} (orange-red) and \nbcg{} 
      	(grey-black) galaxies in lower (left; [11.5,11.70]) and higher 
        (right; [11.7, 11.9]) \mtot{} bins. 
        Other formats are in consistent with the right figure of 
        Fig.\ref{fig:prof_m100_A}.
        The difference in median profiles is more significant in higher \mtot{} bin.}
      \label{fig:prof_m100_C}
  \end{figure*}
%% ------------------------------------------------------------------------------------ %%

    At the similar redshift with our sample, we compare with mass density 
    profile extracted from stacked images of massive ETGs at $0.25 < z < 0.50$ 
    using HST/ACS images (\citealt{Patel2013}).  
    These galaxies are selected at constant cumulative number density, and 
    as progenitors of $z=0$ massive ETGs.  
    The average stellar mass is $\sim 10^{11.2} M_{\odot}$ according to 
    their estimates.  
    However, given the shallower depth of the ACS images and the usage of the 
    BC03 model (each could cause -0.1 dex mass offset compared to this work), 
    this sample should be comparable to the \nbcg{} sample in the
    lowest mass bin, and the comparison of their median profiles does reflect
    that.  
    Moreover, since the profile from HST/ACS image is not bothered by PSF 
    within 1 Kpc, the comparison illustrates that the HSC median profiles 
    in our redshift range are indeed not affected by seeing outside the 
    shaded region. 
    In fact, as we suggested, our estimate is quite conservative, the profiles 
    within 5 kpc should be reliable.

    Besides these observations, we also compare with the predicted average 
    stellar mass density profile for central galaxies of massive halos 
    ($13.5 < \log M_{200,c} < 14.0$) in simulation (\citealt{Cooper13}). 
    This is achieved by combining the detailed assembly information of halos 
    from dark-matter-only simulation with semi-analytic model and 
    particle-tagging technique (\citealt{Cooper10}).
    Although the physical resolution and detailed baryonic processes are 
    not as ``realistic'' as state-of-art hydro-simulation, it is very 
    efficient in providing rough stellar mass distributions for large number 
    of massive halos. 
    Without any adjustment, the profile is quite similar to the \nbcg{}
    sample in the $11.5 < \log M_{\star} < 11.7$ bin within inner 20 kpc; 
    but it predicts a too significant outer envelope for its average halo mass.
    Without going into details to find the cause of such discrepancy in this 
    work, the comparison already suggests that the comparison between deep 
    observed mass density profiles with the ones from modern simulations is 
    going to be quite interesting, especially in the outskirt.

    We summarize the median profiles in Fig.\ref{fig:avg_prof} in Table.~1,
    and all the above median profiles are available  
    \href{https://github.com/dr-guangtou/hsc_cenHighMh/tree/master/profiles}{here}.

%% ------------------------------------------------------------------------------------ %% 
  \begin{figure*}[t!]
      \centering 
      \includegraphics[width=\textwidth]{fig/redbcg_prof_m10C}
      \caption{\alexie{There is a lot of redundancy in the Figures. I'm going to suggest to bring all the main figures together in a 4 panel figure?}. Comparisons of \mden{} profiles of \rbcg{} (orange-red) and \nbcg{} 
          (grey-black) samples that are matched using \minn{} instead of \mtot{}. 
          We perform the match at $11.2<\log (M_{\star,10\ \mathrm{kpc}}/M_{\odot})<11.6$.
          \textbf{Left figure} shows the matching results in the same format with 
          the left figure of Fig.\ref{fig:prof_m100_A}.  
          The only difference is \mtot{} is replaced by \minn{}. ~~~
          \textbf{Middle figure} summarizes the profile comparisons in the same format 
          with the right figure of Fig.\ref{fig:prof_m100_A}.~~~
          \textbf{Right figure} is very similar to the middle one, except we show the 
          profiles that are normalized using the \minn{} value to highlight the 
          scatters of \mden{} at different radius. 
          \alexie{the tow figures actually look pretty similar. I think most people will not notice the difference. Pick one and go with that. I guess I have a preference for the unnormalized one. Your mathced sampeles take care of the reason we made the normalized versions. (mean mass was different between samples.)}
          \song{Ok, will only use the unnormalized ones later; The point of normalized one is to 
          reduce the effect of the \mstar{} width of the mass bin; it is useful when I 
          discuss the dispersion of \mden{} at outer radius; I could simply emphasis that 
          in the text, or put some figures in the Appendix....how many appendix is too
          many}}
      \label{fig:prof_m10_A}
  \end{figure*}
%% ------------------------------------------------------------------------------------ %%

%% ------------------------------------------------------------------------------------ %%

\subsubsection{Matched Samples Using \mtot{} \alexie{This it not the right title for this section!! The point here is not to make matched samples. The point here is to look at the enviromental signature.}}
    \label{sssec:sbp_mtot}

\alexie{Most of the text below describing the procedure for making the mathced samples can go into an appendix. It distracts from your main results! And people have been making mathced samples for years!! Most of the text in between the red text can be moved else where or simplified.}

  \textcolor{red}{FROM HERE}

  As stellar mass is considered the most important parameter in determining the 
    structure and other properties of massive galaxies, it is crucial to make sure 
    the comparison is conducted at ``fixed stellar mass''. 
    And, since we directly compare the stellar mass distributions across very large 
    range in radius, it is also important to make sure that differences in physical 
    size of PSF (for more discussion on this issue, please see Appendix\ref{app:C}) 
    and imaging depth do not affect the result.    
    In practice, we achieve these by matching the \rbcg{} and \nbcg{} samples in 
    the distributions of \mstar{} and redshift in the mass range where both 
    samples have decent completeness.  
    Since the \rbcg{} sample is much smaller in size, we match the \nbcg{} sample to it 
    by searching for the $N$ nearest neighbours on the $M_{\star}$-redshift 
    plane using the KDTree algorithm provided by \texttt{scikit-learn} Python 
    library (\citealt{scikit-learn}). 
    The quality of the matched sample is evaluated by not just the median values 
    of stellar mass and redshift, but also the kernel density distributions (KDE) 
    of these two properties.
    For \mstar{}, we use a Gaussian kernel with width equals to the typical 
    uncertainty of the mass estimate (0.06-0.10);
    and for redshift, we choose 0.025 as the width of the Gaussian kernel. 
    As we only keep unique \nbcg{} galaxies in the matched sample, we manually adjust 
    the value of $N$ to achieve the most similar distribution. 
    In case that the distribution of stellar mass or redshift for \rbcg{} is 
    bi-model, we also try to split the sample into two, and match them separately.
    Normally $N$ is between 3 to 8.

    \mstar{} within different radius traces different physical processes and 
    different epochs in the assembly history.  
    In this section, we match the two samples using our proxy of ``total'' stellar 
    mass, which is the \mstar{} within 100 kpc radius.  
    This choice reflects the stellar mass distributions we can safely measure using 
    model-independent approach given the current depth of the image and the accuracy 
    of sky background subtraction.
    For galaxies at $11.6 \leq \log(M_{\star}/M_{\odot}) < 11.9$, 100 kpc equals to 
    5-8 times of their effective radius, hence should enclose most their stellar mass.
    Also, for a fraction of massive \rbcg{} galaxies, the intra-group/cluster light
    (IGL/ICL) could dominate their mass profiles outside 100 kpc (\addref), and 
    complicate the comparison with the \nbcg{} galaxies.  
    We will discuss these issues later in Section \ref{sec:discussion}, but in 
    conclusion, the proxy of total \mstar{} and the potential ICL component should 
    not affect the results.

    \textcolor{red}{TO HERE}

    \alexie{This is more how you should start your paragraph! It can be as simple as "We match the sample (see appendix) and show the results in Figure xx.}
    \song{Got it! will update later}
    We show the results of the comparison in Fig.\ref{fig:prof_m100_A}. 
    The left panel highlights the properties of the matched \rbcg{} and \nbcg{} 
    samples.
    Given the typical uncertainty of stellar mass, the two samples are well 
    matched in \mtot{}; and their redshift distributions are broadly 
    consistent with each other.  
    We also check their distributions of $k$-corrected colors ($g-r$), and confirm 
    that they have very similar average color distributions as well.  
    For the 45 \rbcg{} galaxies, we have 229 unique \nbcg{} galaxies in the 
    matched sample.
    On the right side, we plot the individual stellar mass density (\mden{}) profile 
    of \rbcg{} and \nbcg{} galaxies along with the median profiles of both samples. 
    The uncertainties of the median profiles are derived using 5000 times of 
    bootstrap resampling.
    As one can see, we can comfortably trace the \mstar{} distribution of these massive 
    galaxies out to 100 kpc at individual level, which gives us huge advantage in 
    studying the statistical properties of their outer envelopes. 
    Mass profile out to such large radius was only available for large sample of 
    massive galaxies through image stacking method (\addref) which suffers from certain
    systematic issues. 
    Although the \mden{} profiles of many galaxies extend well beyond the 100 kpc radius, 
    we notice that some profiles show sign of unphysical truncation that is caused 
    by inaccurate background subtraction.  
    Therefore, we will not include profiles outside 100 kpc in the comparison, even 
    though the median profiles within 200 kpc still behave normally.

    Once we match the distributions of \mtot{}, the \mden{} profiles of massive 
    galaxies in the \rbcg{} and \nbcg{} samples \textbf{greatly overlap with each 
    other over the entire radius range}. 
    At fixed ``total'' stellar mass, massive central galaxies of more massive 
    halos do not form a unique population comparing to the ones from less massive 
    halos in term of the stellar mass distributions.

%% ------------------------------------------------------------------------------------ %% 
  \begin{figure*}[t!]
      \centering 
      \includegraphics[width=16.0cm]{fig/redbcg_m100_m10_compare}
      \caption{Three-color images for a subsample of massive galaxies at similar redshift
      		($\sim 0.4$), and with similar \mstar{} inside inner 10 kpc (indicated by the 
            dash-line cicle on top-left figure; 
            $11.2<\log (M_{\star,10\ \mathrm{kpc}}/M_{\odot})<11.3$). 
            However, they span quite a large range in \mstar{} within 100 kpc 
            (highlighted by the solid bar on the top-left figure). 
            From top-left to bottom-right, their \mtot{} increases from 
            $10^{11.2} M_{\odot}$ to $10^{11.7} M_{\odot}$.
            The three galaxies from \rbcg{} sample are highlighted via red frames.}
      \label{fig:m100_m10_color}
  \end{figure*}
%% ------------------------------------------------------------------------------------ %%

    At the same time, thanks to the high quality individual profile from HSC image, 
    we can also spot interesting differences in the median profiles of the \mtot{} 
    matched \rbcg{} and \nbcg{} samples.  
    In details, it seems that central galaxy in more massive halos tends to have 
    a slightly flattened inner density profile and possess more significant outer 
    envelope.   
    To highlight such subtle structural features, we show the difference between 
    the two median profiles and its uncertainty in the bottom panel. 
    Given the noticeable uncertainties of \mtot{} and $\lambda$, and the current small 
    sample sizes, we also perform statistical test to confirm the robustness of the
    result: we mix the $N_{\mathrm{r}}$ \rbcg{} galaxies and the $N_{\mathrm{n}}$ \nbcg{}
    ones together, then randomly draw $N_{\mathrm{r}}$ galaxies with putting-back, 
    compute the difference between the median profile of this sample and the \nbcg{} 
    sample, and repeat this process 2000 times.  
    Comparing with the statistical distributions of the differential profiles from 
    this test (grey shaded region), we can now conclude that the structural 
    differences between \rbcg{} and \nbcg{} samples are much more significant than 
    any statistical fluctuation given the current samples.  
    And, apparently, the ``transition radius'' where the median profiles of \rbcg{} 
    and \nbcg{} sample have the same \mden{} is $\sim 15$-20 kpc, which is quite 
    close to the expected $R_{\mathrm{e}}$ of ETGs at this stellar mass.

    To further test the robustness of this interesting result, we repeat above 
    procedures to samples with slightly different definition.  
    The results are shown in Fig.\ref{fig:prof_m100_B}: 
    (a). At higher redshift, the profile is more affected by the seeing at inner 
    region, and by the imaging depth in the outskirt.  
    Therefore, in the left panel, we compare the \rbcg{} and \nbcg{} samples at 
    $0.3 \leq z < 0.4$, and the results are very similar.
    (b). We also make the comparison using a \rbcg{} sample with lower average 
    halo mass ($\lambda \geq 20$), the results are still stable (middle panel). 
    We also check whether the structural differences are driven by small fraction 
    of \rbcg{} galaxies in very massive clusters.  
    After removing the \rbcg{} galaxies in clusters with $\lambda \geq 40$, the 
    differences are as significant as above.  
    (c). As noted before, the \mtot{} is just a proxy of the total stellar mass. 
    Given the differences we find, at fixed \mtot{}, \rbcg{} galaxy tends to have 
    more prominent outer envelope outside 100 kpc than the \nbcg{} one. 
    Hence their real ``total'' stellar mass could also be slightly larger, and 
    result in bias in the comparison.  
    In the right panel, we make the comparison using the maximum stellar mass from 
    the integration of the \mden{} profile instead of the \mtot{}.
    In this case, the differences between the median profiles become slightly 
    less significant as expected.  
    But the systematic trends, especially the differences in the low mass density 
    outskirt are still very stable.  
    All these tests confirm that we robustly detect subtle, but systematic 
    structural dependence on halo mass of massive galaxies.   
    The fact that it requires such careful comparison to reliably reveal any 
    potential halo mass dependence already suggests that the role environment
    played in shaping these massive galaxies is not crucially important.  
    At the same time, the systematic differences we find, especially the more 
    extended stellar envelope of massive central galaxy in more massive halos, 
    seem to be consistent with the expectation of richer (minor) merging history
    in denser environment. 
    We will discuss these more in Section \ref{sec:discussion}.

    We also explore the possible \mtot{} dependence of such difference.  
    Limited by the small sample we have, we can only afford to separate the 
    samples into two \mtot{} bins 
    ($11.5 \leq \log (M_{\star,\ 10\mathrm{kpc}}/M_{\odot}) < 11.7$ and 
     $11.7 \leq \log (M_{\star,\ 10\mathrm{kpc}}/M_{\odot}) < 11.9$).  
    After performing the same procedures of sample matching and comparison, 
    we show the results in Fig.\ref{fig:prof_m100_C}.  
    Although the smaller sample leads to larger statistical uncertainties, 
    it is still interesting to see that the structural differences seem to 
    be more significant in the higher \mtot{} bin.  
    For the lower \mtot{} bin, the difference in the inner region becomes quite 
    uncertain, while there is still evidence that suggest difference in the 
    outskirt.  
    This result also has interesting implications in studying the assembly 
    of massive galaxies, and is worth further investigation with larger sample
    in the future.

%% ------------------------------------------------------------------------------------ %%

%% ------------------------------------------------------------------------------------ %% 
  \begin{figure}[bt!]
      \centering 
      \includegraphics[width=\columnwidth]{fig/redbcg_prof_m10B}
      \caption{Comparisons of unnormalized \mden{} profiles of 
      	  $\log (M_{\star,100\ \mathrm{kpc}}/M_{\odot}>11.5$ \rbcg{} (orange-red) 
      	  and \nbcg{} (grey-black) samples that are matched using \minn{}.
          Except for the difference in \mtot{}-limit of both samples, other formats are 
          the same with the middle figure of Fig.\ref{fig:prof_m10_A}.
          \song{The reason I did this test is because the \minn{} matched \nbcg{} 
          sample without any constraint will have a lot of galaxies with 
          \mtot{}$<10^{11.5} M_{\odot}$ in the sample; They are in the 
          \mtot{}-incomplete range, and it is harder to argue that they have similar 
          mass in the `in-situ' component with the \rbcg{} ones. 
          If you think this is weird, another choice is to do the match at \minn{} in 
          between [11.4, 11.6], so that most galaxies have 
          \mtot{}$>10^{11.5} M_{\odot}$.}}
      \label{fig:prof_m10_B}
  \end{figure}
%% ------------------------------------------------------------------------------------ %%

\subsubsection{Matched Samples Using \minn{}}
    \label{sssec:sbp_minn}

    \song{Got it! will update later}
    \alexie{This paragraph goes in the discussions section. The results section should be "We now do this to the data and here is the Figure"} As suggested by the popular two-phase formation scenario of massive galaxies, 
    the inner 5-10 kpc of these galaxies (within 0.5-1$xR_{\mathrm{e}}$) is 
    dominated by stars formed in intensive dissipative processes at high redshift, 
    and is less affected by the succeeding merging events.  
    This picture is broadly supported by the redshift evolution of \mden{} profile
    of massive galaxies (\addref), structure decomposition of nearby massive galaxies 
    (\addref), and their radial distributions of stellar population (\addref).      
    Certainly, major merger can induce violent relaxation process that modifies the 
    stellar mass distribution even in the central region, and must happen to 
    these massive galaxies in last 6-8 Gyrs, it may not invalidate the above assumption.
    For galaxy this massive, the last major merger event is highly likely to be
    non-dissipative, and the redistribution of stars is basically according to their 
    binding energy in the host halo.  
    The inner region of the final merger remnant is still mostly composed 
    of the ``in-situ'' stars from the two progenitors.

    Therefore, it is interesting to compare the \rbcg{} and \nbcg{} samples that are 
    matched using mass within the inner region.  
    In this work, we choose to use the mass within inner 10 kpc (\minn{}). 
    As we can see in previous profile comparison, \minn{} is not affected by the 
    physical size of PSF out to $z=0.5$.  
    We use the same procedure to match the \nbcg{} galaxies to the \rbcg{} ones using 
    \minn{} (within $11.2 \leq \log (M_{\star,\ 10\mathrm{kpc}}/M_{\odot}) < 11.6$) 
    and redshift ($0.3 \leq z < 0.5$).  
    For the 56 galaxies in this \rbcg{} sample, we achieve excellent matching 
    result using 375 \nbcg{} galaxies (see Fig.\ref{fig:prof_m10_A}).  
    Using the same method, we show their individual profiles, and the comparison of 
    their median \mden{} profiles in Fig.\ref{fig:prof_m10_A}.  
    Unlike the \mtot{} matched samples, here we see very striking differences that 
    are already very significant even without any statistical test.     
    As we can see, when matched using \minn{}, the \mden{} profiles of \rbcg{} and 
    \nbcg{} galaxies have quite similar \mden{} profiles in the inner 5-10 kpc, 
    in great contrast of the differences at outer region where the \rbcg{} galaxies 
    show much more extended envelope.      
    Besides the difference in the median profiles, it is also worth discussing 
    the distributions of individual \mden{} profiles. 
    Combining the \rbcg{} and \nbcg{} samples together, it is easy to notice that 
    the dynamical range of \mden{} at fixed radius is much larger in the outskirt. 
    This becomes even more clear when we further normalize each profile using 
    the \minn{} value (see Fig.\ref{fig:prof_m10_A}).   
    For both \rbcg{} and \nbcg{} samples, we find that their outer \mden{} profiles 
    clearly show much larger diversity comparing to the inner region.

    Putting together, these results indicate that massive central galaxies with 
    similar \minn{} and inner mass profiles can span a huge range in total stellar 
    mass, and can have very different mass distributions in the outskirt. 
    We further highlight this by showing the 3-color images of randomly selected 
    massive central galaxies (see Fig.\ref{fig:m100_m10_color}) that 
    have similar \minn{} 
    ($11.1 \leq \log (M_{\star,\ 10\mathrm{kpc}}/M_{\odot}) < 11.2$)
    but different \mtot{}.   
    From the top-left to the bottom-right, the \mtot{} of these galaxies increases
    from $10^{11.22} M_{\odot}$ to $10^{11.75} M_{\odot}$.

    Since the two samples are matched using \minn{}, the \nbcg{} sample starts to 
    include galaxies with 
    $11.2 \leq \log (M_{\star,\ 100\mathrm{kpc}}/M_{\odot}) < 11.6$.  
    On one hand, they are likely to be the central galaxies in halos that are 
    much smaller than the ones in \nbcg{} sample, which still mostly consists of 
    galaxies with $\log (M_{\star,\ 100\mathrm{kpc}}/M_{\odot}) \geq 11.6$.  
    Hence, above results may hint that the outer \mden{} profile of central galaxy 
    could have a clear dependence on the halo mass.
    This is a very interesting clue that deserves more investigations including 
    comparison with simulations.  
    But, on the other hand, we need to recognize that the mass completeness of 
    our current \nbcg{} sample is still not satisfying; and in this lower \mtot{} 
    range, small fraction of satellite contamination should be expected even 
    though we already excluded satellites in known cluster using the \redm{} catalog. 
    To naively test the robustness of these results, we match the two samples using 
    \minn{} again, but only include galaxies with 
    $\log (M_{\star,\ 100\mathrm{kpc}}/M_{\odot}) \geq 11.5$.  
    Comparison between these two more complete, purer samples of massive 
    central galaxies at fixed \minn{} reveals very similar behavior
    \footnote{In fact, their median profiles become even more similar in the 
    inner region} (see Fig.\ref{fig:prof_m10_B}), supporting the robustness of the 
    results.  
    And, we also try to match the two samples using mass within 5 and 15 kpc 
    instead, the results are also very similar.

%% ------------------------------------------------------------------------------------ %%

\subsection{Explore Different Scaling Relations}
    \label{ssec:scaling}

%% ------------------------------------------------------------------------------------ %% 
  \begin{figure}[bt!]
      \centering 
      \includegraphics[width=\columnwidth]{fig/redbcg_mass_r50}
      \caption{The \mstar{}-size relations for \rbcg{} (orange-red circles) and \nbcg{} 
          (grey dots) galaxies. 
          Here we use the \mtot{} as proxy of total \mstar{}, and use the 
          $R_{\mathrm{50}}$ derived from the 1-D profile as size estimate. 
          The size and color coding for \rbcg{} galaxies are the same with the right panel 
          of Fig.\ref{fig:smf1}.  
          And two vertical lines highlight the $11.6<$\logms{}$<11.9$ mass range.
          We also compare these distributions with the relation for nearby central 
          galaxies of massive groups from \citet{HCompany13} (green, long dashed line). 
          The \mstar{} and $R_{\mathrm{e}}$ are estimated via different methods in these
          works. 
          We also illustrate best-fit \mstar{}-$R_{50}$ relations for \rbcg{} (red, solid 
          line)and \nbcg{} (grey, dash line) galaxies that have \logmtot{}$\geq 11.6$.
          The shaded regions in lighter color show the 1-$\sigma$ uncertainties of these 
          relations.
          Please refer to the text for detailed comparisons.}
      \label{fig:mass_r50}
  \end{figure}
%% ------------------------------------------------------------------------------------ %%

\subsubsection{$M_{\ast}-$size relation}
    \label{sssec:mass_size}

    The scaling relation between \mstar{} and effective radius (or half-light radius) 
    was the focus of past investigations of the structural evolution (e.g. \addref) and 
    the environmental dependence of structures for massive galaxies (e.g. \addref{}). 
    So far, the results are quite confusing as no clear evidence of environmental 
    dependence has been found at low redshift using various definitions of ``size''. 
    Also, at high redshift, the results can be even controversial
    (e.g. \addref).  
    As the deep HSC images have already help us reveal subtle, but systematic structural
    differences in massive central galaxies that may depend on the environment, we should 
    expect to see such evidence on the mass-size relation. 
    Here, we briefly discuss the mass-size relations of \rbcg{} and \nbcg{}
    galaxies (Fig.\ref{fig:mass_r50}).

    We choose to use the \mtot{} as the proxy of ``total'' \mstar{}, and use the 
    half-light radius ($R_{\mathrm{50}}$) that is defined as the radius enclosing 50\% of 
    luminosity within 100 kpc as the ``size'' to evaluate the relation.
    And, we use the curve-of-growth of luminosity from the \texttt{Ellipse} run on 
    $i$-band images to estimate the $R_{\mathrm{50}}$ here.
    In principle, such approach is more sensitive to issues like the smearing effect of 
    PSF and the choice of aperture for maximum luminosity than the 2-D fitting method.
    For the massive galaxies in this work, their apparent sizes are always sufficiently 
    large at $0.3 \geq z \geq 0.5$ so that the $R_{\mathrm{50}}$ measurement is not 
    bothered by PSF.   
    And, the 1-D method has the advantages of being less affected by accuracy of background
    subtraction, structural details, and model assumptions.

    We show the distributions of \rbcg{} and \nbcg{} galaxies on the mass-size plane.  
    On top of the clear correlations for both sample, the distributions of $R_{\mathrm{50}}$ 
    for \rbcg{} and \nbcg{} samples greatly overlap with each other over the \mtot{} range
    between $10^{11.6}$ and $10^{11.9} M_{\odot}$. 
    Meanwhile, it seems that the \rbcg{} galaxies do have a slightly higher 
    $R_{\mathrm{50}}$. 
    For the \mtot{} matched samples used in Fig.\ref{fig:prof_m100_A}, the median 
    $\log (R_{\mathrm{50}}/\mathrm{kpc})$ values for the \rbcg{} and \nbcg{} are 
    $1.22\pm 0.02$ ($\sim 17.4\pm 0.75$ kpc) and $1.09\pm 0.01$ 
    ($\sim 12.3\pm0.3$ kpc), which confirm the above observation and is consistent with 
    the differences seen in Fig.\ref{fig:prof_m100_A}. 
    However, all previous works that we are aware of did not match the mass (and redshift) 
    distributions between different samples, and often focus on the form of the 
    \logms{}-$\log R_{\mathrm{e}}$ relation.  
    Here, we try to derive the best-fit parameters for the 
    \logmtot{}-$\log (R_{\mathrm{50}}/\mathrm{kpc})$ relations of galaxies with 
    \logmtot{}$\geq 11.6$ using MCMC-sampling method via the ensemble sampler 
    \texttt{emcee} (\citealt{Emcee}). 
    Using initial guesses based on maximum likelihood estimates, and assuming reasonable 
    flat priors, we derive best-fit relation for the massive \rbcg{} galaxies:

    \begin{equation}
        \begin{aligned}
        \log (R_{\mathrm{50}}/\mathrm{kpc}) = & (0.73\pm0.13) \times \log (M_{\star, 100\ \mathrm{kpc}}/M_{\odot}) \\ & -(7.49\pm1.56)
        \end{aligned}
    \end{equation}

    \noindent And for the massive \nbcg{} galaxies, we have:

    \begin{equation}
        \begin{aligned}
        \log (R_{\mathrm{50}}/\mathrm{kpc}) = & (0.68\pm0.06) \times \log (M_{\star, 100\ \mathrm{kpc}}/M_{\odot}) \\ & -(6.88\pm0.75)
        \end{aligned}
    \end{equation}

    \noindent As shown in Fig.\ref{fig:mass_r50}, at the high-mass end, the two relations 
    have consistent slopes within their uncertainties, but different normalizations. 
    Extending the fitting to the full \logmtot{} range does not change this result. 
    We also explore mass-size relations for our samples using \mstar{} within different 
    radius (e.g. 120, 150 kpc) and the $R_{\mathrm{50}}$ derived within these apertures, 
    and find consistent results.

    We qualitatively compare our results with the one for $z\leq 0.09$ central ETGs in  
    $14\le$\logmh{}$<15$ halos from \citealt{HCompany13} (HC13; green solid line).
    HSC13 use the group catalog by \citet{Yang2007} to estimate \logmh{}. 
    They estimate the 2-D $R_{\mathrm{e}}$ using single-\ser{} model fitting to SDSS images,
    and derive \mstar{} based on SED fitting using the BC03 (\citealt{BC03}) synthetic 
    population model.
    We empirically convert their \mstar{} from the Kroupa IMF to the Chabrier one 
    used in this work by applying a constant -0.03 shift to their values. 
    We also increase the \mstar{} in HC13 by $+0.1$dex to account for the systematic 
    difference between their BC03 and our FSPS model (see Appendix \ref{app:B}). 
    In HC13, the authors found no difference among the mass-size distributions for 
    central galaxies in halos across $12.5\le$\logmh{}$<15.0$. 
    Despite all the systematic differences, the mass-size distributions for both \rbcg{} 
    and \nbcg{} galaxies follow the HC13 relation reasonably well (with slightly
    shallower slopes) even at \logmtot{}$< 11.6$ where our samples start to become 
    incomplete.

    In short, we confirm that the massive central galaxies in this work generally 
    follow the mass-size relation of massive ETGs at low-$z$.
    And, more importantly, we find evidence of differences in the mass-size relations for 
    massive central galaxies of halos with different \mhalo{}. 
    The difference we saw at \logmtot{}$\geq 11.6$ is in line with the conclusion that
    central galaxies in denser environment have slightly larger size comparing to the 
    ones from less-dense environment at fixed \mstar{}.

    This suggests that the previous, inconclusive results on whether the 
    mass-size relation of massive ETGs depends on environment could be due to 
    either the inadequate depth of the image, or the inappropriate model assumptions 
    adopted during the fitting (or both).  
    It would be interesting to revisit this issues using 2-D models based on deep HSC 
    images as the next step.
    It is worth noting that our results are based on massive galaxies at $z\sim 0.4$, 
    while the sample of HC13 has $z<0.09$.
    Although here we can not address the interesting issue of redshift evolution of 
    mass-size relation, the ``evolutionary path'' of massive central galaxies from 
    different environments on this mass-size plane is certainly an intriguing issue. 
    If the \rbcg{} galaxies have a richer merging history than the \nbcg{} ones, 
    perhaps we should expect more significant differences on their mass-size 
    relations and average \mden{} profiles at $z\sim 0$ using deeper image. 
    On the other hand, if the \nbcg{} galaxies could assemble more \mstar{} through minor 
    mergers in the last 3-4 Gyrs, they could ``catch-up'' with the \rbcg{} galaxies 
    a little bit on the mass-size plane. 
    Potentially, we could make some predictions on this issue using the frequency of 
    tidal features and properties of satellites around our \rbcg{} and \nbcg{} 
    galaxies in the near future.

%% ------------------------------------------------------------------------------------ %%

%% ------------------------------------------------------------------------------------ %% 
  \begin{figure*}[t]
      \centering 
      \includegraphics[width=\textwidth]{fig/redbcg_m100_m10_new}
      \caption{\textbf{(a)}: Distributions of galaxies on the 
          \mtot{}-$\Delta (\log M_{\star})_{100\mathrm{kpc}-10\mathrm{kpc}}$ plane. 
          All \rbcg{} (open red boxes) and \nbcg{} (light, grey dots) galaxies 
          are shown in the background. 
          The \mtot{}-matched samples between $11.6<$\logms{}$<11.9$ (between two 
          vertical lines) are highlighted (\rbcg{}: solid red boxes; 
          \nbcg{}: bigger grey dots). 
          We also show the median values (solid triangular and hexagon), and the
          distributions of $\Delta (\log M_{\star})_{100\mathrm{kpc}-10\mathrm{kpc}}$
          (both histograms and kernel density distributions on the right).
          The best scaling relations for \rbcg{} (red, solid) and 
          \nbcg{} (grey, dashed) galaxies along with their uncertainties 
          are illustrated as well.~~
          \textbf{(b)}: Distributions of galaxies on the 
          $\Delta (\log M_{\star})_{100\mathrm{kpc}-10\mathrm{kpc}}$-\minn{} plane.~~
          \textbf{(c)}: Distributions of galaxies on the 
          $\Delta (\log M_{\star})_{100\mathrm{kpc}-10\mathrm{kpc}}$-\mtot{} plane.
          But, this time, we highlight the \rbcg{} and \nbcg{} samples that are 
          matched using \minn{} instead of \mtot{}.~~
          \textbf{(c)}: Relations between \mtot{} and \minn{}. 
          Other formats of figure (b), (c), and (d) are very similar with figure (a).}
      \label{fig:m100_m10_A}
  \end{figure*}
%% ------------------------------------------------------------------------------------ %%

\subsubsection{Relations between stellar mass within different physical apertures}
    \label{sssec:m100_m10}

    We just show that the mass-size relation of massive central galaxies could also 
    depend on the mass of their host halos, and here we explore relationships between 
    \mstar{} within different apertures as alternative tools to help
    us better understand the relationship between halo mass and assembly history of 
    massive galaxies.  
    Here we focus on \mtot{}, \minn{}, and also the logarithmic mass differences 
    between them ($\Delta \log M$, reflecting \mstar{} ratio). 
    Comparing to the mass-size relation, such relations are not affected by the ambiguous 
    definition of ``size''; and comparing to the detailed \mden{} profiles, 
    it is much easier to apply these relations to larger samples and compare with 
    results from simulations.

    We show the results in Fig.\ref{fig:m100_m10_A}.  
    In panel (a), we plot the \mtot{} with $\Delta \log M$ for both the\rbcg{} 
    and \nbcg{} samples. 
    As expected, these two parameters well trace each other over the entire 
    \mtot{} range, indicating that more massive central galaxies have larger 
    fraction of total \mstar{} in the outer part. 
    Suggested in Fig.\ref{fig:prof_m100_C}, the structural differences related to 
    \mhalo{} seem to be more significant in higher \mstar{} bin. 
    We therefore highlight the \mtot{}-matched \rbcg{} and \nbcg{} galaxies at 
    $11.7 \leq$\logmtot$< 11.9$. 
    Although their $\Delta \log M$ distributions still greatly overlap within 
    this \mtot{} range, the median values (big, outlined symbols with error bars),
    and their kernel-density distributions clearly point out systematic difference. 
    Consistent, but slightly smaller, difference also exists for \mtot{}-matched 
    samples at $11.6 \leq$\logmtot$< 11.9$. 
    At the same time, the comparisons in Fig.\ref{fig:prof_m100_A} and 
    Fig.\ref{fig:prof_m100_C} also suggest that the \mtot{}-matched \rbcg{} galaxies 
    have lower \minn{} than the \nbcg{} sample, which is clearly shown by panel (b)
    where we plot $\Delta \log M$ with \minn{} for the \mtot{}-matched samples.
    For the \nbcg{} sample, the correlation between between $\Delta \log M$ and \minn{}
    is much weaker than the one in panel (a), yet it is still visible.  
    However, for our \rbcg{} sample, we see no overall correlation between 
    $\Delta \log M$ and \minn{}.  
    We also construct $\Delta \log M$ using logarithmic mass differences between 
    100 and 30 (or 50) kpc, and the results are very similar.  
    In fact, the average differences between the \mtot{}-matched \rbcg{} and \nbcg{} 
    samples become more significant when mass within 30 (or 50) kpc is used.        
    These reflect the structural differences we found in Fig.\ref{fig:prof_m100_A}: 
    at fixed \mtot{}, central galaxies with higher \mhalo{} tend to have even higher 
    fraction of \mstar{} stored between 10 and 100 kpc.

    In panel (c), we show the relationship between $\Delta \log M$ and \mtot{}, 
    and highlight the \minn{}-matched samples presented in Fig.\ref{fig:prof_m10_A} 
    and Fig.\ref{fig:prof_m10_B}. 
    Using the histograms of \mtot{} and their median values of the \minn{}-matched
    samples, we clearly identify the impact from the much more extended outer envelope 
    hosted by \rbcg{} galaxies.

    Finally, in panel (d), we plot \mtot{} against \minn{} directly.  
    The median values and statistical distributions of \minn{} for the \mtot{}-matched 
    samples highlight the structural difference at fixed \mtot{} again.  
    All the \mhalo{}-dependence of structures we discussed above are clearly 
    displayed here along the directions of both x- and y-axis.
    Firstly, within the \mtot{} range we explored, the \rbcg{} galaxies tend to have 
    lower mass within the inner 10 kpc; and such difference seems to become more 
    significant at higher \mtot{}. 
    As seen from the median \mden{} profiles, this is due to the shallower overall
    profile of \rbcg galaxy, and indicates that the slopes of both inner and outer
    \mden{} profiles of massive central galaxies may depend on \mhalo{}. 
    Secondly, for massive central galaxies that form similar amount of stars through
    the dissipative processes during the ``in-situ'' phase (approximated by \minn{}), 
    their \mtot{} distributions at lower redshift depend on \mhalo{} in a way that the
    ones in more massive halos can develop much more prominent stellar envelope thanks 
    to the more complex mass assembly history.

    But, more importantly, it is intriguing to see that the distributions of \rbcg{} and
    \nbcg{} galaxies on this 2-D plane seem to be offset with each other. 
    At least at the high mass end, the central galaxies in more and less massive halos 
    seem to follow relations between \mtot{} and \minn{} with similar slopes, but 
    different normalizations.  
    Instead of subjectively grouping galaxies into different bins of \mtot{} or \minn{}, 
    such scaling relation brings us a more generic view of the environmental 
    dependence of structure, and could become an interesting tool to investigate the 
    connection between halo mass, assembly history, and the structure of massive galaxies.
    On Fig.\ref{fig:m100_m10_A} (d), we illustrate the \logmtot{}-\logminn{} relations 
    at \logmtot$\geq 11.6$ that are derived using MCMC sampling method.  
    For \rbcg{} galaxies, the best-fit relation is:

    \begin{equation}
        \begin{aligned}
        \log (M_{\star, 10\ \mathrm{kpc}}/M_{\odot}) = & (0.48\pm0.06) \times \log (M_{\star, 100\ \mathrm{kpc}}/M_{\odot}) \\ & +(5.72\pm0.75).
        \end{aligned}
    \end{equation}

    \noindent While for the \nbcg{} ones within the same \mtot{} range, we have:

    \begin{equation}
        \begin{aligned}
        \log (M_{\star, 10\ \mathrm{kpc}}/M_{\odot}) = & (0.56\pm0.03) \times \log (M_{\star, 100\ \mathrm{kpc}}/M_{\odot}) \\ & +(4.82\pm0.30).
        \end{aligned}
    \end{equation}

    The same relation using the full \mtot{} range will results in a even shallower 
    slope for the \rbcg{} galaxies. 
    And, replacing the \minn{} with \mstar{} within 5 or 15 kpc apertures does not change 
    our conclusion. 
    In fact, the relation using $M_{\star, 5\ \mathrm{kpc}}$ shows more significant 
    \mhalo{} dependence at the high-mass end. 
    Unfortunately, due to the impact from PSF, the $M_{\star, 5\ \mathrm{kpc}}$ values 
    are slightly less reliable than the \minn{} ones. 
    Samples at lower redshift, images from space telescope, and 2-D modeling method can 
    all help us pushing this investigation further insider smaller physical aperture 
    that perhaps can provide us better surrogate of \mstar{} from the ``in-situ'' phase.

    Besides the systematic offset, it seems that the \rbcg{} galaxies could have 
    larger scatter of \minn{} at high-mass end. 
    Limited by our capability to estimate the mass and other properties of halos 
    at individual level, we can not further quantify this potential \mhalo{}-dependence 
    here.
    Within the \rbcg{} sample, we do not find any evidence of dependence on the richness 
    ($\lambda$) of the cluster even the \redm{} richness should traces \mhalo{}
    reasonably well.
    In the current sample, most of the \rbcg{} galaxies are centrals of clusters with 
    $30 \leq \lambda \leq 40$, and only small fraction comes from more massive clusters.  
    Together with the small sample size, the uncertainties of richness from shallow SDSS
    images, and the intrinsic scatter of the $\lambda$-\mhalo{} relation may conspire to 
    hide more detailed \mhalo{}-dependence.

    The evidence of \mhalo{}-dependence of structure in massive galaxies could be caused 
    by different processes, it is therefore of great interest for us to dig deeper by 
    comparing the galaxies at higher redshift or from numerical simulations (e.g. see 
    Fig.1 of \citealt{Wellons2016b}) on the \mtot{}-\minn{} or similar planes.

%% ------------------------------------------------------------------------------------ %%

%% ------------------------------------------------------------------------------------ %% 
  \begin{figure}[bt!]
      \centering 
      \includegraphics[width=\columnwidth]{fig/redbcg_discussion_1}
      \caption{Comparison of the average \mden{} profiles of massive galaxies in the 
          two lower \mtot{} bins from Fig.\ref{fig:avg_prof} with other observations 
          and simulations, focusing on the inner region. 
          The format is similar with the right panel of Fig.\ref{fig:avg_prof}.
          We include: 
          1) the average \mden{} profile of inner component from structure 
          decomposition of massive elliptical galaxies at $z\sim 0$ from 
          \citet[][Cyan, solid]{Huang2013a}; 
          2) the average \mden{} profile of massive galaxies at $1.0 < z < 1.5$ 
          from \textit{HST} observations in \citet[][Red, dashed]{Patel2013}; 
          3) the average \mden{} profile of the ``in-situ'' stellar components in 
          simulated massive halos from \citet[][Purple, dot-dashed]{Cooper13};
          4) the \mden{} profile of the very massive cD galaxy at $z\sim 1.1$ 
          discovered by \citet[][Yellow, dashed]{Liu2013} in the Hubble 
          Ultra-Deep Field.}
      \label{fig:discussion_1}
  \end{figure}
%% ------------------------------------------------------------------------------------ %%

\section{Discussions}
    \label{sec:discussion}

    \alexie{In some point of the paper (in the results section). We need to say what the difference is between M100 and the mass that you would get if you extrapolated outwards. How different is M100 potentially from a "total mass"?}
    \song{Yes, I can say a few words about that. I will send you with a figure that shows 
    the mass differences between 150 kpc and 100 kpc; At 150 kpc, we are already pushing 
    it to region where the photometry is uncertain, and affected by background subtraction.
    Most of the current works on ICL are in region smaller than this. 
    The key is 1) the mass increase is small, $<0.1$ dex, same level with the mass 
    uncertainty; 2) still depends on \mtot{}, more massive ones have more 
    extended outer profiles, hence more massive in outer regions.  
    Using 2-D fitting will give you higher stellar mass, as it extrapolates to 
    very large radius, but you have to assume at that radius, the profile follows
    the model defined at inner region, which is kind of dangereous as we show here.
    I could say this is worth exploring in the future.}
    In the past decade, the mass assembly and structural evolution of massive ETGs 
    remains as an active topic.  
    More and more convincing observations of massive galaxies at high redshift reveal that
    the progenitors of massive elliptical galaxies are quite different in structure
    comparing with their local decedents (e.g. \addref{}).  
    Collectively, they need to increase their ``size'' ($R_{\mathrm{e}}$) by a factor of 
    2 to 4 (e.g. \addref{}) without changing their inner \mden profile (e.g. \addref{})
    and central stellar velocity dispersion (e.g. \addref{}) much in the last 8-10 Gyrs.  
    Owing mainly to better numerical simulations, the challenge of ``size''-evolution 
    seems to be eased by the promising ``two-phase'' scenario under the framework of 
    hierarchical formation of galaxy (e.g.\addref{}). 
    Under this picture, the inner region of massive galaxies mostly consist of stars 
    formed within the halo of the main progenitor through intense dissipative processes 
    at high redshift.  
    Later, after the active star formation is quenched in these massive systems, their 
    mass assembly history becomes dominated by accretions of satellites.  
    Based on the prediction of merger tree and current understanding of satellite 
    population around massive galaxies, most these merging events are likely to be 
    minor ones, and without involving much gas dissipation.  
    Since such minor, dry merger basically redistributes stars onto the outer envelope 
    of the central galaxy, it becomes the most promising solution to the ``size-evolution''
    puzzle.  
    However, it raises additional questions at the same time. 
    The most pressing one would be the expected ``environmental''-dependence of structures 
    of massive central galaxies. 
    As minor merger is expected to be more frequent in denser environment, the mass assembly 
    history and \mstar{} distribution of central galaxies should show dependence on \mhalo{}
    or even other properties of the host halo.  
    For instance, centrals in more and less massive halos could have different 
    slopes and/or scatters in their mass-size relation (e.g. \addref{}).  
    Yet, no clear evidence of such dependence has been found. 
    In this work, taking advantage of the deep photometry from the HSC survey and looking 
    into the detailed \mden{} profiles, we clearly see evidence that the \textbf{structure
    of massive central galaxies, both the inner and outer parts, does show dependence on 
    the halo mass}. 
    Here, we briefly discuss the physical causes of this dependence, potential caveats, 
    and more scientific implications.

    Firstly, the comparisons of the \mden{} profiles suggest that \textbf{\mhalo{} does 
    not play a central role in shaping the structure of massive central galaxies}.  
    It appears that massive central galaxies at \logms{}$\geq 11.5$ were shaped by 
    same physical processes in more and less massive halos.  
    At both fixed \mtot{} and \minn{}, centrals from cluster-level halos do not form a 
    unique population. 
    Instead, their individual \mden{} profiles greatly overlap with the ones in halos of 
    small group. 
    At similar \mtot{}, even their median \mden{} profiles only show quite subtle 
    differences. 
    For the \rbcg{} sample, this is similar to the results of \citep{Zhao2015} that 
    a considerable fraction of nearby BCGs (34\%) does not fall into the ``cD'' class.
    From the smooth distributions of \mden{} profiles on Fig.\ref{fig:prof_m100_A} 
    and Fig.\ref{fig:prof_m10_A}, it is even harder to draw a clear difference 
    between ``cD'' galaxy and ``normal elliptical''.

    At the same time, \textbf{the \mhalo{}-dependence we found do fit the expectations 
    of the ``two-phase'' scenario at first glance}. 
    For central galaxies with similar stellar mass at the end of the first phase, 
    their subsequent, accretion-dominated mass assembly naturally relates to \mhalo{}, 
    which determines the number and mass distributions of satellites. 
    Therefore, centrals of more massive halos should accumulate more extended stellar 
    envelope during the second phase, exactly as we find in the \mden{} profile 
    comparison of the \minn{}-matched samples (see Fig.\ref{fig:prof_m10_A}).
    Under the ``two-phase'' formation scenario, the stellar content of massive galaxies
    can be broadly separated into the ``in-situ'' and ``accreted'' components. 
    At fixed \mtot{}, the differences of median \mden{} profiles could reflect 
    different relative fraction of these two components. 
    Recent simulations indicate that (1) the fraction of accreted stars could have 
    a steep correlation with total stellar mass; (2) for \logms$\geq 11.5$ galaxies,
    the accreted component starts to dominate within the $R_{\mathrm{e}}$.  
    Given that (1) the ``transition radius'' between the two median profiles is around
    15-20 kpc (see Fig.\ref{fig:prof_m100_A}); and (2) the differences become more 
    significant at higher mass end, it is likely the \mhalo{}-dependence of structure
    at fixed total stellar mass is driven by the higher fraction of accreted stars 
    in more massive halos.  
    To confirm this inference, in Fig.\ref{fig:discussion_1}, we compare the median 
    \mden{} profiles of the \mtot{}-matched \rbcg{} and \nbcg{} samples at 
    $11.5\leq$\logms{}$<11.7$ with the median profiles of 
    (1) massive ETGs at $1.0 < z < 1.5$ that are considered the progenitors of 
    $\sim 10^{11.5} M_{\odot}$ ETGs at $z=0$ (\citealt{Patel2013}).  
    After this redshift range, their mass growth should be dominated by mergers.  
    (2) inner components of $z=0$ $10^{11.4} M_{\odot}$ ellipticals from 2-D decomposition
    (\citealt{Huang2013a}). 
    \citep{Huang2013b} showed that these structures are quite similar to the compact 
    progenitors at higher redshift in many aspects.  
    (3) the ``in-situ'' components of simulated galaxies in massive halo of 
    \citep{Cooper13} using particle-tagging technique (the inner few kpc is quite 
    uncertain due to the resolution).  
    Regardless the small difference in median \mstar{} and details in \m2l{} estimates, 
    the comparison first confirms that \minn{} is indeed not a bad proxy of the in-situ 
    mass; meanwhile, it also highlights that structural difference between \rbcg{} and 
    \nbcg{} are driven by region that is not dominated by ``in-situ'' stars. 
    Since most of these samples in comparison are not as extreme as the massive 
    \rbcg{} galaxies, we also compare with a uniquely massive BCG at high redshift: 
    a $10^{11.4} M_{\odot}$ BCGs with distinctive envelope at $z\sim 1.1$ 
    (\citealt{Liu2013}).  
    It is interesting to see that its \mden{} profile follows the median profile of 
    \rbcg{} nicely until 20 kpc; then it becomes much steeper in the outskirt.  
    This is quite consistent with the expectation that the inner part of the BCGs 
    should be well developed at $z\sim 1$, but the extended stellar envelope is still 
    assembling.

    Also, it is worth noting that the relative flatten inner \mden{} profile of \rbcg{}
    at fixed \mtot{} could be an interesting feature, although we currently lack of the 
    resolution to investigate region $\leq 3$ kpc.  
    It is known that major-merger induced coalesce of super-massive black hoes (SMBHs)
    can create flattened inner profile (e.g. \citealt{Milosavljvi2002}), and recent 
    discoveries of massive BCGs with very large core 
    (a few kpc; e.g. \citealt{Postman2012, LopezCruz2014}) further complicate the picture.
    On Fig.\ref{fig:m100_m10_A}, there are a few \rbcg{} galaxies with quite low 
    \minn{} at high \mtot{} end. 
    After examining their images, we conclude the low \minn{} is not caused by problematic
    photometry or exceptionally bad seeing; they could also be massive BCGs with very 
    flattened inner \mden{} profiles.  
    Besides the SMBH-merger theory, strong adiabatic expansion induced by strong AGN 
    feedback was proposed to modify the inner mass distribution (e.g. \citealt{Fan2008}). 
    The potential impact of these processes to massive galaxies, and their dependence on 
    halo mass would be an interesting topic to investigate in the future.

	\song{I agree that this can be part of the results; but the profiles here are not as 
    reliable as the \mden{} profiles for technical reason; I put it here as a teaser, as 
    it is better to deal with them using 2-D fitting.}
\alexie{This paragraph shoud go in the results section. Indeed, here you are presenting a result that you measured. And what is the main message / lesson that we learn from this paragraph and figure? We sort of jumped rather abruptly from the enviromental dependence to ellipticity and color profles. Transition needs more motivation. A discussion of the implcations of this figure could go in the discussion section.}    As a start, we only focused on the 1-D \mden{} profile derived from $i$-band image.  
    We should not forget that the multi-band, 2-D HSC images of these massive galaxies 
    actually contain much more useful information, particularly the shapes and colors 
    of different stellar components.  
    Strictly speaking, to extract these information reliably, we need to rely on better 
    background subtraction and 2-D modelling method.  
    But, as a motivation, we shows initial results from 1-D analysis in 
    Fig.\ref{fig:discussion_2}.  
    The ellipticity profile is extracted from \texttt{Ellipse} run on $i$-band image 
    with free geometric parameters. 
    After reducing the $g$- and $r$-band images of our galaxies in consistent way with 
    the $i$-band ones, we simply apply the $i$-band isophotes to images in these two 
    bands without matching the seeing to extract color profiles. 
    All the color profiles are corrected for Galactic extinction and $k$-correction 
    using \texttt{iSEDFit} results. 
    We show the median ellipticity, $g-r$, and $g-i$ color profiles of the 
    \mtot{}-matched (\minn{}-matched) samples of \rbcg{} and \nbcg{} galaxies on the 
    left (right) panel; and also compare them with median profiles from stacking 
    analysis of SDSS galaxies with similar \mstar{} 
    (\citealt{LaBarbera2010, Tal2011, DSouza2014}) excluding the inner and outer regions 
    affected by seeing and uncertainty of background subtraction.  
    For the \mtot{}-matched samples on the left, we notice that the median ellipticity 
    profiles follow the stacked ones well between 5 to 30-50 kpc, but are not plateaued
    at larger radius.  
    Since they are the median of individually measured ellipticity profiles, they do 
    not suffer the systematics in stacking analysis, and are better at reflecting the 
    intrinsic profiles.  
    In consistent with the 2-D decomposition results in \citep{Huang2013a}, the 
    projected \mstar{} distributions in massive galaxies become increasingly elongated 
    at larger radius.  
    We notice no significant difference between the \rbcg{} and \nbcg{} samples, 
    except for maybe some hints that centrals in more massive halos becomes slightly 
    more elliptical at very large radius.  
    For the median color profiles, we see shallow and negative gradients of both 
    $g-r$ and $g-i$ colors as expected. 
    More importantly, the \rbcg{} and \nbcg{} samples show very consistent color 
    gradients, which confirms that using average color to derive \m2l{} of both 
    samples will not bias the comparison of \mden{} profiles.  
    Without conversion between the HSC and SDSS filter systems, our median color 
    profiles appear to be little steeper than the stacked ones from SDSS.  
    The exactly cause of this difference is not clear, although it could also be 
    due to the systematics in the stacking process. 
    As for the \minn{}-matched samples on the right, it becomes very clear that the 
    extended envelopes hosted by \rbcg{} galaxies have much higher ellipticity, and 
    redder $g-i$ colors than the outskirts of \nbcg{} ones at similar \minn{}.  
    All these preliminary results point out that, by including these information 
    into consideration, we will have a more complete view of the connection between 
    halo/environment and structure of massive central galaxies.

    \song{Ok, I'll see how to make it happen; this sounds abrupt since I used 
    to have several subsections in the discussion section.}
    \alexie{The beginning part of this paragraph sounds to me like the summary section. You could move part of this to summary and put it in bullet piont format.}. In the end, we want to emphasis once more \textbf{the importance of deep image 
    and appropriate method of photometry in studying the structure and assembly 
    history of galaxies}. 
    With the help of deep HSC images, we reliably reveal the extended stellar 
    structures around large sample of massive central galaxies out to more than 
    100 kpc.  
    And, using model-independent method, we found that 
    (1) previous de~Vaucouleur or single-\ser{} modeling method on shallower SDSS
    images underestimates the total mass of massive galaxies; 
    (2) even using deep HSC images, photometric method like \cmodel{} also tends 
    to underestimate the mass. 
    For these massive central galaxies, their extended mass distributions are 
    simply beyond the capability of the simply modeling method.  
    In addition, since the slope and mass contribution of the outer envelope 
    depend on not only the total stellar mass, but also likely the halo mass. 
    Applying the same oversimplified model to galaxies across a large range of 
    mass and environment could result in both underestimated and biased mass 
    estimates that can stop us getting accurate stellar mass function at the 
    high-mass end and studying potential environmental dependence of structures. 
    Moreover, these results also raise crucial questions like: 
    \textbf{What is the definition of ``total'' stellar mass of galaxy? 
    And, what is the most informative definition of stellar mass in the studies 
    of galaxy evolution?}.  \alexie{I think we can expand the discussion around this topic. We'd like to say somehting about the impact of total luminosity on the SMF. How about moving the left hand of Figure 3 here? What is the current volume of your data set compared to the volume of SDSS? I think we should emphasize the impact of stellar mass estimate on the SMF, especialy the large impact it has for BCGs}.
    \song{As mentioned earlier, I can expand the discussion easily with the help 
    from another figure, comparing with SMF from other people's work?//}

    \alexie{I thought of another interesting aspect to add to the discussion. This I think is perhaps a nice angle to discuss. The total luminosity of a galaxy has an impact on abundance matching. Do you have the "luminosity" that you would get for each galaxy if you assumed a single sersic fit for example? If so, we could do the following. Assume a one-to-one mapping between halo mass and your M100. Then plot Mhalo versus the stellar mass inferred from either cmodel of the single sersic fit. Then we can quote the scatter in $\log(M_*)$ at fixed Mhalo that this introduces. This figure will be very relevent for the Santa Barabara workshop!}
    \song{We do not have results from single Sersic models; at least not reliable enough 
    to be used here, mostly because single Sersic works terribly for these galaxies.
    We do have \mcmodel{}, but sorry that I failed to follows the logic here, basically this is 
    the difference between \mtot{} and \mcmodel{} expressed in a different way, and I am not 
    sure the scatter is due to the scatter of the mhalo-mstar relation, as the cModel 
    ones are underestimated due to technical reasons; We can bin the \rbcg{} galaxies 
    into a narrow $\lambda$ bin thoug, e.g. $30< \lambda < 40$, and see the scatter of 
    \mcmodel{}, \minn{}, and \mtot{}. We can certaintly shows that the scatter is larger 
    for mass using larger radius.  I think this point is important for the discussion
    of the intrinsic scatter of the mhalo-mstar relation, and at very similar mhalo,
    the outer structure may trace other properties of the halo. Let me know what do 
    you think.\\}

    \alexie{Finally, another quick discussion paragraph could be related to finding BCGs. It would be interesting to have a plot of the magnitude gap between Pcen1 and Pcen 2 (the most likely and the second most likely central) with cmodel versus your estimates. It would be nice to show that having a better luminosty estimate will enable you to identify the BCG better (is this true?)}.
    \song{This is interesting, and I don't know if it is true. It requires more works as 
    I do not have the profiles for all PCEN2 galaxies, I only do it when they are more 
    massive than the threshold.  Can we save this for later, I mean, after we send around 
    the draft.  Although I do compare the structures of really massive satellites in 
    these clusters with the \nbcg{} galaxies, and the results show little difference,
    basically shows that the structure difference could be unique for central galaxies. 
    I originally show a figure for this, and will send you along the e-mail.}

    In this work, we choose mass enclosed by different, but fixed physical 
    apertures, and find promising scientific applications for massive galaxies. 
    But, we recognize that they may not be the best choice. 
    For instance, central galaxies of massive halos like BCGs are known to be
    surrounded by diffuse stellar component that does not follow the potential 
    of the central galaxy (the intra-cluster light, or ICL; e.g. \addref{}).
    In theory, it can contribute a large amount of stellar mass out to very 
    large radius (e.g. \addref{}).  
    Using the definition of \mtot{}, we can not separate the potential 
    contribution of ICL in our \rbcg{} galaxies; and this could make complicate
    the discussion of environmental dependence.  \alexie{I am not sure I agree here. What is the "ICL"? How can is physically be defined? Can one robustly distinguish an "ICL" component from the galaxy component and how? Is the ICL similar to the "cD" enveloppe in the sense that we use these words all the time, but in practice, these are just one extrema of a continous galaxy population?}
    Since most of the \rbcg{} galaxies in our sample belong to low-mass cluster 
    like the Virgo or Fornax, we do not expect the ICL component can makes a 
    smooth and significant contribution to the \mden{} profiles (e.g. \addref{}).
    Even removing all \rbcg{} galaxies in $\lambda > 40$ clusters will not change 
    our findings.  
    However, it is still important to ask whether should we include the ICL when 
    estimating the mass of central galaxies in massive halos.  \alexie{But what is the ICL?}
    \song{The observational definiiton is very vague; and the theoretical one is 
    very hard to achieve through observation, basically the stars that follows the 
    potential of the clsuter.}
    And, if the answer is no, which is the best way to separate the ICL component? 
    In literature, most works studied ICL by subtracting the model for inner part
    of the central galaxy (e.g. \addref{}).  
    As we show in this work, when extended stellar envelopes are revealed around 
    massive galaxies in large range of stellar and halo mass, we may have to 
    reconsider this approach \alexie{I agree with this phrase, this is good to point out.}.
    \song{I think I understand your point here. The thing with ICL is that any 
    conclusion so far is quite uncertain, so I can reorganize this paragraph, 
    point out the interesting question first, discuss it a little bit, but also 
    shows a caveat that say ICL could play a role here, and if it does, how could 
    they change our results.}

%% ------------------------------------------------------------------------------------ %% 
  \begin{figure*}[t!]
      \centering 
      \includegraphics[width=17.0cm]{fig/redbcg_discussion_2}
      \caption{\alexie{Seems to me that this goes in the results section?} Comparisons of the radial profiles for ellipticity and optical colors 
      	for \mtot{}-matched (left figure) and \minn{}-matched (right figure) of 
        \rbcg{} (orange-red) and \nbcg{} (grey-black) samples. 
        The formats are very similar to the figure for \mden{} profile 
        (e.g. right figure of Fig.\ref{fig:prof_m100_A)}. 
        On each side, from top to bottom, we show the elliptical profile, and the 
        color profile using $g-r$ and $g-i$ color. 
        Since the individual profile of these properties has very large uncertain in the 
        outskirt, we block the region at large radii where we think the median profiles 
        are still unreliable using a dark, shaded box.
        We also compare these results with previous results based on SDSS images. 
        They are: 
        1) Ellipticity profile of stacked image of large sample of red galaxies at 
        $z\sim 0.4$ from \citet{Tal2011}, with and without PSF correction 
        (solid and dotted blue lines on top panel); 
        2) Ellipticity and $g-r$ color profiles from stacking analysis of nearby massive 
        galaxies with high concentration index ($C>2.6$) in \citet[][blue dash lines 
        on the top and middle panels]{DSouza2014}; 
        3) average $g-r$ and $g-i$ color profiles for a large sample of nearby 
        elliptical galaxies in \citet[][blue, solid lines on the middle and bottom 
        panels]{LaBarbera2010}.}
      \label{fig:discussion_2}
  \end{figure*}
%% ------------------------------------------------------------------------------------ %%

%\subsection{Impact from ``Fossil Galaxies''}
%
%    ``Fossil galaxy'' is the most massive galaxy in a ``fossil group/cluster''
%    system.
%    According to the commonly adopted definition, ``fossil'' systems are the 
%    group or cluster with high X-ray luminosity and large magnitude gap between 
%    the two brightest galaxies (\addref: Ponman1994; Jones2003). 
%    Both observations and simulations suggest that they could represent the 
%    final stage of the hierarchical merging process in a group/cluster mass
%    halo that is formed at very early epoch (e.g.\ \addref: 
%    Khosroshahi2004, 2007; DOnghia2005; Dariush2007).
%    Under this interpretation, at fixed halo mass, the central galaxy of 
%    a fossil system typically has larger stellar mass than the one in normal 
%    system (e.g.\ Harrison2012), as it has ``consumed'' most satellites 
%    early on; 
%    However, the ``low richness'' of a fossil system makes it difficult for 
%    red-sequence finder like \redm{}~to reliably identify 
%    (e.g.\ Figure~17 of Sadibekova2014).  
%    If some massive ETGs in our \nbcg{} sample are actual the centrals 
%    galaxies of fossil clusters, it could bias our conclusions about the 
%    relation between structure and ``environment''.  
%    
%    Although it is possible that fossil systems formed earlier than normal 
%    clusters, the difference in detailed merging history is still unclear, 
%    hence it is hard to imagine the structural differences between fossil 
%    central galaxies and regular BCGs with same \mstar{}.  
%    Naively speaking, more frequent mergers for fossil centrals should result 
%    in more extended stellar envelope.  
%    However, the distributions of epoch and mass-ratio of those mergers also 
%    matter. 
%    
%    The lack of any difference in stellar population properties may suggest 
%    It is still not clear how the mass assembly history of fossil central

%% ------------------------------------------------------------------------------------ %%

\section{Summary and Future Plans}
    \label{sec:summary}
\alexie{In this section, put a bullet list summarizing our main findings.}
	\song{Ok, will do}

    In this work, we use the deep $i$-band images from the Subaru HSC survey to 
    study the role played by halo mass in determining the mass assembly history and 
    stellar structures of very massive galaxies--the most impressive outcomes of 
    the hierarchical formation process.
    We map the stellar mass distributions of a large sample of massive central 
    galaxies at $0.3 < z < 0.5$ out to $>100$ kpc. 
    With the help of this impressive dataset, we uncover the very extended stellar 
    envelope around massive central galaxies with great accuracy. 
    We also reveal that, the ``total'' \mstar{} of massive galaxies can be significantly 
    underestimated if shallow image and/or inappropriate model is used, which can 
    lead to systematics in the study of stellar mass function and other important 
    topics.

    More importantly, we discover subtle, but systematic, and robust structural 
    differences in massive galaxies that depend on \mhalo{} through comparisons
    of their \mden{} profiles. 
    On average, at fixed \mstar{} within 100 kpc, the massive central galaxies in 
    more massive ($M_{200,m}$ \gax $1.6\times 10^{14} M_{\odot}$) halos show shallower 
    inner \mden{} profile within $\sim 15$-$20$ kpc, and a more prominent outer 
    envelope comparing to the ones in less massive 
    ($M_{200,c}$ \gax $8.7\times 10^{13} M_{\odot}$) halos.
    Also, at the same \mstar{} within inner 10 kpc--a proxy of the \mstar{} assembled 
    through the ``in-situ'' channel, we show that the ones in more massive halos 
    possess much more extended stellar halos in the outskirts.  
    Such differences are broadly consistent with the prediction of the ``two-phase''
    formation scenario of massive galaxies: the ones in more massive halos have 
    experienced more minor, dry mergers that accumulate more stars at large radius. 
    Meanwhile, we also suspect the differences in the inner profiles at fixed \mstar{} 
    or \mhalo{} may help us understand more about the impact from major merger or 
    AGN feedback. 
    We further suggest that the relation between \mtot{} and \minn{}\footnote{or other 
    even better proxies of total \mstar{} and the mass formed in the ``in-situ'' phase.} 
    is potentially interesting in diagnosing the assembly history of massive galaxies 
    and their host halos.

	%\subsection{Future Improvements}
    %\label{ssec:future}
    These results also highlight the advantages of HSC survey in studying the evolution 
    of massive galaxies.
    By the finish of this paper, the HSC survey will have already doubled its sky 
    coverage to $\sim 200$ deg$^2$, and can provide a much larger sample of massive 
    central galaxies. 
    At the same time, we will work on better data reduction of HSC images by 
    improving the deblending in crowded region and the accuracy of background 
    subtraction as they can help provide more reliable SED fitting results, and 
    push the photometric limit toward to even lower surface brightness level.

    Right now, we still rely on the \redm{} catalog from SDSS and spectroscopic 
    redshift from SDSS and BOSS.  
    Running \redm{} algorithm or similar group/cluster finder using HSC catalogs 
    will greatly help us extend the discussion of halo mass to lower mass and 
    higher redshift regimes. 
    Better identifications and measurements of satellites in massive halos will 
    also open up the possibility to study ``environment'' within the halo: e.g. 
    Compare the structures of satellites in massive halos and centrals in less 
    massive ones at similar stellar mass.

    Another key technical improvement would be to apply carefully designed 2-D 
    photometric modeling to these massive galaxies.  
    Using the same principle in \citealt{Huang2013a}, we will gradually build up 
    the complexity, and explore different models to fit these massive galaxies 
    better.  
    These 2-D models can help us removing the impact from seeing, and dealing with 
    the uncertainties of sky subtraction, so that we can measure the stellar mass 
    distributions to both smaller and larger radii. 
    Applying these models to multiband images with reasonable constraints (e.g. 
    \citealt{Huang2016}), we can characterize color gradients, and to achieve better
    estimates of total stellar mass.

    Primarily designed as a cosmological survey, the deep imaging depth and 
    excellent $i$-band seeing of HSC can greatly improve the weak lensing analysis 
    of massive galaxies. 
    Therefore, our next major goal will be to investigate trends between halo 
    properties and galaxy properties by measuring the weak lensing signal for samples 
    binned by properties such as \minn{}, \mtot{}, cluster richness, and magnitude gap.
    Joining the stellar mass distributions and the weak lensing signals together, 
    we can explore tracers that are sensitive to a range of physical processes 
    (strength and type of AGN feedback, merger rate, typical mass ratio of mergers
    \etal). 
    To help us have better physical insights, we will directly compare our results 
    with predictions from cosmological hydro-simulations such as 
    Illustris (\addref{}), Eagle (\addref{}), or \textit{Horizon-AGN}.

%% ------------------------------------------------------------------------------------ %%

\acknowledgements

  % HSC part
  The Hyper Suprime-Cam (HSC) collaboration includes the astronomical communities of Japan
  and Taiwan, and Princeton University.  The HSC instrumentation and software were
  developed by the National Astronomical Observatory of Japan (NAOJ), the Kavli Institute
  for the Physics and Mathematics of the Universe (Kavli IPMU), the University of Tokyo,
  the High Energy Accelerator Research Organization (KEK), the Academia Sinica Institute
  for Astronomy and Astrophysics in Taiwan (ASIAA), and Princeton University.  Funding was
  contributed by the FIRST program from Japanese Cabinet Office, the Ministry of
  Education, Culture, Sports, Science and Technology (MEXT), the Japan Society for the
  Promotion of Science (JSPS), Japan Science and Technology Agency (JST),  the Toray
  Science  Foundation, NAOJ, Kavli IPMU, KEK, ASIAA, and Princeton University.

  % SDSS part
  Funding for SDSS-III has been provided by the Alfred P. Sloan Foundation, the
  Participating Institutions, the National Science Foundation, and the U.S.  Department of
  Energy. The SDSS-III web site is http://www.sdss3.org.  SDSS-III is managed by the
  Astrophysical Research Consortium for the Participating Institutions of the SDSS-III
  Collaboration including the University of Arizona, the Brazilian Participation Group,
  Brookhaven National Laboratory, University of Cambridge, University of Florida, the
  French Participation Group, the German Participation Group, the Instituto de Astrofisica
  de Canarias, the Michigan State/Notre Dame/JINA Participation Group, Johns Hopkins
  University, Lawrence Berkeley National Laboratory, Max Planck Institute for
  Astrophysics, New Mexico State University, New York University, Ohio State University,
  Pennsylvania State University, University of Portsmouth, Princeton University, the
  Spanish Participation Group, University of Tokyo, University of Utah, Vanderbilt
  University, University of Virginia, University of Washington, and Yale University.

  % Pan-STARRS1 part
  The Pan-STARRS1 Surveys (PS1) have been made possible through contributions of the 
  Institute for Astronomy, the University of Hawaii, the Pan-STARRS Project Office, 
  the Max-Planck Society and its participating institutes, the Max Planck Institute 
  for Astronomy, Heidelberg and the Max Planck Institute for Extraterrestrial Physics, 
  Garching, The Johns Hopkins University, Durham University, the University of Edinburgh, 
  Queen's University Belfast, the Harvard-Smithsonian Center for Astrophysics, the Las 
  Cumbres Observatory Global Telescope Network Incorporated, the National Central 
  University of Taiwan, the Space Telescope Science Institute, the National Aeronautics 
  and Space Administration under Grant No. NNX08AR22G issued through the Planetary 
  Science Division of the NASA Science Mission Directorate, the National Science 
  Foundation under Grant No. AST-1238877, the University of Maryland, and Eotvos 
  Lorand University (ELTE).

  % LSST software
  This paper makes use of software developed for the Large Synoptic Survey 
  Telescope. We thank the LSST Project for making their code available as free 
  software at http://dm.lsstcorp.org.

  % Software
  This research made use of:
  \href{http://www.stsci.edu/institute/software_hardware/pyraf/stsci\_python}{\texttt{STSCI\_PYTHON}},
      a general astronomical data analysis infrastructure in Python. 
      \texttt{STSCI\_PYTHON} is a product of the Space Telescope Science Institute, 
      which is operated by AURA for NASA;
  \href{http://www.scipy.org/}{\texttt{SciPy}},
      an open source scientific tools for Python (\citealt{SciPy});
  \href{http://www.numpy.org/}{\texttt{NumPy}}, 
      a fundamental package for scientific computing with Python (\citealt{NumPy});
  \href{http://matplotlib.org/}{\texttt{Matplotlib}}, 
      a 2-D plotting library for Python (\citealt{Matplotlib});
  \href{http://www.astropy.org/}{\texttt{Astropy}}, a community-developed 
      core Python package for Astronomy (\citealt{AstroPy}); 
  \href{http://scikit-learn.org/stable/index.html}{\texttt{scikit-learn}},
      a machine-learning library in Python (\citealt{scikit-learn}); 
  \href{http://www.astroml.org/}{\texttt{astroML}}, 
      a machine learning library for astrophysics (\citealt{astroML});
  \href{https://ipython.org}{\texttt{IPython}}, 
      an interactive computing system for Python (\citealt{IPython});
  \href{https://github.com/kbarbary/sep}{\texttt{sep}} 
      Source Extraction and Photometry in Python (\citealt{PythonSEP});
  \href{https://jiffyclub.github.io/palettable/}{\texttt{palettable}},
      color palettes for Python;
  \href{http://dan.iel.fm/emcee/current/}{\texttt{emcee}}, 
      Seriously Kick-Ass MCMC in Python.

%%%%%%%%%%: Bibliographic Section %%%%%%%%%%

\bibliography{redbcg}

%%%%%%%%%%: Appendix Section %%%%%%%%%%%%

\appendix

%% ------------------------------------------------------------------------------------ %%

\section{A. Extraction of 1-D surface brightness profile} 
    \label{app:A}

%% ------------------------------------------------------------------------------------ %% 
    \figurenum{A1}
    \begin{figure*}[hbt!]
        \centering 
        \includegraphics[width=\textwidth]{fig/redbcg_ellipse_tech}
        \caption{\textbf{Left:} Example of the object-mask built for the \texttt{Ellipse}
            run for a typical massive galaxy in the sample. 
            All the shaded regions are masked out. 
            The three dash lines (red, inner one and two blue ones) around the target 
            at the center outlines the three radius we defined using the flux radius 
            of the target.  
            We increase the mask size for objects detected in different regions 
            separated by these apertures (which are outlined by solid, elliptical 
            apertures with different colors) using slightly different criteria.~~
            \textbf{Middle:}: The flux profile that is zoomed into the near-zero flux 
            range (top panel), and the curve-of-growth of the magnitude for the example
            massive galaxy.  
            To highlight the importance of correcting sky background, we show the profiles 
            using both images with (red, solid line) and without (black, dash line) 
            ad-hoc background correction. 
            On the top panel, besides the horizontal line that highlights the zero flux 
            level, we also show the uncertainty of the sky background estimate via 
            grey-shaded region.  
            On the bottom panel, two horizontal lines indicate the magnitudes 
            corresponding to total (solid) and half flux (dash) using the 
            background-corrected profile. ~~
            \textbf{Right:} compare the 1-D surface brightness profiles for the same 
            example galaxy extracted using different mask 
            (smaller masking region: red, dash line; larger masks: blue, dash line), 
            or different \texttt{Ellipse} settings
            (more aggressive pixel-clipping: cyan, dash line; 
             larger step in radius: green, dash line; 
             using mean flux along the isophote instead of median: purple, dash line)
            with the one using the default configuration (black, solid line).}
        \label{figure:A1}
    \end{figure*}
%% ------------------------------------------------------------------------------------ %%

%% ------------------------------------------------------------------------------------ %%

\section{B. Estimate average {$M_{\star}/L_{\star}$} using \texttt{iSEDFit}} 
    \label{app:B}

%% ------------------------------------------------------------------------------------ %% 
    \figurenum{B1}
    \begin{figure*}[bt!]
        \begin{center}
        \includegraphics[width=\textwidth]{fig/redbcg_isedfit_example.pdf}
        \caption{Left: example output figure from \texttt{iSEDFit} that shows the SED 
        	fitting results. 
            The open-boxes show the observed fluxes in 5-band, and the solid, blue-dots
            show the best-fitted results, along with the high-resolution spectrum for
            this model recontructed using the synthetic spectra from \texttt{FSPS}. 
            Top-left corner shows the best-fit stellar population parameters, and 
            bottom-right corner shows the ID, redshift of this object, and reduced 
            $\chi^2$ of the best-fit model.~~~
            Right: the Posterior distributions of a few key parameters.
            From top-left to bottom right are: 1) stellar mass (\logms{}); 2) age of the 
            population (mass and star-formation rate weighted) in Gyr; 3) star formation 
            rate ($\log\ \mathrm{SFR}\ (M_{\odot}/\mathrm{yr})$; instant one and the one 
            averaged over the previous 100 Myr; 4) stellar metallicity 
            ($\mathrm{Z}/\mathrm{Z}_{\odot}$); 5) dust extinction ($\mathrm{A}_V$ in mag);
            6) birthrate parameter ($\log\ b$; averaged over previous 100 and 1000 Myr).}
        \label{fig:ised}
        \end{center}
    \end{figure*}
%% ------------------------------------------------------------------------------------ %%

%% ------------------------------------------------------------------------------------ %% 
    \figurenum{B2}
    \begin{figure*}[hbt!]
        \begin{center}
        \includegraphics[width=\textwidth]{fig/redbcg_isedfit_2.pdf}
        \caption{Relationships between \mstar{} and certain stellar population 
            parameter using \texttt{iSEDfit}. 
            The left figures are for the \rbcg{} sample, while the right ones are 
            for the \nbcg{} sample. 
            They are both color-coded using the mass-weighted stellar age. 
            The four stellar population properties shown here are: 
            1) mass-weighted stellar age in Gyr (top-left); 
            2) SFR-weighted stellar age in Gyr (bottom-left); 
            3) mass-weighted stellar metallicity in unit of solar value (top-right);
            4) dust extinction value in magnitude (bottom-right).}
        \label{fig:ised_2}
        \end{center}
    \end{figure*}
%% ------------------------------------------------------------------------------------ %%

%% ------------------------------------------------------------------------------------ %% 
    \figurenum{B3}
    \begin{figure*}[hbt!]
        \begin{center}
        \includegraphics[width=\textwidth]{fig/redbcg_isedfit_3.pdf}
        \caption{Comparisons of \mstar{} estimated by \texttt{iSEDFit} using different
            assumptions. 
            In each figure, we plot the default \mstar{} values against their 
            differences with the results using other models. 
            The top (bottom) row shows comparisons for the \rbcg{} (\nbcg{}) sample. 
            The results involved here are labeled: 
            (A): default one; 
            (B): using the Salpeter IMF instead of the Chabrier one (Left column);
            (C): using the BC03 synthetic population model instead of the FSPS one
                 (Middle column);
            (D): assume that there is no dust extinction (Right column). 
            On each figure, a horizontal line shows the zero difference level, along
            with the grey shaded region that highlights the typical uncertainty of 
            the \mstar{}-differences. 
            We also show the histogram of the \mstar{}-differences on the right.}
        \label{fig:ised_3}
        \end{center}
    \end{figure*}
%% ------------------------------------------------------------------------------------ %%

    We adopt flat distribution between 0.5 to 14.0 Gyrs as prior for the look-back 
    time when the star formation turned on. 
    The exponential delayed time-scale ($\tau$) is allowed to change between 
    0.1 to 3.0 with equal probability.  
    The chance of random star burst is set at 0.2 for every 2 Gyrs. 
    The duration of the star burst is draw from a logarithmic distribution 
    between 0.03 to 0.3 Gyr; and the mass fraction formed in the burst is from 
    a logarithmic distribution between 0.01 and 1.0.

    In principle, these choices of priors could leave systematic effects in 
    the estimate of stellar mass (e.g.\ \citealt{Bernardi2016b}). 
    For low-$z$ massive ETGs like the ones in our sample, the details form of 
    SFH, importance of random star burst, and the dust extinction should not be 
    major concern.  
    However, the choices of stellar population models and IMF can still 
    change the results systematically. 
    More discussions on this can be found in Appendix~A. 
    In short: 
    \begin{enumerate}
        \item Both \texttt{FSPS+MILES} and \texttt{BC03} (\citealt{BC03} models 
            still have difficulties recovering the optical color involving filters 
            at the red end (e.g~$i-$Y), which could relate to the challenge for 
            modern stellar population models to reproduce the optical color-color 
            relation of red-sequence galaxies (e.g.\ \citealt{MIUSCAT2}), or the 
            shallower photometry of HSC-Y band data. 
        \item The \texttt{BC03} provides slightly better overall ${\chi}^2$ and 
            systematically smaller \mstar{} than the \texttt{FSPS+MILES} models.  
            However, it is possible that the \texttt{BC03} model tends to 
            underestimate the \m2l{} for a fraction of them as the estimate 
            stellar age is unrealistically young for red, massive galaxies.  
            Therefore, we still use the \texttt{FSPS+MILES} model as the fiducial 
            one.  But, switching to the \texttt{BC03} model will not change any 
            key result in this work. 
        \item The usage of \citet{Salpeter1955} IMF results in systematically 
            higher \mstar{} (on average $+0.25$ dex of \logms{}).  
            Although there are multiple lines of evidence that favor Salpeter 
            or even more ``bottom-heavy'' IMF in the most massive ETGs 
            (e.g.\ \citealt{Conroy2012}; \citealt{Cappellari2012}), we still 
            present the main results using Chabrier IMF to accommodate galaxies 
            with lower \mstar{} in the sample, and to be as consistent as possible 
            with a few other works.  
            Also, the choice of IMF does not impact our results qualitatively.     
    \end{enumerate}

    Besides the priors for stellar population properties, different treatments of 
    the light profile and accuracy of sky background subtraction are also important
    for the estimate of \mstar{} as they strongly impact the estimate of total luminosity
    (e.g.\ \citealt{Bernardi2013} and \citealt{DSouza2015}).  
    At the depth of SDSS, the default \texttt{cModel} photometry is already shown to 
    be not very accurate at high-\mstar{} end (e.g.\ \citealt{Meert2015}; 
    \citealt{Bernardi2016a}) as it does not capture the extended envelope of these 
    galaxies. 
    As for HSC, it is much more challenging for \texttt{cModel} to recover the total 
    luminosity of massive ETGs since the stellar envelope becomes even more 
    extended for neither de~Vaucouleur or exponential model to reproduce.  
    However, the HSC \texttt{cModel} under the force-photometry mode can measure 
    \textbf{average} color of the galaxy with great accuracy (e.g.\ Huang\etal~in prep.~), 
    and provide us a reliable SED to estimate the \textbf{average} \m2l{}. 
    Considering this, we separate the process of estimating the total \mstar{} of 
    massive ETGs in our sample into two steps:

    \begin{enumerate}
        \item Firstly, using the redshift and \texttt{cModel} magnitude in five bands, 
            we derive an initial estimate of the \mstar{} of the galaxy.  
            More importantly, we derive the \textbf{average} \m2l{} in $i$-band of the 
            galaxy, and use the best-fit SED to provide $k$-correction to the photometry.  
        \item In the next section, we will derive better estimate of total \mstar{} using 
            the more accurate total luminosity in $i$-band from integration of carefully 
            measured 1-D surface brightness profile, and the average \m2l{} from the SED
            fitting.  
    \end{enumerate}

    The basic results from \texttt{iSEDFit} are summarized in Figure~2, where we compare
    the relations between initial estimates of \mstar{} and (both luminosity and 
    star formation weighted) stellar age, metallicity, and dust extinction. 
    As expected, most galaxies in our samples are 
    $\log(M_{\star,\ \mathrm{ini}}/M_{\odot}) \geq 11.2$ massive galaxies that have 
    old age, high metallicity ($1.5 \times Z_{\odot}$ is the highest metallicity allowed), 
    and low dust extinction.  
    Given that we only have photometry from five optical bands, degeneracies among 
    age, metallicity, and dust extinction are naturally expected.  
    We will not use them for any scientific reason in this work, but to show that they
    behave reasonably.

%% ------------------------------------------------------------------------------------ %%

\section{C. Comparison of \mden{} profiles using \mstar{} from the GAMA survey}
    \label{app:C}

%% ------------------------------------------------------------------------------------ %% 
\figurenum{C1}
\begin{figure*}[hbt!]
    \centering 
    \includegraphics[width=\textwidth]{fig/redbcg_prof_gama_new}
    \caption{\textbf{Left:} comparison of \mstar{} estimated by the GAMA survey and 
    the ones in this work for the same galaxies. 
    We plot the \logmgama{} against the difference between \logmtot{} and \logmgama{}. 
    The format is very similar to the right panel of Fig. \ref{fig:smf1}, except that
    we color-code the \rbcg{} galaxies (orange-red) using the \ser{} index measured by 
    GAMA using \galfit{} on SDSS image. 
    The two vertical lines highlight the mass range $11.4 \leq$\logmgama{}$<11.9$ that
    is used for the comparison later.~~
    \textbf{Middle and Right figures:} we compare the \mden{} profiles of \rbcg{} 
    (orange-red) and \nbcg{} (grey-black) galaxies using the samples matched on the 
    \mgama{}-$z$ plane at $11.4 \leq$\logmgama{}$<11.9$ and $0.28 \leq z < 0.4$. 
    The formats for figure (b) and (c) are very similar to Fig.\ref{fig:prof_m100_A}.}
    \label{figure:C1}
\end{figure*}
%% ------------------------------------------------------------------------------------ %%

    As the sky coverage of this release greatly overlaps with the GAMA survey,
    we start the detailed comparison with an external check using the stellar
    mass estimated by \citep{Taylor2011}. 
    The purpose is still to investigate the impact of deeper photometry on 
    the analysis of structure of massive galaxies.  
    The stellar masses of GAMA galaxies are initially derived through careful 
    optical-SED fitting (BC03 model; Chabrier IMF) using the PSF-matched 
    aperture photometry.  
    Then they were corrected for the total luminosity estimated using single-\ser{}
    2-D model to multi-band images (\citealt{Kelvin2012}).  
    We separate the \rbcg{} and \nbcg{} galaxies that also have 
    spec-$z$ and stellar mass in GAMA DR2 (\citealt{Liske2015}).
    Most of galaxies in these subsamples have $z < 0.40$.

    % Move the section about matching to Mtot matched section

    In Fig.~10, we show the results of matches for subsamples using 
    $M_{\star, \mathrm{GAMA}}$.  
    To increase the number of available \rbcg{} to match, we loosen 
    its selection criteria to $\lambda \ge 20$ and $P_{\mathrm{CEN}} \ge 0.6$.
    Based on the distributions of the two samples at the $M_{\star, \mathrm{GAMA}}$-$z$ 
    plane, and their overlapped region, we match them in two bins of 
    $M_{\star, \mathrm{GAMA}}$ that have slightly different redshift ranges. 
    %% moved
    For the samples in two $M_{\star, \mathrm{GAMA}}$, the matches appear to be
    very good.

    In Fig.~11, we show the individual profiles, median profiles along with their 
    uncertainties derived using 5000 times bootstrap resampling, and the relative 
    difference between the profile of the \rbcg{} and \nbcg{} 
    subsamples.  
    It is very clear that, for both mass bins, the \rbcg{} has a much 
    more extended outer envelope, while its profile inside $\sim 10$ kpc is 
    very close to the \nbcg{} one.  
    This is inconsistent with the matched distributions of their ``total'' 
    stellar mass, and reveals potential issue with the stellar mass estimated 
    by GAMA.
    We can reproduce very similar results using the luminosity density profiles 
    (with or without $k$-correction) of these matched samples, which suggests 
    that the main problem does not lie in the estimate of \m2l{}, but in the 
    measurement of total luminosity using single-\ser model.  
    This highlights the importance of deep optical images in studying the 
    structure of massive galaxies.

        In cases of these $M_{\star, \mathrm{GAMA}}$-matched samples, the differences
    of median profiles are very robust.

%% ------------------------------------------------------------------------------------ %%

\section{D. Comparisons of \mden{} profiles in different redshift bins}
    \label{app:D}

%% ------------------------------------------------------------------------------------ %% 
\figurenum{D1}
\begin{figure}[htb!]
    \centering 
    \includegraphics[width=8.5cm]{fig/average_mass_profiles_fsps1_B}
    \caption{Comparison of \mden{} profiles of \rbcg{} (orange-red) and \nbcg{} 
        (grey-black) at $11.6 \le$\logmtot$< 11.9$ in redshift bins of 
        $0.3\leq z<0.4$ (solid lines) and $0.4\leq z<0.5$ (dash lines). 
        We show the individual profile in the background using much thinner line, 
        and highlight the median profiles using thicker line and darker color.
        Other formats are exactly the same with the left figure of 
        Fig.\ref{fig:avg_prof}.}
    \label{fig:D1}
\end{figure}    
%% ------------------------------------------------------------------------------------ %%

    As the massive galaxies in our samples span quite a bit in both redshift and \mtot{}, 
    it is important to evaluate the impact from PSF on the \mden{} profiles at different 
    redshift. 
    In Fig.\ref{fig:D1}, we group the \rbcg{} and \nbcg{} galaxies at  
    $11.6 \le$\logmtot$< 11.9$ into two $z$ bins ($0.3\leq z<0.4$ and 
    $0.4\leq z<0.5$). 
    We did not match the \mtot{} distributions in two $z$ bins, so slight difference in 
    average \mtot{} can be expected. 
    With the same seeing condition, the \mden{} profile of galaxy at higher redshift is 
    more vulnerable to the PSF smearing effect in the center, while suffers more in the 
    outskirt due to cosmological dimming and background noise.  
    Fig.\ref{fig:D1} clearly shows that, although the median profiles of the same sample 
    in two redshift bins follow each other well outside 10 kpc, they start to differ in 
    the central 2-3 kpc as the one from higher redshift bin has more flattened profile 
    due to PSF smearing effect.  
    Meanwhile, we notice that the median profiles of \rbcg{} and \nbcg{} in the 
    same $z$ bin are very similar to each other in the region affected by seeing.   
    This comparison confirms the radius range that is sensitive to seeing and 
    redshift distribution is well constrained within the grey-shaded area.
    It also suggests that it is important to match the redshift distributions between 
    samples before we can compare their \mden{} profiles or other properties at inner 
    region, as we did in this work. 
    In the outskirt, within 100-150 kpc, we see no difference in the median \mden{} 
    profiles caused by redshift distributions. 
    We therefore conclude that it is safe to study the \mden{} profile within this 
    radius range for massive galaxies at $z<0.5$ using HSC images.

%% ------------------------------------------------------------------------------------ %%

%% ------------------------------------------------------------------------------------ %% 
%% Table.1 
%% ------------------------------------------------------------------------------------ %% 
\clearpage
\include{table1}

%% ------------------------------------------------------------------------------------ %%

%% ------------------------------------------------------------------------------------ %% 
%%%%%%%%%%%%: End of the File %%%%%%%%%%%%
%% ------------------------------------------------------------------------------------ %%

\end{CJK*}

\clearpage

\label{lastpage}

\end{document}
%% ------------------------------------------------------------------------------------ %%