### 2017-03-24 draft/ms_v10.tex
Close to finishe

% Version 0; preprint format; Written by SH
% Version 1; emulateapj format; Edited by SH and AL
% Version 2; emulateapj format; Edited by SH and AL
% Version 3; emulateapj format; Edited AL
% Version 4; emulateapj format; Edited SH
% Version 5; MNRAS format; Edited SH

\documentclass[a4paper,fleqn,usenatbib]{mnras}

% Packages
\usepackage{deluxetable}
\usepackage{newtxtext,newtxmath}
\usepackage[T1]{fontenc}
\usepackage{ae,aecompl}
\usepackage{amssymb, amsmath}
\usepackage{graphicx}
\usepackage{natbib}
\usepackage{url}
\usepackage{hyperref}
\usepackage{float}
\usepackage[usenames, dvipsnames]{color}

% Package Settings
\hypersetup{colorlinks=true,
            citecolor=cyan,
            linkcolor=cyan,
            filecolor=magenta,      
            urlcolor=cyan}
\urlstyle{same}

% Figure extention
\DeclareGraphicsExtensions{.pdf,.png,.jpg}

%%%%%%%%%%%%: User Defined Commands %%%%%%%%%%%%

% Song Huang's definition 
\def\arcsec{{\prime\prime}}
\def\arcmin{{\prime}}
\def\degree{{\circ}}
\def\h{\hskip -3 mm}
\def\aa{{A\&A}}
\def\aas{{ A\&AS}}
\def\aj{{AJ}}
\def\al{$\alpha$}
\def\bet{$\beta$}
\def\amin{$^\prime$}
\def\annrev{{ARA\&A}}
\def\apj{{ApJ}}
\def\apjs{{ApJS}}
\def\asec{$^{\prime\prime}$}
\def\deg{$^{\circ}$}
\def\ddeg{{\rlap.}$^{\circ}$}
\def\dsec{{\rlap.}$^{\prime\prime}$}
\def\cc{cm$^{-3}$}
\def\flamb{erg s$^{-1}$ cm$^{-2}$ \AA$^{-1}$}
\def\flux{erg s$^{-1}$ cm$^{-2}$}
\def\fnu{erg s$^{-1}$ cm$^{-2}$ Hz$^{-1}$}
\def\hst{{\textit{HST}}}
\def\kms{km s$^{-1}$}
\def\lamb{$\lambda$}
\def\lax{{$\mathrel{\hbox{\rlap{\hbox{\lower4pt\hbox{${\sim}$}}}\hbox{$<$}}}$}}
\def\gax{{$\mathrel{\hbox{\rlap{\hbox{\lower4pt\hbox{${\sim}$}}}\hbox{$>$}}}$}}
\def\simlt{\lower.5ex\hbox{$\; \buildrel < \over {\sim} \;$}}
\def\simgt{\lower.5ex\hbox{$\; \buildrel > \over {\sim} \;$}}
\def\micron{{$\mu$m}}
\def\mnras{{MNRAS}}
\def\nat{{Nature}}
\def\pasp{{PASP}}
\def\perang{\AA$^{-1}$}
\def\peryr{yr$^{-1}$}
\def\reference{\noindent\pp}
\def\refindent{\par\noindent\parskip=2pt\hangindent=3pc\hangafter=1 }
\def\sb{mag~arcsec$^{-2}$}
\def\lsun{$L_\odot$} 
\def\msun{$M_\odot$}
\def\sigs{$\sigma_*$}
\newcommand{\lt}{<}
\newcommand{\gt}{>}

\def\etal{{\ et al.~}}
\def\galfit{{\tt GALFIT}}
\def\ser{{S\'{e}rsic\ }}
\def\redm{\texttt{redMaPPer}}
\def\cmodel{\texttt{cModel}}
% Samples
\def\rbcg{\texttt{cenHighMh}}
\def\nbcg{\texttt{cenLowMh}}
\def\redbcg{{$\lambda \ge 30$}}
\def\nonbcg{{$\lambda < 20$}}
% Mass related 
\def\mstar{{$M_{\star}$}}
\def\mhalo{{$M_{\mathrm{halo}}$}}
\def\logms{{$\log (M_{\star}/M_{\odot})$}}
\def\logmh{{$\log (M_{\mathrm{halo}}/M_{\odot})$}}

\def\minn{{$M_{\star,10\mathrm{kpc}}$}}
\def\meff{{$M_{\star,15\mathrm{kpc}}$}} 
\def\mtot{{$M_{\star,100\mathrm{kpc}}$}}
\def\mout{{$M_{\star,150\mathrm{kpc}}$}}
\def\mmax{{$M_{\star,\mathrm{Max}}$}}
\def\mgama{{$M_{\star,\mathrm{GAMA}}$}}
\def\mcmodel{{$M_{\star,\mathrm{cModel}}$}}

\def\logminn{{$\log (M_{\star,10\mathrm{kpc}}/M_{\odot})$}}
\def\logmtot{{$\log (M_{\star,100\mathrm{kpc}}/M_{\odot})$}}
\def\logmout{{$\log (M_{\star,150\mathrm{kpc}}/M_{\odot})$}}
\def\logmmax{{$\log (M_{\star,\mathrm{Max}}/M_{\odot})$}}
\def\logmgama{{$\log (M_{\star,\mathrm{GAMA}}/M_{\odot})$}}
\def\logmcmodel{{$\log (M_{\star,\mathrm{cModel}}/M_{\odot})$}}

\def\m2l{{$M_{\star}/L_{\star}$}}
\def\mden{{$\mu_{\star}$}}

% Commenting:
\newcommand{\xxx}[1]{\textcolor{red}{\textbf{XXX}}}
\newcommand{\todo}[1]{\textcolor{red}{\textbf{TODO:~#1}}}
\newcommand{\plan}[1]{\textcolor{cyan}{#1}}
\newcommand{\addref}{{\textcolor{red}{REF}}}
\newcommand{\note}[2]{\textcolor{blue}{\textbf{[Comment (#1): #2]}}}
\newcommand{\song}[1]{\textcolor{magenta}{\textbf{[Song: #1]}}}
\newcommand{\alexie}[1]{\textcolor{blue}{\textbf{[Alexie: #1]}}}
\newcommand{\kevin}[1]{\textcolor{green}{\textbf{[Kevin: #1]}}}
\newcommand{\update}[1]{\textcolor{Bittersweet}{#1}}
%\newcommand{\update}[1]{#1}

%%%%%%%%%%%%: Title and Affiliations %%%%%%%%%%%%

%% ------------------------------------------------------------------------------------ %% 
\title[Structure and Halo of Massive Galaxies]{A Detection of the Environmental 
       Dependence of the Stellar Profiles and Halos of Massive Central Galaxies}

\author[S. Huang et al.]{
        Song Huang,$^{1,2}$\thanks{E-mail: song.huang@ipmu.jp (SH)}
        Alexie Leauthaud,$^{2,1}$
        Kevin Bundy,$^{3,1}$
        Jenny Greene,$^{4}$
        Yen-Ting Lin,$^{5}$
        \newauthor
        Rachel Mandelbaum,$^{6}$
        Michael Strauss,$^{4}$
        and HSC collaboration$^{7}$   
        \\
        $^{1}$Kavli-IPMU, The University of Tokyo Institutes for Advanced Study, 
              the University of Tokyo (Kavli IPMU, WPI), Kashiwa 277--8583, Japan\\
        $^{2}$Department of Astronomy and Astrophysics, University of California 
              Santa Cruz, 1156 High St., Santa Cruz, CA 95064, U.S.A\\
        $^{3}$UCO/Lick Observatory, University of California, Santa Cruz,
              1156 High Street, Santa Cruz, CA 95064, USA\\
        $^{4}$Department of Astrophysical Sciences, Peyton Hall,
              Princeton University, Princeton, NJ 08540, USA \\
        $^{5}$Academia Sinica Institute of Astronomy and Astrophysics, 
              P.O. Box 23--141, Taipei 10617, Taiwan\\
        $^{6}$McWilliams Center for Cosmology, Department of Physics, 
              Carnegie Mellon University, Pittsburgh, PA 15213, USA\\
        $^{7}$Another Department, Different Institution, Street Address, City Postal 
              Code, Country
        }   
%% ------------------------------------------------------------------------------------ %% 
\date{Accepted XXX. Received YYY; in original form ZZZ}        
\pubyear{2017}

%%%%%%%%%%%%: Header and Version %%%%%%%%%%%%
\begin{document}

\label{firstpage}
\pagerange{\pageref{firstpage}--\pageref{lastpage}}

\maketitle

%%%%%%%%%%%%: Abstract and Keywords %%%%%%%%%%%%

\begin{abstract}

    We use ${\sim}100$ deg$^2$ of deep ($>28.5$ \sb{} in $i$-band), high-quality 
    (0.6\asec seeing) imaging data from the Hyper Suprime-Cam (HSC) survey 
    to investigate the halo mass dependence of the surface mass density profiles 
    and outer stellar envelopes of massive galaxies. 
    Our sample consists of ${\sim}7000$ massive (\mstar{}$\geq 11.5$) central galaxies 
    at $0.3 < z < 0.5$. 
    The depth of the HSC survey reaches ${\sim}4$ magnitudes fainter than SDSS and 
    enables us to directly trace stellar mass distributions to 100 kpc without 
    requiring stacking. 
    For the first time, we conclusively show that at fixed stellar mass, massive central 
    galaxies display mass density profiles that are not self similar and which 
    present subtle but systematic differences that correlate with \mhalo{}. 
    On average, massive central galaxies in more massive halos have shallower inner 
    stellar mass density profiles (within ${\sim}10$-$20$ kpc) and more prominent outer 
    envelopes. 
    These differences translate into a clear environmental dependence of the mass-size 
    relation: central galaxies in more massive halos have larger $R_{\mathrm{50}}$ 
    at fixed \mtot{}. 
    Shallow imaging will miss light at the outskirts of galaxies and this has an 
    important impact on measurements of the galaxy mass function. 
    Furthermore, because galaxies in larger halos have more extended envelopes, 
    the amount of light that is missed will be halo mass dependent. 
    Finally, at fixed total stellar mass, there is a large (but continuous) diversity 
    in the amplitude of stellar profiles at 100 kpc. 
    We argue that ``cD'' galaxies represent the tail end of a distribution and 
    are not a distinct class of massive galaxies. 
    Our results are broadly consistent with the two phase formation scenario of 
    massive galaxies in which outer envelopes are built up at late times from the 
    accretion of small systems. 
    Our mass profiles are provided in a convenient tabulated format in order to 
    facilitate comparisons with predictions from numerical simulations of galaxy 
    formation.

\end{abstract}

\begin{keywords}
    galaxies: elliptical and lenticular, cD --
    galaxies: formation --
    galaxies: photometry -- 
    galaxies: structure -- 
    galaxies: surveys
\end{keywords}

%%%%%%%%%%%%: Main Text %%%%%%%%%%%%

%% ------------------------------------------------------------------------------------ %% 
    \begin{figure*}
        \centering 
        \includegraphics[width=\textwidth]{fig/redbcg_sdss_compare}
        \caption{
            A comparison between the depth and imaging quality of SDSS and the HSC wide 
            layer for a sample of nearby massive elliptical galaxies at $0.2 < z < 0.5$.  
            These images are generated using $gri$ band images with an arcsinh stretch 
            \citep{Lupton2004}. 
            The HSC \texttt{WIDE} layer is $3.0$-$4.0$ magnitudes deeper than SDSS.
            This added depth is critical in order map the outskirts of ETGs out to 
            large radii.
            }
        \label{fig:sdss_compare}
    \end{figure*}
%% ------------------------------------------------------------------------------------ %%

%% ------------------------------------------------------------------------------------ %% 
  \begin{figure*}
      \centering 
      \includegraphics[width=\textwidth]{fig/redbcg_m100_m10_compare}
      \caption{
          Three-color images for a subsample of massive galaxies at $z{\sim}0.4$. 
          All of these galaxies have the same mass within a 10 kpc aperture 
          ($11.2<\log (M_{\star,10\ \mathrm{kpc}}/M_{\odot})<11.3$). 
          The dash-line circle at the top-left figure indicates 10 kpc.
          Galaxies are rank-ordered from top to bottom and from left to right by 
          their \mstar{} within a 100 kpc aperture which varies from $10^{11.2} M_{\odot}$ 
          to $10^{11.7} M_{\odot}$. 
          At fixed ``inner'' mass ($M_{\star,10}$), massive galaxies display significant
          diversity in their outer profiles. 
          Red boxes indicate galaxies in halos more massive than 
          ${\sim} 10^{14.0} M_{\odot}$. 
          }
      \label{fig:m100_m10_color}
  \end{figure*}
%% ------------------------------------------------------------------------------------ %%

%% ------------------------------------------------------------------------------------ %% 
\section{Introduction}
    \label{sec:intro}

    Massive galaxies are unique laboratories for studying both cosmology and galaxy 
    formation. 
    Simulations of structure formation within the context of the $\Lambda$-CDM 
    cosmological model make predictions for the hierarchical growth of dark matter halos 
    and galaxies (e.g. \citealt{Baugh1996, DeLucia2006}), but there are many open 
    questions regarding the star-formation history, mass assembly process, and 
    structural evolution of massive galaxies. 
    Massive galaxies are thought to grow according to a ``two-phase'' formation scenario 
    (e.g. \citealt{Oser2010, Oser2012}). 
    In this scenario, the progenitors of $z{\sim} 0$ massive early-type galaxies (ETGs) 
    undergo a rapid growth phase at $z{\sim} 2$ triggered either by disk instabilities 
    or gas rich mergers (e.g. \citealt{Hopkins2008, Dekel2009}). 
    Observationally, the progenitors of ETGs are thought to correspond to the so-called 
    ``red nuggets'' (e.g. \citealt{Damjanov2009}) which have smaller effective radii 
    ($R_{\mathrm{e}}$; e.g. \citealt{Trujillo2006, vanDokkum2008, Cimatti2008}), 
    slightly higher central velocity dispersion and stellar mass density 
    (\mden{}; e.g. \citealt{vandeSande2011, Belli2014}), more disk-like morphologies 
    (e.g. \citealt{vanderWel2011}) than their local descendants 
    (e.g. \citealt{Bezanson2009, vanDokkum2010}).

    Following an initial epoch of early formation, feedback from stars and/or AGN 
    (e.g. \citealt{Sijacki2007, Fabian2012}) efficiently quench star formation in 
    massive galaxies and a large fraction of these massive progenitors become quiescent 
    at $z>1$ (e.g. \citealt{Bezanson2009, Kriek2016}). 
    The second phase in the ``two-phase'' formation scenario is driven by 
    non-dissipative processes such as dry mergers with other galaxies 
    (e.g. \citealt{Naab2006, Khochfar2006}). 
    According to the ``two-phase'' scenario, minor mergers drive the growth of massive 
    galaxies at late times and can help explain the significant increase
    of $R_{\mathrm{e}}$ (e.g. \citealt{Newman2012, vdWel2014}) as well as the build-up 
    of the outer stellar halos of ETGs (e.g.  \citealt{Szomoru2012, Patel2013}).

    According to this two-phase formation scenario, the late time growth of massive ETGs 
    is intrinsically tied to the growth of their host dark matter halos 
    (e.g. \citealt{Leauthaud2012, Behroozi2013, Shankar2013}). 
    Massive ETGs that live at the center of more massive halos should experience 
    a different mass assembly history and should be subject to a larger merger rate 
    compared to ETGs that live in smaller dark matter halos. 
    A natural prediction of the two-phase formation scenario is therefore that the 
    structural properties of massive galaxies should depend on their ``environment''
    \footnote{There are different observational definitions of ``environment''. 
    In this paper, we use the term ``environment'' and halo mass interchangeably}.

    Previous work has shown that the half-light radii of massive ETGs is tightly 
    correlated with \mstar{} (e.g. \citealt{Shen2003, Guo2009}). 
    Models based on the ``two-phase'' formation scenario naturally predict that this 
    mass--size relation should exhibit an environmental dependence: the effective radii 
    of massive ETGs in more massive halos should be larger than those in less massive
    halos at fixed \mstar{} (e.g. \citealt{Shankar2013, Shankar2014}).

    However, while hints of an environmental dependence in the mass-size relation have 
    been spotted at high-$z$ (e.g. \citealt{Papovich2012, Lani2013, Delaye2014}; but 
    also see \citealt{Rettura2010}), there is no solid evidence for this effect at 
    low-$z$ (e.g. \citealt{Nair2010, HCompany13}; but also see \citealt{Yoon2017}). 
    The absence of an environmental dependence in the mass-size relation is an 
    outstanding challenge to the two-phase formation scenario of ETGs.

    From an observational perspective, we hypothesize that the lack of a detection of 
    this effect may be related a combination of the methodologies employed, as well as 
    the lack of deep high-quality imaging for large numbers of massive ETGs. 
    In particular, massive ETGs have very steep inner stellar density profiles, 
    extremely extended stellar envelopes, and radial-dependent isophotal shapes which 
    make accurate measurements of \mstar{} and $R_{\mathrm{e}}$ very challenging
    (e.g. \citealt{Bernardi2013, DSouza2014}), 
    especially in the presence of shallow imaging data such as the Sloan Digital Sky 
    Survey (SDSS; e.g. \citealt{SDSSDR7, SDSSDR12}).

    In this paper, we take advantage of new high-resolution (median 
    FWHM${\sim} 0.6^{\arcsec}$ in $i$-band) and deep ($i$-band surface brightness 
    limit $> 28.5$\sb) imaging from the Hyper-Suprime Cam (HSC) Subaru Strategic 
    Program \citep[SSP,][]{HSCDR1} to revisit the 
    question of the environmental dependence in the mass-size relation. 
    The goal of this paper is to compare the \mden{} profiles and structural scaling 
    relations of massive central galaxies in both low and high mass dark matter halos. 
    Fig~\ref{fig:sdss_compare} displays a comparison of the imaging quality of massive 
    ETGs using SDSS versus using the new HSC data. 
    The deep imaging depth and excellent seeing conditions of HSC images make them 
    perfect for mapping the \mstar{} distributions of massive galaxies out to very 
    large radii. 
    With the help of the \redm{} cluster catalog (e.g.\ \citealt{Rykoff2014}; 
    \citealt{Rozo2015b}; please see Section~\ref{ssec:redmapper} for details), we 
    select a large sample (${\sim} 15000$) of massive central 
    galaxies with $0.3 < z < 0.5$ using ${\sim} 100$ square degree of data from the 
    HSC wide layer.

    Motivated by the ``two-phase'' scenario, and also due to the difficulty in 
    providing appropriate model to describe the \mstar{} distribution of massive 
    galaxy from its center to the very outskirt, we focus on \mstar{} within two 
    different physical apertures in this work:

    \begin{itemize}

        \item The stellar mass within 10 kpc (hereafter noted \minn{}). 
            According to the two-phase scenario, the ``in-situ'' star-formation quickly 
            builds up the inner, dense core of massive ETGs.  
            Based on recent observations and simulations (e.g.~\citealt{vanDokkum2010}; 
            \citealt{RodriguezGomez2016}), the in-situ component of massive ETGs is 
            mostly contained within one effective radius ($R_{\mathrm{e}}$, or 5-10 kpc). 
            We use \minn{} as a proxy for the mass forged during the ``in-situ''
            phase. 
            Given the quality of the HSC data, we can reliably measure \minn{} over our 
            redshift range ($1.0^{\arcsec}$ in radius equals 4.4 and 6.1 kpc at redshift
            0.3 and 0.5).

        \item The stellar mass within 100 kpc (hereafter noted \mtot{}). 
            For our galaxy sample, a 100 kpc aperture corresponds to 5-10 
            $\times R_{\mathrm{e}}$. 
            We show in section \ref{ssec:mtotal} that most of the mass for these ETGs 
            can be captured by a 100 kpc radius and that \mtot{} is a good proxy for the 
            ``total'' \mstar{}. 
            Although not perfect, we argue that our measurements of \mtot{} (which are 
            actually measuring the light directly out to 100 kpc) are a better tracer 
            of total \mstar{} than model-dependent results from shallower data 
            (such as SDSS) which rely on extrapolating the light profiles of galaxies 
            out to large radii.

       \end{itemize}
%% ------------------------------------------------------------------------------------ %%

    Massive central galaxies in this work show an intriguing diversity on the plane 
    defined by these two metrics.  
    We preview this using Fig~\ref{fig:m100_m10_color}, which shows a subsample of 
    massive galaxies at $z{\sim} 0.4$ that share very similar value of \minn{}, but 
    exhibit a large diversity in their outer profiles. 
    Assuming that galaxies at fixed \minn{} share a similar in-situ formation history, 
    it is of great interest to investigate the relation between their present day halo 
    mass and the extend of their outer envelopes. 
    We note that although a circular aperture is shown on Fig~\ref{fig:m100_m10_color}, 
    for our mass measurements, we always use elliptical apertures that follow the 
    average isophotal shape. Finally, given our narrow mass and redshift range, we 
    can safely ignore significant mass growth and structural evolution in this work
    (no star formation, lower merger rate \etal~e.g.\ \citealt{Bellstedt2016},
    \citealt{Inagaki2015}; but also see \citealt{Bai2014}).

    This paper is organized as follows. 
    Section \ref{sec:data} presents our data and our sample selection. 
    Section \ref{sec:ellipse} describes our procedure for extracting 1-D surface 
    brightness profiles. 
    Section \ref{sec:mstar} describes how we estimate stellar mass. 
    Our main results are presented in Section \ref{sec:result} and discussed in 
    Section \ref{sec:discussion}. 
    Section \ref{sec:summary} presents our summary and conclusions.

    Magnitudes use the AB system (\citealt{Oke1983}), and are corrected for Galactic 
    extinction using calibrations from \citet{Schlafly11}.
    In this work, we assume $H_0$ = 70~km~s$^{-1}$ Mpc$^{-1}$, ${\Omega}_{\rm m}=0.3$, 
    and ${\Omega}_{\rm \Lambda}=0.7$.
    Stellar mass is noted \mstar{} and has been derived using a Chabrier Initial Mass 
    Function (IMF; \citealt{Chabrier2003}).     
    $M_{\mathrm{halo}}$ denotes halo mass in general whereas $M_{\rm 200b}$ is 
    explicitly defined as $M_{\rm 200b}\equiv M(<r_{\rm 200b})=200\bar{\rho} 
    \frac{4}{3}\pi r_{\rm 200b}^3$ where $r_{\rm 200b}$
    is the radius at which the mean interior density is equal to 200 times
    the mean matter density ($\bar{\rho}$).

%% ------------------------------------------------------------------------------------ %%

%% ------------------------------------------------------------------------------------ %% 
    \begin{figure*}
        \centering 
        \includegraphics[width=\textwidth]{fig/redbcg_ellipse_example}
        \caption{
            Left: example of the 1-D surface brightness and ellipticity profile 
            for a massive galaxy at $z=0.23$ in the $i$-band using \texttt{Ellipse}. 
            In this work, we always show the radial profile using a $R^{1/4}$ scaling 
            on the x-axis. 
            By using this scale, the de Vaucouleurs profile will appear as a straight 
            line on this figure. 
            We also plot the relative brightness profile of the PSF model normalized 
            at the central surface brightness of the galaxy to highlight the region 
            affected by seeing. 
            The grey shaded region highlights the region most affect by seeing 
            ($r<3$ kpc).
            On the top panel, the dash line shows the ellipticity used for the final 
            isophote. 
            Right: the three color image of this galaxy with isophotes 
            extracted by \texttt{Ellipse}. 
            The thick dotted line highlights the isophote with 
            $\mu_{i}{\sim} 28.5$~\sb.
            }
        \label{fig:ellipse}
    \end{figure*}
%% ------------------------------------------------------------------------------------ %%

%% ------------------------------------------------------------------------------------ %% 
\section{Data And Sample Selection}
    \label{sec:data}

\subsection{The Hyper Suprime-Cam Survey}
    \label{ssec:hsc}

    The Subaru Strategic Program (SSP, \citealt{MiyazakiInPrep}) makes use of the new 
    prime-focus camera, the Hyper Suprime-Cam (HSC;~\citealt{Miyazaki2012}), on the 
    8.2-m Subaru telescope at Mauna Kea. 
    The ambitious multi-layer HSC survey takes advantage of the large field of 
    view (FoV;~1.5 deg in diameter) of this camera and will cover ${\sim} 1400$ deg$^2$ 
    in 5 broad bands ($grizY$) to a limiting depth of $r {\sim} 26$ mag 
    in the \texttt{WIDE} layer. 
    This work is based on the internal data release \texttt{S15B}, which covers 
    ${\sim} 100$ deg$^2$ in all 5-band to full \texttt{WIDE} depth.  
    The regions covered by this release overlap with a number of spectroscopic surveys 
    (e.g.\ SDSS/BOSS: \citealt{Eisenstein2011}, \citealt{Alam2015}; 
    GAMA: \citealt{Driver2011}, \citealt{Liske2015}).

    The HSC \texttt{WIDE} survey is about $3.0$-$4.0$ magnitudes deeper in the $i$-band 
    than SDSS. 
    Combined with the excellent imaging resolution (the median $i$-band seeing is 
    0.6\asec), and the wide area, the HSC survey represents a tremendous data set to 
    perform a large statistical study of the surface brightness profiles of ETGs out to 
    large radii.     
    Fig~\ref{fig:sdss_compare} illustrates the quality of HSC imaging compared to SDSS 
    for three low redshift ETGs, and shows that HSC survey data are well suited for 
    mapping the stellar distribution of massive galaxies out to large radii.

	HSC $i$-band images typically have the best seeing compared to other bands because 
	of strict requirements driven by weak lensing science. 
    We will therefore use the $i$-band images to measure the stellar distributions of 
    massive galaxies.

%% ------------------------------------------------------------------------------------ %% 
\subsection{HSC Data Processing}
    \label{sec:pipeline}

    The full details of the HSC data processing can be found in \citet{BoschInPrep}
    and are briefly summarized here. 
    The HSC SSP data are processed with \texttt{hscPipe 4.0.1}, a derivative of the 
    Large Synoptic Survey Telescope (LSST) pipeline (e.g.\ \citealt{Juric2015}; 
    \citealt{Axelrod2010}), modified for HSC. 
    \texttt{hscPipe} first performs a number of tasks at the single exposure level 
    (bias subtraction, flat fielding, background modeling, object detection and 
    measurements). 
    Astrometric and photometric calibrations are performed at the single exposure level. 
    \texttt{hscPipe} then warps different exposures on to a common World Coordinate 
    System (WCS) and combines them into coadded images. 
    At this stage, \texttt{hscPipe} updates the images with a better astrometric and 
    photometric calibration using stars that are common among exposures.

    The pixel scale of the combined images is $0.168^{\arcsec}$. 
    Photometric calibration is based on data from the Panoramic Survey Telescope 
    and Rapid Response System (Pan-STARRS) 1 imaging survey 
    (\citealt{Schlafly2012}, \citealt{Tonry2012}, \citealt{Magnier2013}). 
    To achieve consistent deblending and photometry across all bands, \texttt{hscPipe} 
    performs multi-band post-processing at the coadd level. 
    First, \texttt{hscPipe} performs object detection on coadd images in each band 
    independently and records the flux peak and the above-threshold region 
    (referred as a ``footprint'') for each source. 
    Next, ``footprints'' and peaks from different bands are merged together before     
    performing deblending and measurements. 
    Finally, \texttt{hscPipe} selects a reference band for each object based on the 
    $S/N$ in different bands (for most galaxies in this work, the reference band is 
    the $i$-band). 
    After fixing the centroids, shape, and other non-amplitude parameters of each 
    object in this reference catalog, \texttt{hscPipe} perform forced photometry 
    on the coadd image in each band. 
    This forced photometry approach is optimized to yield accurate galaxy colors.

    For each galaxy, \texttt{hscPipe} measures a \cmodel{} magnitude using an approach 
    that is similar to SDSS (\citealt{BoschInPrep}). 
    However, as opposed to SDSS, the HSC \cmodel{} is based on forced photometry which 
    means that it can accurately measure both the \textit{fluxes and colors of galaxies}. 
    The HSC \cmodel{} algorithm fits the flux distribution of each object using a 
    combination of a de~Vaucouleur and an exponential component and accounts for the PSF. 
    The performance of this algorithm has been tested using synthetic objects 
    (\citealt{SynPipeInPrep}), and the results indicate that, generally speaking, 
    the HSC \cmodel{} photometry is accurate down to $i >25.0$ mag.  
    However, \cmodel{} currently systematically underestimates the total fluxes of 
    massive ETGs with extended stellar distributions. 
    This is caused by an intrinsic limitation of \cmodel{} as it is incapable of
    modeling profiles with extremely extended outskirts, a problem that is exacerbated 
    at the depth of the HSC survey. 
    In addition, at the depth of the HSC survey, accurate deblending in the vicinity of
    large ETGs where satellites and background galaxies often blend with the low surface 
    brightness stellar envelope is a challenging problem. 
    The deblending method implemented by the \texttt{hscPipe} currently tends to 
    ``over-deblend'' the outskirts of bright galaxies and leads to an 
    under-estimation of the the total flux of massive ETGs (this is discussed further in 
    \citealt{BoschInPrep}).  
    For these reasons, our results will be based on custom developed code to measure 
    the luminosities and stellar masses of massive galaxies. 
    We use the HSC \texttt{hscPipe} photometry for two purposes: 
    1) to perform a first broad sample selection, and 2) to estimate the average 
    color of massive galaxies.

%% ------------------------------------------------------------------------------------ %% 
\subsection{Initial Massive Galaxy Sample}
    \label{ssec:initial}

    We begin by using a broad luminosity cut to select an initial sample of massive 
    galaxies at $z < 0.5$ from the HSC photometric catalog. 
    Based on \citep{Leauthaud2016}, galaxies with \logms{}$\geq 11.5$ mostly have 
    $i_{\mathrm{SDSS, cModel}} \leq 21.0$ mag. 
    We therefore select our initial sample to be brighter than 
    $i_{\mathrm{HSC, cModel}} \leq 21.5$\footnote{We neglect small differences between 
    the response curves of SDSS-$i$ and HSC-$i$ filters.}
    and also limit our sample to regions that have reached the required depth of 
    the \texttt{WIDE} survey in the $i$-band.
    %% The SQL search is saved as "sample.sql" 
    %% The master catalog used here is "dr1_wide_galaxy_icmodel_21.5.fits"
    %%   The original catalogs contain 2275477 objects

    We further select extended objects with no deblending errors, with well defined 
    centroids, and with \texttt{cModel} magnitudes in all five bands. 
    After removing objects that have pixels affected by saturation, cosmic-rays, or 
    other optical artifacts\footnote{each criterion removes less than 8\% of the 
    entire sample.}.This sample corresponds to 1760845 galaxies and will be referred 
    as \texttt{hscPho}. 
    %% Saved as "dr1_wide_galaxy_hscPho.fits"

    Here we limit our study to the very high-mass end where the majority of galaxies 
    have either a spectroscopic redshift or a robust red-sequence photo-$z$ from the 
    \redm{} catalog (see Section \ref{ssec:redmapper}).  
    We match the \texttt{hscPho} sample with a spec-$z$ catalog compiled by the HSC 
    team\footnote{It is created by matching HSC objects with all publicly available 
    spectroscopic redshifts (e.g.\ SDSS/BOSS; GAMA). The spec-z quality flag from 
    different catalogs are homogenized into a single flag that indicates secure 
    redshifts.}.  
    To ensure reasonable \mstar{}-completeness at the high-\mstar{} end we focus on
    the redshift range $0.3 \leq z \leq 0.5$. 
    %% This is catalog: "dr1_specz_use.fits"

    For objects without a spectroscopic redshift, we match them with the central 
    galaxies from the \redm{} catalog using a $2.0^{\arcsec}$ matching radius. 
    The matched objects with useful red-sequence photo-$z$ ($z_{\lambda}$) are also 
    included in the final sample of bright galaxies with reliable redshift between 
    $0.3 \leq z \leq 0.5$. The accuracy of the red-sequence photo-$z$ is sufficient 
    (median $|z_{\lambda} - z_{\mathrm{Spec}}| {\sim} 0.01$ for the ones with a 
    spec-$z$) for the goal of this work.
    The \redm{} catalog provides an additional 133 unique redshifts for massive 
    galaxies in our sample.
    %% Estimated from "dr1_redbcg_hsc_sdss_gama_1arcsec.fits"

    In the end, at $0.3 \leq z \leq 0.5$, we gather a sample of 25286 galaxies with 
    reliable redshift information (will be referred as \texttt{hscZ}).
    %% Saved as "dr1_wide_galaxy_icmodel_21.5.fits"
    Among them, the BOSS and SDSS ``legacy'' LRG samples provide the majority of 
    spec-$z$. 
    By comparing with the Stripe 82 Massive Galaxy Catalog
    (\texttt{S82-MGC}; \citealt{Bundy2015}), \citet{Leauthaud2016} have measured the
    \mstar{} completeness of the combined BOSS and SDSS samples. 
    They estimate that the BOSS spec-$z$ sample is about 80\% complete at 
    \logms{}$\geq 11.6$ at $0.3 < z < 0.5$. 
    The GAMA survey, which partially overlaps with the HSC footprints, supplements 
    our \texttt{hscZ} sample by 14\%. 
    Based on \citet{Taylor2011} (e.g.\ their Fig~~6), at $z{\sim} 0.3$, the GAMA 
    sample is 80\% complete down to $10^{10.8}$\msun; but only 80\% complete to 
    $10^{12.0}$\msun at $z{\sim} 0.5$. 
    We can therefore expect our sample to be roughly \mstar{}-complete above 
    \logms{}$\geq 11.5$-$11.6$. 
    This question will be addressed more carefully using the overlap region between 
    HSC and the \texttt{S82-MGC} in Section \ref{ssec:s82}.

%% ------------------------------------------------------------------------------------ %%  
\subsection{The \redm{}{} cluster catalog}
    \label{ssec:redmapper}

    Our study focuses on galaxies located at the center of their dark matter halos. 
    To limit our sample to central galaxies, we use \texttt{v5.10} of the \redm{}{} 
    cluster catalog\footnote{See: http://risa.stanford.edu/redmapper/} 
    (e.g.\ \citealt{Rykoff2014}; \citealt{Rozo2015b}). 
    These authors have developed a well-tested red-sequence cluster finder that has 
    been run on SDSS DR8 (\citealt{SDSSDR8}) photometric data. 
    For each cluster, the catalog provides a photometric redshift ($z_{\lambda}$), a 
    cluster richness ($\lambda$), and identifies the most likely central galaxy (the 
    galaxy with the highest $P_{\mathrm{Cen}}$ value). 
    The \redm{}{} catalog also provides a list of member galaxies for each cluster and 
    their associated membership probabilities. 
    Details about the performance of the \redm{}~cluster catalog can be found in 
    \citet{Rozo2014}, \citet{Rozo2015a}, and \citet{Rozo2015b}. 
    Several studies have published calibrations between the \redm{}~richness estimate, 
    $\lambda$, and halo mass (e.g.\ \citealt{Saro2015, Farahi2016, Simet2016, 
    Melchior2016}). 
    All these studies consistently find that \redm{} clusters with $\lambda > 20$ 
    have $\log (M_{\mathrm{halo}}/M_{\odot}) \geq 14.0$. 
    The \redm{} catalog therefore helps us group massive central galaxies into 
    samples with different average \mhalo{}. 
    We will use the central galaxies of \redm{} clusters 
    ($P_{\mathrm{CEN}} \geq 0.7$) to form a sample of massive central galaxies that 
    live in massive halos.

    Due to the depth and resolution of SDSS, the \redm{} catalog becomes slightly 
    incomplete at lower richness ($\lambda < 40$) at $z > 0.33$. 
    To ensure a clean separation in \mhalo{}, here we only focus on the central 
    galaxies of clusters with $\lambda > 30$. 
    According to \citet{Simet2016}, this $\lambda \geq 30$ sample should correspond to 
    $M_{\mathrm{halo}}>10^{14.2} M_{\odot}$. 
    We have checked that none of the results in this work depend on the choice of this 
    $\lambda$ limit. 
    For massive central galaxies that are not in these cluster-level halos, 
    we unfortunately can not estimate their \mhalo{} individually, but it is safe 
    to assume they will have $\log (M_{\mathrm{halo}}/M_{\odot}) < 14.0$.  
%% ------------------------------------------------------------------------------------ %%

%% ------------------------------------------------------------------------------------ %% 
  \begin{figure*}
      \centering 
      \includegraphics[width=\textwidth]{fig/redbcg_smf_new}
      \caption{
          \textbf{Left:} Difference between \mcmodel{} and \mtot{} for central
      	  galaxies in halos less massive than $10^{14} M_{\odot}$ (grey dots) and 
      	  in halos more massive than $10^{14} M_{\odot}$ (red squares). 
          On average, \mcmodel{} underestimates the total stellar mass of massive 
          galaxies by 0.1 dex while in some cases, the difference can exceed 0.2 dex.
          Vertical histograms indicate the mass difference for galaxies with
          $11.6<$\logmtot{}$<11.9$. 
          \textbf{Right:}: Impact of using \mcmodel{} on the galaxy stellar mass
          function. 
          Dashed lines correspond to the SMF computed using \mcmodel{} as an estimate
          of the total luminosity whereas solid lines correspond to the SMF computed
          using the total luminosity from our 1-D profile. 
          The impact on the SMF can exceed 0.2 dex for massive central galaxies 
          living in halos larger than $\log (M_{\mathrm{halo}}/M_{\odot}) > 14.2$ 
          (red lines).
          }
      \label{fig:smf1}
  \end{figure*}

%% ------------------------------------------------------------------------------------ %%

%% ------------------------------------------------------------------------------------ %% 
\subsection{Massive Central Galaxies in Low and High Mass Halos}
    \label{ssec:mass_central}

    Based on recent constraints of $M_{\star}$-$M_{\mathrm{Halo}}$ relation 
    (e.g.\ \citealt{Leauthaud2012}, \citealt{Behroozi2013}, \citealt{Kravtsov2014}), 
    even though our sample selects very massive galaxies (\logms{}$\geq 11.5$), they 
    live in a fairly broad range of halos masses.
    With the help of the \redm{} catalog, we can broadly separate our sample into two
    groups with \logmh{}$<14.0$ and \logmh{}$>14.2$.

    % We match this sample with the central galaxies of \redm{}~catalogs using a 
    % $1.0^{\arcsec}$ radius, and it results in 704 galaxies.  
    %% Saved as "dr1_redbcg_hsc_sdss_gama_1arcsec.fits"

    We match the \texttt{hscZ} sample with the central galaxies of \redm{} clusters 
    with $\lambda \geq 30$ and $P_{\mathrm{0.7}} \geq 0.7$, and find 164 matched 
    galaxies at $0.3 \leq z \leq 0.5$.
    Our sample of \textbf{central galaxies in more massive halos} will be referred to 
    as the \rbcg{} sample. 
    The median $\lambda$ of the clusters associated with the \rbcg{} sample is 
    ${\sim} 41$, corresponding to halo mass of 
    $M_{\mathrm{halo}}{\sim} 2.2\times 10^{14}$ $h^{-1}$ $M_{\odot}$.
    Only 44 centrals from the \rbcg{} sample live in clusters with $\lambda \geq 50$
    ($M_{\mathrm{halo}} {\sim} 3.0\times 10^{14}$ $h^{-1} M_{\odot}$). 
    Our \rbcg{} sample therefore corresponds to galaxies in moderate mass halos.
    %% Saved as "dr1_redbcg_use_sed5b.fits"

    Out next goal is to construct a sample of central galaxies living in halos with 
    \logmh{}$<14.0$. 
    We first identify and remove all galaxies within a cylindrical region around all 
    \redm{} clusters. 
    We use a radius equal to $R_{\mathrm{200b}}$ and the thickness of the cylinder is 
    set to twice the uncertainty of the photometric redshift error. 
    We convert $\lambda$ to $M_{\mathrm{halo}}$ using the calibration of 
    \citet{Simet2016} and we use the mass-concentration relation from 
    \citet{Diemer2015} to compute $R_{\mathrm{200b}}$. At $0.3 < z < 0.5$, the 
    uncertainty of photo-$z$ is between 0.015 to 0.025, and is enough to exclude 
    cluster members.
    After removing galaxies associated with \redm{} clusters, the remaining galaxies 
    in our sample will be dominated by central galaxies living in halos with 
    \logmh{}$\leq 14.0$. 
    We will refer to this sample of \textbf{central galaxies in less massive halos}
    as the \nbcg{} sample.

    We note that at this stage, our \nbcg{} sample still contains galaxies with a 
    wide range of \mstar{} values. 
    Later on, once we have more accurate \mstar{} estimates, we will further limit this 
    sample to \logms{}$ > 11.5$. 
    At this very high mass end, and given that we have rejected satellites in \redm{} 
    clusters, any further satellite contamination can be safely neglected 
    (e.g. \citealt{Reid2014, Hoshino2015, Saito2016, vanUitert2016}).
    %% Saved as "dr1_nonbcg_use_sed5b.fits"
    % TODO: The actual number waits to be updated}

%% ------------------------------------------------------------------------------------ %%

%% ------------------------------------------------------------------------------------ %% 
\subsection{Summary of Sample Construction}
    \label{ssec:sample}

    Using ${\sim} 100$ deg$^2$ of HSC data, we select a large sample of massive central 
    galaxies with reliable redshift information, and broadly separate them two 
    categories based on $M_{\mathrm{halo}}$.

    The following is a summary of our sample construction.

    \begin{itemize}
        \item \texttt{hscPho} sample: this parent sample consists of bright galaxies 
            with $i_{\mathrm{cModel}} \leq 21.0$, good quality imaging, and reliable 
            \texttt{cModel} photometry in all five HSC bands in the \texttt{S15B} 
            data release. 
        \item \texttt{hscZ} sample: we limit the \texttt{hscPho} sample to galaxies 
            with reliable redshift information. 
        \item \rbcg{} sample: we select a sample of 164 galaxies at 
            $0.3 \leq z \leq 0.5$ that are the central galaxies 
            in $\lambda > 30$ \redm{} clusters. 
            They represent the central galaxies in very massive halos 
            (\logmh{}$\geq 14.2$). 
        \item \nbcg{} sample: after excluding all cluster members from the 
            \texttt{hscZ} sample, we have 14661 bright galaxies at 
            $0.3 \leq z \leq 0.5$. 
            After applying an additional \mstar{} cut (see Section 
            \ref{ssec:isedfit}), this sample corresponds to central galaxies in 
            halos with \logmh{}$<14.0$.  
    \end{itemize}

    We now proceed to measure the luminosities and stellar mass profiles for 
    all galaxies in the \rbcg{} and \nbcg{} samples.
%% ------------------------------------------------------------------------------------ %%

%% ------------------------------------------------------------------------------------ %% 
\section{Measurements of 1-D Surface Brightness Profiles}
    \label{sec:ellipse}

    Massive ETGs imaged at the depth of HSC are not well modeled by simple models such 
    as the de~Vaucouleurs model or a single \ser component model. 
    These models will fail to simultaneously capture the profile in both the inner and 
    the outer region and these models also cannot account for any radial variations in 
    ellipticity and position angle. 
    In principle, massive galaxies can still be described by more complex 
    models (e.g \citealt{Huang2013a, Huang2013b, Oh2017}), but they are still 
    sensitive to the choices of models (e.g. de~Vaucouleurs or \ser{} profile), 
    the number of components, initial guesses of parameters, and the internal 
    degeneracies among different parameters. 
    Background subtraction uncertainty can affect the 2-D model fitting dramatically, 
    especially for massive ETGs in our sample. 
    We therefore perform elliptical isophote fitting using the IRAF \texttt{Ellipse} 
    algorithm (Jedrzejewski 1987) to estimate the total luminosities of massive 
    galaxies and measure their one-dimensional profiles of stellar mass surface 
    density (\mden{}). 
    This 1-D method we adopt is less affected by the issues mentioned above. 
    Also, we will only study galaxies in the radial range where we are not sensitive 
    to either the PSF ($r<3-5$ kpc) or the background subtraction ($r>100$ kpc).

    We prepare large $i$-band cut-outs for each galaxy that extend to 750 kpc in 
    radius, along with a bad pixel mask, and the PSF model. 
    These cut-outs include all of the light of the galaxy and are large enough to 
    also evaluate the background level. 
    We choose to use $i$-band images because they trace the stellar distributions of 
    massive galaxies at $0.3 \leq z \leq 0.5$ reasonably well 
    (the observed i-band corresponds to a rest-frame $g$ or $r$ band), but also 
    because they have better seeing and much lower background levels than the $z$ 
    and Y band images (although these in principle may be better tracers of \mden{}).

    For each cut-out, to overcome the \texttt{hscPipe} ``over-deblending'' issue, we 
    use a customized procedure to detect and appropriately mask out neighbouring 
    objects. 
    Furthermore, \texttt{hscPipe} tends to over-subtract the background around 
    bright objects. 
    To correct for this over-subtraction, we first aggressively mask all objects 
    (including the central massive galaxy), and derive an empirical background 
    correction using \texttt{SExtractor}.
    Then, we run \texttt{Ellipse} on the background-corrected, masked cut-outs 
    following the methodology of \citet{Li2012}. 
    In short, we first fit each isophote using a free centroid and shape 
    (ellipticity and position angle). 
    We then fix the centroid (using the mean flux-weighted centroid) and estimate
    the mean ellipticity and position angles of all isophotes. 
    Finally, we extract a 1-D surface brightness profile along the major axis using 
    the mean ellipticity and position angle. 
    We correct these surface brightness profiles for Galactic extinction and 
    cosmological dimming, and integrate them to various radii to get the luminosity 
    within different physical (elliptical) apertures. 
    Fig~\ref{fig:ellipse} shows an example of the 1-D surface brightness and 
    ellipticity profile for a massive galaxy at $z{\sim}0.2$ and also highlights 
    a few isophotes.

    We test our procedure using different mask sizes, different \texttt{Ellipse} 
    parameters, and with or without our background correction. 
    Based on these tests, we find that our 1-D surface brightness profiles are reliable 
    up to surfaces brightness levels of $>28.5$ \sb. 
    Beyond that, some of our profiles shows signs of truncation and/or large 
    fluctuations which are due to the uncertainty in the background subtraction. 
    We choose to limit our study to surface brightness levels up to ${\sim} 28.5$ \sb. 
    This is a conservative choice but already enables us to measure light profiles 
    out to $100$ kpc on a galaxy-by-galaxy basis (no stacking). 
    For more details about the technical details of how we use \texttt{Ellipse}, 
    please see Appendix~\ref{app:ellipse}.

    For physical (e.g.\ late-stage major merger) and unphysical (e.g.\ nearby foreground 
    galaxy or bright star) reasons, we can not extract reliable 1-D profiles for a small 
    fraction of massive galaxies because they are heavily masked out. 
    This is an intrinsic limitation of the 1-D method. 
    It affects $18/164$ of the \rbcg{} sample, and a smaller fraction of the 
    \nbcg{} sample. 
    It is worth noting that this excludes most major-merging massive galaxies. 
%% ------------------------------------------------------------------------------------ %%

%% ------------------------------------------------------------------------------------ %% 
\section{Stellar Masses and Stellar Mass Density Profiles}
    \label{sec:mstar}

%% ------------------------------------------------------------------------------------ %% 
\subsection{Stellar Masses from SED Fitting}
    \label{ssec:isedfit}

    To convert luminosities into stellar mass, we assume that these massive galaxies 
    can be well described by an average \m2l{}. 
    This is a reasonable assumption considering that they are mostly dominated by 
    old stellar populations and are known to have shallow color gradients. 
    We will further justify this point by measuring their median color profiles in 
    Section \ref{ssec:ell_color}.

    We use the broadband Spectral Energy Distributions (SEDs) fitting 
    (see \citealt{Walcher2011} for a recent review) code 
    \texttt{iSEDFit}\footnote{http://www.sos.siena.edu/~jmoustakas/isedfit/} 
    (\citealt{Moustakas13}) to estimate the average \m2l{} and $k$-corrections using
    5-band HSC \cmodel{} fluxes.
    Although \cmodel{} tends to underestimate the total flux for bright, extended 
    objects, it can still yield accurate \emph{average} colors thanks to the 
    forced-photometry method that takes the PSF convolution into account
    (e.g. \citealt{SynPipeInPrep}).

    \texttt{iSEDFit} takes a simplified Bayesian approach. 
    In short, it first generates a large grid of SEDs from synthetic stellar population 
    models by drawing randomly from the prior distributions of relevant parameters 
    (e.g. age, metallicity, dust extinction, and star formation history).
    Based on these models, it uses the observed photometry and redshift to compute the 
    statistical likelihood, and generates the posterior probability distribution 
    functions (PDF) for each parameter.  
    To get the best estimate of a given parameter, \texttt{iSEDFit} integrates the 
    full PDF over all the other nuisance parameters.
    Then, the median value and the 1-$\sigma$ uncertainty are derived based on the 
    marginalized PDF. 
    Please refer to \citet{Moustakas13} for technical details.

    In this work, we derive average \m2l{} using the Flexible Stellar Population 
    Synthesis\footnote{http://scholar.harvard.edu/cconroy/sps-models}
    (FSPS; \texttt{v2.4}; \citealt{FSPS}, \citealt{Conroy2010}) model based on the MILES
    \footnote{http://www.iac.es/proyecto/miles/pages/stellar-libraries/miles-library.php}
    (\citealt{MILES1}, \citealt{MILES2}) stellar library and assuming a
    \citet{Chabrier2003} IMF between 0.1 to 100 \msun. 
    The star formation history (SFH) is assumed to follow a delayed-$\tau$ model with 
    stochastic star bursts (see Appendix~\ref{app:sed}). 
    This SFH is appropriate for massive galaxies at low redshifts 
    (e.g.\ \citealt{Kauffmann2003}). 
    For stellar metallicity, we assume a flat distribution between 0.004 to 0.03 
    (the highest value allowed by FSPS). 
    We adopt the \citet{Calzetti2000} extinction law with a order two Gamma 
    distribution of $A_{V}$ between 0 to 2 magnitude. 
    A majority of our galaxies are red and quiescent and our results are 
    not very sensitive to parameters related to the SFH or the internal dust extinction. 
    To achieve reasonable sampling across these parameters, we generate 250000 models.

    We construct five band SEDs using the forced-photometry \cmodel{} magnitudes 
    corrected for Galactic extinction. 
    \cmodel{} only accounts for the statistical error on the flux measurement but does
    not account for systematic uncertainties in the model-fitting process. 
    It therefore underestimates the flux errors of bright galaxies. 
    Based on tests using synthetic objects, we find that the average accuracy of
    \cmodel{} for bright galaxy is roughly at the 1\% level in the $r$, $i$, $z$ bands, 
    and slightly worse in the $g$ and Y-band. 
    Therefore, we supply \texttt{iSEDFit} with simplified flux errors assuming 
    $S/N = 100$ for the $riz$ bands, and $S/N = 80$ for the $g$Y bands (on average, 
    images in $gY$ bands are shallower in depth and/or have higher background noise).  
    The typical uncertainty of \logms{} is around 0.08-0.10 dex at \logms${\sim} 11.5$. 
    Please see Appendix~\ref{app:sed} for further details.

%% ------------------------------------------------------------------------------------ %%

%% ------------------------------------------------------------------------------------ %% 
  \begin{figure*}
      \centering 
      \includegraphics[width=\textwidth]{fig/average_mass_profiles_fsps1_A}
      \caption{
          \textbf{Left}: Median \mden{} profiles in three total stellar mass bins. 
              Here we combine the two \rbcg{} and \nbcg{} samples together -- 
              the halo mass dependence of these profiles will be explored in subsequent 
              figures. 
              Thin grey lines show a random subset of individual profiles. 
              The shaded region highlights the region that could be affected by the PSF. 
              Two vertical lines indicate 10 kpc (thin, dotted line) and
              100 kpc (thick, dash line). ~~~ 
          \textbf{Right}: comparison between our \mden{} profiles, previous observations, 
              and simulations. 
              The solid cyan line shows the median profile of massive elliptical 
              galaxies at $z{\sim} 0$ from \citet[][]{Huang2013a}. 
              The red long-dashed line shows the median profile of massive galaxies at 
              $0.25 \leq z < 0.50$ observed by \textit{HST} from \citet[][]{Patel2013}. 
              The purple short-dashed line shows the median radial stellar distributions 
              in massive halos from simulation using the particle tagging method
              (\citealt{Cooper13}).}
      \label{fig:avg_prof}
  \end{figure*}
%% ------------------------------------------------------------------------------------ %%

%% ------------------------------------------------------------------------------------ %% 
  \begin{figure*}
      \centering 
      \includegraphics[width=\textwidth]{fig/redbcg_prof_1}
      \caption{
          The environmental dependence of the stellar mass profiles of massive central 
          galaxies. 
          The left panel corresponds to samples matched by ``total'' stellar mass 
          (\mtot{}). 
          The right panel corresponds to samples matched by ``core-mass'' (\minn{}). 
          Orange and red lines correspond to central galaxies living in halos with 
          \logmh{} $\geq 14.2$. 
          Black and grey correspond to central galaxies living in halos with 
          \logmh{} $\leq 14.0$. 
          Thin lines show the profiles of individual galaxies while thick lines show 
          the median profile. 
          The uncertainty on the median profile is given by the shaded region and is 
          computed via bootstrap resampling. 
          The brown line in the bottom panels show the relative difference between 
          two median profiles 
          ($\Delta = \log(\mu_{\star, \mathrm{cenHighMh}}) - 
          \log(\mu_{\star, \mathrm{cenLowMh}})$). 
          Errors on the difference between the two profiles is computed via bootstrap 
          too. 
          The grey shaded regions show a Monte Carlo test to asses how likely is it to 
          obtain $\Delta$ from random sub-samples of the data. 
          To compute the grey shaded regions, we first mix the two samples 
          (\rbcg{} and \nbcg{}), then draw sub-samples of galaxies from the mixed 
          population and compute $\Delta$ in the same fashion as for our fiducial signal. 
          We repeat this process 5000 times.  
          The dark-grey shaded region (light-grey shaded region) shows the 1-$\sigma$ 
          (3-$\sigma$) fluctuations in $\Delta$ from these 5000 draws.
          }
      \label{fig:prof_1} 
  \end{figure*}
%% ------------------------------------------------------------------------------------ %%

%% ------------------------------------------------------------------------------------ %%        
    \subsection{Stellar Mass Completeness}
    \label{ssec:complete}

    With the help of the \texttt{S82-MGC} (\citealt{Bundy2015}), we investigate the 
    \mstar{} completeness of our samples. 
    The \texttt{S82-MGC} sample matches the deeper SDSS photometric data in the 
    Stripe 82 region (\citealt{Annis2014}) with the near infrared data from the United 
    Kingdom Infrared Telescope Infrared Deep Sky Survey (UKIDSS; \citealt{Lawrence2007}). 
    The deeper photometry and better photo-$z$s make the \texttt{S82-MGC} sample complete 
    to \logms{}$\geq 11.2$ at $z<0.7$ which makes it sufficient to evaluate the 
    completeness of our HSC samples.

    There are 20453 \texttt{S82-MGC} galaxies that are also in the \texttt{hscPho} 
    sample at $0.3 \leq z_{\mathrm{s82}} \leq 0.5$. 
    We compute \mstar{} values for these galaxies using our fiducial 
    procedure\footnote{The \mstar{} using HSC photometry is in excellent with the 
    values derived from the \texttt{S82-MGC} which includes NIR data from UKIDSS; 
    \texttt{S82-MGC} also uses \texttt{iSEDFit} with similar assumptions.}. 
    The right side of Fig~\ref{fig:smf1} compares the SMFs of galaxies from 
    \texttt{S82-MGC} with the SMFs of galaxies from our two 
    samples\footnote{Errors on the SMFs are estimated via bootstrap resampling}. 
    Based on Fig~\ref{fig:smf1}, we conclude that our sample of massive galaxies is 
    reasonably complete down to \logms{}${\sim} 11.6$ at $0.3 \leq z \leq 0.5$. 
    The \rbcg{} sample is less affected by completeness because we can utilize 
    photo-$z$s from \redm{}. 
    However, the completeness of the \nbcg{} sample is still limited by the 
    availability of spec-$z$s from BOSS.

    In this work, we will focus on the range $11.6 \le$ \logmtot{}$ \le 11.9$ 
    where our two samples are complete. 
    When comparing \nbcg{} and \rbcg{} we will also always make sure that the two 
    samples are well matched both in terms of \mstar{} as well as in terms of redshift. 
    Please see the next section and Appendix~\ref{app:match} for more details.
%% ------------------------------------------------------------------------------------ %%

%% ------------------------------------------------------------------------------------ %% 
\subsection{Stellar Mass Corrected for Total Luminosity}
    \label{ssec:mtotal}

    Using the best-fit stellar mass from \texttt{iSEDFit} (noted as as \mcmodel{}), 
    we estimate the average \m2l{} in the $i$-band, then use that \m2l{} to convert our 
    1-D luminosity density profiles into stellar mass density profiles. 
    We also convert our 10 and 100 kpc aperture luminosities into corresponding stellar 
    mass estimates (noted as \minn{} and \mtot{}).

    For the remainder of this paper, we will use \mtot{} as a proxy of ``total'' 
    stellar mass. 
    In Appendix~\ref{app:sedasic}, we briefly summarize the basic statistics of 
    the \rbcg{} and \nbcg{} samples using our estimates of \mtot{} and \minn{}.
    We confirm that both samples follow a similar and very tight ``red-sequence''. 
    Appendix~\ref{app:sed} describes the outputs from our SED fitting process 
    (average stellar age, metallicity, and internal dust extinction). 
    We do not utilize these quantities in this paper, but we have checked that they all 
    behave reasonably for both samples. 
    As expected, the integration of the 1-D profile out to very large radius recovers 
    more luminosity (stellar mass) compared to the \cmodel{}-based estimates. 
    This will be shown in Section \ref{ssec:s82}.

    Our methodology ignores any radial variations in \m2l{}. 
    It is well known that massive ETGs have negative optical color gradients indicating 
    gradients in \m2l{} (e.g.\ \citealt{LaBarbera2012}; 
    \citealt{DSouza2015}). 
    These color gradients are shallow and smooth out to a few times of the effective 
    radius and color gradients at the outskirts are not yet well quantified. 
    More importantly, for the comparison between the \rbcg{} and \nbcg{} samples, 
    our results will be robust as long as there is \emph{no significant difference in 
    the average color gradient between the two samples}. 
    We have confirmed that this is indeed the case and this will be discussed in 
    Section~\ref{ssec:ell_color}.

    Our method for deriving total stellar masses is similar to the 
    method adopted by the GAMA survey (\citealt{Taylor2011, Kelvin2012}). 
    We have compared our \mstar{} values with those from GAMA and the results are given 
    in Appendix~\ref{app:gama}. 
    We find a systematic differences between the two \mstar{} estimates.
    It is likely that the GAMA models fail to capture the flux associated with the 
    outskirts of galaxies due to the limited depth of SDSS.     
%% ------------------------------------------------------------------------------------ %%

%% ------------------------------------------------------------------------------------ %% 
\section{Results}
    \label{sec:result}

%% ------------------------------------------------------------------------------------ %%     
    \subsection{Impact of Missing Light on the Galaxy Stellar Mass Function}
    \label{ssec:s82}

    Different methods for estimating the ``total'' \mstar{} can have a large impact on 
    the high mass end of the galaxy stellar mass function (e.g. \citealt{Bernardi2013, 
    DSouza2014, DSouza2015, Bernardi2016a}). 
    Fig~\ref{fig:smf1} shows the SMF computed using both \mtot{} and \mcmodel{}. 
    As one can be seen from this figure, our careful 1-D approach recovers more flux 
    in massive galaxies than \texttt{cModel}. 
    More importantly, the average difference steadily increases with \mtot{} and also 
    has a dependence on \mhalo{}. 
    At \logms{} $>11.5$, the difference can be larger than 0.2 dex. 
    This difference relates to the intrinsic limitation of the \texttt{cModel} method 
    and also reflects the fact that more massive ETGs tend to have more extended 
    stellar mass distributions (e.g. \citealt{Graham2003}). 
    Generally speaking, the use of \mtot{} instead of the \mcmodel{} shifts the SMF 
    towards the higher-mass end and slightly modifies the high-mass slope. 
    The impact of missing flux is most prominent for the \rbcg{} sample.
    For studies about the SMF or other aspects of galaxy evolution based on 
    \cmodel{}-like photometry or shallower images, these effects can not be neglected.
%% ------------------------------------------------------------------------------------ %%

%% ------------------------------------------------------------------------------------ %%
\subsection{Surface Mass Density Profiles}
    \label{ssec:sbp_compare}

%% ------------------------------------------------------------------------------------ %%
\subsubsection{General Trends and Comparison with Previous Work}
    \label{sssec:sbp_inter}

    Compared to the much discussed ``mass-size'' relation, the \mden{} profiles that 
    we have derived capture the full information about the structural properties of 
    massive galaxies. 
    We argue that by studying the shapes of these galaxy profiles directly, we can 
    by-pass outstanding questions about how to accurately define and measure the 
    ``sizes'' and ``masses'' of these massive galaxies. 
    Fig~\ref{fig:avg_prof} shows the median \mden{} profiles\footnote{The median 
    profiles along with their uncertainties are derived using bootstrap resampling 
    method} of massive central galaxies at $0.3 < z < 0.5$ in three \mtot{} 
    bins\footnote{The sample is clearly not complete in the lowest \mtot{} bin, 
    but its median \mden{} profile is still useful to show the overall trend and to 
    compare with previous work}. 
    As shown in the left panel of Fig~\ref{fig:avg_prof}, we can comfortably trace
    the \mden{} profiles of these massive galaxies out to 100 kpc 
    \textbf{individually}. 
    These median \mden{} profiles have moderate differences in the inner region, 
    but show significant differences in the outskirts with more massive galaxies 
    showing more prominent stellar envelopes. 
    Most of these massive galaxies ($\ge 10^{11.4} M_{\odot}$) are slowly-rotating 
    (e.g.\ \citealt{Cappellari13b}) giant ellipticals with ``boxy'' inner isophotal 
    shapes (e.g.\ \citealt{Kormendy2009}), and slightly flattened inner \mden{} 
    profiles (e.g.\ \citealt{Lauer07}). 
    But in the outskirts, the structure of these galaxies is 
    \emph{clearly not self-similar}. 
    Some of our \mden{} profiles show signs of unphysical truncation and fluctuation 
    related to inaccurate sky subtraction. 
    In this paper, we do not use profiles beyond 100 kpc, even though the median 
    \mden{} profiles for the two most massive bins behave reasonably out to 
    ${\sim} 200$ kpc.

    Fig~\ref{fig:avg_prof} compares our results with previous work. 
    Most previous studies have relied on stacking techniques to measure surface 
    brightness profiles out to large radii.
    So far, the surface brightness profiles of massive galaxies out to such large 
    radii was mostly derived via image stacking technique that suffers from several 
    systematic issues (e.g. \citealt{Tal2011, DSouza2015}; but also see:
    \citealt{Capaccioli2015}); and the \mden{} profiles are even more rarely 
    available.

    \citet{Huang2013a} derived \mden{} profiles for a small sample of very nearby 
    ellipticals (within 100 Mpc; median \logms{} ${\sim} 11.3$ based on shallower 
    images) from the Carnegie-Irvine Galaxy Survey (CGS, \citealt{CGS1}). 
    Because this is a very low redshift sample, these profiles are robust to $r=1$ kpc. 
    The \citet{Huang2013a} profiles show good agreement with ours in the region of 
    overlap and out to 50 kpc. 
    CGS images are deeper than SDSS images in the $r$-band, but the median 
    profile from \citet{Huang2013a} still only reaches ${\sim} 50$ kpc for $z{\sim} 0$ 
    massive galaxies.
    Meanwhile, our deep HSC images can reliably deliver individual \mden{} profiles 
    for $z{\sim} 0.4$ galaxies out to at least 100 kpc.

    \citet{Patel2013} extracted a median \mden{} profile of massive ETGs at 
    $0.25 < z < 0.50$ using stacked \textit{HST}/ACS images. 
    These galaxies are selected at a constant cumulative number density and are 
    thought to be the progenitors of $z=0$ massive ETGs (e.g. \citealt{Leja2013}).  
    The median \mstar{} of the \citet{Patel2013} sample is ${\sim} 10^{11.2} M_{\odot}$ 
    which is lower than our lowest mass bin. 
    However, \citep{Patel2013} use the BC03 SSP model which are roughly 0.1 dex 
    lower than our FSPS masses (Appendix~\ref{app:sed}). 
    Furthermore, the \citep{Patel2013} images are shallower than ours meaning 
    that there masses will be lower than ours because of missing light in the 
    outskirts. 
    Given these two considerations, it is reasonable to roughly compare the 
    \citet{Patel2013} profile with the one in our lowest stellar mass bin. 
    The superb resolution of the \textit{HST}/ACS image means that the profile in
    \citet{Patel2013} is unaffected by the seeing even down to 1 kpc. 
    The good agreement between our profile and the one derived from HST image 
    shows that our profiles are robust at $r > 3$ kpc and that we can accurately 
    measure \minn{}.

    Finally, we also compare with the predicted median \mden{} profile of central 
    galaxies in massive halos ($13.5 < \log M_{200,c} < 14.0$) from cosmological 
    simulation where the \mden{} profiles of galaxies are calculated using the 
    particle-tagging technique (e.g. \citealt{Cooper10}).

    The simulated \mden{} profile is also affected by limited resolution in the 
    center, but it is already quite similar to the median \mden{} profile for 
    the $11.6 <$ \logmtot{} $< 11.8$ bin within 40 kpc. 
    However, compared with our data, this simulation predicts stellar envelope that
    are too prominent at this halo mass. 
    This types of comparisons between our HSC data and predictions from simulations 
    will help to fine-tune simulations and to also investigate the main physical 
    mechanisms that drive the assembly of massive galaxies and the build up of 
    stellar envelopes.

    Table~1 provides tabulated values for the median profiles that are 
    displayed in Fig~\ref{fig:avg_prof}. 
    These profiles are also available  
    \href{https://github.com/dr-guangtou/hsc_redbcg/tree/master/profiles}{here}:
    {\url{https://github.com/dr_guangtou/hsc_redbcg/tree/master/profiles/}}.

%% ------------------------------------------------------------------------------------ %%

%% ------------------------------------------------------------------------------------ %% 
\subsubsection{Environmental Dependence of the Profiles of Massive Galaxies}
    \label{sssec:sbp_mtot}

    We now ask whether or not galaxies from our sample have \mden{} profiles that 
    depend on \mhalo{} (at fixed \mstar{}). 
    For this, we will build matched samples drawing from \rbcg{} and \nbcg{} that are 
    carefully matched in \mstar{} and redshift using a KDTree algorithm 
    (see Appendix~\ref{app:match} and Fig~\ref{fig:match} for details). 
    By matching the redshift distributions within $0.3 < z < 0.5$, we ensure that the 
    \mden{} profiles from both samples are impacted by seeing and background subtraction 
    at similar levels. 
    To match the two samples by stellar mass, we will use two different methods:

    \begin{itemize}

        \item We match the two samples according to \mtot{}. 
            We select galaxies from \rbcg{} with $11.6 < $\logmtot{} $<11.9$ 
            (median \logmtot{} ${\sim} 11.74$). 
            There are 45 \rbcg{} galaxies in this range. 
            We then match this distribution by drawing galaxies from \nbcg{}. 
            There are 229 unique \nbcg{} galaxies in the matched sample. 
            In this \mtot{} range, 100 kpc corresponds to 5-8 times the effective radius 
            and \mtot{} is a good proxy of total stellar mass.

       \item We match the two samples according to \minn{}. 
           We select galaxies from \rbcg{} with $11.2 < $\logminn{}$ < 11.6$ 
            (median \logminn{}${\sim} 11.36$).
            For the 56 galaxies from the \rbcg{} sample, we achieve excellent matches
            in \minn{} and redshift using 375 \nbcg{} galaxies.
            Motivated by recent observations and simulations, we use \minn{} as the 
            proxy of the \mstar{} of the ``in-situ'' component.
            We will discuss more about this assumption in 
            Section~\ref{sec:discussion}.

    \end{itemize}

%% ------------------------------------------------------------------------------------ %%     
    Fig~\ref{fig:prof_1} compares the profiles of galaxies at fixed \mtot{} and at fixed 
    \minn{} for our two matched samples, and shows the main results of this paper which 
    is that we find a clear halo mass dependence in the \mden{} profiles of massive 
    galaxies. 
    The bottom panel of Fig~\ref{fig:prof_1} shows the difference between the two 
    profiles, $\Delta = \log(\mu_{\star, \mathrm{cenHighMh}}) - 
    \log(\mu_{\star, \mathrm{cenLowMh}})$, 
    with errors computed via bootstrap resampling. 
    In Appendix~\ref{app:robust}, we present a variety of tests that verify the 
    robustness of these results against different uncertainties.

    There are several important points to be noticed in Fig~\ref{fig:prof_1}. 
    First, when galaxies are matched by \mtot{}, although we detect a robust 
    environmental dependence, it mainly only becomes prominent in the outskirts 
    at $R>50$ kpc. 
    This partially explains the difficulties for previous works to detect this effect
    on shallower images. 
    Second, we find that central galaxies in more massive dark mater halos display 
    more flattened inner profile and more significant outer stellar envelope than 
    the ones living in less massive halos. 
    Third, the median \mden{} profiles of \rbcg{} and \nbcg{} cross each other 
    at ${\sim} 15$-20 kpc, the expected $R_{\mathrm{e}}$ for ETGs at this \mtot{}. 
    Finally, such systematic difference in structure may become more significant 
    at very high-mass end. 
    We see evidence of this effect when we compare the $Delta$ profiles in two \mtot{} 
    bins (see Fig~\ref{fig:prof_2} and discussion in 
    Appendix~\ref{app:robust})\footnote{Though the lower \mtot{} bin becomes somewhat
    incomplete} and when we discuss the relation between \mtot{} and \minn{} for 
    these two samples (see Fig~\ref{fig:scaling_relation}). 
    This potential interesting result and its implication is worth more investigations 
    later.

    Meanwhile, the subtle environmental dependence of \mden{} profiles for the 
    \mtot{}-matched samples suggests that environment does not play the major role in 
    shaping massive galaxies \textbf{at fixed total stellar mass}.
    Therefore it is interesting to ask, for massive galaxies that \textbf{contain 
    similar \mstar{} in the in-situ component}, how does environment affect their 
    subsequent mass assembly?
    Assuming that the \minn{} is reasonable proxy of in-situ mass, we look into this 
    question by comparing the \mden{} profiles of the \minn{}-matched samples ( 
    Fig~\ref{fig:prof_1}, right).  
    Although the median \mden{} profiles are similar at $r \leq 10$ kpc, the
    central galaxies in more massive halos clearly possess much more prominent outer
    envelope. 
    This result is robust in a series of statistical and systematical tests 
    (see Appendix~\ref{app:robust}.

    At the same time, the uniform slope of the \mden{} profiles within 
    10 kpc is in sharp contrast with the large scatter of the outskirts. 
    It hints that these two regions are shaped by distinctive processes, and may 
    explain their differences in environmental dependency. 
    In simulations, intensive dissipative process helps create 
    self-similar de~Vaucouleur-like ($n{\sim} 4$) \mden{} profile in the inner part 
    of massive galaxy (e.g. \citealt{Hopkins2008}), and the \mhalo{}-dependent
    merging history leads to shallower stellar halo profile for central galaxy 
    of more massive halo (e.g. \citealt{Pillepich2014}). 
    The stochastic nature of the merging process also helps explain the 
    large scatter in the distributions of outer \mden{} profiles. 
    We illustrate this interesting diversity in Fig~\ref{fig:m100_m10_color}, where 
    we show the 3-color images of randomly selected massive central galaxies that 
    have similar \minn{} 
    ($11.1 \leq \log (M_{\star,\ 10\mathrm{kpc}}/M_{\odot}) < 11.2$)
    but different \mtot{}.   
    From the top-left to the bottom-right, the \mtot{} of these galaxies increases
    from $10^{11.22} M_{\odot}$ to $10^{11.75} M_{\odot}$.

    \textbf{In summary, we reliably detect subtle, but systematic \mhalo{}-dependence 
    (environmental dependence) of structure in massive central galaxies}, which seems 
    to be consistent with the expectation of more frequent (minor) mergers in denser 
    environment. 
    These minor, non-dissipative mergers do not alter the inner \mden{} profiles 
    a lot, but are efficient in depositing stars to the outer envelope
    (e.g. \citealt{Hilz2013}, \citealt{Oogi2013})

    The median profiles shown in Fig~\ref{fig:prof_1} are available in Table~1.
%% ------------------------------------------------------------------------------------ %%

%% ------------------------------------------------------------------------------------ %% 
\subsection{Ellipticity and Color Profiles}
    \label{ssec:ell_color}

%% ------------------------------------------------------------------------------------ %% 
  \begin{figure*}
      \centering 
      \includegraphics[width=\textwidth]{fig/redbcg_discussion_2}
      \caption{
          Radial variations in ellipticity and optical colors for massive galaxies. 
          The format of this figure is similar to the right hand side of 
          Fig~\ref{fig:prof_1}. 
          Top panels show the ellipticity profile, middle panels show $g-r$, and lower 
          panels show $g-i$. 
          We compare our results with those from \citet{Tal2011} based on stacking 
          large samples of luminous red galaxies in SDSS at $z{\sim} 0.4$ 
          (solid blue line on top panel). 
          We also compare with the results from a stacking analysis of nearby massive 
          galaxies with high concentration index ($C>2.6$) in 
          \citet[][blue dash lines on the top and middle panels]{DSouza2014}. 
          Finally, we also compare with the average $g-r$ and $g-i$ color profiles 
          from a large sample of nearby elliptical galaxies in \citet[][blue, solid 
          lines on the middle and bottom panels]{LaBarbera2010}.
          }
      \label{fig:ell_color}
  \end{figure*}
%% ------------------------------------------------------------------------------------ %%

%% ------------------------------------------------------------------------------------ %% 
    We now study how the ellipticities (computed via \texttt{Ellipse}) and $k$-corrected 
    optical colors of massive galaxies vary with radius. 
    We also study whether or these radial trends differ between our various samples. 
    These quantities contain interesting information about the physical processes that 
    shape galaxies, but here we study them mainly with two goals in mind:

    \begin{enumerate}

        \item To see whether the ellipticity profile depends strongly on environment, 
            and to see whether or not the fact that we use an using average isophotal 
            shape for our our 1-D \mden{} profiles has any impact on our results.

        \item To verify that the average color gradient does not depend strongly on 
            the environment and that is not steep enough to cause a problem with our 
            choice the assumption that it is reasonable to use average \m2l for 
            comparing massive galaxies. 
            We extract color profiles by applying the isophotes in $i$-band to images 
            in other bands in ``forced photometry'' manner.

    \end{enumerate}

    We compute ellipticities and color profiles in the radial range of 5-50 kpc 
    (this is the range where we can safely ignore differences in sky subtraction and 
    seeing across different bands). 
    We apply Galactic extinction and $k$-correction to both $g-r$ and $g-i$ color 
    profiles; the $k$-correction values are also from \texttt{iSEDFit} results.
    We show the average ellipticity, $g-r$, and $g-i$ color profiles for the 
    \mtot{}-matched (\minn{}-matched) samples of \rbcg{} and \nbcg{} galaxies on the 
    left (right) side of \ref{fig:ell_color}. 
    We also compare our results with previous work based on stacking SDSS galaxies 
    (e.g. \citealt{LaBarbera2010, Tal2011, DSouza2014}).

    We find that in general, the ellipticities of massive galaxies increases with 
    radius from $e\le 0.2$ around 5-10 kpc to $e{\sim} 0.3$ at 40-50 kpc. 
    This is consistent with previous 2-D modeling results from \citet{Huang2013a}. 
    The ellipticity profiles of the \mtot{}-matched samples are very similar at 
    $r<50$ kpc (2-3$\times R_{\mathrm{e}}$ for their \mstar{}). 
    However, for the \minn{}-matched samples, the \rbcg{} galaxies have higher 
    ellipticity than the \nbcg{} ones at $r > 10$ kpc, while their inner ellipticity 
    profiles are very similar.  
    This further hints that massive central galaxies in different halos but with 
    similar ``in-situ'' mass may experience different late-time merging histories. 
    Overall, the relative low ellipticity value and the fact that the ellipticity 
    gradients are shallow lead us to conclude that the use of an average 
    isophotal ellipticity is fine for this present study.

    Previous attempts to constrain ellipticity gradients have relied on image 
    stacking techniques (e.g. \citealt{Tal2011}, \citealt{DSouza2015}). 
    Compared to our measurements, we find that stacked ellipticity profiles tend to 
    underestimate the ellipticity in the outskirt. 
    Ellipticity profiles contain additional information about the assembly histories 
    and in future work, it would be of interest to compare our ellipticity profiles 
    with predictions from hydrodynamic simulations such as Illustris 
    (\citealt{Vogelsberger2014}, \citealt{Genel2014}) or \textit{Horizon-AGN} 
    (\citealt{Dubois2014}).

    In terms of color gradients, we find shallow and negative gradients of median 
    $g-r$ and $g-i$ color profiles for both \rbcg{} and \nbcg{} galaxies. 
    For both \mtot{}- and \minn{}-matched samples, we find no significant difference
    in their median color profiles, which supports the assumption of using single 
    average \m2l{} value to estimate \mden{} profiles.
    The median $g-r$ color profiles are quite consistent with those based on 
    stacked SDSS images. 
    For the $g-i$ profiles, the HSC ones show steeper gradients than the SDSS ones, 
    however, it is important to point out that the SDSS $i$-band suffers from the 
    so-called ``red-halo'' effect of (e.g. \citealt{Wu2005}, \citealt{Tal2011}) that 
    could make the SDSS color appears redder. 
    In contrast, HSC $i$-band image do not suffer from this effect (thanks to thick 
    CCDs), therefore can help provide more accurate color estimates.
%% ------------------------------------------------------------------------------------ %%

%% ------------------------------------------------------------------------------------ %% 
\subsection{The Environmental Dependence of Scaling Relations}
    \label{ssec:scaling}

%% ------------------------------------------------------------------------------------ %% 
  \begin{figure*}
      \centering 
      \includegraphics[width=\textwidth]{fig/redbcg_scaling_relation}
      \caption{
          \textbf{Left}: The mass-size relations for \rbcg{} (orange squares) and 
          \nbcg{} (grey dots) galaxies. 
          Two vertical lines highlight our $11.6<$ \logms{} $<11.9$ mass bin. 
          The red solid line shows the best fit mass-size relation for \rbcg{} and the 
          grey dashed line shows the best fit relation for \nbcg{}. 
          Shaded regions in lighter colors show the 1-$\sigma$ uncertainties
          from MCMC sampling.  
          The green dashed line shows the mass-size relation for nearby central 
          galaxies in massive groups from \citet{HCompany13}. 
          \textbf{Right:} The relations between \mtot{} and \minn{} for the 
          \rbcg{} and \nbcg{} galaxies, along with the best-fit scaling relations for 
          both samples.
          }
      \label{fig:scaling_relation} 
  \end{figure*}
%% ------------------------------------------------------------------------------------ %%

\subsubsection{Mass-Size Relation}
    \label{sssec:mass_size}

    Despite numerous attempts, previous studies have failed to clearly 
    demonstrate the existence of an environmental dependence of mass-size relation
    at low-$z$ (e.g. \citealt{Weinmann2009, Nair2010, HCompany13, Cerbrian2014})

    %Despite being the focus of previous investigations (e.g. \citealt{Shankar2013};
    %\citealt{Leja2013}, \citealt{vdWel2014}), the relation between \mstar{} and 
    %effective radius (or half-light radius; $R_{\mathrm{e}}$ or $R_{\mathrm{50}}$) 
    %has yet to provide clear evidence of environmental dependence of structures of 
    %massive galaxies at low-$z$ (e.g. \citealt{Weinmann2009}, \citealt{Nair2010}, 
    %\citealt{HCompany13}, \citealt{Cerbrian2014}), and sent controversial messages 
    %about the pictures at high-$z$ (e.g. \citealt{MCooper2012}, \citealt{Papovich2012}, 
    %\citealt{Kelkar2015}).

    By using deep HSC images, and by comparing\mden{} profiles directly, 
    we have shown that the structures of massive galaxies do depend on the masses of 
    their host halos (see the left panel of Fig~\ref{fig:scaling_relation}). 
    We now present our results in terms of the more traditional ``mass-size'' relation. 
    For this purpose, we will use \mtot{} as our ``total'' \mstar{} and the radius 
    enclosing 50\% of luminosity within 100 kpc ($R_{\mathrm{50}}$; derived from the 
    $i$-band curve-of-growth) as our ``size''. 
    This definition of ``effective radius'' is more robust against structural details, 
    choice of model, and the background subtraction. 
    Our massive galaxies are large enough that we do not need to worry about the 
    effects of seeing.

    Fig~\ref{fig:scaling_relation} shows the \mstar{}-size relations for both samples. 
    We fit the \logmtot{}-$\log (R_{\mathrm{50}}/\mathrm{kpc})$ relations at 
    \logmtot{}$\geq 11.6$\footnote{We use MCMC and the ensemble sampler 
    \texttt{emcee} (\citealt{Emcee}) with initial guesses are based on maximum 
    likelihood estimates, and we assume reasonable flat priors for parameters.}.
    The best-fit relation for \rbcg{} is:

    \begin{equation}
        \begin{aligned}
        \log (R_{\mathrm{50}}/\mathrm{kpc}) = & (0.73\pm0.13) \times \log (M_{\star, 100\ \mathrm{kpc}}/M_{\odot}) \\ & -(7.49\pm1.56)
        \end{aligned}
    \end{equation}

    \noindent And for \nbcg{}, we find:

    \begin{equation}
        \begin{aligned}
        \log (R_{\mathrm{50}}/\mathrm{kpc}) = & (0.68\pm0.06) \times \log (M_{\star, 100\ \mathrm{kpc}}/M_{\odot}) \\ & -(6.88\pm0.75)
        \end{aligned}
    \end{equation}

    \noindent As shown in Fig~\ref{fig:scaling_relation}, the slopes of the \mtot{}-size 
    relations are similar between the two samples, but the normalization of the 
    \mtot{}-size relation is different for \rbcg{} compared to \nbcg{}. 
    This result is robust against the ranges of \mstar{}, also the definitions of 
    ``total'' \mstar{} and  half-light radius\footnote{Using \mstar{} within 120 or 150 
    kpc, or using the $R_{\mathrm{50}}$ derived within these apertures will not change 
    the results.}.

    We qualitatively compare our results with \citealt{HCompany13} (HC13; green solid line) 
    who studied $z\leq 0.09$ central ETGs in  $14\le$ \logmh{} $<15$ halos. 
    HC13 used the group catalog by \citet{Yang2007} to estimate \mhalo{} after empirically 
    convert their \mstar{} from a Kroupa IMF to a Chabrier IMF by applying a constant 
    -0.05 shift (see \citealt{Bernardi2016}).
    They estimated the 2-D $R_{\mathrm{e}}$ using single-\ser{} model fitting to SDSS 
    images, and derived \mstar{} based on SED fitting using the BC03 (\citealt{BC03}) 
    synthetic population model. 
    We add $+0.1$dex to the HSC13 results to account for the systematic 
    difference between their BC03 and our FSPS model (see Appendix~\ref{app:sed}). 
    The mass-size distributions for both \rbcg{} and \nbcg{} galaxies follow the 
    HC13 relation reasonably well (with slightly shallower slopes) even at 
    \logmtot{}$< 11.6$ where our samples start to become incomplete. 
    In contrast to the present study, HC13 finds no halo mass dependence in the 
    mass-size relation for $12.5\le$ \logmh{} $<15.0$ halos.
    A possible explanation for this difference is that the region that strongly depends 
    on halo mass is the low surface brightness outer envelope where the stars are 
    mostly assembled through multiple mergers. 
    This region is difficult to detect on shallow SDSS images, but the much deeper 
    HSC images can provide accurate measurements of total \mstar{} and size.

    In conclusion, we confirm that massive central galaxies at $0.3 < z < 0.5$
    follow \mstar{}-size relations that depend on the masses of their host halos. 
    Central galaxies in denser environment (higher \mhalo{}) have slightly 
    larger $R_{\mathrm{50}}$ comparing to ones in smaller haloes at fixed 
    \mtot{}. 
%% ------------------------------------------------------------------------------------ %%

%% ------------------------------------------------------------------------------------ %% 
\subsubsection{\mtot{} - \minn{} Relation}
    \label{sssec:m100_m10}

    We also investigate the relationship between \mstar{} for two different
    fixed physical apertures (\mtot{} v.s. \minn{}) as alternative tool to help us 
    understand the relationship between \mhalo{} and the assembly history of massive
    galaxies.   
    Comparing to the mass-size relation, this relation is not affected by the 
    ambiguous definition of ``size'', and is much easier to apply to larger samples 
    and to simulations.

    The right panel of Fig \ref{fig:scaling_relation} compares \mtot{} and \minn{}. 
    We find that our \rbcg{} and \nbcg{} occupy different locations within this plane,
    and follow different best-fit \mtot{} - \minn{} relations. 
    For \rbcg{} galaxies we find:

    \begin{equation}
        \begin{aligned}
        \log (M_{\star, 10\ \mathrm{kpc}}/M_{\odot}) = & (0.48\pm0.06) \times \log (M_{\star, 100\ \mathrm{kpc}}/M_{\odot}) \\ & +(5.72\pm0.75).
        \end{aligned}
    \end{equation}

    \noindent In the same range of \mtot{}, the best-fit relation for \nbcg{} is:

    \begin{equation}
        \begin{aligned}
        \log (M_{\star, 10\ \mathrm{kpc}}/M_{\odot}) = & (0.56\pm0.03) \times \log (M_{\star, 100\ \mathrm{kpc}}/M_{\odot}) \\ & +(4.82\pm0.30).
        \end{aligned}
    \end{equation}

    We have checked that our results are unchanged if we replace the \minn{} with 
    \mstar{} within 5 or 15 kpc apertures. 
    Although the $M_{\star, 5\ \mathrm{kpc}}$ values are less reliable than \minn{} 
    due to the impacts of seeing, the relation using $M_{\star, 5\ \mathrm{kpc}}$ shows 
    more significant \mhalo{} dependence at the high-mass end.
    Also, best-fit relation using the full \mtot{} range results in even shallower 
    slope for the \rbcg{} galaxies.

    Our results suggest that: 
    (1) At fixed \mtot{}, central galaxies in more massive halos tend to have 
    lower fractions of \mstar{} stored in the inner region; 
    (2) For massive galaxies with similar \minn{}, the ones in more massive halos on 
    average have more prominent outer envelopes. 
    The scientific implication of this result is briefly discussed in 
    Section~\ref{sec:discussion}.
%% ------------------------------------------------------------------------------------ %%

%% ------------------------------------------------------------------------------------ %% 
  \begin{figure}
      \centering 
      \includegraphics[width=\columnwidth]{fig/redbcg_discussion_1}
      \caption{Comparison of the median \mden{} profiles of massive galaxies in the 
          two lower \mtot{} bins from Fig~\ref{fig:avg_prof} with other observations 
          and simulations, focusing on the inner region. 
          The format is similar with the right panel of Fig~\ref{fig:avg_prof}.
          We include: 
          1) the median \mden{} profile of inner component from structure 
          decomposition of massive elliptical galaxies at $z{\sim} 0$ from 
          \citet[][Cyan, solid]{Huang2013a}; 
          2) the median \mden{} profile of massive galaxies at $1.0 < z < 1.5$ 
          from \textit{HST} observations in \citet[][Red, dashed]{Patel2013}; 
          3) the median \mden{} profile of the ``in-situ'' stellar components in 
          simulated massive halos from \citet[][Purple, dot-dashed]{Cooper13};
          4) the \mden{} profile of the very massive cD galaxy at $z{\sim} 1.1$ 
          discovered by \citet[][Yellow, dashed]{Liu2013} in the Hubble 
          Ultra-Deep Field.}
      \label{fig:discussion_1}
  \end{figure}
%% ------------------------------------------------------------------------------------ %%

%% ------------------------------------------------------------------------------------ %% 
\section{Discussion}
    \label{sec:discussion}

    By using deep imaging from the HSC survey we have shown that the stellar mass 
    density profiles of massive central galaxies depend systematically on the mass 
    of their host dark matter halos. 
    Here we discuss the physical origin, scientific implications, and potential 
    caveats of this result.

%% ------------------------------------------------------------------------------------ %%

\subsection{The ``Two phase'' Formation Scenario of Massive Galaxies and the 
            Build-Up of Outer Envelopes}

    Our results are consistent with the picture that massive central galaxies form via 
    a ``two-phase'' scenario. 
    According to this theory, the cores of massive galaxies were formed at $z{\sim} 2$ 
    during an intense period of in-situ star formation. 
    The outskirts of massive galaxies are then built up via a more gradual second 
    phase of evolution (the ``ex-situ'' phase) which is dominated by mass growth via 
    the accretion of satellite systems. 
    Non-dissipative minor mergers mostly deposit stars on the outskirts of the 
    centrals and do not have large impact on the central \mden{} profile 
    (e.g. \citealt{Oogi2013, Bedorf2013}). 
    Hence, according to this picture, the outer stellar envelopes of massive centrals 
    should correlate with the masses of their host halos, in qualitative agreement with
    our findings.

    State of the art hydrodynamic simulations suggest that massive dark matter halos 
    have central galaxies that have outer stellar envelopes with shallow slopes 
    (e.g. \citealt{Pillepich2014}). 
    This is consistent with the results in Fig \ref{fig:prof_1}.

    Meanwhile, these simulation also point out that the fraction of accreted stars 
    increases steeply with total \mstar{} and that the ``ex-situ'' component
    starts to dominate the \mden{} profiles around $R_{\mathrm{e}}$ for 
    \logms{}$\geq 11.5$ galaxies. 
    In this work, we find that the median \mden{} profiles of \rbcg{} and \nbcg{} 
    samples start to show differences around 15-20 kpc, which is close to 
    their $R_{\mathrm{e}}$. 
    Furthermore, the environmental dependence of our median \mden{} profiles 
    becomes more significant in our highest \mtot{} bin (see Fig \ref{fig:prof_4}). 
    These results all align with the picture that the \mhalo{}-dependence of the 
    \mden{} profiles of massive galaxies is driven by the variations in the 
    ``ex-situ'' component at large radii.

    We now investigate to what degree our assumption that the inner 10 kpc \mstar{}
    corresponds to an in-situ mass.    
    Fig \ref{fig:discussion_1} compares the median \mden{} profiles of the 
    combined \rbcg{} and \nbcg{} samples in two \mtot{} bins 
    ($11.4\leq$\logmtot{}$<11.6$ and $11.6\leq$\logmtot{}$<11.8$) with 
    (1) the median \mden{} profiles of massive ETGs at $1.0 < z < 1.5$ in
    \citealt{Patel2013} since they are considered as the progenitors of 
    ${\sim} 10^{11.5} M_{\odot}$ ETGs at $z=0$.    
    (2) the inner components of $z{\sim} 0$ ellipticals from 2-D decomposition 
    (\citealt{Huang2013a}).
    As proxy of the in-situ components, \citet{Huang2013b} showed they are 
    structurally similar to the compact ``red nuggets'' at high-$z$. 
    (3) the ``in-situ'' components of simulated central galaxies in massive halos 
    from \citet{Cooper13} (the inner ${\sim} 5$ kpc is quite uncertain due to the 
    resolution).  
    The comparison confirms that \minn{} should be reasonable proxy of the 
    ``in-situ'' mass\footnote{We convert these \mden{} profiles to the same 
    Chabrier IMF; but there are still differences in median \mstar{} and 
    details in the \m2l{} estimates}.  
    More importantly, it shows that the \mden{} profile at 
    $\mathrm{R} > 15$-20 kpc is likely dominated by stars that are not related 
    to the ``in-situ'' processes.

    We also compare with a uniquely massive BCG at high redshift: 
    a ${\sim} 10^{11.4} M_{\odot}$ BCG with distinctive ``cD''-like envelope at 
    $z{\sim} 1.1$ (\citealt{Liu2013}).  
    Its \mden{} profile follows the median profile of $11.6\leq$\logmtot{}$<11.8$ 
    massive galaxies nicely inside 10-15 kpc, but becomes much steeper in the outskirt.  
    This unique object may present an interesting case where the inner ``core'' of a 
    massive BCGs is already in pace at $z{\sim} 1$ when its extended envelope is still
    under construction.
    More BCGs like this at $0.5 < z < 1.0$ can shed light on this intriguing 
    process, and help us confirm whether the environmental dependence of structure 
    has already emerged within massive galaxies at higher redshift 
    (e.g. \citealt{Papovich2012})

    At the same time, we also find \rbcg{} galaxy tends to have lower \mstar{} within 
    inner 5-10 kpc and shallower inner \mden{} profile than the \nbcg{} one at 
    fixed \mtot{}. 
    Although the HSC images do not have sufficient resolution to investigate this in 
    details, it is worth pointing out that this difference could be physically 
    interesting. 
    Observations of compact quiescent galaxies at high-$z$ often reveal higher 
    \mden{} in the central region than their descendants, and certain process need to  
    be responsible for reducing the central \mden{} (e.g. strong adiabatic expansion 
    induced by powerful AGN feedback; \citealt{Fan2008}).
    Coalesce of super-massive black hoes (SMBHs) at the end of major merger, which is 
    common in the assembly history of massive galaxies, can also flatten the central 
    \mden{} profile (e.g. \citealt{Milosavljevi2002}).
    On the right side of Fig \ref{fig:scaling_relation}, we find a few \rbcg{} galaxies 
    with very low \minn{} at high \mtot{} end that are apparently not bothered by 
    problematic photometry or exceptionally bad seeing.  
    They may be similar to the recently discovered massive BCGs with very large core 
    (a few kpc; e.g. \citealt{Postman2012, LopezCruz2014}) resulting from SMBH mergers.
    The impacts of these processes, and their dependence on \mhalo{} is an interesting 
    topic to investigate in the future.
%% ------------------------------------------------------------------------------------ %%

%% ------------------------------------------------------------------------------------ %% 
  \begin{figure*}
      \centering 
      \includegraphics[width=\textwidth]{fig/redbcg_discussion_5}
      \caption{
          \textbf{Left}: Relation between \mtot{} and the logarithmic mass difference 
          between \mmax{} and \mtot{} for both \rbcg{} (red square) and \nbcg{} 
          (grey dot) galaxies.  The format is very similar to the left figure of 
          Fig~\ref{fig:smf1}.~~
          \textbf{Right}: The stellar mass volume-density distributions of massive 
          galaxies after combining the \rbcg{} and \nbcg{} samples together, using 
          both \mtot{} (black solid line) and \mmax{} (black dash line). 
          The uncertainty for distribution using \mtot{} is shown in grey shaded region.  
          Here, we also compare the distributions with SMF from previous works: 
          (a): SDSS galaxies at $z{\sim} 0.1$ from \citet{Bernardi2013}; the \mstar{} 
          is based on photometry from 2-D \ser{}$+$Exponential model fitting; 
          (b): SDSS galaxies at $z{\sim} 0.1$ from \citet{Moustakas13} based on 
          improved SDSS \texttt{cModel} photometry; 
          (c): \texttt{S82-MGC} galaxies at $0.15 < z< 0.43$ from 
          \citet{Leauthaud2016} based on PSF-matched SDSS-UKIDSS photometry.
          Please see text for more details of the comparison.
          }
      \label{fig:discussion_2}
  \end{figure*}
%% ------------------------------------------------------------------------------------ %%

%% ------------------------------------------------------------------------------------ %%

\subsection{Scatter in Outer Stellar Enveloppes and ``cD'' Galaxies}

   The thin lines in Figure \ref{fig:prof_1} show the \mden{} density profiles of 
   individual galaxies. 
   In general, we find that any given galaxy has a relatively smooth profile, even out 
   to large radii, however, on a galax-by-galaxy basis there is a large (but continuous) 
   variation in the amplitude of \mden{} at large scales. 
   Traditionally, the so-called ``cD'' galaxies are defined to have large stellar 
   envelopes and are found at the hearts of massive clusters (\citealt{Matthews1964, 
   Schombert1988}). 
   However, until present, it has not been clear whether ``cD'' galaxies are a distinct 
   and special population, or if they are simply the continum of a smoothly varying 
   population. 
   Fig \ref{fig:prof_1} shows that with the help of our deep HSC images, we can directly 
   map out the variations in \mden{} profiles at large scales. 
   Instead of a bi-model distribution, we find that the outskirts of massive galaxies 
   display a smooth but continuous variation. 
   We can hence conclude that ``cD'' galaxies are simply the extremes of a continuous 
   population but that do not necessarily form a unique a special class of galaxies. 
   This suggests that it might not be very physically meaningful to separate these 
   massive galaxies into two arbitrary morphological classes (``cD'' versus non-``cD''). 
   This conclusion is similar to those of \citep{Zhao2015} who found that a large 
   fraction of nearby BCGs (34\%) does not fall into the ``cD'' class. 
   Our finding that massive galaxies display a diverse range of outer stellar 
   envelope is also supported by HST observations of BCGs at $0.3 < z <0.9$ 
   by \citealt{Bai2014}.

%% ------------------------------------------------------------------------------------ %%

%% ------------------------------------------------------------------------------------ %% 
\subsection{What is the ``Total Mass'' of Massive Galaxies?}

    Even with deep HSC images, it is non trivial to clearly define and measure the
    ``total'' \mstar{} for these massive systems. 
    As discussed in Section~\ref{ssec:mtotal}, our 1-D approach has certain advantages 
    over the 2-D modeling method and in principle, should capture a the majority of 
    the stellar mass. 
    However, even though we extend to 100 kpc, there is no evidence that the profile 
    drops to zero on these scales. 
    We now estimate how much extra mass is ``hidden'' beyond the 100 kpc radius. 
    For this, we integrate the \mden{} profiles out to the detection limit in 
    surface brightness profile that is defined by the background noise.
    We will call this the ``maximum mass'' of the galaxy (\mmax{}). 
    For obvious reasons, we find that \mmax{} is larger than \mtot{} with an average 
    difference of about ${\sim}0.04$ dex. 
    The left hand side of Fig \ref{fig:discussion_2} shows differences between \mtot{}
    and \mmax{}. 
    On a galaxy-by-galaxy basis, the difference \mmax{}-\mtot{} depends on the slope 
    of the stellar envelope. 
    In this work, we have opted to in general use \mtot{} rather than \mmax{} as the 
    uncertainty on \mmax{} is larger. 
    However, we have checked that none of our results are changed by replacing \mtot{} 
    via \mmax{}.

    How to define and measure the ``total'' mass of a galaxy becomes a critical 
    question when studying the high-mass end of the galaxy SMF. 
    The issues discussed above are currently one of the primary limitations in 
    determinations of the high-mass end of the SMF (\citealt{Bernardi2016}). 
    Fig~\ref{fig:smf1} illustrated how improved \mstar{} measurement can 
    significantly modifies the SMF. 
    On the right hand side of Fig~\ref{fig:discussion_2}, we further highlight that 
    the shape of SMF at high-mass end is quite sensitive to even very small 
    change of \mstar{}. 
    We also compare the SMFs of our HSC samples with the following previous studies:

    \begin{itemize}

        \item The SMF derived from a SDSS-\textit{GALEX} sample of $z{\sim} 0.1$ 
            galaxies from \citet{Moustakas13}. 
            Here, the total luminosity is based on the SDSS \cmodel{} magnitude. 
            The \m2l{} is derived via SED fitting using \texttt{iSEDfit} with similar 
            assumptions as in our work.

        \item The SMF for $z{\sim} 0.1$ SDSS galaxies from \citet{Bernardi2013}. 
            The total luminosity is based on 2-D \texttt{SerExp} models 
            (\ser{}$+$ Exponential disk model; integrated to infinity) that 
            recovers much more flux compared to SDSS \cmodel{} magnitudes. 
            The \m2l{} is also from SED fitting assuming a Chabrier IMF.

        \item The SMF for $0.15 < z < 0.30$ galaxies from the \texttt{S82-MGC} sample
            \citep{Leauthaud2016} where the \mstar{} is derived using 
            \texttt{iSEDfit} on PSF-corrected aperture photometry from 
            \texttt{S82}.

    \end{itemize}

%% ------------------------------------------------------------------------------------ %% 
    It is clear that different analyses lead to very different slopes at high-mass end. 
    Here, we do not try to evaluate whether the differences displayed in 
    Fig~\ref{fig:discussion_2} are dominated by differences in photometry versus 
    assumptions about stellar population models (e.g. \citealt{Bernardi2016}). 
    However, given that the results that we are comparing assume the same IMF and adopt 
    similar assumptions about stellar population models, we hypothesize that the 
    differences in Fig~\ref{fig:discussion_2} are dominated by differences in photometry. 
    The goal of our of present paper is not to present and discuss a careful evaluation
    of the high-mass end of the SMF -- this will be presented in a forthcoming paper 
    once the HSC survey has accumulated more data. 
    However, this is clearly a topic that warrants further investigation, and careful 
    photometry measurements from HSC data will shed further light on these questions.

    Finally, we emphasize that in this present study we do not attempt to separate the 
    light of the galaxy with a potential (and distinct) ``intra-cluster''
    light component (ICL; e.g. \citealt{Carlberg1997, Lin2004, Gonzalez2005, 
    Mihos2005}). 
    Instead, we adopt the view that the light of the main galaxy and the 
    ``intra-cluster''component are a smooth and continuous distribution. 
    A reliable isolation of a physically meaningful ICL component using photometry 
    alone is extremely difficult, if not impossible. 
    At the imaging depth of HSC, we find no evidence for a clearly distinction between 
    an ``inner component'' and an ``intra-cluster'' component. 
    Instead, most of our profiles are smoothly varying out to 100 kpc with little 
    evidence for a distinct secondary component. 
    We do note however, that our \rbcg{} sample contains only a small fraction of 
    extremely massive clusters where strong evidence of photometric ICL component is 
    often found (\mhalo{}$\simgt 10^{15}M_{\star}$). 
    At the typical \mhalo{} of our \rbcg{} clusters, one good nearby example would be 
    the Virgo cluster, whose ultra-deep images (e.g. \citealt{Mihos2016}) show
    that the stellar halo of the BCG can smoothly extends beyond 100 kpc, while the 
    ICL only becomes important outside that radius and is dominated by individual 
    tidal features.

%% ------------------------------------------------------------------------------------ %%

%% ------------------------------------------------------------------------------------ %% 
\section{Summary and Conclusions}
    \label{sec:summary}

    In this work, we study how environment (halo mass) affects the stellar mass density 
    profiles of massive central galaxies using deep images from the Subaru HSC 
    survey. 
    With the help of this high-quality and wide area data set, we directly map the 
    stellar mass distributions of ${\sim}7000$ massive central galaxies at 
    $0.3 < z < 0.5$ out to $>100$ kpc without resorting to stacking techniques. 
    We group massive central galaxies into two categories based on their host halo 
    mass (\mhalo{}$\simgt 10^{14.2} M_{\odot}$ and \mhalo{}$\simlt 10^{14} M_{\odot}$). 
    Our main results are:

    \begin{enumerate}
        \item We find that the ``total'' \mstar{} of these massive galaxies can be 
            significantly underestimated with shallow imaging data such as SDSS and/or 
            imperfect model assumption (e.g. the \texttt{cModel} or single-\ser{}) 
            are used. 
            In contrast to previous work, our results do not depend on stacking or any 
            parametric models. 
            Moreover, the level of such underestimation could also depend on the 
            stellar mass and on halo mass. 
            This should be carefully taken into account when discussing topics such 
            as the evolution of the galaxy SMF.

        \item We show that the \mden{} profiles of massive galaxies are generally quite 
            smooth out to 10 kpc but that there is a large but continuous scatter in the 
            amount of light in their outskirts. 
            We argue that ``cD'' galaxies simply represent the tail end of a 
            distribution but that they are not a distinct class of massive galaxies.

        \item At fixed \mstar{} evaluate within 100 kpc, we find that massive central 
            galaxies display \mden{} profiles that are not self-similar and which 
            present subtle but systematic differences that correlate with \mhalo{}.
            On average, massive central galaxies in more massive halos have slightly 
            shallower inner \mden{} profiles and more prominent outer envelope.

        \item We also study how the \mden{} profiles vary with \mhalo{} at fixed 
            ``in-situ'' mass where it is approximated by the \mstar{} within 10 kpc. 
            At fixed \minn{} we find that central galaxies in more massive halos 
            possess much more extended stellar envelope in the outskirts than the ones 
            in smaller halos. 
            At the same time the ellipticity, and optical color profiles inside 10 kpc 
            show little to no dependence on environment.

        \item The \mstar{}-size relation also exhibits a clear environmental 
            dependence and central galaxies in more massive halos have larger 
            $R_{\mathrm{50}}$ at fixed \mtot{}. 
            We suggest that the relation between \mtot{} and \minn{}\footnote{or any 
            better proxy of total \mstar{} and the mass formed in the ``in-situ'' phase.} 
            can be a useful tool in diagnosing the assembly history of these massive 
            galaxies and the connection with the assembly of their dark matter halos.

    \end{enumerate}

    These results highlight the advantages of wide area, deep, and high-quality imaging 
    for studying the evolution of massive galaxies. 
    Upon finishing this work, the HSC survey has already doubled its sky coverage to 
    ${\sim} 200$ deg$^2$, and provides a much larger sample of massive central galaxies. 
    In the near future, we will extend this work to lower \mtot{} by using photometric 
    redshifts, and we will also apply 2-D photometry method (e.g.\ \citealt{Huang2013a}) 
    to take advantages of the multi-wavelength nature of the HSC survey 
    (e.g. \citealt{Huang2016}). 
    Our current work can also be combined with weak lensing measurements of the dark 
    matter halos of massive galaxies and physical insights into the assembly histories 
    of these galaxies can be gained by comparing with cosmological hydro-simulations 
    such as Illustris (\citealt{Vogelsberger2014}, \citealt{Genel2014}), 
    EAGLE (\citealt{Schaye2015}, \citealt{Crain2015}), or \textit{Horizon-AGN} 
    (\citealt{Dubois2014}).

%% ------------------------------------------------------------------------------------ %%

%\acknowledgements
\section*{Acknowledgements}

  % Personal 
  SH thanks F.S.Liu for sharing the data from his work.

  % HSC part
  The Hyper Suprime-Cam (HSC) collaboration includes the astronomical communities of 
  Japan and Taiwan, and Princeton University.  The HSC instrumentation and software were
  developed by the National Astronomical Observatory of Japan (NAOJ), the Kavli Institute
  for the Physics and Mathematics of the Universe (Kavli IPMU), the University of Tokyo,
  the High Energy Accelerator Research Organization (KEK), the Academia Sinica Institute
  for Astronomy and Astrophysics in Taiwan (ASIAA), and Princeton University.  
  Funding was contributed by the FIRST program from Japanese Cabinet Office, the Ministry 
  of Education, Culture, Sports, Science and Technology (MEXT), the Japan Society for 
  the Promotion of Science (JSPS), Japan Science and Technology Agency (JST), the
  Toray Science Foundation, NAOJ, Kavli IPMU, KEK, ASIAA, and Princeton University.

  % SDSS part
  Funding for SDSS-III has been provided by the Alfred P. Sloan Foundation, the
  Participating Institutions, the National Science Foundation, and the U.S.  Department of
  Energy. The SDSS-III web site is http://www.sdss3.org.  SDSS-III is managed by the
  Astrophysical Research Consortium for the Participating Institutions of the SDSS-III
  Collaboration including the University of Arizona, the Brazilian Participation Group,
  Brookhaven National Laboratory, University of Cambridge, University of Florida, the
  French Participation Group, the German Participation Group, the Instituto de Astrofisica
  de Canarias, the Michigan State/Notre Dame/JINA Participation Group, Johns Hopkins
  University, Lawrence Berkeley National Laboratory, Max Planck Institute for
  Astrophysics, New Mexico State University, New York University, Ohio State University,
  Pennsylvania State University, University of Portsmouth, Princeton University, the
  Spanish Participation Group, University of Tokyo, University of Utah, Vanderbilt
  University, University of Virginia, University of Washington, and Yale University.

  % Pan-STARRS1 part
  The Pan-STARRS1 Surveys (PS1) have been made possible through contributions of the 
  Institute for Astronomy, the University of Hawaii, the Pan-STARRS Project Office, 
  the Max-Planck Society and its participating institutes, the Max Planck Institute 
  for Astronomy, Heidelberg and the Max Planck Institute for Extraterrestrial Physics, 
  Garching, The Johns Hopkins University, Durham University, the University of Edinburgh, 
  Queen's University Belfast, the Harvard-Smithsonian Center for Astrophysics, the Las 
  Cumbres Observatory Global Telescope Network Incorporated, the National Central 
  University of Taiwan, the Space Telescope Science Institute, the National Aeronautics 
  and Space Administration under Grant No. NNX08AR22G issued through the Planetary 
  Science Division of the NASA Science Mission Directorate, the National Science 
  Foundation under Grant No. AST-1238877, the University of Maryland, and Eotvos 
  Lorand University (ELTE).

  % LSST software
  This paper makes use of software developed for the Large Synoptic Survey 
  Telescope. We thank the LSST Project for making their code available as free 
  software at http://dm.lsstcorp.org.

  % Software
  This research made use of:
  \href{http://www.stsci.edu/institute/software_hardware/pyraf/stsci\_python}{\texttt{STSCI\_PYTHON}},
      a general astronomical data analysis infrastructure in Python. 
      \texttt{STSCI\_PYTHON} is a product of the Space Telescope Science Institute, 
      which is operated by AURA for NASA;
  \href{http://www.scipy.org/}{\texttt{SciPy}},
      an open source scientific tools for Python (\citealt{SciPy});
  \href{http://www.numpy.org/}{\texttt{NumPy}}, 
      a fundamental package for scientific computing with Python (\citealt{NumPy});
  \href{http://matplotlib.org/}{\texttt{Matplotlib}}, 
      a 2-D plotting library for Python (\citealt{Matplotlib});
  \href{http://www.astropy.org/}{\texttt{Astropy}}, a community-developed 
      core Python package for Astronomy (\citealt{AstroPy}); 
  \href{http://scikit-learn.org/stable/index.html}{\texttt{scikit-learn}},
      a machine-learning library in Python (\citealt{scikit-learn}); 
  \href{http://www.astroml.org/}{\texttt{astroML}}, 
      a machine learning library for astrophysics (\citealt{astroML});
  \href{https://ipython.org}{\texttt{IPython}}, 
      an interactive computing system for Python (\citealt{IPython});
  \href{https://github.com/kbarbary/sep}{\texttt{sep}} 
      Source Extraction and Photometry in Python (\citealt{PythonSEP});
  \href{https://jiffyclub.github.io/palettable/}{\texttt{palettable}},
      color palettes for Python;
  \href{http://dan.iel.fm/emcee/current/}{\texttt{emcee}}, 
      Seriously Kick-Ass MCMC in Python;
  \href{http://bdiemer.bitbucket.org/}{\texttt{Colossus}}, 
      COsmology, haLO and large-Scale StrUcture toolS (\citealt{Colossus}).

%%%%%%%%%%: Bibliographic Section %%%%%%%%%%

\bibliographystyle{mnras}
\bibliography{redbcg}

%%%%%%%%%%: Appendix Section %%%%%%%%%%%%

\appendix

%% ------------------------------------------------------------------------------------ %%

\section{A. Basic Statistical Properties of the Sample} 
	\label{app:sedasic}

    On the top-left panel of Fig~\ref{fig:sample_stats}, we show the \mtot{}-color 
    relations using the $k$-corrected $g-r$ color. 
    Both samples follow the same tight ``red-sequence'' with little contamination 
    from the ``blue cloud''.
    At fixed \mtot{}, we see little offset in color distributions of the two 
    samples, suggesting that both samples consist of quiescent galaxies with 
    similar average stellar population properties.  
    This is consistent with previous result that suggests the average stellar 
    population of massive central galaxy does not depend on \mhalo{} 
    (e.g.\ \citealt{Park2007}).  
    In this work, we focus on the \mstar{} range of $11.6 \le$\logmtot{}$\le 11.9$, 
    where both samples have acceptable completeness, and their \mtot{} distributions 
    greatly overlap (see the normalized distributions of \mtot{} in the bottom-left 
    panel of Fig~\ref{fig:sample_stats}). 
    As for the \minn{} distributions, the two samples overlap the most within 
    $11.2 \le$\logmtot{}$\le 11.6$, but now they show quite different
    distributions (bottom-right figure).

    The redshift distributions also show small difference
    (upper-right panel) even in the high-\mtot{} bin, where the redshift distribution 
    of the \nbcg{} sample skews toward higher-$z$ end due to the contribution of BOSS 
    spec-$z$.
    Since this could bias the comparison of \mden{} profiles and other properties 
    (please see Appendix\ref{app:redshift} for more details), we address this via matching 
    the two samples in both mass and redshift distributions carefully
    (see Appendix~\ref{app:match}).

%% ------------------------------------------------------------------------------------ %% 
  \begin{figure*}
      \centering 
      \includegraphics[width=16.5cm]{fig/redbcg_sample_stats}
      \caption{
          \textbf{Top-left}: The \logms{}-$g-r$ color relation of the \rbcg{} 
          (red circle; detaield format is the same with the right panel of 
          Fig~\ref{fig:smf1}) and \nbcg{} (grey dots) samples.
          We apply the $k$-corrections from \texttt{iSEDFit} fitting to the colors.~~          
          \textbf{Top-right}: the histograms of the redshift for the \rbcg{} and 
          \nbcg{} galaxies in both \logmtot$>11.2$ and $11.6<$\logmtot{}$<11.9$
          mass bins.
          The vertical lines highlights the $0.3\leq z \leq 0.5$ redshift range.~~
          \textbf{Bottom-left}: the histograms of \mtot{} for the \rbcg{} (orange-red) 
          and \nbcg{} (grey-black) samples at both $z>0.2$ (step-filled histogram) and 
          $0.3 \leq z \leq 0.5$ (stepped histogram). 
          The vertical lines in both top-left and bottom-left figures highlight the 
          $11.6<$\logmtot{}$<11.9$ mass range that will be used in the comparison of 
          the \mtot{}-matched samples.~~
          \textbf{Bottom-right}: the histograms of \minn{} in similar format. 
          Here the vertical lines highlight the 
          $11.2<$\logminn{}$<11.6$ mass range that is used for comparison.
      }
      \label{fig:sample_stats}
  \end{figure*}
%% ------------------------------------------------------------------------------------ %%

%% ------------------------------------------------------------------------------------ %% 
\section{B. Extraction of 1-D surface brightness profile} 
    \label{app:ellipse}

    Here we briefly discuss a few technical issues related to the measurements of the 
    1-D surface brightness profiles around massive galaxies.

    To derive reliable 1-D profile, it is important to mask out all the irrelevant 
    objects around the target.
    At the depth of the HSC images, this becomes a quite difficult task, especially 
    for massive galaxies with extended outer profiles and more satellites. 
    Right now, the \texttt{hscPipe} tends to ``over-deblend'' around bright object, 
    we therefore perform \texttt{SExtractor}-like background subtraction and object 
    detection using the \texttt{SEP} Python library to generate the necessary masks.
    Combining two different local background models and $S/N$ thresholds, we obtain 
    the centroid, shape, and radius that enclose 90\% of flux for each object, even 
    the one that is close to the center of bright galaxy (left panel of 
    Fig~\ref{fig:ell_tech}), and generate its initial elliptical mask based on these
    information. 
    We first create the mask that covers all contaminating objects around the target 
    after adaptively increasing the sizes of their masks according to their brightness 
    and distance to the target center. 
    Generally speaking, we mask out the very bright objects or objects in the 
    outskirt of the target a little more aggressively as they could impact the 
    low surface brightness regions more significantly. 
    We also create masks that are less and more aggressive to test their impacts
    on the surface brightness profiles.

    Next, we aggressively mask out all objects on the cut-out image and use this mask
    to evaluate the background level using the distribution of pixel values after 
    median smoothing the masked image within $6x6$ pixels.
    This provides us a global estimate of background level along with its uncertainty. 
    Given the typical RMS of the background pixels, the HSC \texttt{WIDE} images 
    should reach to $> 29$ \sb surface brightness level in $i$-band.  
    However, the global background level is often negative around our massive 
    galaxies, indicating systematic issue in background subtraction.  
    In current \texttt{hscPipe}, the background on each CCD is modeled with a 
    Chebyshev-polynomial that is fit to the smoothed image after excluding pixels 
    with $S/N >5$.
    This algorithm performs much better than the SDSS version 
    (e.g.\ see \citealt{Blanton2011}), yet still over-subtracts sky around bright 
    objects and results in unphysical truncation in the surface brightness profile.
    We attempt to empirically correct this using the background model generated by 
    the \texttt{SExtractor} algorithm on the masked image 
    ($200x200$ pixels background box size, and 6 pixels median filtering size of 
    sky boxes).
    This model can account for the slightly over-subtracted background at large scale,
    and also reduces the impact from the low surface brightness ``wings'' of bright 
    nearby objects. 
    After the correction, we can see the improvement from both the remeasured global 
    background value (closer to 0) and the surface brightness profile (middle panel 
    of Fig~\ref{app:ellipse}; the negative intensity and the turn-over of the 
    curve-of-growth in the outskirt of the ``Original'' profile are successfully 
    corrected). 
    Such ad hoc correction is not the best way to achieve accurate background 
    subtraction, but the surface brightness profile within 100 kpc is not very
    sensitive to the over-subtraction of sky as while.

    The procedure to derive 1-D surface brightness profile from the 
    background-corrected, contamination-masked images is already described in 
    Section~\ref{sec:ellipse} briefly. 
    In practice, the profile at very low surface brightness level is sensitive to 
    several \texttt{Ellipse} configurations.
    After some tests, we choose to use 0.1 dex in logarithm as the step in semi-major 
    axis length between successive ellipses, and we use the median pixel value over the
    elliptical annulus after rejecting outlying pixels via $3\sigma$-clipping three
    times.
    We make the above choices to make the final profile less affected by any nearby
    object, and also test the differences between the profiles derived using larger
    step, or mean value on the annulus, or fewer times of $\sigma$-clipping. 
    Generally speaking, the surface brightness profile is very robust against these
    changes.  
    On the right panel of Fig~\ref{fig:ell_tech}, we compare the surface brightness
    profiles for an example massive galaxy using different masks and \texttt{Ellipse}
    parameters. 
    The profile within 100 kpc is very stable, and the only meaningful difference 
    we can spot is caused by the less aggressive object mask in the very outskirt.

    We should also mention that we run \texttt{Ellipse} allowing for more 
    sophisticated shapes than a simple ellipse (4th Fourier modes that can make 
    isophote more ``disky'' or ``boxy'', e.g.\ \citealt{Kormendy2009}) to fit the 
    isophote better.
    We also apply the isophotes from $i$-band images to other bands in 
    ``force-photometry'' mode \texttt{Ellipse} run to get rough estimates of 
    color profiles.  
%% ------------------------------------------------------------------------------------ %%

%% ------------------------------------------------------------------------------------ %% 
    \begin{figure*}
        \centering 
        \includegraphics[width=\textwidth]{fig/redbcg_ellipse_tech}
        \caption{
            \textbf{Left:} Example of the object-mask built for the \texttt{Ellipse}
            run for a typical massive galaxy in the sample. 
            All the shaded regions are masked out. 
            The three dash lines (red, inner one and two blue ones) around the target 
            at the center outlines the three radius we defined using the flux radius 
            of the target.  
            We increase the mask size for objects detected in different regions 
            separated by these apertures (which are outlined by solid, elliptical 
            apertures with different colors) using slightly different criteria.~~
            \textbf{Middle:}: The flux profile that is zoomed into the near-zero flux 
            range (top panel), and the curve-of-growth of the magnitude for the example
            massive galaxy.  
            To highlight the importance of correcting sky background, we show the profiles 
            using both images with (red, solid line) and without (black, dash line) 
            ad-hoc background correction. 
            On the top panel, besides the horizontal line that highlights the zero flux 
            level, we also show the uncertainty of the sky background estimate via 
            grey-shaded region.  
            On the bottom panel, two horizontal lines indicate the magnitudes 
            corresponding to total (solid) and half flux (dash) using the 
            background-corrected profile. ~~
            \textbf{Right:} compare the 1-D surface brightness profiles for the same 
            example galaxy extracted using different mask 
            (smaller masking region: red, dash line; larger masks: blue, dash line), 
            or different \texttt{Ellipse} settings
            (more aggressive pixel-clipping: cyan, dash line; 
             larger step in radius: green, dash line; 
             using mean flux along the isophote instead of median: purple, dash line)
            with the one using the default configuration (black, solid line).
            }
        \label{fig:ell_tech}
    \end{figure*}
%% ------------------------------------------------------------------------------------ %%

%% ------------------------------------------------------------------------------------ %% 
\section{C. Estimate average {$M_{\star}/L_{\star}$} using \texttt{iSEDFit}} 
    \label{app:sed}

%% ------------------------------------------------------------------------------------ %% 
    \begin{figure*}
        \begin{center}
        \includegraphics[width=\textwidth]{fig/redbcg_isedfit_example.pdf}
        \caption{
            Left: example output figure from \texttt{iSEDFit} that shows the SED 
        	fitting results. 
            The open-boxes show the observed fluxes in 5-band, and the solid, blue-dots
            show the best-fitted results, along with the high-resolution spectrum for
            this model reconstructed using the synthetic spectra from \texttt{FSPS}. 
            Top-left corner shows the best-fit stellar population parameters, and 
            bottom-right corner shows the ID, redshift of this object, and reduced 
            $\chi^2$ of the best-fit model.~~~
            Right: the Posterior distributions of a few key parameters.
            From top-left to bottom right are: 1) stellar mass (\logms{}); 2) age of the 
            population (mass and star-formation rate weighted) in Gyr; 3) star formation 
            rate ($\log\ \mathrm{SFR}\ (M_{\odot}/\mathrm{yr})$; instant one and the one 
            averaged over the previous 100 Myr; 4) stellar metallicity 
            ($\mathrm{Z}/\mathrm{Z}_{\odot}$); 5) dust extinction ($\mathrm{A}_V$ in mag);
            6) birthrate parameter ($\log\ b$; averaged over previous 100 and 1000 Myr).
            }
        \label{fig:ised}
        \end{center}
    \end{figure*}
%% ------------------------------------------------------------------------------------ %%

%% ------------------------------------------------------------------------------------ %% 
    \begin{figure*}
        \begin{center}
        \includegraphics[width=\textwidth]{fig/redbcg_isedfit_2.pdf}
        \caption{
            Relationships between \mstar{} and other stellar population 
            parameters using the results from \texttt{iSEDfit}. 
            The \textbf{left} figures are for the \rbcg{} sample, and the 
            \textbf{righ}t ones are 
            for the \nbcg{} sample. 
            They are both color-coded using the redshift of the object, while 
            their sizes reflect the magnitude differences in $i$-band between 
            the observed and best-fit value. 
            The four stellar population properties are: 
            1) \textbf{Top-left}: mass-weighted stellar age in Gyr; 
            2) \textbf{Bottom-left}: \m2l{} in $i$-band; 
            3) \textbf{Top-right}: mass-weighted stellar metallicity in unit of Solar 
            value;
            4) \textbf{bottom-right}: dust extinction values in magnitude.
            }
        \label{fig:ised_2}
        \end{center}
    \end{figure*}
%% ------------------------------------------------------------------------------------ %%

%% ------------------------------------------------------------------------------------ %% 
    \begin{figure*}
        \begin{center}
        \includegraphics[width=\textwidth]{fig/redbcg_isedfit_3.pdf}
        \caption{
            Comparisons of \mstar{} estimated by \texttt{iSEDFit} using different
            assumptions. 
            In each figure, we plot the \mstar{} from default model against their 
            differences with other models. 
            The top (bottom) row shows the comparisons of the \rbcg{} (\nbcg{}) sample. 
            The four models involved here are labeled as: 
            (A): default one; 
            (B): using the Salpeter IMF instead of the Chabrier one 
                (\textbf{Left column});
            (C): using the BC03 synthetic population model instead of the FSPS one
                (\textbf{Middle column});
            (D): assume that there is no dust extinction (\textbf{Right column}). 
            On each figure, the grey shaded region highlights the uncertainty 
            of the $\log($\logms{}$)$ using typical \mstar{} uncertainties from 
            both sides. 
            For each comparison, we highlight the median $\chi^{2}$ value, and the fraction
            of galaxies with $\chi^{2} > 10.0$ at the top. 
            We also show the histogram of the \mstar{}-differences on the right panel.
            }
        \label{fig:ised_3}
        \end{center}
    \end{figure*}
%% ------------------------------------------------------------------------------------ %%

%% ------------------------------------------------------------------------------------ %%

    In Section~\ref{ssec:isedfit}, we briefly explain the SED fitting procedure and 
    the priors used.   
    In Fig~\ref{fig:ised_1}, we show an example of the \texttt{iSEDFit} output by 
    visualizing the 5-band HSC SED on top of the best-fit model along with the PDF 
    of the key parameters. 
    Although we only use the best-fit \m2l{} in this work, it is necessary to make 
    sure the model is reasonable. 
    We show the relations between \mstar{} and a few key stellar population parameters 
    derived by \texttt{iSEDFit} in Fig~\ref{fig:ised_2}. 
    Although degeneracies among these parameters are almost inevitable using only 
    five broad-band photometry, most massive galaxies in both \rbcg{} (left) and 
    \nbcg{} (right) samples have old age, high metallicity ($1.5 \times Z_{\odot}$ is 
    the highest metallicity allowed by the adopted \texttt{FSPS} SSP models), and low 
    dust extinction as expected

    Meanwhile, \mstar{} measurement based on SED fitting heavily depends on the 
    adopted SSP model, the form of IMF, dust extinction law, and details in 
    the assumption of SFH (e.g.\ \citealt{Bernardi2016b}). 
    For massive galaxies in this sample, the form of the SFH\footnote{We choose 
    to use the delayed-$\tau$ model for SFH; we adopt flat distribution between 
    0.5 to 14.0 Gyrs as the prior for the look-back time when the star formation 
    turned on. 
    The exponential delayed time-scale ($\tau$) is allowed to change between 
    0.1 to 3.0 with equal probability}, the contribution from random star 
    burst\footnote{The chance of random star burst is set at 0.2 for every 2 Gyrs. 
    The duration of the star burst is draw from a logarithmic distribution 
    between 0.03 to 0.3 Gyr; and the mass fraction formed in the burst is from 
    a logarithmic distribution between 0.01 and 1.0.} do not affect the results. 
    On the other hand, the choices of SSP model, IMF, and dust extinction do impact
    the SED fitting results systematically, and we explore it using a few 
    additional tests (see Fig~\ref{fig:ised_3}):

    \begin{enumerate}

        \item Choosing the \citet{Salpeter1955} IMF results in systematically 
            higher \mstar{} (on average $+0.25$ dex of \logms{}) for both \rbcg{}
            and \nbcg{} galaxies (Left panels).
            Although there are multiple lines of evidence that favor Salpeter 
            or even more ``bottom-heavy'' IMF in massive galaxies 
            (e.g.\ \citealt{Conroy2012}; \citealt{Cappellari2012}), we still 
            present the main results using Chabrier IMF to accommodate galaxies 
            with lower \mstar{} in the sample, and to be as consistent as possible 
            with a few previous works.  
            This choice of IMF does not change the main results qualitatively.

        \item \mstar{} based on the \texttt{BC03} models are systematically lower 
            than the ones based on \texttt{FSPS+MILES} models (middle panels)
            The difference shows a large scatter, and can be as large as 0.4 dex,
            although it does show any \mstar{} dependency. 
            We notice that the \texttt{BC03} provides better overall ${\chi}^2$. 
            This could relates to the higher upper limit of stellar metallicity 
            ($2 \times \mathrm{Z}_{\odot}$) allowed by the \texttt{BC03} model 
            that help fit the shape of the SED in the red-end better.  
            However, the \texttt{BC03} model could underestimate the 
            stellar age for a fraction of massive galaxies ($< 3$--4 Gyrs), and 
            leads to underestimate of  \m2l{} for them too.
            It is worth noting that, both \texttt{FSPS} and \texttt{BC03} 
            models still have difficulties recovering SED at the red-end (between $z$ 
            and Y-band), and reproducing the optical color-color relations for 
            red-sequence galaxies (e.g.\ \citealt{MIUSCAT2}).
            In this work, we decide to keep using the \texttt{FSPS+MILES} model as 
            the fiducial one.  
            However, using the results based on \texttt{BC03} model will not 
            only change the main conclusions here, but also results in more 
            significant differences in the median \mden{} profiles.

        \item On the right panels of Fig~\ref{fig:ised_3}, we also compare with the
            SED fitting results without considering the dust extinction. 
            This choice leads to slightly smaller \mstar{} values as expected. 
            The impact is very small for the \rbcg{} galaxies, and becomes larger 
            at the low-\mstar{} end of the \nbcg{} sample. 
            But it still will not change any of the important results in this work.

    \end{enumerate}
%% ------------------------------------------------------------------------------------ %%

%% ------------------------------------------------------------------------------------ %% 
\section{D. Comparison of \mden{} profiles using \mstar{} from the GAMA survey}
    \label{app:gama}

%% ------------------------------------------------------------------------------------ %% 
\begin{figure*}
    \centering
    \includegraphics[width=\textwidth]{fig/redbcg_prof_gama_new}
    \caption{
        \textbf{Left:} comparison of \mstar{} estimated by the GAMA survey and 
        the \mtot{} using HSC images in this work. 
        We plot the \logmgama{} against the difference between \logmtot{} and \logmgama{}. 
        The format is very similar to the left panel of Fig~\ref{fig:smf1}. 
        The two vertical lines highlight the mass range $11.4 \leq$\logmgama{}$<11.8$ 
        that is used for the comparison.~~
        \textbf{Right:} we compare the \mden{} profiles of \rbcg{} (orange-red) and 
        \nbcg{} (grey-black) galaxies using the samples matched on the 
        \mgama{}-$z$ plane at $11.4 \leq$\logmgama{}$<11.9$ and $0.28 \leq z < 0.4$. 
        The format is very similar to the ones in Fig~\ref{fig:prof_1}.}
    \label{fig:gama}
\end{figure*}
%% ------------------------------------------------------------------------------------ %%

    The GAMA survey greatly overlaps with the HSC survey, and it provides carefully 
    measured \mstar{} for large sample of galaxies (\citealt{Taylor2011}) that help 
    produce many interesting results (e.g. \citealt{Bauer2013, Ferreras2017}).
    They use 2-D single-\ser{} model to correct the total luminosity of the galaxy 
    (\citealt{Kelvin2012}), and derive the \m2l{} through optical-SED fitting 
    (BC03 model; Chabrier IMF) based on the PSF-matched aperture photometry. 
    Since the \ser{} model is generally more flexible than the \texttt{cModel} one, 
    it is therefore interesting to compare with the \rbcg{} and \nbcg{} galaxies 
    that also have spec-$z$ (at $z < 0.40$) and \mstar{} in GAMA DR2 
    (\citealt{Liske2015}) and see the impact of deep photometry again.

    We summarize the results in Fig~\ref{fig:gama}.  
    On the left panel, we compare the differences between \mtot{} and \mgama{}. 
    HSC survey on average recovers more \mstar{} at high-\mstar{} end, which is 
    consistent with the expectation from deeper photometry, although the 
    systematic differences in the estimates of \m2l{} could play a role here. 
    Meanwhile, it is interesting see that, above \logmtot{}$> 11.8$, \mgama{} 
    becomes increasingly larger than \mtot{}, and most of these massive 
    galaxies have very high \ser{} index from the 2-D fitting. 
    This suggests that the single-\ser{} model is no longer an appropriate one to 
    describe very massive galaxies as it tends to over-estimate the \mstar{} the 
    inner and/or outer regions.

    To verify the cause of the difference in \mstar{}, we further select samples 
    of \rbcg{} and \nbcg{} galaxies with matched \mgama{} and redshift 
    distributions (at $11.4 <$\logmgama{}$<11.8$; see Appendix \ref{app:match}), 
    and compare their \mden{} profiles (right panel). 
    Although these two subsamples are equally massive according to results from 
    GAMA survey, it is clear that the \rbcg{} galaxy has much more extended 
    outer envelope, even though its median \mden{} profile is very similar 
    to the \nbcg{} sample at $< 10$ kpc. 
    We can reproduce very the same trend with the luminosity density profiles 
    (with or without $k$-correction), suggesting that the inaccurate \ser{} 
    model definitely leads to under-estimate of \mstar{}.

%% ------------------------------------------------------------------------------------ %%

\section{E. Comparisons of \mden{} profiles in different redshift bins}
    \label{app:redshift}

%% ------------------------------------------------------------------------------------ %% 
\begin{figure}
    \centering 
    \includegraphics[width=8.2cm]{fig/redbcg_avg_prof_z}
    \caption{
        Comparison of \mden{} profiles of \rbcg{} (orange-red) and \nbcg{} 
        (grey-black) at $11.6 \le$\logmtot$< 11.9$ in redshift bins of 
        $0.3\leq z<0.4$ (solid lines) and $0.4\leq z<0.5$ (dash lines). 
        We show the individual profile in the background using much thinner line, 
        and highlight the median profiles using thicker line and darker color.
        Other formats are exactly the same with the left figure of 
        Fig~\ref{fig:avg_prof}.}
    \label{fig:avg_prof_z}
\end{figure}    
%% ------------------------------------------------------------------------------------ %%

    Given the redshift range for our samples, it is important to evaluate 
    the impacts from the physical extend of seeing and the imaging depth on the \mden{} 
    profiles at different redshift. 
    Under the same seeing, the \mden{} profile of galaxy at higher-$z$ is more 
    vulnerable to the PSF smearing effect at the center. 
    It is also harder to reach to the same \mden{} level under the same imaging depth 
    due to cosmological dimming and background noise.

    In Fig~\ref{fig:avg_prof_z}, we group the \rbcg{} and \nbcg{} galaxies within 
    $11.6 \le$\logmtot$< 11.9$ into two $z$ bins ($0.3\leq z<0.4$ and $0.4\leq z<0.5$),
    and compare their \mden{} profiles. 
    In two redshift bins, the median \mden{} profiles from the same sample follow each 
    other very well outside 10 kpc, but become visibly different in the central 3-4 kpc,
    where the effect from seeing kicks in. 
    Meanwhile, the median \mden{} profiles of \rbcg{} and \nbcg{} in the same $z$ bin 
    are identical in the central region, which indicates similar average seeing 
    conditions.       
    This confirms that \mden{} profile at $> 5$ kpc is safe from the impacts of seeing 
    and difference in redshift.
    More importantly, it also suggests that, once the redshift distributions are 
    carefully matched, the difference of \mden{} profile is likely to be physical 
    even in the central region.

%% ------------------------------------------------------------------------------------ %%

\section{F. Match the \rbcg{} and \nbcg{} samples in \mstar{} and redshift distributions}
    \label{app:match}

%% ------------------------------------------------------------------------------------ %% 
\begin{figure*}
    \centering 
    \includegraphics[width=\textwidth]{fig/redbcg_match}
    \caption{
        \textbf{Left figure} shows the details of the \mtot{}-matching process, 
        corresponding to the results shown in the left figure of   
        On the top panel, we show the overall distributions of \rbcg{} (light orange boxes) 
        and \nbcg{} (light grey dots) galaxies on the \mtot{}-$z$ plane.  
        And, we match the two sample in the \mtot{}-$z$ space outlined by the shaded region.
        We highlight the \rbcg{} galaxies in this region using bigger boxes in red frames, 
        whose size reflects the $P_{\mathrm{Cen}}$ value.  
        We also color-code them using the richness ($\lambda$) of the host cluster. 
        The matched \nbcg{} galaxies are highlighted using darker color and bigger dots. 
        To further evaluate the matching results, we show the distributions of \mtot{} 
        (middle panel) and redshift (bottom panel) separately. 
        On both panesl, we show the histograms along with their kernel density 
        distributions.  
        And, on the top of each panel, two sets of short vertical lines highlight the median 
        value (solid) and the inter-quartile (dash) of each distribution.~~~
        \textbf{Right figure} shows the similar matching results for the \minn{}-matched
        samples used for the right figure of Fig~\ref{fig:prof_1}.
        The format is exactly the same as the left one, except the \minn{} replaces the 
        \mtot{} in the top and middle panels.}
    \label{fig:match}
\end{figure*}
%% ------------------------------------------------------------------------------------ %%

    As explained earlier, it is important to make sure the two samples have similar 
    distributions in both \mstar{} and redshift before comparing their median \mden{} 
    profiles.  
    Here we briefly describe the procedure used in this work. 
    Since the \rbcg{} sample is smaller in size, we always match the \nbcg{} sample to 
    it by searching for the $N$-nearest neighbours on the $M_{\star}$-redshift plane 
    using the KDTree algorithm in the \texttt{scikit-learn} Python library 
    (\citealt{scikit-learn}), and evaluate the quality of the match using the 
    distributions of both parameters (as shown in Fig~\ref{fig:match}).

    As we only keep the unique \nbcg{} galaxies in the matched sample, we manually 
    adjust the value of $N$ to achieve the best match. 
    When the redshift distribution of the \rbcg{} sample becomes bi-model, we also try 
    to split the sample into two redshift bins and match them separately. 
    Typically $N$ is between 3 to 8.
    In Fig~\ref{fig:match}, we demonstrate this procedure using the results for 
    the \mtot{}-matched (Left) and the \minn{}-matched samples in Fig~\ref{fig:prof_1}
    (Right), and the two samples are well matched in the distributions of \mtot{}
    (or \minn{}) and redshift.  
    For all the comparisons of \mden{} profiles in this work, we match the samples 
    in the same way, and make sure the match has the same quality.

%% ------------------------------------------------------------------------------------ %% 
\section{G. Robustness of the \mden{} Differences} 
	\label{app:robust}

    In Fig~\ref{fig:prof_1}, we compare the \mden{} profiles of \mtot{}- and 
    \minn{}-matched samples of \rbcg{} and \nbcg{} galaxies, and here we test the 
    robustness of the results using a few extra tests that are illustrated in
    Fig~\ref{fig:prof_2}, Fig~\ref{fig:prof_3}, and Fig~\ref{fig:prof_4}, and 
    are briefly described here:

    \begin{enumerate}

        \item In Fig~\ref{fig:prof_2}, we group the samples into two \mtot{} bins. 
            Given the small sample size, we extend slightly toward lower \mtot{} range 
            ($11.5 \leq \log (M_{\star,\ 10\mathrm{kpc}}/M_{\odot}) < 11.7$ and 
             $11.7 \leq \log (M_{\star,\ 10\mathrm{kpc}}/M_{\odot}) < 11.9$). 
            Although the smaller sample leads to larger statistical uncertainties, 
            we can still see similar structural differences in both \mtot{} bins, 
            and the difference becomes more significant in the higher \mtot{} bin.  
            For the lower \mtot{} bin, the difference in the inner region becomes 
            quite uncertain, while the difference in the outskirt is still visible. 
            This potentially suggests that the environmental dependence of structure 
            also varies with \mstar{}, an important implication deserves more 
            investigations in the future.

        \item On the left panel of Fig~\ref{fig:prof_3}, we match the \rbcg{} and 
            \nbcg{} samples in a lower redshift bins ($0.30 < z < 0.42$).
            Despite the larger uncertainties due to smaller samples, we find the 
            results are the same.

        \item On the middle panel of Fig~\ref{fig:prof_3}, we includes \rbcg{} 
            galaxies in poorer clusters ($20 < \lambda < 30$), which should result 
            in overlapped \mhalo{} distributions with the \nbcg{} samples 
            considering the typical uncertainty of $\lambda$.
            This makes the difference in the inner region slightly less significant, 
            but the overall results are the same.

        \item On the right panel of Fig~\ref{fig:prof_3}, in stead of using \mtot{}, 
            we use the \mmax{}--the maximum \mstar{} by integrating the \mden{} 
            profiles to the largest radius allowed.  
            The \mmax{} values are less reliable than \mtot{} due to the 
            uncertainty of background subtraction and contamination from nearby 
            bright objects, but they can serve as different estimates of the ``total''
            \mstar{} of these galaxies.
            As shown in Section~\ref{sec:discussion}, they on average increase
            the \mstar{} by a little bit and affect the \rbcg{} more.
            The differences in the \mden{} profiles still remain very similar.

    \end{enumerate}

    We also test the robustness of the \mtot{}-matched results using the samples with 
    only spectroscopic redshift, the samples in the three GAMA fields, and the \rbcg{} 
    samples without the ones in very massive halos ($\lambda > 40$).  
    Limited by space, we do not show these results here, but they all verify the 
    robustness of the results.

    For the results from the \minn{}-matched samples:

    \begin{enumerate}

        \item
            We match the two samples using both \minn{} and the \mstar{} within 15 kpc 
            at the same time.  
            This makes the two median \mden{} profiles very similar inside 10-15 
            kpc, while the result in the outskirt remains the same (left panel of 
            Fig~\ref{fig:prof_4}).
            Use \mstar{} within 5 or 20 kpc leads to the same conclusion.

        \item 
            To make sure the two samples are comparable in their overall assembly history,
            we also try to only include the very massive galaxies (\logmtot{}$>11.5$)
            in both samples. 
            This excludes the \nbcg{} galaxies that are much less massive and 
            more ``compact'' in structure. 
            Yet, the results regarding the structural differences remain the same.

    \end{enumerate}

%% ------------------------------------------------------------------------------------ %% 
  \begin{figure*}
      \centering 
      \includegraphics[width=\textwidth]{fig/redbcg_prof_2}
      \caption{
          Comparisons of the \mden{} profiles for \mtot{}-matched \rbcg{} 
          (orange-red) and \nbcg{} (grey-black) galaxies in lower (left; [11.5,11.70]) 
          and higher (right; [11.7, 11.9]) \mtot{} bins. 
          Other formats are in consistent with the right figure of Fig~\ref{fig:prof_1}.
          The difference in median profiles is more significant in higher \mtot{} bin.
          }
      \label{fig:prof_2}
  \end{figure*}
%% ------------------------------------------------------------------------------------ %%

%% ------------------------------------------------------------------------------------ %% 
  \begin{figure*}
      \centering 
      \includegraphics[width=\textwidth]{fig/redbcg_prof_3}
      \caption{
        Comparisons of the \mden{} profiles for \rbcg{} (orange-red) and \nbcg{} 
      	(grey-black) galaxies that are matched using proxies of total \mstar{}. 
        The formats are in consistent with the right figure of Fig~\ref{fig:prof_1}.
        The differences are, here, the samples are matched in slightly differnt ways. 
        From left to right: a) using samples at lower redshift ($0.3 < z < 0.4$); 
        b) using \rbcg{} sample with $\lambda > 20$ instead of 30; 
        c) using \mstar{} within 150 kpc instead of 100 kpc.
        The results are broadly consistent with the one in Fig~\ref{fig:prof_1}.
        }
      \label{fig:prof_3} 
  \end{figure*}
%% ------------------------------------------------------------------------------------ %%

%% ------------------------------------------------------------------------------------ %% 
  \begin{figure*}
      \centering 
      \includegraphics[width=\textwidth]{fig/redbcg_prof_4}
      \caption{
          Comparisons of the \mden{} profiles for \rbcg{} (orange-red) and \nbcg{} 
          (grey-black) galaxies that are matched using the \mstar{} enclosed in the 
          inner region. 
          Left panel shows the results after matching the \minn{} and \meff{} together, 
          and the right panel shows the results when only the \logmtot{}$\ge 11.5$
          \rbcg{} and \nbcg{} galaxies are included.
          Other formats are in consistent with the right figure of Fig~\ref{fig:prof_1}.
          }
      \label{fig:prof_4} 
  \end{figure*}
%% ------------------------------------------------------------------------------------ %%

%% ------------------------------------------------------------------------------------ %% 
%% Table.1 
%% ------------------------------------------------------------------------------------ %% 
\clearpage
\include{table1}
\clearpage
%% ------------------------------------------------------------------------------------ %%

%% ------------------------------------------------------------------------------------ %%  
\bsp
\label{lastpage}
\end{document}
%% ------------------------------------------------------------------------------------ %% 
%%%%%%%%%%%%: End of the File %%%%%%%%%%%%
%% ------------------------------------------------------------------------------------ %%