### 2017-06-13 draft/paper1_rachel.tex
Comments from Rachel

% Version 0; preprint format; Written by SH
% Version 1; emulateapj format; Edited by SH and AL
% Version 2; emulateapj format; Edited by SH and AL
% Version 3; emulateapj format; Edited AL
% Version 4; emulateapj format; Edited SH
% Version 5; MNRAS format; Edited SH
% Version 6; MNRAS format; Paper I edited by JG
% Version 7; MNRAS format; Edited by SH
% Version 10; MNRAS format; Edited by AL and JG
% Version 11; MNRAS format; Edited by SH

\documentclass[a4paper,fleqn,usenatbib]{mnras}

% Packages
\usepackage{deluxetable}
\usepackage{newtxtext,newtxmath}
\usepackage[T1]{fontenc}
\usepackage{ae,aecompl}
\usepackage{amssymb, amsmath}
\usepackage{graphicx}
\usepackage{natbib}
\usepackage{url}
\usepackage{hyperref}
\usepackage{float}
\usepackage[usenames, dvipsnames]{color}

% Package Settings
\hypersetup{colorlinks=true,
            citecolor=MidnightBlue,
            linkcolor=MidnightBlue,
            filecolor=magenta,      
            urlcolor=cyan}
\urlstyle{same}

% Figure extention
\DeclareGraphicsExtensions{.pdf,.png,.jpg}

%%%%%%%%%%%%: User Defined Commands %%%%%%%%%%%%

% Song Huang's definition 
\def\arcsec{{\prime\prime}}
\def\arcmin{{\prime}}
\def\degree{{\circ}}
\def\h{\hskip -3 mm}
\def\aa{{A\&A}}
\def\aas{{ A\&AS}}
\def\aj{{AJ}}
\def\al{$\alpha$}
\def\bet{$\beta$}
\def\amin{$^\prime$}
\def\annrev{{ARA\&A}}
\def\apj{{ApJ}}
\def\apjs{{ApJS}}
\def\asec{$^{\prime\prime}$}
\def\deg{$^{\circ}$}
\def\ddeg{{\rlap.}$^{\circ}$}
\def\dsec{{\rlap.}$^{\prime\prime}$}
\def\cc{cm$^{-3}$}
\def\flamb{erg s$^{-1}$ cm$^{-2}$ \AA$^{-1}$}
\def\flux{erg s$^{-1}$ cm$^{-2}$}
\def\fnu{erg s$^{-1}$ cm$^{-2}$ Hz$^{-1}$}
\def\hst{{\textit{HST}}}
\def\kms{km s$^{-1}$}
\def\lamb{$\lambda$}
\def\lax{{$\mathrel{\hbox{\rlap{\hbox{\lower4pt\hbox{${\sim}$}}}\hbox{$<$}}}$}}
\def\gax{{$\mathrel{\hbox{\rlap{\hbox{\lower4pt\hbox{${\sim}$}}}\hbox{$>$}}}$}}
\def\simlt{\lower.5ex\hbox{$\; \buildrel < \over {\sim} \;$}}
\def\simgt{\lower.5ex\hbox{$\; \buildrel > \over {\sim} \;$}}
\def\micron{{$\mu$m}}
\def\mnras{{MNRAS}}
\def\nat{{Nature}}
\def\pasp{{PASP}}
\def\perang{\AA$^{-1}$}
\def\peryr{yr$^{-1}$}
\def\reference{\noindent\pp}
\def\refindent{\par\noindent\parskip=2pt\hangindent=3pc\hangafter=1 }
\def\sb{mag~arcsec$^{-2}$}
\def\lsun{$L_\odot$} 
\def\msun{$M_\odot$}
\def\sigs{$\sigma_*$}
\newcommand{\lt}{<}
\newcommand{\gt}{>}

\def\etal{{\ et al.~}}
\def\galfit{{\tt GALFIT}}
\def\ser{{S\'{e}rsic\ }}
\def\redm{\texttt{redMaPPer}}
\def\cmodel{\texttt{cModel}}
% Samples
\def\rbcg{\texttt{cenHighMh}}
\def\nbcg{\texttt{cenLowMh}}
\def\redbcg{{$\lambda \ge 30$}}
\def\nonbcg{{$\lambda < 20$}}
% Mass related 
\def\mstar{{$M_{\star}$}}
\def\mhalo{{$M_{\mathrm{halo}}$}}
\def\logms{{$\log (M_{\star}/M_{\odot})$}}
\def\logmh{{$\log (M_{\mathrm{halo}}/M_{\odot})$}}

\def\minn{{$M_{\star,10\mathrm{kpc}}$}}
\def\meff{{$M_{\star,15\mathrm{kpc}}$}} 
\def\mtot{{$M_{\star,100\mathrm{kpc}}$}}
\def\mout{{$M_{\star,150\mathrm{kpc}}$}}
\def\mmax{{$M_{\star,\mathrm{Max}}$}}
\def\mgama{{$M_{\star,\mathrm{GAMA}}$}}
\def\mcmodel{{$M_{\star,\mathrm{cModel}}$}}

\def\logminn{{$\log (M_{\star,10\mathrm{kpc}}/M_{\odot})$}}
\def\logmtot{{$\log (M_{\star,100\mathrm{kpc}}/M_{\odot})$}}
\def\logmout{{$\log (M_{\star,150\mathrm{kpc}}/M_{\odot})$}}
\def\logmmax{{$\log (M_{\star,\mathrm{Max}}/M_{\odot})$}}
\def\logmgama{{$\log (M_{\star,\mathrm{GAMA}}/M_{\odot})$}}
\def\logmcmodel{{$\log (M_{\star,\mathrm{cModel}}/M_{\odot})$}}

\def\m2l{{$M_{\star}/L_{\star}$}}
\def\s2n{{$\mathrm{S}/\mathrm{N}$}}
\def\mden{{$\mu_{\star}$}}

% Commenting:
\newcommand{\xxx}[1]{\textcolor{red}{\textbf{XXX}}}
\newcommand{\todo}[1]{\textcolor{red}{\textbf{TODO:~#1}}}
\newcommand{\plan}[1]{\textcolor{cyan}{#1}}
\newcommand{\addref}{{\textcolor{red}{REF}}}
\newcommand{\note}[2]{\textcolor{blue}{\textbf{[Comment (#1): #2]}}}
\newcommand{\song}[1]{\textcolor{magenta}{\textbf{[Song: #1]}}}
\newcommand{\alexie}[1]{\textcolor{blue}{\textbf{[Alexie: #1]}}}
\newcommand{\jenny}[1]{\textcolor{Bittersweet}{\textbf{[Jenny: #1]}}}
\newcommand{\kevin}[1]{\textcolor{green}{\textbf{[Kevin: #1]}}}
\newcommand{\rachel}[1]{\textcolor{cyan}{\textbf{[Rachel: #1]}}}
\newcommand{\update}[1]{\textcolor{PineGreen}{#1}}
%\newcommand{\update}[1]{#1}

%%%%%%%%%%%%: Title and Affiliations %%%%%%%%%%%%

%% ------------------------------------------------------------------------------------ %% 
\title[Mass Dependent Stellar Halo in Massive Galaxies]{Tracing the Stellar Halos of 
	   Massive Galaxies to 100 kpc at $0.3<z<0.5$ using Hyper Suprime-Cam}

\author[S. Huang et al.]{
        Song Huang,$^{1,2}$\thanks{E-mail: song.huang@ipmu.jp (SH)}
        Alexie Leauthaud,$^{2,1}$
        Jenny E. Greene,$^{4}$
        Kevin Bundy,$^{3,1}$
        \newauthor
        Yen-Ting Lin,$^{5}$
        Masayuki Tanaka,$^{5}$
        Rachel Mandelbaum,$^{7}$
        Satoshi Miyazaki,$^{5,8}$
        \newauthor
        Yutaka Komiyama,$^{5,8}$
        HSC collaboration 
        \\
        $^{1}$Kavli-IPMU, The University of Tokyo Institutes for Advanced Study, 
              the University of Tokyo (Kavli IPMU, WPI), Kashiwa 277--8583, Japan\\
        $^{2}$Department of Astronomy and Astrophysics, University of California 
              Santa Cruz, 1156 High St., Santa Cruz, CA 95064, U.S.A\\
        $^{3}$UCO/Lick Observatory, University of California, Santa Cruz,
              1156 High Street, Santa Cruz, CA 95064, USA\\
        $^{4}$Department of Astrophysical Sciences, Peyton Hall,
              Princeton University, Princeton, NJ 08540, USA \\
        $^{5}$National Astronomical Observatory of Japan, 2--21--1 Osawa, Mitaka, 
              Tokyo 181--8588, Japan\\
        $^{6}$Academia Sinica Institute of Astronomy and Astrophysics, 
              P.O. Box 23--141, Taipei 10617, Taiwan\\
        $^{7}$McWilliams Center for Cosmology, Department of Physics, 
              Carnegie Mellon University, Pittsburgh, PA 15213, USA\\
        $^{8}$SOKENDAI (The Graduate University for Advanced Studies), Mitaka,
              Tokyo, 181--8588, Japan
        }   
%% ------------------------------------------------------------------------------------ %% 
\date{Accepted XXX. Received YYY; in original form ZZZ}        
\pubyear{2017}

%%%%%%%%%%%%: Header and Version %%%%%%%%%%%%
\begin{document}

\label{firstpage}
\pagerange{\pageref{firstpage}--\pageref{lastpage}}

\maketitle

%%%%%%%%%%%%: Abstract and Keywords %%%%%%%%%%%%

\begin{abstract} 
    Massive galaxies display extended light profiles that can reach out to several 
    hundreds of kilo parsecs. 
    These stellar halos are a key fossil record of galaxy assembly history. Using data that is both wide (${\sim}100$ deg$^2$) and deep ($>28.5$ \sb{} in 
    the $i$-band), we present a systematic study of the stellar halos of  a sample of more than $7000$ galaxies at $0.3 < z < 0.5$. 
Our study is based on high-quality (0.6\asec\ seeing) imaging data from the Hyper 
    Suprime-Cam (HSC) survey, which reaches ${\sim}4$ magnitudes fainter than SDSS and enables us to directly trace mass profiles to 100 kpc 
    without resorting to stacking. Shallow images and/or over-simplified parametric models cannot capture the extended stellar halo and result in underestimated stellar masses.
    Our extended profiles reveal a number of mass-dependent trends. The outer parts of our galaxies have higher ellipticity, and this ellipticity increases with stellar mass, but we find no difference in color gradients with mass. As in previous work, we see a clear trend towards more extended outer profiles in more massive galaxies. Thus, when this extended light is not properly accounted for, the stellar mass function is systematically underestimated at the highest masses. Our results are broadly consistent with the two phase formation scenario of massive galaxies in which outer envelopes are built up at late times from minor merging. We provide surface mass surface mass density profiles in a convenient tabulated format to facilitate comparisons with predictions from numerical simulations of galaxy formation.\end{abstract}

%massive central galaxies display surface mass density profiles that are not 
%    self-similar, with the outer-slope becoming shallower and the fraction of mass 
%    in the outskirt increasing with total stellar mass (xxxxxxxxx).

\begin{keywords}
    galaxies: elliptical and lenticular, cD --
    galaxies: formation --
    galaxies: photometry -- 
    galaxies: structure -- 
    galaxies: surveys
\end{keywords}

%%%%%%%%%%%%: Main Text %%%%%%%%%%%%

%% ------------------------------------------------------------------------------------ %% 
    \begin{figure*}
        \centering 
        \includegraphics[width=\textwidth]{fig/redbcg_sdss_compare}
        \caption{
            A comparison between the depth and imaging quality of SDSS and the HSC wide 
            layer for a sample of nearby massive elliptical galaxies at $0.2 < z < 0.5$.  
            These images are generated using $gri$ band images with an arcsinh stretch 
            \citep{Lupton2004}. 
            The HSC \texttt{WIDE} layer is $3.0$-$4.0$ magnitudes deeper than SDSS.
            This added depth is critical in order map the outskirts of ETGs out to 
            large radii.
\rachel{Should replace the quote marks with arcsec symbols in the scale on these plots.}
}
        \label{fig:sdss_compare}
    \end{figure*}
%% ------------------------------------------------------------------------------------ %%

%% ------------------------------------------------------------------------------------ %% 
\section{Introduction}
    \label{sec:intro}

    Simulations of structure formation within the context of the $\Lambda$-CDM 
    cosmological model make predictions for the hierarchical growth of dark matter 
    halos and galaxies (e.g. \citealt{Baugh1996, DeLucia2006}), but there are many 
    open questions regarding the star-formation history, mass assembly process, and 
    structural evolution of massive galaxies. 
    Massive galaxies are thought to grow according to a ``two-phase'' formation 
    scenario (e.g. \citealt{Oser2010, Oser2012}). 
    In this scenario, the progenitors of $z{\sim} 0$ massive early-type galaxies 
    (ETGs) undergo a rapid growth phase at $z{\sim} 2$ triggered either by disk 
    instabilities or gas-rich mergers (e.g. \citealt{Hopkins2008, Dekel2009}). 
    Observationally, the progenitors of ETGs are thought to correspond to the 
    population of massive quiescent galaxies at high redshift that have smaller 
    average effective radii ($R_{\mathrm{e}}$; e.g. \citealt{Trujillo2006, 
    vanDokkum2008, Cimatti2008}).  These high redshift galaxies also have slightly higher central velocity dispersions and stellar mass densities (\mden{}; e.g. \citealt{vandeSande2011, Belli2014}), and 
    more disk-like morphologies (e.g. \citealt{vanderWel2011}) than low redshift galaxies (e.g. \citealt{Bezanson2009, vanDokkum2010}) with similar stellar 
    mass.

    Following this initial epoch of early formation, feedback from stars and/or AGN 
    (e.g. \citealt{Sijacki2007, Fabian2012}) efficiently quenches star formation in 
    massive galaxies. 
    A large fraction of these massive progenitors are already quiescent by $z=1$ 
    (e.g. \citealt{Bezanson2009, Kriek2016}). 
    The second phase of their assembly is driven by non-dissipative processes such 
    as dry mergers with other galaxies (e.g. \citealt{Naab2006, Khochfar2006}). The ``two-phase'' formation of massive galaxies is  designed to both explain the observed increase in the effective radii ($R_{\mathrm{e}}$) of massive galaxies (e.g. 
    \citealt{Newman2012, vdWel2014}) and the build-up of stellar 
    halos  (e.g. \citealt{Szomoru2012, Patel2013}). According to this formation mechanism, minor mergers dominate the mass assembly of massive galaxies at late times (e.g. \citealt{Hilz2012, Hilz2013, Oogi2013, Bedorf2013, Laporte2013}).

    Both numerical simulations (e.g. \citealt{Oser2010}) and 
    semi-analytic models (SAM; e.g. \citealt{LeeYi2013, LeeYi2017}) agree that the 
    mass fraction in the accreted component should increase with total galaxy stellar mass (e.g. \citealt{Lackner2012, Cooper2013, 
    Qu2017}). For instance, recent results from the Illustris\footnote{
    \texttt{http://www.illustris-project.org/}} simulation 
    (\citealt{Vogelsberger2014}, \citealt{Genel2014}) predicts that the fraction of  
    accreted stars increases significantly and reaches $f_{\mathrm{accreted}}>0.5$ at \logms{}$>11.5$ (\citealt{RodriguezGomez2016}). \rachel{Given that Illustris gets the SMF totally wrong at high stellar mass, is there a reason we should be quoting their results?  Perhaps better to wait for Illustris-TNG?  Or give some caveat here that it's not clear that the physics that shapes the most massive galaxies is a realistic implementation?}
%% ------------------------------------------------------------------------------------ %%

    Given the success of this two-phase scenario in explaining trends like overall size 
    growth, it is time to confront this model with additional observations,  in particular, the detailed surface mass density profiles of low redshift massive galaxies. Early studies based on one-dimensional light profiles suggested that the profiles of nearby ETGs are well described by single-\ser{} profiles (e.g. \citealt{Kormendy2009}; except for the
    most central region) and that the \ser{} index increases with total luminosity 
    (e.g. \citealt{Graham2013}). However,  later work revealed a more complicated  situation  showing  that  ETGs  belong to two famillies, those that follow single-\ser{} law, versus those that significant deviation from the single-\ser{} profile \citep{Schombert2015}. \rachel{You should say what differed in these later studies; e.g., use of deeper/better data?} Two-dimensional analyses have found that the stellar distributions of massive ETGs are often better fit by multiple-component models 
    (e.g. \citealt{Huang2013a, Oh2017}).  \citet{Huang2013b} further suggest a connection between the multi-component 
    nature of massive galaxies and their two-phase assembly history. 
    To further confront the two-phase scenario requires very deep observations of 
    large samples of massive ETGs to correctly estimate their total stellar masses
    (e.g. \citealt{Bernardi2013, DSouza2014}) as well as to quantify the amplitude and scatter among outer envelopes (e.g. \citealt{Capaccioli2015, Iodice2016, 
    Iodice2017}).

    Until now,  studies of the profiles of massive galaxies have been  either  been conducted  using large samples but with shallow imaging, for example the Sloan Digital Sky Survey (SDSS; e.g. \citealt{SDSSDR7, SDSSDR12}), or with deeper imaging  but smaller sample sizes (e.g cite CLASH AND FRONTIER FIELDS??). In this paper, we take advantage of new high-quality (median  seeing of
    FWHM${\sim} 0.6^{\arcsec}$ in the $i$-band) and deep ($i$-band surface brightness 
    limit $> 28.5$~\sb) images from the Hyper Suprime-Cam (HSC) Subaru Strategic 
    Program \citep[SSP,][]{HSCDR1} to characterize the light profiles of massive galaxies out to 100 kpc. The deep imaging depth and excellent seeing conditions of HSC images make them 
    perfect for mapping the \mstar{} distributions of massive galaxies out to very 
    large radii. We select a large sample (${\sim} 7000$) of massive central galaxies at $0.3 < z < 0.5$ using ${\sim} 100$ deg$^2$ of data from the HSC wide layer.

 %       With the help of reliable spectroscopic and photometric redshift 
 %   (e.g. the ones for \redm{} clusters, \citealt{Rykoff2014}; \citealt{Rozo2015b}),
 %   we select a large sample (${\sim} 7000$) of massive central galaxies at 
 %   $0.3 < z < 0.5$ using ${\sim} 100$ deg$^2$ of data from the HSC wide layer.

    In this paper, we use this sample to 
    (a) reliably measure individual surface mass density (\mden{}) profiles of 
    massive galaxies out to 100 kpc, 
    (b) investigate the dependence of their outer stellar halos on total stellar 
    mass, and 
    (c) examine the implications in terms of  evaluating the high mass end of the galaxy stellar mass function (SMF). In a second paper in this series, we will investigate the environmental (dark matter halo mass) dependence of the sizes of massive ETGs (Huang\etal in prep.).

    This paper is organized as follows. 
    Section \ref{sec:data} presents our data and sample selection. 
    Section \ref{sec:ellipse} describes our procedure for extracting 1-D surface 
    brightness profiles. 
    Section \ref{sec:mstar} describes how we estimate stellar mass. 
    Our main results are presented in Section \ref{sec:result} and discussed in 
    Section \ref{sec:discussion}. 
    Section \ref{sec:summary} presents our summary and conclusions. \rachel{Outline does not mention section 5.}

    Magnitudes use the AB system (\citealt{Oke1983}), and are corrected for Galactic 
    extinction using calibrations from \citet{Schlafly11}.
    In this work, we assume $H_0$ = 70~km~s$^{-1}$ Mpc$^{-1}$, ${\Omega}_{\rm m}=0.3$, 
    and ${\Omega}_{\rm \Lambda}=0.7$.
    Stellar mass is denoted \mstar{} and has been derived using a Chabrier Initial Mass 
    Function (IMF; \citealt{Chabrier2003}).

    $M_{\mathrm{halo}}$ denotes halo mass in general whereas $M_{\rm 200b}$ is 
    explicitly defined as $M_{\rm 200b}\equiv M(<r_{\rm 200b})=200\bar{\rho} 
    \frac{4}{3}\pi r_{\rm 200b}^3$ where $r_{\rm 200b}$
    is the radius at which the mean interior density is equal to 200 times
    the mean matter density ($\bar{\rho}$).

%% ------------------------------------------------------------------------------------ %%

%% ------------------------------------------------------------------------------------ %% 
    \begin{figure*}
        \centering 
        \includegraphics[width=\textwidth]{fig/redbcg_ellipse_example}
        \caption{
            Left: example of the 1-D surface brightness and ellipticity profile 
            for a massive galaxy at $z=0.23$ in the $i$-band using \texttt{Ellipse}. 
            In this work, we always show the radial profile using a $R^{1/4}$ scaling 
            on the x-axis. \rachel{So should the units be kpc$^{1/4}$?}
            By using this scale, the de Vaucouleurs profile will appear as a straight 
            line on this figure.  
            We also plot the relative brightness profile of the PSF model normalized 
            at the central surface brightness of the galaxy to highlight the region 
            affected by seeing. \rachel{`most strongly affected by seeing'?  It's not the case that the seeing is totally irrelevant outside this region.  The `most affect' in the sentence below does not quite make sense either.}
            The grey shaded region highlights the region most affect by seeing 
            ($r<3$ kpc).
            On the top panel, the dashed line shows the ellipticity used for the final 
            isophote. 
            Right: the three color image of this galaxy with isophotes 
            extracted by \texttt{Ellipse}. 
            The thick dotted line highlights the isophote with 
            $\mu_{i}{\sim} 28.5$~\sb.
            }
        \label{fig:ellipse}
    \end{figure*}
%% ------------------------------------------------------------------------------------ %%

%% ------------------------------------------------------------------------------------ %% 
\section{Data And Sample Selection}
    \label{sec:data}

\subsection{The Hyper Suprime-Cam Survey}
    \label{ssec:hsc}

    The Subaru Strategic Program (SSP, \citealt{HSC_SSP, HSC_DR1}) makes use of the 
    new prime-focus camera, the Hyper Suprime-Cam (HSC;~\citealt{Miyazaki2012}), on the 
    8.2-m Subaru telescope at Mauna Kea. 
\rachel{Need to cite the HSC survey camera paper, Miyazaki et al in prep.}
The ambitious multi-layer HSC survey takes advantage of the large field of 
    view (FoV;~1.5 deg in diameter) of this camera and will cover $>1000$ deg$^2$ 
    of sky in 5 broad bands ($grizy$) to a limiting depth of $r {\sim} 26$ mag 
    in the \texttt{WIDE} layer. 
    This work is based on the internal data release \texttt{S15B}, which covers 
    ${\sim} 100$ deg$^2$ in all 5-band to full \texttt{WIDE} depth.  
\rachel{Need to say how this relates to the release that is available to the public and described in the HSC DR1 paper, for the sake of reproducibility.}
The regions covered by this release overlap with a number of spectroscopic surveys 
    (e.g.\ SDSS/BOSS: \citealt{Eisenstein2011}, \citealt{Alam2015}; 
    GAMA: \citealt{Driver2011}, \citealt{Liske2015}).

    The HSC \texttt{WIDE} survey is about $3.0$-$4.0$ magnitudes deeper in terms of the $i$-band surface brightness limit than SDSS. Combined with the excellent imaging resolution (the median $i$-band seeing is 
    0.6\asec) and the wide area, the HSC survey represents a tremendous dataset to 
    perform large statistical studies of the surface brightness profiles of massive
    galaxies out to their very outskirts.  \rachel{This language seems a bit too informal for a journal article.  Perhaps `\dots represents an ideal dataset to perform statistical studies\dots'?}   
    Fig~\ref{fig:sdss_compare} illustrates the quality of HSC imaging compared to SDSS 
    for three low redshift ETGs, and shows that HSC survey data are well suited for 
    mapping the stellar distribution of massive galaxies out to large radii.

	HSC $i$-band images typically have the best seeing compared to other bands because 
	of strict requirements driven by weak lensing science. 
    We will therefore use $i$-band images to measure the stellar distributions of 
    massive galaxies.

%% ------------------------------------------------------------------------------------ %% 
\subsection{HSC Data Processing}
    \label{sec:pipeline}

    The full details of the HSC data processing can be found in \citet{Bosch2017}
    and are briefly summarized here. 
    The HSC SSP data are processed with \texttt{hscPipe 4.0.2}, a derivative of the 
    Large Synoptic Survey Telescope (LSST) pipeline (e.g.\ \citealt{Juric2015}; 
    \citealt{Axelrod2010}), modified for HSC. 
    \texttt{hscPipe} first performs a number of tasks at the single exposure level 
    (bias subtraction, flat fielding, background modeling, object detection and 
    measurements). 
    Astrometric and photometric calibrations are performed at the single exposure 
    level. 
    \texttt{hscPipe} then warps different exposures on to a common World Coordinate 
    System (WCS) and combines them into coadded images. 
    At this stage, \texttt{hscPipe} updates the images with a better astrometric and 
    photometric calibration using stars that are common among exposures.

    The pixel scale of the combined images is $0.168$\asec{}. 
    Photometric calibration is based on data from the Panoramic Survey Telescope 
    and Rapid Response System (Pan-STARRS) 1 imaging survey 
    (\citealt{Schlafly2012}, \citealt{Tonry2012}, \citealt{Magnier2013}). 
    To achieve consistent deblending and photometry across all bands, \texttt{hscPipe} 
    performs multi-band post-processing at the \texttt{coadd} level. 
    First, \texttt{hscPipe} performs object detection on \texttt{coadd} images in 
    each band independently and records the flux peak and the above-threshold region 
    (referred as a \texttt{footprint}) for each source. 
    Next, \texttt{footprints} and peaks from different bands are merged together before     
    performing deblending and measurements. 
    Finally, \texttt{hscPipe} selects a reference band for each object based on the 
    $S/N$ in different bands (for most galaxies in this work, the reference band is 
    the $i$-band). 
    After fixing the centroids, shape, and other non-amplitude parameters of each 
    object in this reference catalog, \texttt{hscPipe} perform forced photometry 
    on the \texttt{coadd} image in each band. 
    This forced photometry approach is optimized to yield accurate galaxy colors.

    For each galaxy, \texttt{hscPipe} measures a \cmodel{} magnitude using an approach 
    that is similar to SDSS (\citealt{Bosch2017}). 
    However, as opposed to SDSS, the HSC \cmodel{} is based on forced photometry which 
    means that it can accurately measure both the \textit{fluxes and colors of galaxies}. 
    The HSC \cmodel{} algorithm fits the flux distribution of each object using a 
    combination of a de~Vaucouleur and an exponential component and accounts for the PSF. 
    The performance of this algorithm has been tested using synthetic objects 
    (\citealt{SynPipe}), and the results indicate that, generally speaking, 
    the HSC \cmodel{} photometry is accurate down to $i >25.0$ mag.  
    However, \cmodel{} currently systematically underestimates the total fluxes of 
    massive ETGs with extended stellar distributions. 
    This is caused by an intrinsic limitation of \cmodel{} as it is incapable of
    modeling profiles with extremely extended outskirts, a problem that is exacerbated 
    at the depth of the HSC survey. 
    In addition, at the depth of the HSC survey, accurate deblending in the vicinity of
    large ETGs where satellites and background galaxies often blend with the low surface 
    brightness stellar envelope is a challenging problem. 
    The deblending method currently implemented in \texttt{hscPipe} tends to 
    ``over-deblend'' the outskirts of bright galaxies and leads to an 
    under-estimation of the the total flux of massive ETGs (this is discussed further in 
    \citealt{Bosch2017}).  
    For these reasons, our results will be based on custom-developed code to measure 
    the luminosities and stellar masses of massive galaxies. 
    We use the HSC \texttt{hscPipe} photometry for two purposes: 
    1) to perform a first broad sample selection, and 2) to estimate the average 
    color of massive galaxies.

%% ------------------------------------------------------------------------------------ %% 
  \begin{figure*}
      \centering 
      \includegraphics[width=\textwidth]{fig/redbcg_mass_diff_new}
      \caption{              
          \textbf{Left:} Difference between \mcmodel{} and \mtot{} for massive
      	  galaxies (grey dots).
      	  The running-median of the mass difference is shown by large red hexagons. On average, \mcmodel{} underestimates the total stellar mass of massive 
          galaxies by 0.1 dex while in some cases, the difference can exceed 0.2 dex.
          Vertical histograms indicate the mass difference for all galaxies (shaded 
          histogram) and for the ones with \logmtot{}$>11.6$ (empty histogram).
          \textbf{Right:} Difference between \mmax{} and \mtot{} in the same format. 
          The average difference is small (0.02 dex) and with no clear mass--dependence. \rachel{It might be worth explicitly noting the different vertical axis scales in the caption.  I did not notice at first, and if you do not notice then you might draw the wrong conclusion.}}
      \label{fig:mass_diff}
  \end{figure*}

%% ------------------------------------------------------------------------------------ %%

%% ------------------------------------------------------------------------------------ %% 
\subsection{Initial Massive Galaxy Sample}
    \label{ssec:initial}

    We begin by using a broad luminosity cut to select an initial sample of massive 
    galaxies at $z < 0.5$ from the HSC photometric catalog. 
\rachel{You do not seem to explicitly calculate luminosities, but rather to make a flux (not luminosity) cut.  So should we call it a `broad flux cut'?}
Based on \citet{Leauthaud2016}, $i_{\mathrm{SDSS, cModel}} \leq 21.0$ mag can 
    define a sample that includes almost all \logms{}$\geq 11.5$ galaxies.    
    We therefore perform an initial conservative selection of massive galaxies
    with $i_{\mathrm{HSC, cModel}} \leq 21.5$\footnote{We neglect small differences
    between the response curves of the SDSS-$i$ and HSC-$i$ filters.}. We also limit our sample to regions that have reached the required depth of 
    the \texttt{WIDE} survey in the $i$-band. \rachel{As defined in DR1 paper?}
    %% The SQL search is saved as "sample.sql" 
    %% The master catalog used here is "dr1_wide_galaxy_icmodel_21.5.fits"
    %%   The original catalogs contain 2275477 objects

    We further select extended objects with no deblending errors, with well defined 
    centroids, and with useful \texttt{cModel} magnitudes in all five bands. 
    After removing objects that have pixels affected by saturation, cosmic-rays, or 
    other optical artifacts\footnote{each criterion removes less than 8\% of the 
    entire sample.}, this sample corresponds to 1760845 galaxies and will be referred 
    as \texttt{hscPho}. 
    %% Saved as "dr1_wide_galaxy_hscPho.fits"

    Here we limit our study to the very high-mass end where the majority of galaxies 
    have either a spectroscopic redshift or a robust red-sequence photo-$z$ from the 
    \redm{} galaxy cluster catalog\footnote{See: http://risa.stanford.edu/redmapper/} 
    (e.g.\ \citealt{Rykoff2014}; \citealt{Rozo2015b}).  
\rachel{You require it to be a redmapper cluster member?  Or do you mean it must be part of the redmagic galaxy sample?  I do not fully understand the logic in using redmapper rather than redmagic.}
We match the \texttt{hscPho} sample with a spec-$z$ catalog compiled by the HSC 
    team\footnote{It is created by matching HSC objects with all publicly available 
    spectroscopic redshifts (e.g.\ SDSS/BOSS; GAMA). 
    The spec-$z$ quality flag from different catalogs are homogenized into a single 
    flag that indicates secure redshifts. \rachel{Is there a paper or a note at the data release page that we could reference so readers can check out how this was defined?}}.  
    To ensure reasonable \mstar{}-completeness at the high-\mstar{} end we focus on
    the redshift range $0.3 \leq z \leq 0.5$. 
    %% This is catalog: "dr1_specz_use.fits"

    Objects without a spectroscopic redshift are matched with central 
    galaxies from the \redm{} SDSS DR8 (\citealt{SDSSDR8}) catalog using a $2.0^{\arcsec}$ matching radius. Matched objects with a red-sequence photo-$z$ ($0.3 \leq z_{\lambda} < 0.5$) are included in our sample. The accuracy of the red-sequence photo-$z$ is sufficient (median 
    $|z_{\lambda} - z_{\mathrm{Spec}}| {\sim} 0.01$ for the ones with spec-$z$) 
    for the goal of this work.
    The \redm{} catalog provides an additional 133 unique redshifts for massive 
    galaxies in our sample.
    %% Estimated from "dr1_redbcg_hsc_sdss_gama_1arcsec.fits"

    In total, at $0.3 \leq z \leq 0.5$, our sample consists of 25286 galaxies with 
    reliable redshift information (referred as \texttt{hscZ}).
    %% Saved as "dr1_wide_galaxy_icmodel_21.5.fits"
    The majority of our redshifts come from the BOSS and SDSS ``legacy'' LRG samples. The GAMA survey provides an additional $14$\% of spec-$z$.
\rachel{This is only of order 1\% of the original sample.  Do we understand why?  Of particular concern are cuts that affect survey homogeneity; e.g., in the area covered by GAMA vs.\ the area not covered by GAMA, might the sample look different?  What will that do to this study?}
We choose the redshift range of $0.3 \leq z \leq 0.5$ to make sure that 
    (1) The inner region of massive galaxies can be resolved, and \mstar{} within 
    10 kpc can be reliably measured; 
    (2) The background noise and cosmological dimming are not major issues and the 
    \mden{} profile can be measured out to $>100$ kpc; 
    (3) Redshift evolution in the stellar population properties can be largely 
    ignored.  
    Also at higher redshift, the completeness of the spec-$z$ sample starts to decline. And finally, the over-subtraction of the background level becomes a more serious issue at lower redshifts.  We will now describe our one-dimensional photometric analysis  (\S \ref{sec:ellipse}) and our stellar mass estimates (\S \ref{sec:mstar}). We will then define the final sample in \S \ref{sec:final}.

%    By comparing with the Stripe 82 Massive Galaxy Catalog
%    (\texttt{S82-MGC}; \citealt{Bundy2015}), \citet{Leauthaud2016} have measured the
%    \mstar{} completeness of the combined BOSS and SDSS samples. 
%    They estimate that the BOSS spec-$z$ sample is about 80\% complete at 
%    \logms{}$\geq 11.6$ at $0.3 < z < 0.5$. 
%    The GAMA survey, which partially overlaps with the HSC footprints, supplements 
%    our \texttt{hscZ} sample by 14\%. 
%    Based on \citet{Taylor2011} (e.g.\ their Fig~~6), at $z{\sim} 0.3$, the GAMA 
%    sample is 80\% complete down to $10^{10.8}$\msun; but only 80\% complete to 
%    $10^{12.0}$\msun at $z{\sim} 0.5$. 
%    We can therefore expect our sample to be roughly \mstar{}-complete above 
%    \logms{}$\geq 11.5$-$11.6$. 
%    This question will be addressed more carefully using the overlap region between 
%    HSC and the \texttt{S82-MGC} in Section \ref{ssec:s82}.

%% ------------------------------------------------------------------------------------ %%  
%\subsection{The \redm{}{} cluster catalog}
%    \label{ssec:redmapper}
%    
%    For now, we wish to focus on galaxies located at the center of their dark matter halos. 
%    To limit our sample to central galaxies, we use \texttt{v5.10} of the \redm{}{} 
%    cluster catalog\footnote{See: http://risa.stanford.edu/redmapper/} 
%    (e.g.\ \citealt{Rykoff2014}; \citealt{Rozo2015b}). 
%    These authors have developed a well-tested red-sequence cluster finder that has 
%    been run on SDSS DR8 (\citealt{SDSSDR8}) photometric data. 
%    For each cluster, the catalog provides a photometric redshift ($z_{\lambda}$), a 
%    cluster richness ($\lambda$), and identifies the most likely central galaxy (the 
%    galaxy with the highest $P_{\mathrm{Cen}}$ value). 
%    The \redm{}{} catalog also provides a list of member galaxies for each cluster and 
%    their associated membership probabilities. 
%    Details about the performance of the \redm{}~cluster catalog can be found in 
%    \citet{Rozo2014}, \citet{Rozo2015a}, and \citet{Rozo2015b}. 
%    Several studies have published calibrations between the \redm{}~richness estimate, 
%    $\lambda$, and halo mass (e.g.\ \citealt{Saro2015, Farahi2016, Simet2016, 
%    Melchior2016}). 
%    All these studies consistently find that \redm{} clusters with $\lambda > 20$ 
%    have $\log (M_{\mathrm{halo}}/M_{\odot}) \geq 14.0$. 
%    The \redm{} catalog therefore helps us group massive central galaxies into 
%    samples with different average \mhalo{}. 
%    We will use the central galaxies of \redm{} clusters 
%    ($P_{\mathrm{CEN}} \geq 0.7$) to form a sample of massive central galaxies that 
%    live in massive halos.
%% ------------------------------------------------------------------------------------ %%

%% ------------------------------------------------------------------------------------ %% 
%\subsection{Massive Central Galaxies in Low and High Mass Halos}
%    \label{ssec:mass_central}
%
%    One of the goals of our work is to investigate the halo dependence of the outer 
%    envelope. Although there is a correlation between $M_{\star}$ and $M_{\mathrm{Halo}}$,
%    based on recent constraints of $M_{\star}$-$M_{\mathrm{Halo}}$ relation 
%    (e.g.\ \citealt{Leauthaud2012}, \citealt{Behroozi2013}, \citealt{Kravtsov2014}), 
%    even though our sample selects very massive galaxies (\logms{}$\geq 11.5$), they 
%    With the help of the \redm{} catalog, we can broadly separate our sample into two
%    groups with \logmh{}$<14.0$ and \logmh{}$>14.2$.
%    live in a fairly broad range of halos masses.
%    
%    % We match this sample with the central galaxies of \redm{}~catalogs using a 
%    % $1.0^{\arcsec}$ radius, and it results in 704 galaxies.  
%    %% Saved as "dr1_redbcg_hsc_sdss_gama_1arcsec.fits"
%    
%    We match the \texttt{hscZ} sample with the central galaxies of \redm{} clusters 
%    with $\lambda \geq 30$ and $P_{\mathrm{0.7}} \geq 0.7$, and find 164 matched 
%    galaxies at $0.3 \leq z \leq 0.5$.
%    Our sample of \textbf{central galaxies in more massive halos} will be referred to 
%    as the \rbcg{} sample. 
%    The median $\lambda$ of the clusters associated with the \rbcg{} sample is 
%    ${\sim} 41$, corresponding to halo mass of 
%    $M_{\mathrm{halo}}{\sim} 2.2\times 10^{14}$ $h^{-1}$ $M_{\odot}$.
%    Only 44 centrals from the \rbcg{} sample live in clusters with $\lambda \geq 50$
%    ($M_{\mathrm{halo}} {\sim} 3.0\times 10^{14}$ $h^{-1} M_{\odot}$). 
%    Our \rbcg{} sample therefore corresponds to galaxies in moderate mass halos.
%    %% Saved as "dr1_redbcg_use_sed5b.fits"
%    
%    Out next goal is to construct a sample of central galaxies living in halos with 
%    \logmh{}$<14.0$. 
%    We first identify and remove all galaxies within a cylindrical region around all 
%    \redm{} clusters. 
%    We use a radius equal to $R_{\mathrm{200b}}$ and the thickness of the cylinder is 
%    set to twice the uncertainty of the photometric redshift error. 
%    We convert $\lambda$ to $M_{\mathrm{halo}}$ using the calibration of 
%    \citet{Simet2016} and we use the mass-concentration relation from 
%    \citet{Diemer2015} to compute $R_{\mathrm{200b}}$. At $0.3 < z < 0.5$, the 
%    uncertainty of photo-$z$ is between 0.015 to 0.025, and is enough to exclude 
%    cluster members.
%    After removing galaxies associated with \redm{} clusters, the remaining galaxies 
%    in our sample will be dominated by central galaxies living in halos with 
%    \logmh{}$\leq 14.0$. 
%    We will refer to this sample of \textbf{central galaxies in less massive halos}
%    as the \nbcg{} sample.
%   
%    Due to the depth and resolution of SDSS, the \redm{} catalog becomes slightly 
%    incomplete at lower richness ($\lambda < 40$) at $z > 0.33$. 
%    To ensure a clean separation in \mhalo{}, here we only focus on the central 
%    galaxies of clusters with $\lambda > 30$. 
%    According to \citet{Simet2016}, this $\lambda \geq 30$ sample should correspond to 
%    $M_{\mathrm{halo}}>10^{14.2} M_{\odot}$. 
%    We have checked that none of the results in this work depend on the choice of this 
%    $\lambda$ limit. 
%    For massive central galaxies that are not in these cluster-level halos, 
%    we unfortunately can not estimate their \mhalo{} individually, but it is safe 
%    to assume they will have $\log (M_{\mathrm{halo}}/M_{\odot}) < 14.0$.  
%    {\bf I am still a bit confused why stellar mass completeness depends on Mh??}
%
%    We note that at this stage, our \nbcg{} sample still contains galaxies with a 
%    wide range of \mstar{} values. 
%    Later on, once we have more accurate \mstar{} estimates, we will further limit this 
%    sample to \logms{}$ > 11.5$. 
%    At this very high mass end, and given that we have rejected satellites in \redm{} 
%    clusters, any further satellite contamination can be safely neglected 
%    (e.g. \citealt{Reid2014, Hoshino2015, Saito2016, vanUitert2016}).
%    %% Saved as "dr1_nonbcg_use_sed5b.fits"
%    
% 
%%% ------------------------------------------------------------------------------------ %%

%% ------------------------------------------------------------------------------------ %% 
%\subsection{Summary of Sample Construction}
%    \label{ssec:sample}
%
%    Using ${\sim} 100$ deg$^2$ of HSC data, we select a large sample of massive central 
%    galaxies with reliable redshift information, and broadly separate them two 
%    categories based on $M_{\mathrm{halo}}$.
%    
%    The following is a summary of our sample construction.
%    
%    \begin{itemize}
%        \item \texttt{hscPho} sample: this parent sample consists of bright galaxies 
%            with $i_{\mathrm{cModel}} \leq 21.0$, good quality imaging, and reliable 
%            \texttt{cModel} photometry in all five HSC bands in the \texttt{S15B} 
%            data release. 
%        \item \texttt{hscZ} sample: we limit the \texttt{hscPho} sample to galaxies 
%            with reliable redshift information. 
%        \item {\bf Select only central galaxies using RedMapper: total number in full 
%        sample?}
%        \item We further divide the sample into an \rbcg{} sample and a \nbcg{} sample, 
%          with the former roughly having \logmh{}$\geq 14.2$ and the latter having 
%           \logmh{}$<14.0$.
%
%          : we select a sample of 164 galaxies at 
%            $0.3 \leq z \leq 0.5$ that are the central galaxies 
%            in $\lambda > 30$ \redm{} clusters. 
%            They represent the central galaxies in very massive halos 
%            (\logmh{}$\geq 14.2$). 
%        \item \nbcg{} sample: after excluding all cluster members from the 
%            \texttt{hscZ} sample, we have 14661 bright galaxies at 
%            $0.3 \leq z \leq 0.5$. 
%            After applying an additional \mstar{} cut (see Section 
%            \ref{ssec:isedfit}), this sample corresponds to central galaxies in 
%            halos with \logmh{}$<14.0$.  
%    \end{itemize}
%    
%    We now proceed to measure the luminosities and stellar mass profiles for 
%    all galaxies in the \rbcg{} and \nbcg{} samples.
%% ------------------------------------------------------------------------------------ %%

%% ------------------------------------------------------------------------------------ %% 
\section{Measurements of 1-D Surface Brightness Profiles}
    \label{sec:ellipse}

    The surface brightness profiles of massive ETG are not well modeled by the 
    de~Vaucouleurs or single-\ser law, especially at the imaging depth of HSC.
    These models will fail to simultaneously describe the profile in both the inner 
    and the outer regions and also cannot account for any radial variations in 
    ellipticity and position angle. 
    In principle, massive galaxies can still be described by more complex 
    models (e.g \citealt{Huang2013a, Huang2013b, Oh2017}), but they are still 
    sensitive to the choice of a particular model (e.g. de~Vaucouleurs or \ser{} profile), 
    the number of components, initial guesses of parameters, and internal 
    degeneracies among different parameters. 
    Background subtraction uncertainties can also affect the 2-D model fitting method, especially for the massive ETGs that make up our sample.

    We therefore perform elliptical isophote fitting using the IRAF \texttt{Ellipse} 
    algorithm (Jedrzejewski 1987) to estimate the total luminosities of massive 
    galaxies and to measure their one-dimensional stellar mass surface 
    density profiles (\mden{}). 
    This 1-D method is less affected by the issues mentioned above. 
    Also, we will only study galaxies in the radial range where we are not 
    sensitive to either the PSF ($r<3-5$ kpc) or the background subtraction 
    ($r>100$ kpc). \rachel{Here too I think we have to be careful about how we say this.  The PSF is not irrelevant outside of the inner part of the profile.  It's just less important.  Perhaps say `\dots where we are less sensitive to the PSF or the background subtraction\dots'}

    We prepare large $i$-band postage-stamps for each galaxy that extend to 750 kpc in 
    radius, along with a bad pixel mask and the PSF model. 
    These postage-stamps include all of the light of the galaxy and are large enough to 
    also evaluate the background level. We choose to use $i$-band images because they trace the stellar distributions of 
    massive galaxies at $0.3 \leq z \leq 0.5$ reasonably well 
    (the observed $i$-band corresponds to a rest-frame $g$ or $r$ band), but also 
    because they have better seeing and much lower background levels than the $z$ 
    and $y$ band images (although these in principle may be better tracers of 
    \mden{}).

    For each cut-out, to overcome the \texttt{hscPipe} ``over-deblending'' issue, 
    we use a customized procedure to detect and aggressively mask out
    neighbouring objects. 
    Furthermore, \texttt{hscPipe} tends to over-subtract the background around 
    bright objects. 
    To improve the background subtraction, we first aggressively mask 
    out all objects (including the central massive galaxy), and derive an 
    empirical background correction using \texttt{SExtractor}.
\rachel{Need to describe what settings were used and, more importantly, how these were validated.  Otherwise there is no reason to trust them over the \texttt{hscPipe} results!  You state below that you've tested this procedure, but without enough information for readers to judge.  Perhaps a key plot or two could go in the appendix?}
Then, we run \texttt{Ellipse} on the background-corrected, masked cut-outs 
    following the methodology of \citet{Li2012}. 
    In short, we first fit each isophote using a free centroid and shape 
    (ellipticity and position angle). 
    We then fix the centroid (using the mean flux-weighted centroid) and estimate
    the mean ellipticity and position angles of all isophotes. 
    Finally, we extract a 1-D surface brightness profile along the major axis using 
    the mean ellipticity and position angle. 
    We correct these surface brightness profiles for Galactic extinction and 
    cosmological dimming, and integrate them to various radii to get the luminosity 
    within different physical (elliptical) apertures. 
    Fig~\ref{fig:ellipse} shows an example of the 1-D surface brightness and 
    ellipticity profile for a massive galaxy at $z{\sim}0.2$ and also highlights 
    a few isophotes.

    We test our procedure using different mask sizes, different \texttt{Ellipse} 
    parameters, and with or without our background correction. 
    Based on these tests, we find that our 1-D surface brightness profiles are reliable 
    up to surfaces brightness levels of $i{\sim}28.5$ \sb. 
    Beyond that, some of our profiles shows signs of truncation and/or large 
    fluctuations which are due to either the uncertainty in the background 
    subtraction or the unmasked flux from other objects.
    We choose to limit our study to surface brightness levels up to ${\sim} 28.5$ \sb. 
    This is a conservative choice but already enables us to measure light profiles 
    out to $100$ kpc on a galaxy-by-galaxy basis (no stacking). 
    For more technical details of the \texttt{Ellipse} procedure, please see 
    Appendix~\ref{app:ellipse}.

%% ------------------------------------------------------------------------------------ %%

    For physical (e.g.\ late-stage major merger) and nuisance (e.g.\ nearby foreground 
    galaxy or bright star) reasons, we cannot extract reliable 1-D profiles for a small 
    fraction of massive galaxies because they are heavily masked out. 
    This is an intrinsic limitation of the 1-D method, and it removes ${\sim}10$\% of 
    the sample.
    %It affects $18/164$ of the \rbcg{} sample, and a smaller fraction of the 
    %\nbcg{} sample. 
    It is worth noting that this excludes most massive galaxies undergoing major mergers.

    Given our exquisite profiles that extend to 100 kpc, the definition and meaning 
    of ``total'' magnitude (and stellar mass) becomes nuanced.
    With the help of the \m2l{} estimated in the next section (\S \ref{sec:mstar}), 
    we integrate the profile to a range of radii, and estimate the stellar mass 
    within these different apertures.  \rachel{You should at least clarify whether it's a radius-dependent \m2l{} here or not, so readers know whether both luminosity and \m2l{} depend on radius, or just luminosity.}
    Motivated by the two-phase scenario, we will consider two benchmark physical 
    apertures throughout this work:

    \begin{itemize}

        \item \textbf{The \mstar{} within the inner 10 kpc} 
            (hereafter noted \minn{}). 
            According to the two-phase scenario, the \textit{in situ} star-formation phase quickly builds up the inner, dense core of massive ETGs.  
            Based on recent observation (e.g.~\citealt{vanDokkum2010}) and 
            simulations (e.g. \citealt{RodriguezGomez2016}), the \textit{in situ} 
            component dominates the \mstar{} within one effective radius 
            ($R_{\mathrm{e}}$, or 5-10 kpc) of $z{\sim}0$ massive ETGs.
            We therefore use \minn{} as proxy for the mass forged during the 
            \textit{in situ} phase. 
            Given the quality of the HSC data, we can reliably measure \minn{} over 
            our redshift range ($1.0^{\arcsec}$ in radius equals 4.4 and 6.1 kpc 
            at redshifts 0.3 and 0.5 respectively).  
\rachel{Earlier you said you do not measure the profiles on the very smallest scales, below 3-5 kpc, so how are scales in that range handled when measuring \mstar{} within 10 kpc?}
It is worth noting that, at the very high-\mstar{} end, accreted stars may make a significant contribution to the mass within 10 kpc 
            (e.g. \citealt{RodriguezGomez2016}). We will further discuss the justification of this assumption in 
            \ref{ssec:twophase}.

        \item \textbf{The stellar mass within 100 kpc} 
            (hereafter noted \mtot{}). 
            For our galaxy sample, a 100 kpc aperture corresponds to 5-10 
            $\times R_{\mathrm{e}}$. 
            We show in section \ref{ssec:mtotal} that most of the \mstar{} for 
            these ETGs can be captured by a 100 kpc radius and that \mtot{} is 
            a good proxy for the ``total'' \mstar{}. 
            Although not perfect, we argue that our measurements of \mtot{} (which are 
            actually measuring the light directly out to 100 kpc) are a better tracer 
            of total \mstar{} than model-dependent results from shallower data 
            (such as SDSS) which rely on extrapolating the light profiles of galaxies 
            out to large radii.

       \end{itemize}

%% ------------------------------------------------------------------------------------ %%

%% ------------------------------------------------------------------------------------ %% 
\section{Stellar Masses and Mass Density Profiles}
    \label{sec:mstar}

%% ------------------------------------------------------------------------------------ %% 
\subsection{Stellar Masses from SED Fitting}
    \label{ssec:isedfit}

    To convert luminosities into \mstar{}, we assume that these massive galaxies 
    can be well described by an average \m2l{}. 
    This is a reasonable assumption considering that they are mostly dominated by 
    old stellar populations and are known to have only shallow color gradients. 
    We will further justify this point by measuring their median color profiles in 
    Section \ref{ssec:ell_color}.

    We use the broadband Spectral Energy Distributions (SEDs) fitting 
    (see \citealt{Walcher2011} for a recent review) code 
    \texttt{iSEDFit}\footnote{http://www.sos.siena.edu/~jmoustakas/isedfit/} 
    (\citealt{Moustakas13}) to estimate the average \m2l{} and $k$-corrections using
    5-band HSC \cmodel{} fluxes.
    Although \cmodel{} tends to underestimate the total flux for bright, extended 
    objects, it can still yield accurate \emph{average} colors thanks to the 
    forced-photometry method that takes the PSF convolution into account
    (e.g. \citealt{SynPipe}).

    \texttt{iSEDFit} takes a simplified Bayesian approach. 
    In short, it first generates a large grid of SEDs from synthetic stellar 
    population models by drawing randomly from the prior distributions of relevant
    parameters (e.g. age, metallicity, dust extinction, and star formation history).
    Based on these models, it uses the observed photometry and redshift to compute 
    the statistical likelihood, and generates the posterior probability distribution 
    functions (PDF) for each parameter.  
    To get the best estimate of a given parameter, \texttt{iSEDFit} integrates the 
    full PDF over all the other nuisance parameters.
    Then, the median value and the 1-$\sigma$ uncertainty are derived based on the 
    marginalized PDF. 
    Please refer to \citet{Moustakas13} for technical details.

    In this work, we derive average \m2l{} using the Flexible Stellar Population 
    Synthesis\footnote{http://scholar.harvard.edu/cconroy/sps-models}
    (FSPS; \texttt{v2.4}; \citealt{FSPS}, \citealt{Conroy2010}) model based on the 
    MILES\footnote{http://www.iac.es/proyecto/miles/pages/stellar-libraries}
    (\citealt{MILES1}, \citealt{MILES2}) stellar library and assuming a
    \citet{Chabrier2003} IMF between 0.1 to 100 \msun. 
    The star formation history (SFH) is assumed to follow a delayed-$\tau$ model with 
    stochastic star bursts (see Appendix~\ref{app:sed}). 
    This SFH is appropriate for massive galaxies at low redshifts 
    (e.g.\ \citealt{Kauffmann2003}). 
    For stellar metallicity 
    ($[\mathrm{M}/\mathrm{H}]=\log\ \mathrm{Z}/\mathrm{Z}_{\odot}$), we assume a flat 
    distribution between 0.004 to 0.03 (the highest value allowed by FSPS).
    We adopt the \citet{Calzetti2000} extinction law with a order two Gamma 
    distribution of $A_{V}$ between 0 to 2 magnitude. 
    The majority of our galaxies are red and quiescent ones so the results are not 
    very sensitive to parameters related to the SFH or the internal dust extinction. 
    To achieve reasonable sampling across these parameters, we generate 250000 models.

    We construct five band SEDs using the forced-photometry \cmodel{} magnitudes 
    corrected for Galactic extinction. 
    \cmodel{} only accounts for the statistical error on the flux measurement but 
    does not account for systematic uncertainties in the model-fitting process. 
    It therefore underestimates the flux errors of bright galaxies. 
    For this work, we supply \texttt{iSEDFit} with simplified flux errors assuming 
    $S/N = 100$ for the $riz$ bands, and $S/N = 80$ for the $g$ and $y$ band 
    (on average, images in $gy$ bands are shallower in depth and/or have higher 
    background noise).  
    Although these are still conservative estimates of the systematic uncertainties 
    involved in the model-fitting, tests using synthetic galaxies show that \cmodel{}
    provides excellent measurements of five-band colors, which are crucial for 
    reliable \m2l{} estimates.
\rachel{I'm a little surprised at the statement that 1\% error is a conservative estimate of systematic uncertainties.  It definitely seems like a conservative estimate of {\em statistical} error.  Can you justify that all systematics are really below 1\%?}
The typical uncertainty of \logms{} is around 0.06-0.08 dex at \logms${\sim} 11.5$.

    In Appendix~\ref{app:sed}, we briefly summarize the basic statistics of 
    the sample, including average stellar age, metallicity, and internal dust 
    extinction along with their relation with \mtot{}. 
    All these properties behave reasonably for massive galaxies in this sample. 
    Using the $k$-corrected optical color, we can also confirm that the sample follows 
    a tight ``red-sequence''.  
    Please see Appendix~\ref{app:sed} for further details.

%% ------------------------------------------------------------------------------------ %%

%% ------------------------------------------------------------------------------------ %% 
\subsection{``Total'' Stellar Masses}
    \label{ssec:mtotal}

    Using the best-fit \mstar{} from \texttt{iSEDFit} (noted as as \mcmodel{}), 
    we estimate the average \m2l{} in the $i$-band, then use that \m2l{} to convert 
    our 1-D luminosity density profiles into stellar mass density (\mden{}) profiles. 
    We also convert our 10 and 100 kpc aperture luminosities into corresponding stellar 
    mass estimates (noted as \minn{} and \mtot{}).

    For the remainder of this paper, we will use \mtot{} as a proxy of ``total'' 
    stellar mass. 
    As expected, the integration of the 1-D profile out to very large radius recovers 
    more luminosity (stellar mass) compared to the \cmodel{}-based estimates 
    (Fig \ref{fig:mass_diff}).
    At the high-\mstar{} end (e.g. \logmtot{}$>11.6$), the average difference is larger 
    than 0.1 dex and can be as large as 0.2--0.3 dex.  
    More importantly, the differences between \mtot{} and \mcmodel{} clearly show a 
    dependence on total stellar mass as \mcmodel{} tends to miss more light in 
    more massive galaxies.  
    This relates to the mass-dependent nature of the stellar halos of massive 
    galaxies and the intrinsic limitation of \cmodel{} method, which we will discuss
    more in the next section.
    Such differences also have important implications for estimates of the stellar 
    mass function and for studies of the environment-dependence of galaxy structure. These topics will be discussed in Section \ref{sec:discussion}.

    Although 100 kpc is already a very large radius that should enclose the majority 
    of stars that belong to these massive galaxies, we know that their \mden{} 
    profiles extend beyond 100 kpc without showing any signs of truncation. 
\rachel{This bold assertion is practically begging for justification.  Is this based on past work we can reference, or a part of this paper that we can forward reference?}
Therefore, for massive galaxies, even \mtot{} should be only considered a lower
    limit on the ``total'' \mstar{}. 
    In Fig \ref{fig:mass_diff}, we integrate the \mden{} of each galaxy to the  edge of the postage-stamp and call this mass \mmax{}. The right panel of Fig \ref{fig:mass_diff} compares \mmax{} and \mtot{}.  Uncertainties in the background subtraction, and the impact of neighbouring objects, means that \mmax{} is much more uncertain than \mtot{}. Nontheless, we do see that \mmax{} is larger than than \mtot{}. However, the mass differences are on average very small ($\sim0.02-0.03$ dex) and 
    do not show a strong mass dependence.  
    This confirms that, at the current depth of HSC images, \mtot{} can be used as 
    good proxy of ``total'' stellar mass.

    Our methodology ignores any radial variations in \m2l{}. 
    It is well known that massive ETGs have negative optical color gradients 
    indicating gradients in \m2l{} (e.g.\ \citealt{Carollo1993, Davies1993, 
    LaBarbera2012, DSouza2015}). However, these color gradients are shallow and smooth out to a few times the effective 
    radius (color gradients at larger radii are not yet well quantified). Because the gradients are shallow, using an average \m2l{} is unlikely to bias our results on \mstar{}
    measurements. \rachel{I can see why this statement might be true for the total stellar mass estimates, but might the stellar mass within 10 kpc be biased?  Can we at least say the typical sign of this bias or a rough magnitude?} Color gradients will be discussed more in Section~\ref{ssec:ell_color}. 
    In summary, our results about the mass-dependence of \mden{} profiles should not be affected by the asumption of a constant  \m2l{} ratio because optical color gradients in our sample do not show a dependence on  stellar mass.

    %Our results looking at the outer envelopes as a function of mass and environment 
    %will be robust as long as there is \emph{no significant difference in 
    %the average color gradient between sub-samples}.
    %We have confirmed that this is indeed the case (see Section~\ref{ssec:ell_color}).

    %We now compare our masses with those from the GAMA survey. Our method for deriving total \mstar{} is similar to the method adopted by the 
    %GAMA survey (\citealt{Taylor2011, Kelvin2012}), where the average \m2l{} is from 
    %SED fitting of PSF-matched aperture photometry and the total luminosity comes from 
    %sing-\ser{} fits to SDSS images.  
    %To avoid differences in stellar population assumptions, we directly compare the 
    %total luminosity within 100 kpc and the GAMA single-\ser{} luminosity for the same
    %galaxies.

    %At high-\mstar{} end, the HSC 1-D profiles still recover more light than the 
    %single-\ser{} model on average.
    %It is likely that the GAMA models fail to capture the flux associated with the 
    %outskirts of galaxies due to the limited depth of SDSS.  
    %Meanwhile, there are also many cases where the high-\ser{} index of GAMA model leads 
    %to over-estimated total luminosity. 
    %We will discuss more of this comparison in the next paper of this series 
    %(Huang\etal in prep.).

    %We have compared our \mstar{} values with those from GAMA and the results are given 
    %in Appendix~\ref{app:gama}.

%% ------------------------------------------------------------------------------------ %%

%% ------------------------------------------------------------------------------------ %%        
\subsection{Stellar Mass Completeness}
    \label{ssec:complete}

    With the help of the \texttt{S82-MGC} (\citealt{Bundy2015}), we investigate the 
    \mstar{} completeness of our samples. 
    The \texttt{S82-MGC} sample matches the deeper SDSS photometric data in the 
    Stripe 82 region (\citealt{Annis2014}) with the near infrared data from the United 
    Kingdom Infrared Telescope Infrared Deep Sky Survey (UKIDSS; 
    \citealt{Lawrence2007}). 
    The deeper photometry and better photo-$z$s \rachel{Are `deeper' and `brighter' here compared to SDSS?  It isn't deeper than HSC, is it?  Should be clear what we are comparing here.} make the \texttt{S82-MGC} sample 
    complete to \logms{}$\geq 11.2$ at $z<0.7$ which makes it sufficient to evaluate 
    the completeness of our HSC sample.   
	By comparing with the Stripe 82 Massive Galaxy Catalog
    (\texttt{S82-MGC}; \citealt{Bundy2015}), \citet{Leauthaud2016} have measured the
    \mstar{} completeness of the combined BOSS and SDSS samples. 
    They estimate that the BOSS spec-$z$ sample, which is the main source of redshifts for our sample, is about 80\% complete at 
    \logms{}$\geq 11.6$ at $0.3 < z < 0.5$. The GAMA survey is 80\% complete down to 
    $10^{10.8}$\msun at $z{\sim} 0.3$, but is only 80\% complete to 
    $10^{12.0}$\msun{} at $z{\sim} 0.5$ according to \citet{Taylor2011} (e.g.\ 
    their Fig.~6).

    There are 20453 \texttt{S82-MGC} galaxies that are also in the \texttt{hscPho} 
    sample at $0.3 \leq z_{\mathrm{s82}} \leq 0.5$.  Because the \texttt{S82-MGC} uses  cmodel  magnitudes,  in this section we use  \mcmodel{} for consistency with the  \texttt{S82-MGC} catalog.  The \texttt{S82-MGC} also uses \texttt{iSEDFit} with similar assumptions as used in this work and we find excellent agreement between HSC  \mcmodel{} and the mass derived in  the \texttt{S82-MGC} which includes NIR data from UKIDSS;

    Figure~\ref{fig:mass_complete} compares the number density distributions of galaxies from \texttt{S82-MGC} with those from our sample\footnote{We do not apply any statistical corrections for completeness and hence we do not use the term ``stellar mass function'' to avoid confusion;
    Errors on the distributions are estimated via bootstrap resampling.}. 
    Based on Fig.~\ref{fig:mass_complete}, we conclude that our sample of massive 
    galaxies is reasonably complete down to \logmcmodel{}${\sim} 11.5$ at 
    $0.3 \leq z \leq 0.5$. 
    %The \rbcg{} sample is less affected by completeness because we can utilize 
    %photo-$z$s from \redm{}. 
    %However, the completeness of the \nbcg{} sample is still limited by the 
    %availability of spec-$z$s from BOSS.    
    Given the average difference between \mtot{} and \mcmodel{}, we will focus on galaxies with \logmtot{}$> 11.6$ where our sample shows good completeness. In our discussion section, we will also show results for massive galaxies with $11.4 \le$\logmtot{}$<11.6$ but we caution that our sample is incomplete in this  lower mass bin.
\rachel{I am a bit lost as to the primary source of this incompleteness.  Is it due to limitations inherited from CMASS, or something inherent to the data (no NIR) or procedure used to select the galaxies in this work?}

    %When comparing \nbcg{} and \rbcg{} we will also always make sure that the two 
    %samples are well matched both in terms of \mstar{} as well as in terms of redshift. 
    %Please see the next section and Appendix~\ref{app:match} for more details.

%% ------------------------------------------------------------------------------------ %%

%% ------------------------------------------------------------------------------------ %%  
\section{The Final Sample}
    \label{sec:final}

	With  \mstar{} estimates in hand, we now use the \redm{} galaxy cluster catalog to  create a sample of massive central galaxies. The ``central'' galaxy is defined as the galaxy that lives in the 
	center of its own dark matter halo.  
	In contrast, a galaxy in a sub-halo that is orbiting within the virial radius of a 
	more massive halo is referred to as a ``satellite''. 
	%The central galaxy has its unique importance in the study of galaxy evolution and 
	%galaxy-halo coevolution.
	To better understand the connection between the structures of massive 
	galaxies and the assembly history of both their stars and dark matter halos, we wish to focus on central 
	galaxies in this work.

\subsection{Candidate massive central galaxies}
    \label{ssec:redmapper}

    The robust identification of the central galaxy is not an easy task (e.g. \citealt{Yang2007}). In this paper, we use the \redm{} catalog create a sample of massive central galaxies with minimal contamination from satellite galaxies.
\rachel{What about those that are centrals in massive groups?  Also, the \redm{} catalog is not volume-limited to $z=0.5$, so what are the implications for this work?  It wasn't clear to me if that was part of the incompleteness shown in figure 4 since I was not sure if that was after the final selection or not.}

%    Given the smaller fraction of satellite galaxies at high-\mstar{} end (e.g. 
%    \citealt{Reid2014, Hoshino2015, Saito2016}), 
%    we further assume that the massive ones (\logmtot{}$>11.4$) among them are likely 
%    to be the central galaxies of these smaller halos.

    We use \texttt{v5.10} of the \redm{} cluster  catalog\footnote{See: http://risa.stanford.edu/redmapper/} 
    (e.g.\ \citealt{Rykoff2014}; \citealt{Rozo2015b}). 
    These authors have developed a well-tested red-sequence cluster finder that has 
    been run on SDSS DR8 (\citealt{SDSSDR8}) photometric data. 
    For each cluster, the catalog provides a photometric redshift ($z_{\lambda}$), a 
    cluster richness ($\lambda$), and identifies the most likely central galaxy (the 
    galaxy with the highest $P_{\mathrm{Cen}}$ value). 
    The \redm{}{} catalog also provides a list of member galaxies for each cluster and 
    their associated membership probabilities. 
    Details about the performance of the \redm{}~cluster catalog can be found in 
    \citet{Rozo2014}, \citet{Rozo2015a}, and \citet{Rozo2015b}.

    Several studies have published calibrations between the \redm{}~richness estimate, 
    $\lambda$, and halo mass (e.g.\ \citealt{Saro2015, Farahi2016, Simet2016, 
    Melchior2016}). 
    All these studies consistently find that \redm{} clusters with $\lambda > 20$ 
    have $\log (M_{\mathrm{halo}}/M_{\odot}) \geq 14.0$. 
    Therefore the central galaxies of these \redm{} clusters\footnote{We only use the 
    central galaxies with $P_{\mathrm{CEN}} \geq 0.7$} form a sample of massive central 
    galaxies that live in massive halos as described above.
\rachel{Intrinsic scatter in \mhalo{} at fixed $\lambda$ is pretty big, though, so we should be careful about this assertion.}
Such information is also very useful in studying the relationship between galaxy
    structure and environment (Huang\etal in prep.).
%% ------------------------------------------------------------------------------------ %%

%% ------------------------------------------------------------------------------------ %% 
  \begin{figure}
      \centering 
      \includegraphics[width=\columnwidth]{fig/redbcg_completeness}
      \caption{
          Evaluation of the \mstar{} completeness of the HSC massive galaxy sample.
          We compare the volume number density function of the massive galaxies 
          for this work (black line) with the one of a much more complete sample
          from the S82-MGC catalog (green line).  
          The associated uncertainties derived from bootstrap resampling are shown in 
          shaded regions. 
          The vertical grey line highlight the \logms{}$=11.6$ limit.  
          Below it, the HSC massive galaxy sample becomes significantly incomplete in 
          stellar mass. 
\rachel{Is this figure before or after the final cuts in section 5?  The figure is referenced in section 4, so I wasn't sure.}
}
      \label{fig:mass_complete}
  \end{figure}     
%% ------------------------------------------------------------------------------------ %%

%% ------------------------------------------------------------------------------------ %% 
%\subsection{Massive Central Galaxies in Low and High Mass Halos}
%    \label{ssec:mass_central}

    %One of the goals of our work is to investigate the halo dependence of the outer 
    %envelope. Although there is a correlation between $M_{\star}$ and $M_{\mathrm{Halo}}$,
    %based on recent constraints of $M_{\star}$-$M_{\mathrm{Halo}}$ relation 
    %(e.g.\ \citealt{Leauthaud2012}, \citealt{Behroozi2013}, \citealt{Kravtsov2014}), 
    %even though our sample selects very massive galaxies (\logms{}$\geq 11.5$), they 
    %With the help of the \redm{} catalog, we can broadly separate our sample into two
    %groups with \logmh{}$<14.0$ and \logmh{}$>14.2$.
    %live in a fairly broad range of halos masses.

    % We match this sample with the central galaxies of \redm{}~catalogs using a 
    % $1.0^{\arcsec}$ radius, and it results in 704 galaxies.  
    %% Saved as "dr1_redbcg_hsc_sdss_gama_1arcsec.fits"

    % TODO: Move this to Paper II
    %To ensure a clean separation in \mhalo{}, here we only focus on the central 
    %galaxies of clusters with $\lambda > 30$. 
    %According to \citet{Simet2016}, this $\lambda \geq 30$ sample should correspond to 
    %$M_{\mathrm{halo}}>10^{14.2} M_{\odot}$. 
    %We have checked that none of the results in this work depend on the choice of this 
    %$\lambda$ limit.

    %The median $\lambda$ of the clusters associated with the \rbcg{} sample is 
    %${\sim} 41$, corresponding to halo mass of 
    %$M_{\mathrm{halo}}{\sim} 2.2\times 10^{14}$ $h^{-1}$ $M_{\odot}$.
    %Only 44 centrals from the \rbcg{} sample live in clusters with $\lambda \geq 50$
    %($M_{\mathrm{halo}} {\sim} 3.0\times 10^{14}$ $h^{-1} M_{\odot}$). 
    %Our \rbcg{} sample therefore corresponds to galaxies in moderate mass halos.

    After matching the \texttt{hscZ} sample with the central galaxies of \redm{} 
    clusters with $\lambda \geq 20$ and $P_{\mathrm{0.7}} \geq 0.7$ \rachel{What is $P_{\mathrm{0.7}}$?  Does not seem to have been defined.}, we find 164 
    matched galaxies at $0.3 \leq z \leq 0.5$.
    This sample of \textbf{central galaxies in more massive halos} will be referred to 
    as the \rbcg{} sample. 
    Due to the depth and resolution of SDSS, the \redm{} catalog becomes slightly 
    incomplete at lower richness ($\lambda < 40$) at $z > 0.33$, but the main results 
    of this work are not affected by this.
    %% Saved as "dr1_redbcg_use_sed5b.fits"

    As the next step, we want to construct a sample of central galaxies living in halos 
    with \logmh{}$<14.0$. 
    To achieve this, we identify and remove all galaxies within a cylindrical region 
    around all \redm{} clusters. 
    We use a radius equal to $R_{\mathrm{200b}}$ and the thickness of the cylinder is 
    set to twice the value of the photometric redshift uncertainty of each cluster ().
\rachel{Need to fill in empty parenthesis in previous sentence.}

    We convert $\lambda$ of each cluster to $M_{\mathrm{halo}}$ using the calibration 
    of \citet{Simet2016} and we use the mass-concentration relation from 
    \citet{Diemer2015} to compute $R_{\mathrm{200b}}$. 
    At $0.3 < z < 0.5$, the uncertainty of photo-$z$ is between 0.015 to 0.025, and 
    is enough to exclude cluster members.

    After removing galaxies associated with \redm{} clusters, the remaining galaxies 
    in our sample will be dominated by central galaxies living in halos with 
    \logmh{}$< 14.0$. 
    We will refer to this sample of \textbf{galaxies in less 
    massive halos} as the \nbcg{} sample. Because we have rejected satellites in \redm{} clusters, and because above our galaxy mass cut most satellites reside in halos with $\log{(M_{halo})}>14$ (e.g. \citealt{Reid2014, 
    Hoshino2015, Saito2016, vanUitert2016}), galaxies in the \nbcg{} sample should  be dominated by central galaxies.
    \rachel{There is a statistical distribution of halo masses that satellites at this stellar mass live in, based on the papers you cited; can you say more quantitatively what fraction you expect to be satellites in less massive halos?  Right now it all sounds kind of hand-wavey.}

    %% Saved as "dr1_nonbcg_use_sed5b.fits"
%% ------------------------------------------------------------------------------------ %%

%% ------------------------------------------------------------------------------------ %% 
\subsection{Summary of Sample Construction}
    \label{ssec:sample}

    Using ${\sim} 100$ deg$^2$ of HSC data, we select a large sample of massive central 
    galaxies with reliable redshift information, and broadly separate them into two 
    categories based on $M_{\mathrm{halo}}$.

    The following is a summary of our sample construction. \rachel{Would be valuable to remind the reader in what sections these samples were defined, and how many galaxies are in each one.}

    \begin{itemize}
        \item \texttt{hscPho} sample: this parent sample consists of bright galaxies 
            with $i_{\mathrm{cModel}} \leq 21.0$, good quality imaging, and reliable 
            \texttt{cModel} photometry in all five HSC bands in the \texttt{S15B} 
            data release. 
        \item \texttt{hscZ} sample: we limit the \texttt{hscPho} sample to galaxies 
            with reliable redshift information. 
        \item With the help of the \redm{} cluster catalog, we further select 
            candidates of massive central galaxies in a range of \mhalo{}.
            We broadly divide the sample into central galaxies living in halos
            with \logmh{}$\geq 14.0$ (\rbcg{}) and \logmh{}$<14.0$ (\nbcg{}). 
    \end{itemize}

    The  division of our sample into two halo mass bins is mainly relevent for Paper II.  For this paper, we only consider the halo mass dependence on our sample when we evaluate impact of  mass estimates on the SMF in Section \ref{ssec:smf}.

\rachel{Could we show the redshift distribution of the second and third of these samples?}

%% ------------------------------------------------------------------------------------ %%

%% ------------------------------------------------------------------------------------ %% 
  \begin{figure*}
      \centering 
      \includegraphics[width=\textwidth]{fig/redbcg_mass_smf}
      \caption{
          \textbf{Left:} Impact of using \mtot{} on the galaxy stellar mass function. 
          Dashed lines correspond to the SMF computed using \mcmodel{} whereas solid 
          lines correspond to the SMF computed using \mtot{}. 
          Here we separate our HSC sample into  centrals in halos more massive than
          \logmh{}$\sim14.2$ (red lines) and centrals in halos with \logmh{}$<14.0$ 
          (black lines).
          The impact on the SMF can exceed 0.2 dex for massive central galaxies in 
          very massive halos.
          \textbf{Right}: The stellar mass volume density distributions of massive 
          HSC galaxies, using both \mtot{} (black solid line) and \mmax{} (black dotted 
          line).  
          The grey shaded region shows the resampling error on the HSC SMF plus an additional 20\%
          uncertainty to account for the fact that  we do not include satellite galaxies and that we fail to extract a 1-D profile for xxx\% of our galaxies. These issues will be adressed in forthcoming work. We compare our results with previous studies: 
          (a): SDSS galaxies at $z{\sim} 0.1$ from \citet{Bernardi2017} with \mstar{} 
          values based on photometry from 2-D \ser{}$+$Exponential model fitting 
          (purple); 
          (b): SDSS galaxies at $z{\sim} 0.1$ from \citet{Moustakas13} based on 
          improved SDSS \texttt{cModel} photometry (blue); 
          (c): \texttt{S82-MGC} galaxies at $0.15 < z< 0.43$ from 
          \citet{Leauthaud2016} based on PSF-matched SDSS-UKIDSS photometry (green).
          \rachel{Need to state explicitly what the vertical line is.  I assume this is the stellar mass below which the sample is presumed to be too incomplete for proper derivation of the SMF.  If that's true, should we even be showing masses below that?  Or perhaps the entire left side can be shaded, instead of just having a line there?  Also, are any completeness corrections applied at any value of stellar mass?}}
      \label{fig:smf}
  \end{figure*}
%% ------------------------------------------------------------------------------------ %%

%% ------------------------------------------------------------------------------------ %% 
\section{Results}
    \label{sec:result}

%% ------------------------------------------------------------------------------------ %%     
\subsection{Impact of Missing Light on the Galaxy Stellar Mass Function}
    \label{ssec:smf}

    The Stellar Mass Function (SMF) and its evolution are critical to our understanding of galaxy evolution. On average, the integration of our carefully derived non-parametric \mden{} profile 
    to 100 kpc (\mtot{}) recovers more light (hence mass) than the \texttt{cModel} method. 
    At \logmtot{} $>11.6$, the difference can sometimes be larger than 0.2 dex. 
    More importantly, the average difference steadily increases with \mtot{}. 
    These differences relate to the intrinsic limitation of the \texttt{cModel} 
    method and also reflects the fact that more massive ETGs tend to have more 
    extended stellar mass distributions (e.g. higher-\ser{} index, 
    \citealt{Graham2003}). Hence,  the determination of the ``total luminosities'' of massive galaxies can result in  a significant impact 
    on the high mass end of the galaxy stellar mass function (e.g. 
    \citealt{Bernardi2013, DSouza2014, DSouza2015, Bernardi2017}).

%    \footnote{To do that, one needs to 
%    separate the differences in the measurements of luminosity and \m2l{}, and 
%    take the uncertainties of \mstar{} measurements into account}

In this paper, we have only measured light profiles for central galaxies. Also,  we cannot extract  a 1-D profile for xxxxxx\% of galaxies (because contamination from neighbouring galaxies is too severe). These two effects will lead to a ${\sim}20$\% uncertainty in  the amplitude of our SMF.  \rachel{Mass-dependent, or really 20\% at all masses?} These are limitations that we are currently working to address and a forthcoming paper will  address them. Our goal in this paper is therefore not  to  attempt a detailed comparison between different SMFs
    \citep[e.g.,][]{Bernardi2013, Bernardi2017}. Instead, our goal in this section is more simply to characterize the  impact of various \mstar{} measurements  on the high-mass end of the SMF.

     Figure \ref{fig:smf} displays the impact of missing light and of different definitions of \mstar{} on the SMF. The left panel of Fig.~\ref{fig:smf} shows the SMF\footnote{Our SMFs turn over at the low mass end as we have not applied any completeness corrections here. \rachel{Why is this a footnote?  It's a major feature of the plot, not a minor detail.}} computed using \mcmodel{} and \mtot{} for galaxies that live in low and high mass halos.  Figure \ref{fig:smf} shows that the impact of missing flux is more severe for  galaxies in high mass halos. In Paper II it will be demonstrated that this occurs because  galaxies in more massive halos have larger extended  wings than those in low mass halos (Huang\etal in prep.). Figure \ref{fig:smf} also shows that  the use of the HSC \texttt{cModel} magnitude leads to a severe underestimation of the  high mass end of the SMF. 
\rachel{It would be useful if we can say whether this is primarily because of the adoption of simple parametric profiles, or due to background subtraction failures.  For example, if the adoption of simple parametric profiles is the primary cause of this difference, then that is a statement that might be more widely applicable than just the HSC pipeline.  If it's really just because our background subtraction algorithm is bad, then the implications are more limited.  Do we have the information that would allow us to make a statement about this?  If not, we should at minimum reiterate here the possible causes of this discrepancy.}

     %According to \citet{Wright2017}, the $z{\sim}0.1$ GAMA SMF based on a single-\ser{} model from \citet{Kelvin2012} suffers from similar issues.
    %appears to have the same issue (\citealt{Wright2017}),
    %    although the \ser{} model
    %also can overestimate the luminosity at high-mass end\footnote{Due to 
    %overestimated \ser{} index and/or effective radius.}.

        As discussed in Section~\ref{ssec:mtotal}, even with our deep HSC images,
    it is non-trivial to clearly define and measure the ``total'' \mstar{} for 
    these massive systems, and the \mden{} profiles of these galaxies often extend 
    well beyond 100 kpc. Is our fiducial 100 kpc radius sufficient in order to capture  most of the flux associated with these galaxies, or does one need to integrate out to even larger radii? To answer this question,  we attempt to capture more flux by integrating our \mden{} profiles out to the detection 
    limit that is defined by the background noise. We will refer to this mass as \mmax{}. Although the \mden{} profiles become more uncertain at $>100$ kpc, \mmax{} helps us quantify how much extra mass may be contained at radii greater than 100 kpc.  The right hand side of Figure \ref{fig:smf} shows the impact of using \mmax{} instead of  \mtot{} on the SMF.  We find that  a 100 kpc radius captures a majority of the galaxy mass and that the impact of going from \mtot{} to \mmax{} is relatively small

    In Figure \ref{fig:smf} we also compare our results  with the following previous studies:

    \begin{itemize}

        \item The SMF for $0.15 < z < 0.30$ galaxies from the \texttt{S82-MGC} sample
            \citep{Leauthaud2016} where \mstar{} is derived using 
            \texttt{iSEDfit} on PSF-corrected aperture photometry from 
            \texttt{S82}. They adopt the same IMF, dust model, and star-formation model as we do, but use the \texttt{BC03} model. Based on the test results from Appendix \ref{app:sed}, we shift the \citet{Leauthaud2016} SMF by 0.08 dex to add their data to Figure~\ref{fig:smf}.      
        \item The SMF derived from a SDSS-\textit{GALEX} sample of $z{\sim} 0.1$ 
            galaxies from \citet{Moustakas13}.  Total luminosities from \citet{Moustakas13} are based on SDSS \cmodel{} magnitude. 
            They use a \m2l{} that is derived via SED fitting using \texttt{iSEDfit} with similar 
            assumptions as here.
        \item The observed SMF for $z{\sim} 0.1$ SDSS galaxies from 
            \citet{Bernardi2017}.  Total luminosities in  \citet{Bernardi2017} are based on 2-D \texttt{SerExp} models 
            (\ser{}$+$ Exponential disk model; integrated to infinity) that 
            recover more flux compared to SDSS \cmodel{} magnitudes 
            (\citealt{Bernardi2013, Meert2015}). 
            They use a  \m2l{} that is adopted from the SED fitting results from the ``dusty'' 
            model in \citet{Mendel2014}.  \alexie{CLARIFY HERE SSp model}.   
    \end{itemize}

    As mentioned above, our current SMF is subject to an additional 20\% uncertainty that will be addressed in forthcoming work. Hence,  we do not attempt a detailed comparison  with respect to previous work on the galaxy SMF, reserving this  for  a future study. Here, we simply note that the  HSC SMF derived using \mtot{} is close to the one derived by \citet{Bernardi2017} using SDSS data at $z{\sim}0.1$ and assuming \texttt{SerExp} models.  Our  sample and the \citet{Bernardi2017} sample have different redshifts. Without a more in-depth study, we cannot ascertain wheter or not  the apparent agreement between the HSC SMF and the  \citet{Bernardi2017} SMF indicates a global lack of evolution in the galaxy SMF, or  that other effects are at play (difference between \mtot{} masses and \texttt{SerExp} masses) that have cancelled redshift evolution.  In future work, it will be interesting to perform a more consistent analysis to search for redshift evolution in the galaxy SMF and to study whether  or not massive galaxies in HSC are well described by \texttt{SerExp} models.

    % TODO: Move this to Paper II
    %In this work, we have opted to in general use \mtot{} rather than \mmax{} as the 
    %uncertainty on \mmax{} is larger. 
    %However, we have checked that none of our results are changed by replacing \mtot{} 
    %via \mmax{}.

%    The high-mass end of SMF is crucial for tracing the evolution of massive galaxies.  It is also  very sensitive to

%    And due to its sensitivity to physical process like the AGN-feedback, it is also 
%    important in calibrating simulation and semi-analytic model.

%    Here, using individual \mden{} profiles, we confirm that shallow images and/or
%    simplified models often lead to underestimated \mstar{} at high-mass end that 
%    can not be ignored.

%% ------------------------------------------------------------------------------------ %%

%% ------------------------------------------------------------------------------------ %%
\subsection{Surface Mass Density Profiles}
    \label{ssec:sbp_compare}

%% ------------------------------------------------------------------------------------ %% 
  \begin{figure*}
      \centering 
      \includegraphics[width=\textwidth]{fig/average_mass_profiles_fsps1_A}
      \caption{
          \textbf{Left}: Median \mden{} profiles in three total stellar mass bins. 
              Thin grey lines show a random subset of individual profiles.  The scatter between the thin grey lines reflects the true scatter in the profiles of massive galaxies (not measurement error). The shaded region highlights the region that could be affected by the PSF. \rachel{`most strongly' affected?}
              Two vertical lines indicate 10 kpc (thin, dotted line) and
              100 kpc (thick, dashed line). ~~~ 
          \textbf{Right}: comparison between our \mden{} profiles, previous observations, 
              and simulations. 
              The solid cyan line shows the median profile of massive elliptical 
              galaxies at $z{\sim} 0$ from \citet[][]{Huang2013a}. 
              The red long-dashed line shows the median profile of massive galaxies at 
              $0.25 \leq z < 0.50$ observed by \textit{HST} from \citet[][]{Patel2013}. 
              The purple short-dashed line shows the median radial stellar distributions 
              in massive halos from simulation using the particle tagging method
              (\citealt{Cooper2013}).}
      \label{fig:avg_prof}
  \end{figure*}
%% ------------------------------------------------------------------------------------ %%

%% ------------------------------------------------------------------------------------ %%
\subsubsection{General Trends and Comparison with Previous Work}
    \label{sssec:sbp_inter}

    Previous work on the structural evolution of massive galaxies has often focused on scaling relations such as the ``\mstar{}-size'' relation.  We argue that  comparing \mden{} 
    profiles  directly  captures more information than the \mstar{}-size relation and  has the advantage that it bypasses difficult questions 
    about how to accurately define and measure galaxy ``sizes'' and ``masses''.

    Fig.~\ref{fig:avg_prof} shows the median \mden{} profiles\footnote{The median 
    profiles along with their uncertainties are derived using bootstrap resampling. \rachel{Why is this a footnote?  It's a major feature of the plot, not a minor detail.  Same comment applies to footnote below this.}} of massive central galaxies at $0.3 < z < 0.5$ in three \mtot{} 
    bins\footnote{Note that our sample is not complete in our lowest \mtot{} bin.}. 
    As shown in the left panel of Fig~\ref{fig:avg_prof}, we can comfortably trace
    the \mden{} profiles of these massive galaxies out to 100 kpc 
    \textbf{individually}.  At large scales, some of our \mden{} profiles show signs of unphysical truncation and fluctuation 
    related to inaccurate sky subtraction. In this paper, we do not use profiles beyond 100 kpc, even though the median 
    \mden{} profiles for the two most massive bins behave reasonably out to 
    ${\sim} 200$ kpc.

       From Figure~\ref{fig:avg_prof} we can see  the galaxies in our sample have homogeneous profiles on small radial scales. The amplitude of \mden{} increases with galaxy mass on 10 kpc scales but the slope of  \mden{} remains similar.  From previous work on this topic, we already know that the inner regions of massive elliptical galaxies display relatively uniform structural (e.g. \mden{} profile, isophotal shape, 
    \citealt{Lauer07, Kormendy2009, Schombert2015} and kinematic (e.g. 
    \citealt{Cappellari13b}) properties.  However,  Figure~\ref{fig:avg_prof} reveals  a significant diversity in the outer envelopes of massive galaxies.  Given the \s2n{} of HSC images at these surface brightness levels, the scatter shown in Figure~\ref{fig:avg_prof} corresponds to \textbf{intrinsic scatter in the stellar envelopes of massive galaxies.}. Importantly, Figure~\ref{fig:avg_prof} shows that the global mass profiles of galaxies at these masses are  \textbf{clearly not 
    self-similar} out to 100 kpc.

In the right-hand side of Fig.~\ref{fig:avg_prof}, we compare our 
\mden{} profiles with results from previous  work. Deep \mden{} profiles of massive galaxies are rarely available. Even in the nearby universe, it is not trivial to map the low surface brightness  outskirts of massive galaxies (e.g., 
    \citealt{Capaccioli2015, Iodice2016, Iodice2017, Spavone2017, Mihos2017}). There are also not many galaxies that are this massive in the local universe. For example, the central galaxy of the COMA cluster, NCG xxx has a galaxy mass of xxxx and is smaller than any of the galaxies considered in this sample.

Most previous studies have often focused on surface brightness profiles instead of mass density profiles.  Results can also depend on the stacking technique or the model  used to extract the profile (e.g., \citealt{Tal2011, DSouza2015}). \citet{Huang2013a} derived \mden{} profiles for a small sample of very nearby 
    ellipticals (within 100 Mpc; median \logms{} ${\sim} 11.3$) based on relatively shallow 
    images from the Carnegie-Irvine Galaxy Survey (CGS, \citealt{CGS1}).  This sample is a very low redshift ($z=xxx$), and so the \citet{Huang2013a} galaxies are accurate to  smaller scales (down to $r=1$ kpc) than our HSC sample.  Our \mden{} profiles show good agreement with the \citet{Huang2013a} sample in the radial range of overlap (out to 50 kpc). CGS images are deeper than SDSS images in the $r$-band, but the median 
    profiles from \citet{Huang2013a} still only reach to ${\sim} 50$ kpc for $z{\sim} 0$ (PUT THE EXACT REDSHIFTS HERE)
    massive galaxies.
    Meanwhile, our deep HSC images can reliably deliver individual \mden{} profiles 
    for $z{\sim} 0.4$ galaxies out to at least 100 kpc.

    \citet{Patel2013} extracted a median \mden{} profile for massive ETGs at 
    $0.25 < z < 0.50$ using stacked \textit{HST}/ACS images. 
    These galaxies are selected at a constant cumulative number density and are 
    thought to be the progenitors of $z=0$ massive ETGs (e.g. \citealt{Leja2013}).  
    The median \mstar{} of the \citet{Patel2013} sample is 
    ${\sim} 10^{11.2} M_{\odot}$ which is lower than our lowest mass bin. 
    However, \citet{Patel2013} use the BC03 SSP model which provides masses that are roughly 0.1 dex 
    lower than our FSPS masses (Appendix~\ref{app:sed}). 
    Furthermore, the \citet{Patel2013} images are shallower than ours meaning 
    that their masses will be lower than ours because of missing light in the 
    outskirts. 
    Given these two considerations, it is reasonable to roughly compare the 
    \citet{Patel2013} profile with the one in our lowest stellar mass bin. 
    The superb resolution of the \textit{HST}/ACS image means that the profiles in
    \citet{Patel2013} can be measured down to 1 kpc. 
    The good agreement between our profiles and the ones derived from HST imaging demonstrates that our profiles are robust at $r > 3$ kpc and that we can accurately 
    measure \minn{}.

    Finally, we also compare with the predicted median \mden{} profile of central 
    galaxies in massive halos ($13.5 < \log M_{200,c} < 14.0$) from a cosmological 
    simulation where the \mden{} profiles of galaxies are calculated using the 
    particle-tagging technique (e.g., \citealt{Cooper2010}). 
    The simulated \mden{} profile is affected by the resolution limit of the simulation in the inner region, but  is in good agreement with our median \mden{} profile for 
    the $11.6 <$ \logmtot{} $< 11.8$ bin within 40 kpc. 
    However, outside 40 kpc, the particle tagging method seems to predict a too 
    prominent stellar halo and has much shallower outer slope compared to our data. In future work, we will compare  our HSC profiles with  predictions from more advanced 
    hydrodynamic simulations such as XXX AND XXXXX. These data will help fine-tune simulations and to understand the physical mechanisms that drive the assembly of massive galaxies and the 
    build up of stellar halos.

    Table~1 provides tabulated values for the median profiles that are 
    displayed in Fig.~\ref{fig:avg_prof}. 
    These profiles are also available  
    \href{https://github.com/dr-guangtou/hsc_redbcg/tree/master/profiles}{here}:
    {\url{https://github.com/dr_guangtou/hsc_redbcg/tree/master/profiles/}}.

%% ------------------------------------------------------------------------------------ %%

%% ------------------------------------------------------------------------------------ %% 
\subsection{Ellipticity and Color Profiles}
    \label{ssec:ell_color}

%% ------------------------------------------------------------------------------------ %% 
  \begin{figure*}
      \centering 
      \includegraphics[width=\textwidth]{fig/redbcg_ell_color.pdf}
      \caption{\alexie{Remake the light grey lines in this figure. Pick a smaller number. Say 100. And make the thin grey lines a little darker. Right now it is really hard to get anything useful from the think grey lines.}
          Radial profiles of the ellipticty and optical colors of massive galaxies in our sample. The general format of this figure is similar to Fig~\ref{fig:avg_prof}.  The
          \textbf{left panel} displays ellipticity profiles, 
          the \textbf{Upper-right panel} shows $g-r$ color profiles, and  the
          \textbf{lower-right panel} is for $g-i$ color profiles. 
          We compare our results with those from (1) \citet{Tal2011} based on stacking 
          large samples of luminous red galaxies in SDSS at $z{\sim} 0.4$ 
          (solid red line on the left panel), 
          (2) the results from a stacking analysis of nearby massive 
          galaxies with high concentration index ($C>2.6$) in 
          \citet[][blue dash lines on the left and upper-right panels]{DSouza2014}, 
          (3) and the average $g-r$ and $g-i$ color profiles 
          from a large sample of nearby elliptical galaxies in \citet[][blue, solid 
          lines on both right panels]{LaBarbera2010}.
          For color profiles, we also compare with deep observations of a few nearby 
          massive ETGs: the $g-r$ color profiles of NGC 4472 (\citealt{Mihos2013}) 
          and NGC 4365 (\citealt{Mihos2017}), and the $g-i$ profile of NGC 1399 
          (\citealt{Iodice2016}). \rachel{State explicitly that these are rest-frame colors.}
          }
      \label{fig:ell_color}
  \end{figure*}
%% ------------------------------------------------------------------------------------ %%

%% ------------------------------------------------------------------------------------ %% 
	%Aforementioned results only use the 1-D \mden{} profiles that are baed on single 
	%band image and the flux-weighted isophotal shape, while more useful information 
	%can be gathered from the HSC images of these massive galaxies.
    So far, we have focused only on 1-D \mden{} profiles. We now consider ellipticity and $k$-corrected optical color profiles.
	We extract ellipticity profiles using \texttt{Ellipse} by leaving the shape of each isophote as a free parameter. We also apply the isophotal information derived in the $i$-band directly to other 
	filters and extract 1-D $g-r$ and $g-i$ color profiles. We apply the Galactic extinction correction and the \texttt{iSEDFit} $k$-correction  to the color profiles. Smearing by the PSF wil l make the central isophotal shape rounder than the 
	real value, while seeing differences between filters will bias the central 
	color\footnote{Since the HSC $i$-band always has better seeing, the central color 
	will become redder is seeing effects are not accounted for}. On large scales, it is more difficult to extract reliable ellipticity and 
	color profiles out to 100 kpc: at low surface brightness levels, the isophotal shape 
	becomes unstable and is easily affected by contamination. Color profiles are also more difficult to extract out to large radial scales because  getting the color accurately depends on the background subtraction and the noise levels in both bands. In this paper, we will focus on the median ellipticity and color profiles between 
	8--60 kpc where we can safely ignore the issues described above.

	The ellipticity of the isophotes contains information about the 3-D geometry 
	(e.g. \citealt{Tremblay1995, Tremblay1996, Chang2013, RodriguezPadilla2013, 
	Mitsuda2017}) and kinematics (e.g. \citealt{Cappellari2012, Weijmans2014}) of 
	stars in massive galaxies.  
	The left panel of Fig \ref{fig:ell_color} shows the ellipticity profiles of HSC 
	massive galaxies and highlights the median profiles for the same three \mtot{} 
	bins used in Fig \ref{fig:avg_prof}.  Our results are also compared with previous work based on image stacking techniques (the PSF-removed $i$-band results from \citealt{Tal2011},  and
concentrated galaxies with $11.0<$\logms{}$<11.4$ from \citealt{DSouza2015}).
	As expected, ellipticity profiles from image stacking methods yield results that are more shallow than when  ellipticities can be measured on a galaxy-by-galaxy basis. Uncertainties in how to align galaxies, and the intrinsic isophotal twist can lead to this effect.

	In general, we find that the ellipticities of massive galaxies slowly increase with radius. This trend can even be seen directly "by eye" in HSC images (e.g. Fig \ref{fig:sdss_compare}). More interestingly, the ellipticity profiles vary with  \mtot{}: 
	at ${\sim} 10$ kpc, the median ellipticity ($< 0.2$) is similar for all three redshift bins,  but the ellipticity of the outer stellar halo increases with galaxy mass. Galaxies with  \logms{}$>11.8$ have median ellipticity profiles that become steeper at $>10$ kpc. The ellipticity of the outer profiles steadily increases from $e\le 0.2$ to $e{\sim} 0.4$ at 50-60 kpc.

	This is consistent with studies of nearby massive galaxies, both from 1-D 
	ellipticity profiles and 2-D modeling results \citep{Porter1991, Gonzalez2005, 
	Zibetti2005, Spavone2017; Huang2013a, Oh2017}. However,  to the best of our knowledge, our HSC results are the first to show  clear evidence that:  a) the ellipticities of the stellar halo  depend strong ly on galaxy mass, and b)  the ellipticity of the stellar halo is anti-correlated with the slope of \mden{}.

	If accreted stars dominate the stellar halos of massive galaxies, this 
	mass-dependent ellipticity profile may contain clues about the assembly history of massive galaxies (e.g. average time since last merger, 
	average merger mass ratio).
	Previous simulations often focused on reproducing the rounder average shape 
	and slow-rotating nature of massive ETGs (e.g. \citealt{Wu2014}), and the 
	projected ellipticity profiles and their correlations with the kinematic of 
	stars and merging history have not been well explored.

    Regarding the color profiles, it is well known that massive elliptical galaxies 
    have shallow and negative color gradients that reflect radial variations in their  
    stellar populations (e.g.\ \citealt{Carollo1993, LaBarbera2012}) and hence contain
    information about the assembly history of their stellar halos (e.g. 
    \citealt{Hirschmann2015}). 
    The right panels of Fig \ref{fig:ell_color} show the $k$-corrected $g-r$ and 
    $g-i$ color profiles for all galaxies in our sample, together with the median profiles in three \mtot{} bins. We find the the  median $g-r$ and $g-i$ color  decreases  at larger radii, but there does not appear to be an significant mass dependence in the  gradient of the color profiles. We also compare our results with the stacked color profiles from  \citealt{LaBarbera2010}\footnote{We use the median color profiles of high-mass 
    ETGs; the original profile is in units of $R_{\mathrm{e}}$, we use a typical 
    $R_{\mathrm{e}}=8.0$ kpc to convert it into physical kpc} and 
    \citealt{DSouza2014}.
    The median color profiles from HSC images are systematically steeper than the 
    stacked SDSS ones. 
    Considering differences in the response curves between HSC and SDSS filters, together with the uncertainties of color measurements, the HSC $g-r$ color profiles are  in failry good agreement with  those from SDSS. However, the $g-i$ profiles of HSC galaxies are steeper compared to SDSS. The SDSS $i$-band suffers from the so-called ``red-halo'' effect
    (e.g. \citealt{Wu2005}, \citealt{Tal2011}).  This is due to  the fact that the SDSS $i$-band PSF has more prominent wing than other bands. Because the PSF model does not capture these wings, this artificially distributes more flux to the outskirts, and leads to apparently redder colors in 
    the low surface brightness outskirts of galaxies. Because HSC uses thick CCDs, HSC $i$-band images do not suffer from this effect and can be used to determine galaxy colors with higher accuracy.

%    For the ones from \citet{LaBarbera2010}, uncertainty of $R_{\mathrm{e}}$ 
%    measurements could also ``smooth'' the color gradients.

    Fairly steep color profiles have been observed in several very nearby massive 
    ETGs.  Fig \ref{fig:ell_color} shows the $g-r$ color profiles of 
    NGC 4472 (\citealt{Mihos2013}) and NGC 4365 (\citealt{Mihos2017})\footnote{
    Both are converted from $B-V$ colors}, and the $g-i$ profile of NGC 1399 
    (\citealt{Iodice2016}). These individual profiles  display  similar color gradients as our HSC sample.

    % TODO: Move to discussion
    %Ellipticity profiles contain additional information about the assembly histories 
    %and in future work, it would be of interest to compare our ellipticity profiles 
    %with predictions from hydrodynamic simulations such as Illustris 
    %\citealt{Vogelsberger2014}, \citealt{Genel2014}) or \textit{Horizon-AGN} 
    %(\citealt{Dubois2014}).

    % TODO: Move this to Paper II 
    %The ellipticity profiles of the \mtot{}-matched samples are very similar at 
    %$r<50$ kpc (2-3$\times R_{\mathrm{e}}$ for their \mstar{}). 
    %However, for the \minn{}-matched samples, the \rbcg{} galaxies have higher 
    %ellipticity than the \nbcg{} ones at $r > 10$ kpc, while their inner ellipticity 
    %profiles are very similar.  
    %This further hints that massive central galaxies in different halos but with 
    %similar ``in situ'' mass may experience different late-time merging histories. 
    %Overall, the relative low ellipticity value and the fact that the ellipticity 
    %gradients are shallow lead us to conclude that the use of an average 
    %isophotal ellipticity is fine for this present study.
    %We show the average ellipticity, $g-r$, and $g-i$ color profiles for the 
    %\mtot{}-matched (\minn{}-matched) samples of \rbcg{} and \nbcg{} galaxies on the 
    %left (right) side of \ref{fig:ell_color}. 
    %We also compare our results with previous work based on stacking SDSS galaxies 
    %(e.g. \citealt{LaBarbera2010, Tal2011, DSouza2014}).

    %In terms of color gradients, we find shallow and negative gradients of median 
    %$g-r$ and $g-i$ color profiles for both \rbcg{} and \nbcg{} galaxies. 
    %For both \mtot{}- and \minn{}-matched samples, we find no significant difference
    %in their median color profiles, which supports the assumption of using single 
    %average \m2l{} value to estimate \mden{} profiles.
%% ------------------------------------------------------------------------------------ %%

%% ------------------------------------------------------------------------------------ %% 
  \begin{figure*}
      \centering 
      \includegraphics[width=\textwidth]{fig/redbcg_insitu_accretion}
      \caption{
          \textbf{Left:} Ratio of the fraction of stars between 10 to 100 kpc to the total galaxy mass \mtot{}. We adopt this ratio as a proxy for the fraction of ex-situ stars in our sample. Typical observational uncertainties are shown in the upper left hand corner. The solid grey line shows the predicted relation  derived from the Illustris simulation at $z=0$ 
          (Fig 4 in \citealt{RodriguezGomez2016}). Grey dashed lines correspond to the predicted 68 percentile of the distribution \alexie{CHECK}. \textbf{Right:} comparison between our median \mden{} profiles with the inner component of the structural decomposition of massive elliptical galaxies at $z{\sim} 0$ from \citet[][Cyan, solid]{Huang2013a}.  \alexie{A phrase about what we learn here}.  At high redshifts,  the profiles of massive galaxies  should be dominated by  the in-situ component.  We compare our profiles with the median \mden{} profile of massive galaxies at $1.0 < z < 1.5$ from \textit{HST} observations \citet[][Red, dashed]{Patel2013} and the \mden{} profile of a very massive cD galaxy at $z{\sim} 1.1$ discovered by \citet[][Yellow, dashed]{Liu2013} in the Hubble Ultra-Deep Field. \alexie{Always finish with a phrase saying what we learn from this figure. Are we going to say something about redshift evolution? What is the goal?}.}
      \label{fig:discussion_1}
  \end{figure*}
%% ------------------------------------------------------------------------------------ %%

%% ------------------------------------------------------------------------------------ %% 
\section{Discussion}
    \label{sec:discussion}

    In this paper, we have used  data from the HSC survey that is both simultaneously deep and wide to trace the stellar mass distributions of 
    $0.3 < z < 0.5$ massive galaxies out to $>100$ kpc and to reveal the mass-dependent nature 
    of their stellar halos.
    Here we briefly discuss the scientific implications of our results .

    % TODO: Move to Paper II
    %By using deep imaging from the HSC survey we have shown that the stellar mass 
    %density profiles of massive central galaxies depend systematically on the mass 
    %of their host dark matter halos.

%% ------------------------------------------------------------------------------------ %% 
\subsection{The Formation of Massive Galaxies and the Assembly of Their Outer Halos}
    \label{ssec:twophase}

    %The mass-dependent nature we find in the \mden{} and ellipticity profiles of 
    %massive galaxies are consistent with the two-phase formation scenario. 
    We find that the outer halos of massive elliptical galaxies grow more prominent and 
    more elliptical with increasing stellar mass.
    According to the two-phase formation scenario, the inner 5--10 kpc of these 
    massive central galaxies are formed at $z>1$ during an intense period of 
    \textit{in situ} star formation. 
    The outskirts of massive galaxies are then built up through a more gradual second 
    phase of evolution (the \textit{ex situ} phase) that is dominated by mass assembly via 
    accretion via mergers.
    Non-dissipative mergers, especially the minor ones, mostly deposit stars in the 
    outskirts of centrals and do not have a large impact on the central \mden{} 
    profile (e.g. \citealt{BoylanKolchin2008, Oogi2013, Bedorf2013}).
    Given the stochastic nature of the merging process, it is easy to understand why 
    the \mden{} profiles of massive galaxies are similar in the inner region but show
    larger scatter in the outskirt.

    State of the art hydrodynamic simulations of massive galaxy formation predict 
    that the fraction of accreted stars should strongly increase with stellar 
    mass and that for very massive galaxies, the  ``ex-situ'' may reach up to xxxxxxx \% of the total galaxy mass (e.g.
    \citealt{Oser2010, Cooper2013, Dubois2013, LeeYi2013, Hirschmann2015,
    RodriguezGomez2016}).

    This picture is supported by our observations that massive galaxies display more prominent stellar halos as well as by their  negative color gradients. The fact that the outskirts of these galaxies are slighly bluer than  the inner regions, is  consistent with the picture that  stellar halos are built up from  less massive galaxies \alexie{justify here. Why bluer?}.

    Also according to this picture, the shape of the stellar halo should preserve information
    about the merging history and possibly even about the shape of the dark matter halo. 
    Simulations show the shape of the stellar and dark matter halos are closely correlated (e.g. \citealt{Wu2014}) for slowly-rotating massive ETGs having undergone multiple 
    minor-mergers. The more elongated outer halo and the trend between the ellipticity profile and 
    stellar mass may reflect the orbital properties of infalling satellite. 
    In simulations, satellite orbits become more radial for more massive halos (e.g. \citealt{Murante2007, Wetzel2011, Jiang2015}. Satellites infalling along radial orbits may  help to form elongated  stellar halos. Hence, in principle, these results help explain the trend that we see. However, on contrast, simulation from \citet{Wu2014} predict that more massive 
    galaxies should have \emph{rounder} outskirts which is the opposite trend compared to HSC. This difference warrants further investigation.

    At larger scales, the distribution of satellites galaxies in massive halos is found to  be aligned with the major axis of the central galaxy \alexie{CHECK} (e.g. \citealt{Brainerd2005, 
    Yang2006b, NiedersteOstholt2010, HuangMandelbaum2016}). This alignement signal is found to be stronger for more massive \alexie{satellites???} galaxies (e.g. \citealt{Hirata2007}).
    The fact that more massive central galaxies have steeper ellipticity profiles 
    and become increasingly more elongated in the outskirt may arise because of such alignements. Moreover, the shape of the dark matter halo can be estimated by analyzing satellite distributions and weak lensing profiles (\citealt{ClampittJain2016}). 
    It is interesting to point out that the most recent measurement by 
    \citet{Shin2017} around SDSS clusters show a halo axis-ratio of ${\sim}0.55$, 
    which is only slightly more elongated than the stellar halo of galaxies in our 
    highest \mtot{} bin.

    % TODO: Move to paper II     
    %State of the art hydrodynamic simulations suggest that massive dark matter halos 
    %have central galaxies that have outer stellar envelopes with shallow slopes 
    %(e.g. \citealt{Pillepich2014}). 
    %This is consistent with the results in Fig \ref{fig:prof_1}.

    %Hence, according to this picture, the outer stellar envelopes of massive centrals 
    %should correlate with the masses of their host halos, in qualitative agreement with
    %our findings.

%% ------------------------------------------------------------------------------------ %% 
\subsection{Aperture Masses as Proxies of the \textit{In situ} and Accreted Stars}
    \label{ssec:insitu}

    Recent hydrodynamic simulations of galaxy formation often separate stars in 
    galaxies into \textit{in situ} and \textit{ex situ} components. 
    In simulations, the \textit{in situ} component is often defined as the stars formed 
    in the halo of the ``main progenitor''.  
    The spatial distribution, kinematic, and stellar population properties of both 
    components are important theoretical predictions. Among various predictions,  the faction of  \textit{ex situ} stars and the scaling of the  ex situ fraction with galaxy mass is perhaps the  first aspect to test. However, on the observational side, there is no strightforward way to actually disentangle these two components for massive galaxies. Disk galaxies  are more easily decomposed into multiple components and recently, deep surveys of nearby disk galaxies have started to provide constraints on their stellar halos \alexie{NEED REFERENCE HERE}.

    For more massive ETGs,  results so far mainly depend on multi-component model fitting
    and image stacking analyses (e.g. \citealt{Huang2013a, dSouza2014, Spavone2017}).
    In this work, we propose that \mstar{} computed within different fixed physical apertures are worth exploring as  proxies of the  ex situ fraction. We propose to use the mass within 10 kpc (\minn{}) and 100 kpc (\mtot{}) as proxies for the \textit{in situ} componnt and for the total \mstar{}.

    On the left panel of Fig \ref{fig:discussion_1} we show the relation between \mtot{} and the fraction of mass between 10 and 100 kpc (as a proxy of the mass of the accreted component). We compare this with the  fraction of ex-situ stars predicted by the Illustris simulation 
    (\citealt{RodriguezGomez2016}).  We find that  our proxy for the accreted mass component correlates strongly with  \mtot{} and that this relation is surprisingly consistent with predictions from \citet{RodriguezGomez2016}.  Given the limitations of the Illustris simulation, and the impperfect nature of our proxy for the ex sity fraction proxy, the almost perfect agreeement seen in Figure \ref{fig:discussion_1} may well be a coincidence.  However, it is encouraging to see that our simple proxy for the fraction of ex sity stars seems to match both the slope and the scatter of the predicted relation.  There exist other  hydrodynamic simulations that predict significantly different ex situ fractions compared to Illustris (e.g. 
    \citealt{Lackner2012, Qu2017}).  In future work, we will explore more detailed comparisons between our data and predictions from hydrodynamic simulations and we will  investigate to what degree simple aperture masses may be used to  trace \textit{ex situ} fractions.

    % TODO: Move to Paper II
    %Meanwhile, these simulation also point out that the fraction of accreted stars 
    %increases steeply with total \mstar{} and that the ``ex-situ'' component
    %starts to dominate the \mden{} profiles around $R_{\mathrm{e}}$ for 
    %\logms{}$\geq 11.5$ galaxies. 
    %In this work, we find that the median \mden{} profiles of \rbcg{} and \nbcg{} 
    %samples start to show differences around 15-20 kpc, which is close to 
    %their $R_{\mathrm{e}}$. 
    %Furthermore, the environmental dependence of our median \mden{} profiles 
    %becomes more significant in our highest \mtot{} bin (see Fig \ref{fig:prof_4}). 
    %These results all align with the picture that the \mhalo{}-dependence of the 
    %\mden{} profiles of massive galaxies is driven by the variations in the 
    %``ex-situ'' component at large radii.

   How well justified is our choice of  \minn{} as a proxy of the  in-situ component?  A full investigation is beyond the scope of this paper. Here we simply present a comparison  between our profiles and  a)  several estimates of the in-situ component and b) observations of high redshift massive galaxies that should be dominated by the  \textit{in situ} component.  In particular, we compare with:

    \begin{enumerate}

        \item The median \mden{} profiles of massive ETGs at $1.0 < z < 1.5$ from
            \citealt{Patel2013}. These are considered to be the progenitors of 
            ${\sim} 10^{11.5} M_{\odot}$ ETGs at $z=0$ and their inner region 
            should be dominated by \textit{in situ} stars.

        \item The inner component of $z{\sim} 0$ ellipticals from the 2-D 
            decomposition of \citet{Huang2013a}. \citet{Huang2013b} showed 
            that this inner component is structurally similar to the compact ``red nuggets'' at  high-$z$. 
        \item The \textit{in situ} components of simulated central galaxies in 
            massive halos from \citet{Cooper2013} (the inner ${\sim} 5$ kpc is 
            quite uncertain due to the resolution).  
            These \mden{} profiles are generated using the particle tagging 
            method (see \citealt{Cooper2010})

    \end{enumerate}

    We also compare with a uniquely massive BCG at high redshift: 
    a ${\sim} 10^{11.4} M_{\odot}$ BCG with a distinctive ``cD''-like envelope at 
    $z{\sim} 1.1$ (\citealt{Liu2013}).  This high redshift galaxy has a \mden{} profile that follows the median profile of our $11.6\leq$\logmtot{}$<11.8$ sample nicely inside 10-15 kpc, but becomes steeper in the outskirt. This suggest that the  inner ``core'' of massive BCGs are already in pace at $z{\sim} 1$ while the outer halo is still being assembled.

    These simple comparisons certainly support the idea that \minn{} mainly consists of 
    \textit{in situ} stars whereas mass at $\mathrm{R} > 15$-20 kpc is dominated by accreted 
    component.   Meanwhile, this comparison also shows that the \texttt{in situ} component may extend beyond 10 kpc\footnote{We convert these \mden{} profiles to the same 
    Chabrier IMF; but there are still differences in median \mstar{} and 
    details in the \m2l{} estimates}. Further comparisons with massive galaxies at high redshift and with hydrodynamicl simulations wil help understand which radius is optimal for probing the in situ mass.

%    In line with the prediction from Illustris, this comparison also 
%    suggests that, at very high-mass end, the accreted component can make significant 
%    contribution even within the inner 10--20 kpc. 
%    The nature of these ``accreted'' stars in the center of massive galaxies deserve 
%    more attention. 
%    It is possible that they mainly come from major merger and used to be the 
%    \texttt{in situ} component of another massive galaxy. 
%    It would be interesting see whether there is an alternative, but also physically 
%    motivated definition of \textit{in situ} component that can be more easily examined 
%    by observation.

    % TODO: Move to Paper II
    %More BCGs like this at $0.5 < z < 1.0$ can shed light on this intriguing 
    %process, and help us confirm whether the environmental dependence of structure 
    %has already emerged within massive galaxies at higher redshift 
    %(e.g. \citealt{Papovich2012})

%% ------------------------------------------------------------------------------------ %%

%% ------------------------------------------------------------------------------------ %% 
\subsection{Stellar Halo and Intra-cluster Light}

    Finally, we emphasize that in this work, we do not attempt  to  disentangle the galaxy light from any ``intra-cluster'' light component (ICL; e.g. 
    \citealt{Carlberg1997, Lin2004, Gonzalez2005, Mihos2005}). Instead, we adopt the view that the light of the main galaxy and the 
    ``intra-cluster''component share a smooth and continuous distribution.

    In practice, reliable isolation of a physically meaningful ICL component using 
    photometry alone is extremely difficult, if not impossible. 
    More importantly, the majority of massive galaxies in our samples are not considered 
    as the BCG of a cluster (\mhalo{}$\simgt 10^{14}M_{\star}$) \alexie{CHECK}. 
    Only a small fraction of the BCGs in the \rbcg{} sample belong to very 
    massive clusters where the strongest photometric evidence of an ICL component is 
    often found (\mhalo{}$\simgt 10^{15}M_{\star}$, e.g. \citealt{Presotto2014, 
    Burke2015}). The virgo cluster is a good  local analog of the typical galaxy in our sample.

    Ultra-deep images show that the stellar halo of the BCG smoothly extends beyond 
    100 kpc, while the ICL only becomes important outside that radius and is dominated 
    by many individual tidal features (e.g. \citealt{Mihos2016}).
    Therefore, it is highly unlikely that out main results are driven by the ICL component. \alexie{I'm not sure I agree with these statements. What is the ICL anyway?}

    At the imaging depth of HSC and out to 100 kpc, the \mden{} profiles of massive 
    galaxies vary smoothly and we find no evidence for a distinctive ICL component. 
    Although the rising stellar velocity dispersion in the outskirt of massive BCG
    has been used as evidence of a kinematically separated ICL component (e.g. 
    \citealt{Dressler1979, Carter1999, Kelson2002, Bender2015, Longobardi2015}), 
    we think it is reasonable to treat the stellar halo of BCG and potential ICL as 
    an integrated component. 
    They both carry important information regarding the assembly of central galaxy 
    and its dark matter halo. 
    In recent years, we can start to investigate the origin and evolution of ICL using 
    simulations where physical separation of ICL is possible
    (e.g. \citealt{SommerLarsen2005, Murante2007, Martel2012, Cui2014, Cooper2015, 
    Harris2017}). 
    It will be interesting to perform detailed comparisons between these simulations 
    and deep HSC observations to look into the contributions from different stellar 
    components to the halo of massive galaxy.

%% ------------------------------------------------------------------------------------ %%

%% ------------------------------------------------------------------------------------ %% 
\section{Summary and Conclusions}
    \label{sec:summary}

    In this work, we study how the stellar mass density profiles and other structural
    properties of massive central galaxies depend on their total stellar mass using 
    deep images from the Subaru HSC survey. 
    With the help of this high-quality and wide area data set, we directly map the 
    stellar mass distributions of ${\sim}7000$ massive central galaxies at 
    $0.3 < z < 0.5$ out to $>100$ kpc without resorting to stacking techniques. 
    We group massive central galaxies into two categories based on their host halo 
    mass (\mhalo{}$\simgt 10^{14.0} M_{\odot}$ and \mhalo{}$\simlt 10^{14} M_{\odot}$)
    and three bins of \mtot{}.  
    Our main results are:

    \begin{enumerate}
        \item We find that the ``total'' \mstar{} of these massive galaxies can be 
            significantly underestimated with shallow imaging data such as SDSS and/or 
            oversimplified model assumptions (e.g. the \texttt{cModel} or single-\ser{}). 
            In contrast to previous work, our results do not depend on stacking or any 
            parametric models. 
            Moreover, the degree to which mass is underestimated depends on \mtot{}. Simple model miss more light for massive galaxies because they have more extended enveloppes. There is also a \mhalo{}-dependence of this effect and naive luminosity estimates will  miss more light for BCGs is massive halos compared to centrals in less massive halos. These effects need to be carefully taken into account when discussing topics such as the evolution of the galaxy SMF.     
        \item We show that the \mden{} profiles of massive galaxies are relatively homogenous  within 10-20 kpc. However, there is a large but continuous scatter in outer profiles of massive galaxies. Galaxies with higher \mtot{} show more prominent stellar halos and have 
            shallower outer \mden{} slopes. 
            Assuming that stellar halos are dominated by accreted stars, this result is consistent with the two-phase formation picture of massive galaxies. 
        \item We show that, on average, massive galaxies have positive ellipticity 
            gradients out to at least 60 kpc. 
            The average ellipticity profile also depend on \mtot{}: more massive 
            galaxies tend to have steeper ellipticity gradients and become more 
            elongated in the accreted stellar halo. 
            On the other hand, the average $g-r$ and $g-i$ color gradients do not 
            show clear dependence on \mtot{} within 10-60 kpc.      
    \end{enumerate}

    These results highlight the advantages of wide area, deep, and high-quality imaging 
    for studying the evolution of massive galaxies. 
    At present, the HSC survey has already doubled its sky coverage to 
    ${\sim} 200$ deg$^2$, and provides a much larger sample of massive central galaxies. 
    In the near future, we will extend this work to lower \mtot{} by using photometric 
    redshifts, and we will also apply 2-D photometric methods (e.g.\ \citealt{Huang2013a}) 
    to take advantage of the multi-wavelength nature of the HSC survey 
    (e.g. \citealt{Huang2016}). 
    Our current work can also be combined with weak lensing measurements of the dark 
    matter halos of massive galaxies and physical insights into the assembly histories 
    of these galaxies can be gained by comparing with cosmological hydro-simulations 
    such as Illustris (\citealt{Vogelsberger2014}, \citealt{Genel2014}), 
    EAGLE (\citealt{Schaye2015}, \citealt{Crain2015}), or \textit{Horizon-AGN} 
    (\citealt{Dubois2014}).

%% ------------------------------------------------------------------------------------ %%

\section*{Acknowledgements}

  % Personal 
  SH thanks F.S.Liu for sharing the data from his work.

  % HSC part
  The Hyper Suprime-Cam (HSC) collaboration includes the astronomical communities of 
  Japan and Taiwan, and Princeton University.  The HSC instrumentation and software were
  developed by the National Astronomical Observatory of Japan (NAOJ), the Kavli Institute
  for the Physics and Mathematics of the Universe (Kavli IPMU), the University of Tokyo,
  the High Energy Accelerator Research Organization (KEK), the Academia Sinica Institute
  for Astronomy and Astrophysics in Taiwan (ASIAA), and Princeton University.  
  Funding was contributed by the FIRST program from Japanese Cabinet Office, the Ministry 
  of Education, Culture, Sports, Science and Technology (MEXT), the Japan Society for 
  the Promotion of Science (JSPS), Japan Science and Technology Agency (JST), the
  Toray Science Foundation, NAOJ, Kavli IPMU, KEK, ASIAA, and Princeton University.

  % SDSS part
  Funding for SDSS-III has been provided by the Alfred P. Sloan Foundation, the
  Participating Institutions, the National Science Foundation, and the U.S.  Department of
  Energy. The SDSS-III web site is http://www.sdss3.org.  SDSS-III is managed by the
  Astrophysical Research Consortium for the Participating Institutions of the SDSS-III
  Collaboration including the University of Arizona, the Brazilian Participation Group,
  Brookhaven National Laboratory, University of Cambridge, University of Florida, the
  French Participation Group, the German Participation Group, the Instituto de Astrofisica
  de Canarias, the Michigan State/Notre Dame/JINA Participation Group, Johns Hopkins
  University, Lawrence Berkeley National Laboratory, Max Planck Institute for
  Astrophysics, New Mexico State University, New York University, Ohio State University,
  Pennsylvania State University, University of Portsmouth, Princeton University, the
  Spanish Participation Group, University of Tokyo, University of Utah, Vanderbilt
  University, University of Virginia, University of Washington, and Yale University.

  % Pan-STARRS1 part
  The Pan-STARRS1 Surveys (PS1) have been made possible through contributions of the 
  Institute for Astronomy, the University of Hawaii, the Pan-STARRS Project Office, 
  the Max-Planck Society and its participating institutes, the Max Planck Institute 
  for Astronomy, Heidelberg and the Max Planck Institute for Extraterrestrial Physics, 
  Garching, The Johns Hopkins University, Durham University, the University of Edinburgh, 
  Queen's University Belfast, the Harvard-Smithsonian Center for Astrophysics, the Las 
  Cumbres Observatory Global Telescope Network Incorporated, the National Central 
  University of Taiwan, the Space Telescope Science Institute, the National Aeronautics 
  and Space Administration under Grant No. NNX08AR22G issued through the Planetary 
  Science Division of the NASA Science Mission Directorate, the National Science 
  Foundation under Grant No. AST-1238877, the University of Maryland, and Eotvos 
  Lorand University (ELTE).

  % LSST software
  This paper makes use of software developed for the Large Synoptic Survey 
  Telescope. We thank the LSST Project for making their code available as free 
  software at http://dm.lsstcorp.org.

  % Software
  This research made use of:
  \href{http://www.stsci.edu/institute/software_hardware/pyraf/stsci\_python}{\texttt{STSCI\_PYTHON}},
      a general astronomical data analysis infrastructure in Python. 
      \texttt{STSCI\_PYTHON} is a product of the Space Telescope Science Institute, 
      which is operated by AURA for NASA;
  \href{http://www.scipy.org/}{\texttt{SciPy}},
      an open source scientific tools for Python (\citealt{SciPy});
  \href{http://www.numpy.org/}{\texttt{NumPy}}, 
      a fundamental package for scientific computing with Python (\citealt{NumPy});
  \href{http://matplotlib.org/}{\texttt{Matplotlib}}, 
      a 2-D plotting library for Python (\citealt{Matplotlib});
  \href{http://www.astropy.org/}{\texttt{Astropy}}, a community-developed 
      core Python package for Astronomy (\citealt{AstroPy}); 
  \href{http://scikit-learn.org/stable/index.html}{\texttt{scikit-learn}},
      a machine-learning library in Python (\citealt{scikit-learn}); 
  \href{http://www.astroml.org/}{\texttt{astroML}}, 
      a machine learning library for astrophysics (\citealt{astroML});
  \href{https://ipython.org}{\texttt{IPython}}, 
      an interactive computing system for Python (\citealt{IPython});
  \href{https://github.com/kbarbary/sep}{\texttt{sep}} 
      Source Extraction and Photometry in Python (\citealt{PythonSEP});
  \href{https://jiffyclub.github.io/palettable/}{\texttt{palettable}},
      color palettes for Python;
  \href{http://dan.iel.fm/emcee/current/}{\texttt{emcee}}, 
      Seriously Kick-Ass MCMC in Python;
  \href{http://bdiemer.bitbucket.org/}{\texttt{Colossus}}, 
      COsmology, haLO and large-Scale StrUcture toolS (\citealt{Colossus}).

%%%%%%%%%%: Bibliographic Section %%%%%%%%%%

\bibliographystyle{mnras}
\bibliography{redbcg}

%%%%%%%%%%: Appendix Section %%%%%%%%%%%%

\appendix

%% ------------------------------------------------------------------------------------ %% 
\section{A. Extraction of 1-D surface brightness profile} 
    \label{app:ellipse}

    Here we briefly discuss a few technical issues related to the measurements of the 
    1-D surface brightness profiles around massive galaxies.

    To derive reliable 1-D profile, it is important to mask out all the irrelevant 
    objects around the target.
    At the depth of the HSC images, this becomes a challenging task, especially 
    for massive galaxies with extended outer profiles and many satellites. 
    At this point, the \texttt{hscPipe} tends to over-subtract the background around 
    bright objects.  
    The performance of its deblending process is also not optimized for extended
    objects. 
    For these reasons, we perform \texttt{SExtractor}-like background subtraction and 
    object detection using the \texttt{SEP} Python library to generate the necessary 
    masks.
    Combining two different local background models and $S/N$ thresholds, we obtain 
    the centroid, shape, and radius that enclose 90\% of flux for each object, 
    including the one that is very close to the center of bright galaxy (left panel of 
    Fig~\ref{fig:ell_tech}). 
    Based on these information, we then create the mask that covers all contaminating 
    objects around the target after adaptively increasing the sizes of their masks 
    according to their brightness and distance to the central target. 
    Generally speaking, we mask out bright objects or objects in the outskirt of the 
    image more aggressively to reduce their impact on the surface brightness profiles 
    in the outskirt. 
    We also create masks that are less and more aggressive than the default one to 
    test their impacts on the surface brightness profiles.

    Next, we aggressively mask out all objects on the cut-out image.  
    We then evaluate the background level using the unmasked pixels after median 
    smoothing the masked image using box of $6x6$ pixels.
    This provides estimate of global background level along with its uncertainty. 
    Given the typical background uncertainty, the HSC \texttt{WIDE} image should be 
    able to reach down to $> 29$ \sb surface brightness level in the $i$-band.  
    However, as mentioned, we often find evidence of slightly over-subtracted 
    background for massive galaxies in our sample. 
    In the current \texttt{hscPipe}, the background on each CCD is modeled with a 
    Chebyshev-polynomial that is fit to the smoothed image after excluding pixels 
    with $S/N >5$.
    This algorithm performs much better than the SDSS version 
    (e.g.\ see \citealt{Blanton2011}), yet still over-subtracts background around 
    bright objects and results in unphysical truncation in their surface brightness 
    profiles.
    We empirically correct this issue using the background model generated by 
    the \texttt{SExtractor} algorithm on the masked image 
    ($200x200$ pixels background box size, and 6 pixels median filtering size of 
    sky boxes).
    This model can account for the slightly over-subtracted background at large scale,
    and reduce the impact from the low surface brightness ``wings'' of bright 
    neighbors. 
    We clearly see improvement in both the distributions of background pixels 
    (more symmetric distribution; median value is closer to 0) and the surface 
    brightness profile (middle panel of Fig~\ref{app:ellipse}; the negative intensity 
    and the turn-over of the curve-of-growth in the outskirt of the ``Original'' 
    profile are successfully corrected) after this correction.
    Also, it is worth mentioning that such correction does not often affect the 
    surface brightness profile within 100 kpc.

    The procedure used to derive 1-D surface brightness profile from the 
    background-corrected, contamination-masked images is already described in 
    Section~\ref{sec:ellipse} briefly. 
    In practice, the profile at very low surface brightness level is sensitive to 
    several \texttt{Ellipse} configurations.
    After some tests, we choose to use 0.1 dex in logarithm as the step in semi-major 
    axis length between successive ellipses, and we use the median pixel value over the
    elliptical annulus after rejecting outlying pixels via $3\sigma$-clipping three
    times.
    We make the above choices to make the final profile less affected by any nearby
    object, and also test the differences between the profiles derived using larger
    step, or mean value on the annulus, or fewer times of $\sigma$-clipping. 
    Generally speaking, the surface brightness profile is very robust against these
    changes, especially within 100 kpc. 
    On the right panel of Fig~\ref{fig:ell_tech}, we compare the surface brightness
    profiles for an example massive galaxy using different masks and \texttt{Ellipse}
    parameters. 
    The profile within 100 kpc is very stable, and the only noticable difference 
    is caused by the less aggressive object-mask in the very outskirt.

    We should also mention that we run \texttt{Ellipse} allowing for more 
    sophisticated shapes than simple ellipse (4th Fourier modes that can make 
    isophote more ``disky'' or ``boxy'', e.g.\ \citealt{Kormendy2009}) to fit the 
    isophote better.
    We also apply the isophotes from $i$-band images to other bands in 
    ``force-photometry'' mode \texttt{Ellipse} run to get initial estimates of 
    color profiles.  
%% ------------------------------------------------------------------------------------ %%

%% ------------------------------------------------------------------------------------ %% 
    \begin{figure*}
        \centering 
        \includegraphics[width=\textwidth]{fig/redbcg_ellipse_tech}
        \caption{
            \textbf{Left:} Example of the object-mask built for the \texttt{Ellipse}
            run for a typical massive galaxy in the sample. 
            All the shaded regions are masked out. 
            The three dash lines (red, inner one and two blue ones) around the target 
            at the center outlines the three radius we defined using the flux radius 
            of the target.  
            We increase the mask size for objects detected in different regions 
            separated by these apertures (which are outlined by solid, elliptical 
            apertures with different colors) using slightly different criteria.~~
            \textbf{Middle:}: The zoom-in intensity profile around very low intensity 
            value (top panel), and the curve-of-growth of the enclosed magnitude 
            (bottom panel) of the example galaxy.  
            To highlight the importance of background correction, we show the profiles 
            using both images with (red, solid line) and without (black, dotted line) 
            background correction. 
            On the top panel, besides the horizontal line that highlights the zero flux 
            level, we also show the uncertainty of the sky background estimate using 
            the grey-shaded region.  
            On the bottom panel, two horizontal lines indicate the magnitudes 
            corresponding to total flux (solid) and flux within 100 kpc (dash).~~
            \textbf{Right:} compares the 1-D surface brightness profiles for the same 
            example galaxy using different masks 
            (smaller masking region: red, dash line; larger masks: blue, dash line), 
            or different \texttt{Ellipse} configurations
            (more aggressive pixel-clipping: cyan, dash line; 
             larger step in radius: green, dash line; 
             using mean flux along the isophote instead of median: purple, dash line)
            with the default one (black, solid line).
            }
        \label{fig:ell_tech}
    \end{figure*}
%% ------------------------------------------------------------------------------------ %%

%% ------------------------------------------------------------------------------------ %% 
\section{B. Estimate average {$M_{\star}/L_{\star}$} using \texttt{iSEDFit}} 
    \label{app:sed}

%% ------------------------------------------------------------------------------------ %% 
    \begin{figure*}
        \begin{center}
        \includegraphics[width=\textwidth]{fig/redbcg_isedfit_example.pdf}
        \caption{
            \textbf{Left:} Example of output figure from \texttt{iSEDFit} that shows 
            the SED fitting results. 
            The open-boxes show the observed fluxes in 5-band, and the solid, blue-dots
            show the best-fitted results, along with the high-resolution spectrum for
            this model reconstructed using the synthetic spectra from \texttt{FSPS}. 
            Top-left corner shows the best-fit stellar population parameters, and 
            bottom-right corner shows the ID, redshift of this object, and reduced 
            $\chi^2$ of the best-fit model.~~~
            \textbf{Right:} the Posterior distributions of a few key parameters.
            From top-left to bottom right are: 
            1) stellar mass (\logms{}); 
            2) age of the population (mass and star-formation rate weighted) in Gyr; 
            3) star formation rate ($\log\ \mathrm{SFR}\ (M_{\odot}/\mathrm{yr})$; 
            instant one and the one averaged over the previous 100 Myr; 
            4) stellar metallicity ($\mathrm{Z}/\mathrm{Z}_{\odot}$); 
            5) dust extinction ($\mathrm{A}_V$ in mag);
            6) birthrate parameter ($\log\ b$; averaged over previous 100 and 1000 Myr).
            }
        \label{fig:ised}
        \end{center}
    \end{figure*}
%% ------------------------------------------------------------------------------------ %%

%% ------------------------------------------------------------------------------------ %% 
    \begin{figure*}
        \begin{center}
        \includegraphics[width=12cm]{fig/redbcg_isedfit_2.pdf}
        \caption{
            Relationships between \mstar{} and key stellar population parameters from 
            \texttt{iSEDfit}. 
            The four stellar population properties are: 
            1) \textbf{Top-left}: \mstar{}-weighted stellar population age in Gyr; 
            2) \textbf{Bottom-left}: SFR-weighted age in Gyr; 
            3) \textbf{Top-right}: \mstar{}-weighted stellar metallicity in unit of 
            Solar value;
            4) \textbf{bottom-right}: dust extinction value in $V$-band.
            As expected, most of the HSC massive galaxies are old, metal-rich, and 
            dust-free. 
            }
        \label{fig:ised_2}
        \end{center}
    \end{figure*}
%% ------------------------------------------------------------------------------------ %%

%% ------------------------------------------------------------------------------------ %% 
    \begin{figure}
        \begin{center}
        \includegraphics[width=\columnwidth]{fig/redbcg_isedfit_3.pdf}
        \caption{
            Comparisons of \mstar{} estimated by \texttt{iSEDFit} using different
            model assumptions. 
            In each figure, we plot the \mstar{} from the default model against the
            differences with other models. 
            The four models involved are labeled as: 
            (A): Default model; 
            (B): Using the Salpeter IMF instead of the Chabrier one 
                (\textbf{Top panel});
            (C): Using the BC03 synthetic population model instead of the FSPS one
                (\textbf{Middle panel});
            (D): No dust extinction (\textbf{Bottom panel}). 
            On each panel, the grey shaded region highlights the typical uncertainty 
            of the $\log($\logms{}$)$.            
            For each pair of models, we highlight their median $\chi^{2}$ values and 
            the fraction of galaxies with $\chi^{2} > 10.0$ at the top. 
            On each panel, we also show the histograms of the \mstar{}-differences on 
            the right side.
            }
        \label{fig:ised_3}
        \end{center}
    \end{figure}
%% ------------------------------------------------------------------------------------ %%

%% ------------------------------------------------------------------------------------ %%

    In Section~\ref{ssec:isedfit}, we briefly explain the SED fitting procedure and 
    the priors used.   
    In Fig~\ref{fig:ised_1}, we show an example of the \texttt{iSEDFit} output by 
    visualizing the 5-band HSC SED on top of the best-fit model along with the PDF 
    of the key parameters.

    Although we only use the best-fit \m2l{} in this work, it is necessary to make 
    sure the model is reasonable. 
    We show the relations between \mstar{} and a few key stellar population parameters 
    derived by \texttt{iSEDFit} in Fig~\ref{fig:ised_2}. 
    Degeneracies among these parameters are inevitable based on only five broad-band
    photometry, but as expected, most massive galaxies show old stellar age, high
    stellar metallicity ($1.5 \times Z_{\odot}$ is the highest metallicity allowed by 
    the adopted \texttt{FSPS} SSP models), and low dust extinction.

    Meanwhile, \mstar{} measurement based on SED fitting heavily depends on the 
    adopted SSP model, the form of IMF, dust extinction law, and details in 
    the assumption of SFH (e.g.\ \citealt{Bernardi2017}). 
    For massive galaxies in this sample, the form of the SFH\footnote{We choose 
    to use the delayed-$\tau$ model for SFH; we adopt flat distribution between 
    0.5 to 14.0 Gyrs as the prior for the look-back time when the star formation 
    turned on. 
    The exponential delayed time-scale ($\tau$) is allowed to change between 
    0.1 to 3.0 with equal probability}, the contribution from random star 
    burst\footnote{The chance of random star burst is set at 0.2 for every 2 Gyrs. 
    The duration of the star burst is draw from a logarithmic distribution 
    between 0.03 to 0.3 Gyr; and the mass fraction formed in the burst is from 
    a logarithmic distribution between 0.01 and 1.0.} rarely affect the \mstar{}. 
    But he choices of SSP model, IMF, and dust extinction do systematically impact
    the estimates of \mstar{}, and therefore we look into this with a few 
    additional tests (see Fig~\ref{fig:ised_3}):

    \begin{enumerate}

        \item Choosing the \citet{Salpeter1955} IMF results in systematically 
            higher \mstar{} (on average $+0.25$ dex of \logms{}) for these massive 
            galaxies (top panel).
            Although there are multiple lines of evidence that favor Salpeter 
            or even more ``bottom-heavy'' IMF in massive galaxies 
            (e.g.\ \citealt{Conroy2012}; \citealt{Cappellari2012}), we still 
            present the main results using Chabrier IMF to accommodate galaxies 
            with lower \mstar{} in the sample, and to be as consistent as possible 
            with previous works.  
            This choice of IMF does not change the main results qualitatively.

        \item \mstar{} based on the \texttt{BC03} models are systematically lower 
            than the ones based on \texttt{FSPS+MILES} models (middle panel)
            The difference shows a large scatter, and can be as large as 0.4 dex,
            although it is not \mstar{}-dependent. 
            The \texttt{BC03} results show better average ${\chi}^2$ than the 
            \texttt{FSPS} ones. 
            This relates to the higher upper-limit of stellar metallicity 
            ($2 \times \mathrm{Z}_{\odot}$) allowed by the \texttt{BC03} model, 
            which help fits the shape of the SED in the red-end slightly better.  
            However, the \texttt{BC03} results also show puzzlingly low stellar 
            ages ($< 3$--4 Gyrs) for these massive, red galaxies. 
            This could also lead to underestimated \m2l{} values.
            It is worth noting that, both \texttt{FSPS} and \texttt{BC03} 
            models still have difficulties recovering SED at the very red-end 
            (between $z$ and $y$-band), and reproducing the optical color-color 
            relations for red-sequence galaxies (e.g.\ \citealt{MIUSCAT2}).
            In this work, we decide to keep using the \texttt{FSPS+MILES} model as 
            the fiducial one.  
            Using results based on \texttt{BC03} model will not change any of our 
            conclusions here.

        \item On the bottom panel of Fig~\ref{fig:ised_3}, we compare with the
            SED fitting results without considering the dust extinction. 
            This choice leads to slightly smaller \mstar{} values as expected. 
            Its impact becomes slightly larger at lower \mstar{} end. 
            It will not change any of our conclusions here.

    \end{enumerate}
%% ------------------------------------------------------------------------------------ %%

%% ------------------------------------------------------------------------------------ %% 
%% Table.1 
%% ------------------------------------------------------------------------------------ %% 
\clearpage
\include{table1}
\clearpage
%% ------------------------------------------------------------------------------------ %%

%% ------------------------------------------------------------------------------------ %%  
\bsp
\label{lastpage}
\end{document}
%% ------------------------------------------------------------------------------------ %% 
%%%%%%%%%%%%: End of the File %%%%%%%%%%%%
%% ------------------------------------------------------------------------------------ %%