### 2017-07-26 draft/paper2_v3.tex
Edit introduction

% Version 0; preprint format; Outline of paper II by Song Huang
% Version 1; preprint format; Edited by Jenny Greene
% Version 2; preprint format; Edited by Song Huang
% Version 3; preprint format; Edited by Alexie Leauthaud

\documentclass[a4paper,fleqn,usenatbib]{mnras}

% Packages
\usepackage{deluxetable}
\usepackage{newtxtext,newtxmath}
\usepackage[T1]{fontenc}
\usepackage{ae,aecompl}
\usepackage{amssymb, amsmath}
\usepackage{graphicx}
\usepackage{natbib}
\usepackage{url}
\usepackage{hyperref}
\usepackage{float}
\usepackage[usenames, dvipsnames]{color}

% Package Settings
\hypersetup{colorlinks=true,
            citecolor=cyan,
            linkcolor=cyan,
            filecolor=magenta,      
            urlcolor=cyan}
\urlstyle{same}

% Figure extention
\DeclareGraphicsExtensions{.pdf,.png,.jpg}

%%%%%%%%%%%%: User Defined Commands %%%%%%%%%%%%

% Song Huang's definition 
\def\arcsec{{\prime\prime}}
\def\arcmin{{\prime}}
\def\degree{{\circ}}
\def\h{\hskip -3 mm}
\def\aa{{A\&A}}
\def\aas{{ A\&AS}}
\def\aj{{AJ}}
\def\al{$\alpha$}
\def\bet{$\beta$}
\def\amin{$^\prime$}
\def\annrev{{ARA\&A}}
\def\apj{{ApJ}}
\def\apjs{{ApJS}}
\def\asec{$^{\prime\prime}$}
\def\deg{$^{\circ}$}
\def\ddeg{{\rlap.}$^{\circ}$}
\def\dsec{{\rlap.}$^{\prime\prime}$}
\def\cc{cm$^{-3}$}
\def\flamb{erg s$^{-1}$ cm$^{-2}$ \AA$^{-1}$}
\def\flux{erg s$^{-1}$ cm$^{-2}$}
\def\fnu{erg s$^{-1}$ cm$^{-2}$ Hz$^{-1}$}
\def\hst{{\textit{HST}}}
\def\kms{km s$^{-1}$}
\def\lamb{$\lambda$}
\def\lax{{$\mathrel{\hbox{\rlap{\hbox{\lower4pt\hbox{${\sim}$}}}\hbox{$<$}}}$}}
\def\gax{{$\mathrel{\hbox{\rlap{\hbox{\lower4pt\hbox{${\sim}$}}}\hbox{$>$}}}$}}
\def\simlt{\lower.5ex\hbox{$\; \buildrel < \over {\sim} \;$}}
\def\simgt{\lower.5ex\hbox{$\; \buildrel > \over {\sim} \;$}}
\def\micron{{$\mu$m}}
\def\mnras{{MNRAS}}
\def\nat{{Nature}}
\def\pasp{{PASP}}
\def\perang{\AA$^{-1}$}
\def\peryr{yr$^{-1}$}
\def\reference{\noindent\pp}
\def\refindent{\par\noindent\parskip=2pt\hangindent=3pc\hangafter=1 }
\def\sb{mag~arcsec$^{-2}$}
\def\lsun{$L_\odot$} 
\def\msun{$M_\odot$}
\def\sigs{$\sigma_*$}
\newcommand{\lt}{<}
\newcommand{\gt}{>}

\def\etal{{\ et al.~}}
\def\galfit{{\tt GALFIT}}
\def\ser{{S\'{e}rsic\ }}
\def\redm{\texttt{redMaPPer}}
\def\cmodel{\texttt{cModel}}
% Samples
\def\rbcg{\texttt{cenHighMh}}
\def\nbcg{\texttt{cenLowMh}}
\def\redbcg{{$\lambda \ge 30$}}
\def\nonbcg{{$\lambda < 20$}}
% Mass related 
\def\mstar{{$M_{\star}$}}
\def\mhalo{{$M_{\mathrm{200b}}$}}
\def\logms{{$\log (M_{\star}/M_{\odot})$}}
\def\logmh{{$\log (M_{\mathrm{200b}}/M_{\odot})$}}
\def\logmhalo{{$\log (M_{\mathrm{200b}}/M_{\odot})$}}

\def\minn{{$M_{\star,10\mathrm{kpc}}$}}
\def\meff{{$M_{\star,15\mathrm{kpc}}$}} 
\def\mtot{{$M_{\star,100\mathrm{kpc}}$}}
\def\mout{{$M_{\star,150\mathrm{kpc}}$}}
\def\mmax{{$M_{\star,\mathrm{Max}}$}}
\def\mgama{{$M_{\star,\mathrm{GAMA}}$}}
\def\mcmodel{{$M_{\star,\mathrm{cModel}}$}}

\def\logminn{{$\log (M_{\star,10\mathrm{kpc}}/M_{\odot})$}}
\def\logmtot{{$\log (M_{\star,100\mathrm{kpc}}/M_{\odot})$}}
\def\logmout{{$\log (M_{\star,150\mathrm{kpc}}/M_{\odot})$}}
\def\logmmax{{$\log (M_{\star,\mathrm{Max}}/M_{\odot})$}}
\def\logmgama{{$\log (M_{\star,\mathrm{GAMA}}/M_{\odot})$}}
\def\logmcmodel{{$\log (M_{\star,\mathrm{cModel}}/M_{\odot})$}}

\def\m2l{{$M_{\star}/L_{\star}$}}
\def\s2n{{$\mathrm{S}/\mathrm{N}$}}
\def\mden{{$\mu_{\star}$}}

\def\insitu{{\textit{in situ}}}
\def\exsitu{{\textit{ex situ}}}

% Commenting:
\newcommand{\xxx}[1]{\textcolor{red}{\textbf{XXX}}}
\newcommand{\todo}[1]{\textcolor{red}{\textbf{TODO:~#1}}}
\newcommand{\plan}[1]{\textcolor{cyan}{#1}}
\newcommand{\addref}{{\textcolor{red}{REF}}}
\newcommand{\note}[2]{\textcolor{blue}{\textbf{[Comment (#1): #2]}}}
\newcommand{\song}[1]{\textcolor{magenta}{\textbf{[Song: #1]}}}
\newcommand{\alexie}[1]{\textcolor{blue}{\textbf{[Alexie: #1]}}}
\newcommand{\jenny}[1]{\textcolor{Bittersweet}{\textbf{[Jenny: #1]}}}
\newcommand{\kevin}[1]{\textcolor{green}{\textbf{[Kevin: #1]}}}
\newcommand{\update}[1]{\textcolor{PineGreen}{#1}}

%% ------------------------------------------------------------------------------------ %% 
%% Title and Affiliations 
%% ------------------------------------------------------------------------------------ %%

%\title[Structure and Environment of Massive Galaxies]{
%       A Detection of the Environmental Dependence of the Stelllar and haloes
%       of Massive Central Galaxies}

\title[Structure and Environment of Massive Galaxies]{
       A Detection of the Environmental Dependence of the Sizes and Stellar Haloes
       of Massive Central Galaxies}

\author[S. Huang et al.]{
        Song Huang$^{1,2}$\thanks{E-mail: song.huang@ipmu.jp (SH)},
        Alexie Leauthaud$^{1}$,
        Jenny Greene$^{4}$,
        Kevin Bundy$^{3}$,
        \newauthor
        Yen-Ting Lin$^{5}$,
        Masayuki Tanaka$^{6}$,
        Rachel Mandelbaum$^{7}$,
        Satoshi Miyazaki$^{5,8}$,
        \newauthor
        Yutaka Komiyama$^{5,8}$
        \\
        $^{1}$Department of Astronomy and Astrophysics, University of California 
              Santa Cruz, 1156 High St., Santa Cruz, CA 95064, U.S.A\\
        $^{2}$Kavli-IPMU, The University of Tokyo Institutes for Advanced Study, 
              the University of Tokyo (Kavli IPMU, WPI), Kashiwa 277--8583, Japan\\              
        $^{3}$UCO/Lick Observatory, University of California, Santa Cruz,
              1156 High Street, Santa Cruz, CA 95064, USA\\
        $^{4}$Department of Astrophysical Sciences, Peyton Hall,
              Princeton University, Princeton, NJ 08540, USA \\
        $^{5}$National Astronomical Observatory of Japan, 2--21--1 Osawa, Mitaka, 
              Tokyo 181--8588, Japan\\
        $^{6}$Academia Sinica Institute of Astronomy and Astrophysics, 
              P.O. Box 23--141, Taipei 10617, Taiwan\\
        $^{7}$McWilliams Center for Cosmology, Department of Physics, 
              Carnegie Mellon University, Pittsburgh, PA 15213, USA\\
        $^{8}$SOKENDAI (The Graduate University for Advanced Studies), Mitaka,
              Tokyo, 181--8588, Japan
        }   
%% ------------------------------------------------------------------------------------ %% 
\date{Accepted XXX. Received YYY; in original form ZZZ}        
\pubyear{2017}

%% ------------------------------------------------------------------------------------ %% 
%% Header and Version 
%% ------------------------------------------------------------------------------------ %%

\begin{document}

\label{firstpage}
\pagerange{\pageref{firstpage}--\pageref{lastpage}}

\maketitle

%% ------------------------------------------------------------------------------------ %% 
%% Abstract and Keywords 
%% ------------------------------------------------------------------------------------ %%

\begin{abstract}

    In \citet{hscMassiveI}, we use deep ($>28.5$ \sb{} in $i$-band), high-quality 
    (0.6\asec seeing) images of ${\sim}3000$ massive (\mstar{}$\geq 11.5$) central 
    galaxies at $0.3 < z < 0.5$ from the Hyper Suprime-Cam Subaru Strategic Program 
    to show that more massive central galaxies display more prominent stellar haloes 
    with shallower outer stellar mass density profile (\mden) and higher ellipticity. 
    In this work, we investigate the environment dependence of stellar mass 
    distributions of massive central galaxies using the same dataset.
    Treating stellar mass within a 100 kpc elliptical aperture (\mtot{}) as proxy of 
    total stellar mass, we show that at fixed \mtot{}, massive central galaxies in 
    \mhalo{}$\geq 10^{14}$\msun{} haloes show slightly shallower \mden{} profiles, 
    larger half-light radius, and lower fraction of stellar mass within inner 10 kpc 
    (\minn{}) than the centrals of \mhalo{}$< 10^{14}$\msun{} haloes. 
    At fixed \minn{}--a tracer of the \textit{in situ} stellar mass, central galaxies 
    from more massive haloes host much more significant outer stellar haloes despite 
    their inner \mden{} profiles are very similar with the ones in less massive haloes.
    These results clearly demonstrate that mass--size relation for central galaxies 
    indeed depends on environment at high-\mtot{} end. 
    We also show that the \mtot{}--\minn{} relation can reflect this dependence 
    better and can help diagnose the assembly of massive galaxies and their haloes.
    These results suggest more massive haloes favor more efficient growth of stellar 
    haloes in their central galaxies, which is broadly consistent with the two phase 
    formation scenario of massive galaxies in which their outer envelopes are 
    accumulated through accretions of small systems.

\end{abstract}

\begin{keywords}
    galaxies: elliptical and lenticular, cD --
    galaxies: formation --
    galaxies: photometry -- 
    galaxies: structure -- 
    galaxies: surveys
\end{keywords}

%% ------------------------------------------------------------------------------------ %% 
%% Main Text
%% ------------------------------------------------------------------------------------ %%

%% ------------------------------------------------------------------------------------ %% 
%% Introduction 
%% ------------------------------------------------------------------------------------ %%

\section{Introduction}
    \label{sec:intro}

    A key discovery in the last 
    decade has been the dramatic structural transformation of massive quiescent 
    galaxies \citep[e.g.][]{Trujillo2006, vanDokkum2008, Cimatti2008, Damjanov2009, 
    vanderWel2011, Szomoru2012, Patel2013} from $z \approx 2$ to the present day. 
    These observations suggest that the progenitors of $z{\sim} 0$ massive early-type 
    galaxies (ETGs) need to increase their effective radii ($R_{\rm e}$) by a factor 
    of 2--4 over a time span of 10 Gyrs (e.g. \citealt{Newman2012, vdWel2014}). 
    This observational result spurred the development of the ``two-phase'' formation
    scenario for massive ETGs (e.g. \citealt{Oser2010, Oser2012}),
    in which galaxies form a compact central region at $z\sim 2$ through highly 
    dissipative processes (e.g. gas-rich merger or cold gas-accretion;
    \citealt{Hopkins2008, Dekel2009}). 
    They subsequently assemble extended stellar haloes via dry mergers 
    \citep[e.g,][]{Naab2006, Khochfar2006, Oser2010, Oser2012} which can cause 
    significant size growth at late times. 
    An alternative explanation for size growth, progenitor bias, hypothesizes that 
    larger ETGs were quenched more recently, but this explanation is still under 
    active debate (e.g. \citealt{Newman2012, Carollo2013, Poggianti2013, Belli2015,
    Keating2015, Fagioli2016}).

    There have been multiple observational attempts to test the two-phase 
    formation scenario using galaxies at low redshift, by investigating surface 
    brightness or mass density profiles (e.g. \citealt{Huang2013a, Huang2013b, 
    Oh2017}, optical color gradients (e.g. \citealt{LaBarbera2010, LaBarbera2012}), 
    and stellar population gradients \citep[e.g.,][]{Coccato2010, Coccato2011, 
    Greene2015, Barbosa2016}. 
    These observations are generally consistent with the two-phase scenario, 
    but this picture still face many challenges.
    One of the most outstanding issues is the question of whether or not the 
    fraction of accreted stellar mass displays a dependency on dark matter halos mass.

    According to the two-phase formation scenario, the late time growth of massive 
    ETGs is intrinsically tied to the growth of their host dark matter haloes 
    (e.g. \citealt{Leauthaud2012, Behroozi2013, Shankar2013}). 
    Massive ETGs living at the centers of more massive haloes should be subject to 
    higher merger rate compared to ETGs with similar \mstar{} but living in less 
    massive dark matter haloes. 
    A natural prediction of this picture is therefore that the structures of 
    massive ETGs should depend on their ``environment''\footnote{There are 
    multiple definitions of ``environment'' in the literature.  
    In this work, we use ``environment'' and halo mass interchangeably.}.
    Theoretical models based on the two-phase formation scenario predict an 
    environmental dependence of the stellar mass--effective radius relation for 
    ETGs (\mstar{}--$R_{\rm e}$; 
    e.g., \citealt{Shen2003, Guo2009}): the sizes of massive ETGs in more massive 
    haloes should be larger than those in less massive haloes at fixed \mstar{} 
    (e.g. \citealt{Shankar2013, Shankar2014}). 
    Yet, solid evidence of this effect at low redshift is still missing 
    (e.g. \citealt{Nair2010, HCompany13}; but also see \citealt{Yoon2017}).

    %However, while hints of an environmental dependence in the mass-size relation have 
    %been spotted at high-$z$ (e.g. \citealt{Papovich2012, Lani2013, Delaye2014}; but 
    %also see \citealt{Rettura2010}).

	% Song: Actually the Illustris paper suggests that, for massive galaxies, 
	%       even the inner region is dominated by accreted stars.

    Testing this prediction requires deep imaging of ETGs in order to probe the 
    accreted component (the stellar halo) which can extend to 100 kpc and 
    beyond \citep[eg.,][]{hscMassiveI}. 
    This is very challenging observation as the surface brightness profiles of 
    massive ETGs decline rapidly with typical values of $\mu > 26.0$ \sb{} in 
    $i$-band at 100 kpc and at $z\sim0.3$. 
    In \citet[][Paper I hereafter]{hscMassiveI}, we showed that deep, 
    multi-band imaging from the Subaru Strategic Program (SSP; \citealt{HSC-SSP,
    HSC-DR1}) using Hyper Suprime-Cam (HSC; \citealt{Miyazaki2012}, 
    Miyazaki in~prep.) allows us to extract robust surface stellar
    mass density (\mden{}) profiles for {\it individual} galaxies with 
    \logms{}$>11.4$ at $0.3 < z < 0.5$ and out to 100 kpc. 
    In Paper I, we characterized the stellar mass distributions and mass density 
    profiles of stellar haloes of massive ETGs, and showed that there is a large 
    intrinsic scatter in the \mden{} stellar halos of massive galaxies on 100 kpc 
    scales, and that the fraction of accreted stars residing in the stellar halos 
    increases as a function of \mstar{}. 
    In this follow-up paper, we now address a fundamental prediction of the 
    two-phase formation scenario and conclusively show that the extended stellar 
    halos of massive central galaxies depend on dark matter halo mass. 
    In other terms, we reveal the halo mass dependence of the mass-size relation 
    for massive ETGs.

%    with the help of HSC images of a large sample 
%    of massive ETGs and the \redm{} cluster catalog (e.g.\ \citealt{Rykoff2014}; 
%    \citealt{Rozo2015b}; please see \S \ref{sec:data} for details)

    This paper is organized as follows. 
    In \S \ref{sec:data} we briefly introduce the sample selection and the data 
    reduction processes.  
    Please refer to \citet{hscMassiveI} for more technical details.
    Our main results are presented in \S \ref{sec:result} and discussed in 
    \S \ref{sec:discussion}. Section
    \S \ref{sec:summary} presents our summary and conclusions.

    Magnitudes use the AB system (\citealt{Oke1983}), and are corrected for Galactic 
    extinction using calibrations from \citet{Schlafly11}.
    In this work, we assume $H_0$ = 70~km~s$^{-1}$ Mpc$^{-1}$, ${\Omega}_{\rm m}=0.3$, 
    and ${\Omega}_{\rm \Lambda}=0.7$.
    Stellar mass is denoted \mstar{} and has been derived using a Chabrier Initial Mass 
    Function (IMF; \citealt{Chabrier2003}). Halo mass is defined as $M_{\rm 200b}$ as 
    $M_{\rm 200b}\equiv M(<r_{\rm 200b})=200\bar{\rho} 
    \frac{4}{3}\pi r_{\rm 200b}^3$ where $r_{\rm 200b}$
    is the radius at which the mean interior density is equal to 200 times
    the mean matter density ($\bar{\rho}$). 
    As in \citet{hscMassiveI}, we do not attempt to decompose or distinguish any  
    potential ``intra-cluster'' 
    component (ICL; e.g. \citealt{Carlberg1997, Lin2004, Gonzalez2005, Mihos2005}).

%% ------------------------------------------------------------------------------------ %%

%% ------------------------------------------------------------------------------------ %% 
  \begin{figure*}
      \centering 
      \includegraphics[width=\textwidth]{fig/redbcg_m100_m10_compare}
      \caption{
          Three-color images for a subsample of massive galaxies at $z{\sim}0.4$. 
          All of these massive galaxies have very similar \mstar{} within a 10 kpc 
          elliptical aperture 
          ($11.2<\log (M_{\star,10\ \mathrm{kpc}}/M_{\odot})<11.3$). 
          The dash-line circle at the top-left figure indicates $R=$10 kpc.
          These galaxies are rank-ordered from top to bottom and from left to right 
          by their \mstar{} within a 100 kpc elliptical aperture which varies 
          from $10^{11.2} M_{\odot}$ to $10^{11.7} M_{\odot}$. 
          At fixed ``inner'' mass (\minn{}), massive galaxies display significant
          diversity in their outer profiles. 
          Red boxes indicate galaxies from dark matter haloes that are more massive 
          than ${\sim} 10^{14.0} M_{\odot}$. 
          }
      \label{fig:m100_m10_color}
  \end{figure*}
%% ------------------------------------------------------------------------------------ %%

%% ------------------------------------------------------------------------------------ %% 
  \begin{figure*}
      \centering 
      \includegraphics[width=\textwidth]{fig/redbcg_prof_1}
      \caption{
          The environmental dependence of the stellar mass profiles of massive central 
          galaxies. 
          Left: halo mass dependence of galaxy profiles at fixed ``total'' stellar mass 
          (\mtot{}). 
          Right: halo mass dependence of galaxy profiles at fixed ``inner-mass'' 
          (\minn{}). Orange and red lines correspond to central galaxies living in 
          haloes with \logmh{} $\geq 14.2$. 
          Black and grey lines correspond to central galaxies living in haloes with 
          \logmh{} $\leq 14.0$. 
          Thin lines show the profiles of individual galaxies while thick lines show 
          the median profile. 
          The uncertainty on the median profile is given by the shaded region and is 
          computed via bootstrap resampling. 
          Brown lines in the bottom panels show the relative difference between  
          two median profiles 
          ($\Delta = \log(\mu_{\star, \mathrm{cenHighMh}}) - 
          \log(\mu_{\star, \mathrm{cenLowMh}}$). 
          Errors on the difference between the two profiles are also computed 
          via bootstrap.
          The grey shaded regions show a Monte Carlo test to asses how likely is it to 
          obtain $\Delta$ from random sub-samples of the data. 
          To compute the grey shaded regions, we first mix the two samples 
          (\rbcg{} and \nbcg{}), then draw sub-samples of galaxies from the mixed 
          population and compute $\Delta$ in the same fashion as for our fiducial 
          signal. 
          We repeat this process 5000 times.  
          The dark-grey shaded region (light-grey shaded region) shows the 1-$\sigma$ 
          (3-$\sigma$) fluctuations in $\Delta$ from these 5000 draws.
          }
      \label{fig:prof_1} 
  \end{figure*}
%% ------------------------------------------------------------------------------------ %%

%% ------------------------------------------------------------------------------------ %% 
  \begin{figure*}
      \centering 
      \includegraphics[width=\textwidth]{fig/redbcg_scaling_relation}
      \caption{
          \textbf{Left}: The mass-size relations for \rbcg{} (orange squares) and 
          \nbcg{} (grey dots) galaxies. 
          Two vertical lines highlight our $11.6<$ \logms{} $<11.9$ mass bin. 
          The red solid line shows the best fit mass-size relation for \rbcg{} and the 
          grey dashed line shows the best fit relation for \nbcg{}. 
          Shaded regions in lighter colors show the 1-$\sigma$ uncertainties
          from MCMC sampling.  
          The green dashed line shows the mass-size relation for nearby central 
          galaxies in massive groups from \citet{HCompany13}. 
          \textbf{Right:} The relations between \mtot{} and \minn{} for the 
          \rbcg{} and \nbcg{} galaxies, along with the best-fit scaling relations for 
          both samples.
          }
      \label{fig:scaling_relation} 
  \end{figure*}
%% ------------------------------------------------------------------------------------ %%

%% ------------------------------------------------------------------------------------ %% 
%% Sample Selection and Data Reduction
%% ------------------------------------------------------------------------------------ %% 
\section{Sample Selection and Data Reduction}
    \label{sec:data}

    We refer the reader to Paper I for an in-depth description of the sample selection 
    and data reduction processes. 
    Here, we briefly summarize the main steps.

    We use imaging data from the HSC internal data release 
    \texttt{S15B}, which is very similar to the Public Data Release 1 
    (\citealt{HSC-DR1} and covers ${\sim} 110$ deg$^2$ in all five-band ($grizy$) to 
    the full depth in the Wide field. The data are reduced by \texttt{hscPipe 4.0.2}, a derivative of the 
    Large Synoptic Survey Telescope (LSST) pipeline (e.g.\ \citealt{Juric2015}; 
    \citealt{Axelrod2010}), modified for HSC (\citealt{HSC-PIPE}).
    The pixel scale of the reduced image is $0.168$\asec{}.
    We use $i$-band images for extracting surface brightness profiles. HSC $i$-band images are typically 3--4 mag deeper than SDSS (spell out here and reference ?????)  and have superb seeing conditions (mean $i$-band seeing has FWHM$=0.6$\asec{}).

    In Paper~I, we select a sample of xxxxx(give number) massive galaxies with spectroscopic 
    redshifts or reliable ``red-sequence'' photometric redshifts (\citealt{Rykoff2014}) 
    at $0.3<z<0.5$. Within this redshift range, we have a large enough volume (xxx give the number here) to sample the galaxy mass function above $M-*>11.6$, we can spatially resolve galaxies profiles to $\sim 5$ 
    kpc ($1.0^{\arcsec}$ corresponds to 4.4 and 6.1 kpc at 
    $z=0.3$ and 0.5), and we can safely ignore structural evolution over our redshift range (give time span in Gyr xxx).

    We derive $i$-band surface brightness profiles out to 100 kpc after carefully 
    masking out surrounding neighbors and accounting for the subtraction of the background light. We use the broadband spectral energy distributions (SED) fitting code 
    \texttt{iSEDFit}\footnote{http://www.sos.siena.edu/~jmoustakas/isedfit/} 
    (\citealt{Moustakas13}) to measure \m2l{} ratios and $k$--corrections using five-band forced \cmodel{} magnitudes from \texttt{hscPipe}. We assume a Chabrier IMF \citep{Chabrier2003}, the Flexible Stellar Population 
    Synthesis models\footnote{http://scholar.harvard.edu/cconroy/sps-models}
    (FSPS; \texttt{v2.4}; \citealt{FSPS}, \citealt{Conroy2010}), a Calzetti extinction law \citep{Calzetti2000}, and a simple delayed-$\tau$ model for star 
    formation histories (SFH). Using HSC data, we can measure the mass profiles of massive galaxies to $\sim 100$ kpc and we integrate these profiles within isophotal elliptical apertures at different physical radii. As explained in Paper I, here focus on the two following metric masses:

    \begin{itemize}

        \item stellar mass within 10 kpc (hereafter noted \minn{}) which we use as a proxy for the 
            \textit{in situ} stellar component. This is motivated both by recent observations and simulations 
            (e.g.~\citealt{vanDokkum2010}, \citealt{RodriguezGomez2016}). The value of 10 kpc that we quote here corresponds to the radius of the major axis of the isophotal ellipse (xxx check).
        \item stellar mass within 100 kpc (hereafter noted \mtot{}). We use \mtot{} as a proxy of the total stellar mass. In Paper I we showed that \mtot{} recovers more light compared to HSC \cmodel{} photometry with offsets that can be a large as 0.2 dex in magnitude.           
   \end{itemize}

   We use these two simple metric masses to explore the 
   \mstar{}-dependence of the fraction of accreted stars, and reveal to the diversity of 
   stellar envelopes among massive galaxies. In practice, we also have the full profiles for each galaxy and can also cast our results in terms of the full stellar mass profiles. However, in many cases we find it more useful to display figures using the more simple \minn{} and \mtot{} quantities. In Figure \ref{fig:m100_m10_color}, we highlight the diversity of galaxies as a function of \minn{} and \mtot{}. Figure \ref{fig:m100_m10_color} shows a random subsample of massive galaxies with very similar \minn{} that have large 
   range of \mtot{}. 
   We will use these two apertures masses to guide our comparisons of massive galaxies as a function of environments.

%% ------------------------------------------------------------------------------------ %%    
\subsection{Massive Central Galaxies from Different Environments}
    \label{ssec:cen}

    In this work, we focus on massive galaxies with \mtot{} $>10^{11.6}$~\msun{}. 
    In Paper~I, we demonstrate that this sample is almost mass complete over our full redshift range. In addition to the mass cut above, we also limit our sample to galaxies that live at the centers of their own 
    dark matter haloes-- so-called ``central'' galaxies. We use the \redm{} \texttt{v5.10} \citep{Rykoff2014, Rozo2015b} cluster catalog to construct a central galaxy sample.

    %Central galaxies are important to the study of galaxy-halo connection and follow stellar mass--halo mass relation (SHMR; e.g. \citealt{Leauthaud2012, 
    %Behroozi2013, Kravtsov2014, Tinker2017}) at high-\mstar{}.

    Firstly, we select 68 massive central galaxies from \redm{} clusters 
    with $\lambda \geq 30$, $P_{\mathrm{Cen}} \geq 0.7$, and \logmtot{}$>11.5$ 
    at $0.3 < z < 0.5$ (63/69 have \logmtot{}$>11.6$). 
    This $\lambda$ limit is chosen to mitigate incompleteness in the cluster catalog
    at the high end of our redshift window. The $P_{\mathrm{Cen}}$ limit is 
    selected to exclude low probability central galaxy candidates. 
    Based on the \mhalo{}-$\lambda$ relation calibrated by \citet{Simet2016}, these corresponds to central galaxies living in halos with \mhalo{}$>10^{14.2}$~\msun{}. This calibration is consistent with other works (e.g. \citealt{Saro2015, Farahi2016, 
    Melchior2016, Murata2017}). 
    The median richness of the sample is $\lambda \approx 41$ 
    (\mhalo{}$\approx 2.2 \times 10^{14}$~\msun{}), and there are 
    only 44 central galaxies in clusters with $\lambda>50$ 
    (\mhalo{}$\approx 3.0 \times 10^{14}$~\msun{}).
    We will refer to this sample of \textbf{central galaxies in massive haloes} as 
    the \rbcg{} sample.

    Secondly, we build a complementary sample of central galaxies in lower-mass haloes
    by excluding all galaxies that are in any \redm{} cluster with $\lambda > 20$.
    We convert their $\lambda$ into $M_{\mathrm{200b}}$ using calibration by 
    \citet{Simet2016}. 
    Then we estimate the $R_{\mathrm{200b}}$ of each cluster using the method in 
    \citet{Diemer2015}. We further exclude all galaxies within a cylinder around each cluster, with a radius
    equal to $R_{\mathrm{200b}}$ and a length equal to twice the value of the photometric 
    redshift uncertainty of the cluster (typically around 0.015 to 0.025). This second sample will be dominated by central galaxies living in haloes with
    $M_{\mathrm{200b}} < 10^{14}$~\msun{}, and will refer to this sample as \nbcg{}. There are  833 galaxies with  
\logmtot{}$> 11.6$ in this sample.

%    Initially, we select 3453 galaxies this way with \logmtot{}$> 11.2$ at 
%    $0.3 < z < 0.5$ \alexie{Why do we care about this?}. 
%    Among them, 1564 (833) have \logmtot{}$> 11.5$ (11.6).

    Satellite contamination should be relative low at the high-\mtot{} end
    (e.g. \citealt{Reid2014, Hoshino2015, Saito2016, vanUitert2016}). 
    For instance, the model from \citet{Saito2016} predicts a $\sim 7$\% 
    fraction for galaxies with \logmtot{}$>11.6$ and \logmh$<11.4$ haloes.

    In Appendix~\ref{app:basic}, we show the distributions of redshift, \mtot{}, and 
    \minn{} for the \rbcg{} and \nbcg{} samples. 
    We also compare these two samples on the \mtot{} - rest-frame $g-r$ color plane. 
    Both of them follow the same ``red-sequence'' with only a handful of galaxies displaying bluer colors.

    We note that our method fails to extract 1-D profiles for $\sim10$\% of 
    \rbcg{} and \nbcg{} galaxies due on-going major mergers or projection effects (e.g.\ nearby foreground galaxy or bright stars). However, this small failure rate should not affect any of the results in this paper.

%% ------------------------------------------------------------------------------------ %%

%% ------------------------------------------------------------------------------------ %% 
%% Results 
%% ------------------------------------------------------------------------------------ %% 
\section{Results}
    \label{sec:result}

    As shown in Figure~\ref{fig:m100_m10_color}, massive central galaxies at fixed  \minn{} display a large diversity in their stellar haloes. In Paper~1, we explored the \mstar{}-dependence of the stellar haloes in massive 
    galaxies, and concluded that the general trends are consistent with the 
    predictions of the two-phase formation scenario. In this paper, we now investigate the relation between galaxy profiles, their stellar halos, and dark matter halo mass. Here we use \mtot{} as proxy for the ``total'' \mstar{}
    of massive galaxies and we use \minn{} as a rough proxy for the mass of the 
    \textit{in situ} component.  We remind the reader that although a circular aperture is shown on 
    Fig~\ref{fig:m100_m10_color}, in practice we extract 1-D \mden{} profiles and estimate 
    \mtot{} and \minn{} using elliptical apertures following the average fluxed-weighted isophotal shape.

    % We said this already...
    %Also,given the narrow \mstar
    %{} and redshift range, we ignore \mstar{} growth 
    %and structural evolution.

% Not sure we need this? Seems too off hand
%    (no star formation, lower merger rate \etal~e.g.\ \citealt{Bellstedt2016},
%    \citealt{Inagaki2015}; but also see \citealt{Bai2014}).

%% ------------------------------------------------------------------------------------ %% 
\subsection{Environmental Dependence of the Profiles of Massive Galaxies}
    \label{ssec:sbp_mtot}

    First, we ask whether the \mden{} profiles of massive central galaxies depend on halo mass at fixed stellar mass. To ask this question, however, we must first define which stellar mass exactly is being fixed in the comparison. We will show results both when we fix \mtot{} as well as when we fix \minn{}. We also ensure that all comparisons are performed with a fixed underlying redshift distribution by mathcing samples in redshift in addition to stellar mass
    (please see Appendix~\ref{app:match}, Appendix~\ref{app:redshift}, and Fig~\ref{fig:match} for details).

%    The similar distributions of redshift of the matched samples ensure that the 
%    median \mden{} profiles from both samples are affected by seeing and background 
%    subtraction to the same extent.

%% ------------------------------------------------------------------------------------ %%     
    Fig~\ref{fig:prof_1} compares the \mden{} profiles of massive central galaxies living in low mass dark matter haloes to those living in high mass dark matter haloes at fixed \mtot{} 
    (left panel) and at fixed \minn{} (right panel). This Figure shows the main result of this paper, namely that \textbf{the \mden{} profiles of 
    massive central galaxies show a clear dependence on dark matter halo mass at both fixed \mtot{} and 
    \minn{}.}

    We perform several statistical tests to demonstrate that the differences we see in the median 
    \mden{} profiles are significant. We also conduct a variety of tests that verify the robustness of these results with respect choices regarding \mtot{} bins, our $\lambda$ cut, the redshift range, and the choice of apertures used for the metric masses. Please see Appendix~\ref{app:robust} for further details.

    The key features in Figure Fig~\ref{fig:prof_1} are the following:

    \begin{itemize}

        \item For the \mtot{} matched sample, although there is a clear and robust difference between the two median \mden{} profiles, the overall difference is subtle is mainly only apparent on $R>50$ kpc scales. This may explain why previous attempts to detect this signal using shallower images were not successful. 
        \item At fixed \mtot{}, central galaxies in high mass dark matter haloes display shallower \mden{} profiles compared to those living in low mass dark matter haloes (they have a flatter inner \mden{} profiles and more significant outer stellar haloes. 
        \item The median \mden{} profiles of the \rbcg{} and \nbcg{} samples cross each 
            other at ${\sim} 15$-20 kpc, roughly also the scale of the effective radius $R_{\mathrm{e}}$.      
        \item Finally, the differences between the two sets of profiles becomes more prominent at higher \mtot{}. This can be seen further dividing the sample into two \mtot{} bins (see Fig~\ref{fig:prof_2} and discussion in 
            Appendix~\ref{app:robust}) and when we discuss the relation between \mtot{} and 
            \minn{} (see Fig~\ref{fig:scaling_relation}). This trend has potentially interesting implications \alexie{can we explain in one phrase?} that will be investigated in follow-up work. 
    \end{itemize}

\alexie{lets chat about the following two paragraphs}
    Generally speaking, such subtle environmental dependence of \mden{} profiles in 
    the \mtot{}-matched samples suggests that environment does not play a major role 
    in shaping massive galaxies \textbf{at fixed ``total'' stellar mass}.
    Therefore it is interesting to ask, for massive galaxies that \textbf{contain 
    similar \mstar{} in the \textit{in situ} component}, how does environment affect 
    their subsequent mass assembly?
    Assuming \minn{} is reasonable tracer of the \textit{in situ} mass, we compare the 
    \mden{} profiles of \minn{}-matched samples (right panel of Fig~\ref{fig:prof_1}).  
    Although the median \mden{} profiles are very similar at $R \leq 10$ kpc, the
    massive central galaxies in more massive haloes host much more prominent outer 
    stellar halo.

    Meanwhile, the uniform slope of the \mden{} profile within 10 kpc is in sharp 
    contrast with the large \textit{intrinsic} scatter in the outskirts. 
    It hints that these two regions are shaped by distinctive processes, and may 
    explain why they depend on environment at different levels.  
    In simulations, intensive dissipative process helps create 
    self-similar de~Vaucouleur-like ($n{\sim} 4$) \mden{} profile in the inner part 
    of massive galaxy (e.g. \citealt{Hopkins2008}), and the \mhalo{}-dependent
    merging history leads to shallower stellar halo profile for central galaxy 
    of more massive halo (e.g. \citealt{Pillepich2014}). 
    The scatter of halo assembly process at fixed halo mass and the stochastic 
    nature of the galaxy--galaxy mergers also explain the large \textit{intrinsic} 
    scatter in the outer \mden{} profiles and the diversity seen on the 
    \mtot{}--\minn{} plane (Fig~\ref{fig:m100_m10_color}).

    \textbf{In summary, we reliably detect subtle, but systematic \mhalo{}-dependence 
    (environmental dependence) of \mden{} profiles within 100 kpc in massive central 
    galaxies}. 
    Such dependence seems to be consistent with the expectation of more (minor) 
    mergers in more massive dark matter haloes. 
    Non-dissipative (minor) mergers should not strongly alter the inner \mden{} 
    profiles, but can build up the outer halo efficiently
    (e.g. \citealt{Hilz2013}, \citealt{Oogi2013}).

%% ------------------------------------------------------------------------------------ %%

%% ------------------------------------------------------------------------------------ %% 
\subsection{The Environmental Dependence of Scaling Relations}
    \label{ssec:scaling}

    The previous section showed that the \mden{} profiles of massive galaxies show a dependence on the dark matter mass of their host halos. In this section, we now turn our attention to the more commonly studied mass size relation  (\mstar{}--$R_{\mathrm{e}}$). In addition, we will also consider halo mass dependence in the \mtot{}--\minn{} plane.

%% ------------------------------------------------------------------------------------ %% 
\subsubsection{Mass-Size Relation}
    \label{sssec:mass_size}

    The tight relation between \mstar{} and effective radius (or half-light radius; 
    $R_{\mathrm{e}}$ or $R_{\mathrm{50}}$; e.g. \citealt{Shankar2013, Leja2013, 
    vdWel2014}) is one of the most important scaling relation for ETGs. 
    Despite numerous attempts, previous studies have failed to detect the dark matter halo mass dependence of the \mstar{}--$R_{\mathrm{e}}$ relation at low-$z$ 
    (e.g. \citealt{Weinmann2009, Nair2010, HCompany13, Cerbrian2014}; 
    but see \citealt{Yoon2017}).

    The previous section clearly showed that the profiles of massive galaxies depend on the masses of their host dark matter halos. We now ask whether or not this dependence is also manifest when considered in terms of the more traditionally studied \mstar{}--$R_{\mathrm{e}}$ relation. However, when considering the mass-size relation, there is ambiguity and multiple approaches in terms of defining "sizes". This is especially true in the context of massive galaxies with extended stellar haloes (a reason why we prefer to show profiles directly as in the previous section). In this section, we still use \mtot{} as proxy of ``total'' \mstar{} and we adopt 
    the radius enclosing 50\% of luminosity \alexie{mass?} within 100 kpc ($R_{\mathrm{50}}$; derived 
    from the $i$-band curve-of-growth) as the ``size''. This definition of ``effective radius'' is more robust against structural details, 
    choice of model, and the background subtraction \alexie{compared to what?}. 
    Massive galaxies from this sample are large enough that the impact of seeing is not a concern.

    The left panel of Fig~\ref{fig:scaling_relation} shows the \mtot{}--$R_{\mathrm{50}}$
    relations for galaxies in high and low mass dark matter halos. We fit the \logmtot{}-$\log (R_{\mathrm{50}}/\mathrm{kpc})$ relations at $\log(M_{\star,100})>11.6$ using the \texttt{emcee} MCMC sampler (\citealt{Emcee})\footnote{The initial guesses are based on maximum 
    likelihood estimates, and we assume flat priors for parameters.}.

    The best-fit relation for \rbcg{} is:

    \begin{equation}
        \begin{aligned}
        \log (R_{\mathrm{50}}/\mathrm{kpc}) = & (0.74\pm0.13) \times \log (M_{\star, 100\ \mathrm{kpc}}/M_{\odot}) \\ & -(7.56\pm1.56)
        \end{aligned}
    \end{equation}

    \noindent And for \nbcg{}, we find:

    \begin{equation}
        \begin{aligned}
        \log (R_{\mathrm{50}}/\mathrm{kpc}) = & (0.68\pm0.06) \times \log (M_{\star, 100\ \mathrm{kpc}}/M_{\odot}) \\ & -(6.88\pm0.75)
        \end{aligned}
    \end{equation}

    \noindent As shown in the left panel of Fig~\ref{fig:scaling_relation}, galaxies in low and high mass haloes live on \mtot{}--$R_{\mathrm{50}}$ relations that have similar slopes but different normalizations. This result is robust against the ranges of \mstar{} over which the fit is performed and against the definitions 
    of  ``total'' \mstar{} and half-light radius\footnote{Using \mstar{} within 120 or 
    150 kpc, or using the $R_{\mathrm{50}}$ derived within these apertures does not 
    change the results.}.

    We qualitatively compare our results with \citealt{HCompany13} (HC13; 
    green solid line) who studied $z\leq 0.09$ central ETGs in 
    $12.5 \le$ \logmh{} $< 15$ haloes. 
    HC13 use the group catalog by \citet{Yang2007} to estimate halo mass. They obtain effective radii from  single-\ser{} model fitting to SDSS 
    images and \mstar{} based on SED fitting using the BC03 (\citealt{BC03}) 
    synthetic population model. We convert their \mstar{} estimates from a Kroupa IMF to a Chabrier 
    IMF by applying a constant -0.05 dex shift (see \citealt{Bernardi2016a}). We further add $+0.1$dex to the HSC13 stellar masses to account for a systematic offset between BC03 and our FSPS model (see Paper~1). 
    The distributions of galaxies on the 
    \mtot{}--$R_{\mathrm{50}}$ plane follow the HC13 relation reasonably well 
    (with slightly shallower slopes), even at \logmtot{}$< 11.6$ where our samples 
    start to become incomplete. 
    In contrast to the present study, HC13 finds no environment dependence in the 
    mass--size relation for central galaxies in the range $12.5\le$ \logmh{} $<15.0$. However, HC13 use SDSS which is considerably shallower compared the HSC and hence will not be as sensitive to the outer portions of the profile where we see the largest differences. A more detailed comparison is beyond the scope of this work, but we briefly \alexie{we demsontrate what exactly? chat about this test}
    demonstrate this using the \rbcg{} and \nbcg{} galaxies that overlap with the GAMA
    survey (please see Appendix~\ref{app:gama} for details).

    We matched the two samples using the \mstar{} measured using single-\ser{} 2-D
    model by the GAMA team (\citealt{Kelvin2012}). 
    At fixed \mgama{}, the \rbcg{} galaxies display much more significant stellar 
    halo despite they have very similar inner \mden{} profiles.  
    This shows again that the deeper HSC images has unique advantage in studying the
    connection between the assembly of massive galaxies and their dark matter haloes.

    In conclusion, we confirm that the \mtot{}--$R_{\mathrm{50}}$ relation of massive
    central galaxies at $0.3 < z < 0.5$ depends on halo mass. 
    At fixed \mtot{}, massive central galaxies in denser environments (higher \mhalo{}) 
    have larger $R_{\mathrm{50}}$ comparing to ones in smaller haloes at fixed 
    \mtot{}. 
%% ------------------------------------------------------------------------------------ %%

%% ------------------------------------------------------------------------------------ %% 
\subsubsection{\mtot{} - \minn{} Relation}
    \label{sssec:m100_m10}

    Now we investigate the relationship \mtot{}, \minn{}, and halo mass. Compared to the mass--size relation, the \mtot{}--\minn{} relation is not plagued by the ambiguity of galaxy ``size'', and also enables a more straightforward comparison with numerical simulations.

    The right panel of Fig \ref{fig:scaling_relation} compares \rbcg{} and \nbcg{}
    galaxies on the \mtot{}--\minn{} plane. We find that these two samples are offset when considered in this plane a follow distinct best-fit \mtot{}--\minn{} relations. 
    For \rbcg{} galaxies with \logmtot{}$>11.6$ we find:

    \begin{equation}
        \begin{aligned}
        \log (M_{\star, 10\ \mathrm{kpc}}/M_{\odot}) = & (0.48\pm0.06) \times \log (M_{\star, 100\ \mathrm{kpc}}/M_{\odot}) \\ & +(5.72\pm0.75).
        \end{aligned}
    \end{equation}

    \noindent In the same range of \mtot{}, the best-fit relation for \nbcg{} is:

    \begin{equation}
        \begin{aligned}
        \log (M_{\star, 10\ \mathrm{kpc}}/M_{\odot}) = & (0.56\pm0.03) \times \log (M_{\star, 100\ \mathrm{kpc}}/M_{\odot}) \\ & +(4.82\pm0.30).
        \end{aligned}
    \end{equation}

    These results are robust against the exact choice for the \mtot{} range over which the fit is performed. These results are also unchanged if \minn{} is replaced by mass within a 5 or a 15 kpc aperture, or if \mtot{} is replaced with mass within a 120 or a 150 kpc aperture.

%    The \mtot{}--$M_{\star, 5\ \mathrm{kpc}}$ relation actually suggests clearer 
%    \mhalo{}-dependence at high-\mtot{} end, but the $M_{\star, 5\ \mathrm{kpc}}$ 
%    are more vulnerable to the impact from seeing than \minn{}.

    From Figure \ref{fig:scaling_relation} we can draw the same conclusions as in the previous section, namely that: at fixed \mtot{}, central galaxies in more massive haloes tend to have less galaxy mass in the inner regions but display more prominent outer envelopes. The scientific implication of this result is discussed in 
    Section~\ref{sec:discussion}.

\alexie{This seems too detailed. This also belongs in the discusison section}

    Meanwhile, we find \rbcg{} galaxy tends to have lower \mstar{} within 10--15 kpc 
    and shallower inner \mden{} profile than the \nbcg{} one at fixed \mtot{}. 
    Although the HSC can not resolve the central region of these massive galaxies, 
    such difference could have interesting physical implications. 
    Observations of compact quiescent galaxies at high-$z$ tend to suggest higher 
    \mden{} in the central region than their descendants, hence it is necessary to 
    have certain process to reduce the central \mden{} (e.g. strong adiabatic expansion 
    induced by powerful AGN feedback; \citealt{Fan2008}) to match the inner \mden{} 
    profiles of massive central galaxies at low-$z$.
    Coalesce of super-massive black hoes (SMBHs) at the end of major merger, which is 
    common in the assembly history of massive galaxies, can also flatten the central 
    \mden{} profile via efficient scattering effect (e.g. \citealt{Milosavljevi2002}).
    On the right side of Fig \ref{fig:scaling_relation}, we find a few \rbcg{} galaxies 
    with very low \minn{} at high \mtot{} end that are apparently not bothered by 
    problematic photometry or exceptionally bad seeing.  
    They may be similar to the recently discovered massive BCGs with very large core 
    (a few kpc; e.g. \citealt{Postman2012, LopezCruz2014}) resulting from SMBH mergers.
    The impacts of these processes, and their dependence on \mhalo{} is an interesting 
    topic to investigate in the future.
%% ------------------------------------------------------------------------------------ %%

%% ------------------------------------------------------------------------------------ %% 
  \begin{figure*}
      \centering 
      \includegraphics[width=\textwidth]{fig/redbcg_ell_color_environ}
      \caption{
          Radial variations in ellipticity and optical colors for massive galaxies. 
          The format of this figure is similar to the right hand side of 
          Fig~\ref{fig:prof_1}. 
          Top panels show the ellipticity profile, middle panels show $g-r$, and lower 
          panels show $g-i$. 
          We compare our results with those from \citet{Tal2011} based on stacking 
          large samples of luminous red galaxies in SDSS at $z{\sim} 0.4$ 
          (solid blue line on top panel). 
          We also compare with the results from a stacking analysis of nearby massive 
          galaxies with high concentration index ($C>2.6$) in 
          \citet[][blue dash lines on the top and middle panels]{DSouza2014}. 
          Finally, we also compare with the average $g-r$ and $g-i$ color profiles 
          from a large sample of nearby elliptical galaxies in \citet[][blue, solid 
          lines on the middle and bottom panels]{LaBarbera2010}.
          }
      \label{fig:ell_color}
  \end{figure*}
%% ------------------------------------------------------------------------------------ %%

%% ------------------------------------------------------------------------------------ %% 
\subsection{Ellipticity and Color Profiles}
    \label{ssec:ell_color}

    In Paper~1, we show that the ellipticity of the outer stellar halo increases with \mtot{}, while in contrast, rest--frame color gradients do not display any strong \mtot{}-dependencies. In this paper, we take this analysis one step further to investigate whether or not either of these quantities display a halo mass dependency. We focus on ellipticity and color profiles within 5--60 kpc where we can ignore differences in sky subtraction and seeing across different bands. Galactic extinction and $k$-corrections are applied to both $(g-r)$ and $(g-i)$ color profiles.

    Figure \ref{fig:ell_color} shows the average ellipticity, $g-r$, and $g-i$ color profiles for galaxies living in low and high mass haloes both at fixed \mtot{} and at fixed \minn{}. We also compared our results with previous studies that stacked SDSS images (e.g. \citealt{LaBarbera2010, Tal2011, DSouza2014}). Figure \ref{fig:ell_color} shows that:

    \begin{itemize}
        \item \alexie{discuss here}. Galaxies in low and high mass haloes display similar ellipticity profiles at fixed \mtot{}. There is hint that suggests the central galaxies from more massive 
            haloes become slightly more elongated at $R>50$ kpc than the ones 
            in less massive haloes with similar \mtot{}. 
            However, the uncertainties in the current profiles make it not 
            conclusive.              
        \item For the \minn{}-matched samples, the \rbcg{} galaxies show higher
            average ellipticity than the \nbcg{} ones at $R > 10$ kpc. 
            The average ellipticities in the inner region are very similar. 
            This further suggests that the more elongated outer stellar haloes 
            of massive haloes relate to the accreted stars.         
        \item Galaxies in low and high mass dark matter haloes display similar rest--frame color gradients (at $r<60$ kpc) at fixed \mtot{} and at fixed  \minn{}.    
    \end{itemize}

    Besides these conclusions, the fact that we find  smooth ellipticity profile and its shallow gradient supports 
            the idea of using the flux-weighted average isophotal shape to 
            extract 1-D profiles of massive galaxies.

     The similarity of average rest-frame color profiles between the 
            \rbcg{} and \nbcg{} samples suggests that \mden{} profiles based on 
            single average \m2l{} will not bias our comparisons.

%% ------------------------------------------------------------------------------------ %%

%% ------------------------------------------------------------------------------------ %% 
%% Discussion 
%% ------------------------------------------------------------------------------------ %%

%% ------------------------------------------------------------------------------------ %% 
\section{Discussion}
    \label{sec:discussion}

    %Using the deep images from the HSC survey, we show that the stellar mass 
    %distributions of massive central galaxies at $0.3 < z < 0.5$ depend on their 
    %environment or \mhalo{}. 
    %Here we briefly discuss the physical implications and potential caveats of this 
    %result.

%% ------------------------------------------------------------------------------------ %% 
\subsection{The Role of Environment in the ``Two--Phase'' Formation of 
            Massive Central Galaxies}

    Results presented here are broadly consistent with the ``two-phase'' formation
    scenario of massive central galaxies in a hierarchical universe. 
    Under this picture, intense dissipative processes at $z > 2$ are responsible for 
    the formation of the \textit{in situ} stars in massive central galaxies. 
    After the star formation was quenched, the more gradual second phase of evolution
    is dominated by non-dissipative mergers of satellite galaxies. 
    Since dry minor mergers can effectively deposit \textit{ex situ} stars on to the 
    outskirts of these massive galaxies (e.g. \citealt{Oogi2013, Bedorf2013}), 
    it is expected that the stellar mass distributions of massive galaxies 
    show halo mass dependence, as the central galaxies of more massive haloes should
    experience more mergers.

    \citet{Shankar2014} investigates the environment dependence of galaxy size using 
    a series of hierarchical semi-analytic models. 
    All models suggest that, at fixed stellar mass, the median size of galaxies 
    strongly increases with halo mass. 
    Although the centrals of more massive halos only experience slightly higher 
    number of major mergers (mass ratio $>1$:3; e.g. \citealt{Hirschmann2013}), 
    the number of minor mergers strongly depend on the assumed dynamical friction time. 
    If short dynamical friction time is favored (e.g. \citealt{Newman2012}), 
    centrals of more massive haloes can have up to four times more minor mergers 
    (1:100--1:3). 
    These extra minor mergers alone can lead to moderate halo mass dependence of 
    size. 
    Similarly, \citet{Yoon2017} confirms these theoretical predictions using 
    semi-analytic model from \citet{Guo2011}.

    These predictions are in general agreement with our findings here as we do see 
    differences of \mden{} profiles in the outskirts are a major driver of the 
    halo mass dependence, especially for the \minn{}--matched samples. 
    At the same same, whether or not the difference in number of mergers can explain 
    all the environment dependence is yet to be examined.

    Both semi-analytic models and hydrodynamic simulations still have difficulties 
    reproducing the observed mass--size relation for massive galaxies. 
    Besides the reasons related to physical recipes, it is not straightforward for 
    simulations to predict galaxy size in the same way with observations may also 
    contribute to the problem. 
    Since the definition of apertures mass is more clear, it would be interesting 
    to compare with simulations using the \mtot{}--\minn{} relation instead of the 
    mass--size one.

    More importantly, our results show that the inner region of massive central 
    galaxies also play a role here. 
    Both the \mden{} profiles of \mtot{}--matched samples and the \mtot{}--\minn{}
    plane suggest that, at fixed \mtot{}, centrals of more massive haloes have 
    slightly shallower inner \mden{} slope and lower \mtot{}.  
    Minor mergers can not explain such difference as they should not alter the inner 
    \mden{} profiles. 
    This is also not caused by the choice of outer aperture to estimate ``total''
    \mstar{}: assuming that centrals in more massive haloes have more extended 
    \mden{} profiles, integration out to larger aperture will make the differences 
    between \rbcg{} and \nbcg{} galaxies on the \mtot{}--\minn{} plane more significant. 
    The fact that, at fixed \mtot{}, the overall stellar mass distributions of massive
    central galaxies relate to halo mass may hint us more insights about the connection 
    between the assembly of massive haloes and their central galaxies, and it is worth 
    further investigation.

%% ------------------------------------------------------------------------------------ %%

%% ------------------------------------------------------------------------------------ %% 
\subsection{Comparisons with Mass--Size Relations from Other Works}

    The best-fit mass--size relation derived here suggests that, at $0.3 < z < 0.5$, 
    central galaxies with \logmtot{}$>11.6$ in \logmhalo{}$>14.0$ haloes are on average
    $\sim20$\% larger than the centrals from \logmhalo{}$<14.0$ haloes at fixed \mtot{}.

    Previous works on this topic at lower redshift mostly find no environment dependence
    of mass--size relation (e.g. \citealt{Nair2010, Maltby2010, Cappellari2013, 
    HCompany13}).  
    The discrepancy between these results and ours should be explained by limitation of 
    shallower images, method for estimating galaxy size, differences in \mstar{} and 
    \mhalo{} ranges. 
    \citet{Cerbrian2014} finds that ETGs with \logms{}$>11.5$ are slightly larger in 
    more massive haloes, but they also shows that this environment trend is reversed 
    at lower \mstar{}. 
    \citet{Yoon2017} carefully studies a large sample of $z\sim0.1$ SDSS ETGs using  
    non-parametric method and recovers a clear environment dependence of mass-size 
    relation at \logms{}$>11.2$. 
    They conclude that, at the same \mstar{}, massive ETGs from dense environment 
    are 20--40\% larger than the ones in underdense environment. 
    The definition of environment, estimates of \mstar{} and sizes in \citet{Yoon2017} 
    are different with this work, but the results generally agree with ours.

    Although recent results start to agree that mass--size relation of massive
    ETGs show environment dependence, the situation at higher redshift is complicated. 
    While certain works find environment dependence of mass--size relation for massive 
    quiescent or early-type galaxies that is qualitatively consistent with the 
    low-$z$ results (e.g. \citealt{Papovich2012, Lani2013, Strazzullo2013, 
    Delaye2014}), there are still observations that suggest otherwise (e.g. 
    \citealt{Rettura2010, Raichoor2012, Kelkar2015, Allen2015}). 
    Some of this confusion can be explained by smaller sample size, special environment
    (high-$z$ massive cluster), or impact of massive star-forming galaxies. 
    But it also reflects the intrinsic difficulty of measuring \mstar{} and size 
    of high-$z$ galaxies (e.g. \cite{Price2017}).
    As we suggested earlier, with the help of deeper observations at higher redshift, 
    it would be beneficial to directly compare the rest-frame surface brightness 
    or the stellar mass density profiles instead of using mass--size relation alone 
    (e.g. \citealt{Szomoru2012, Patel2013, Buitrago2017, Hill2017}).
%% ------------------------------------------------------------------------------------ %%

%% ------------------------------------------------------------------------------------ %% 
%% Summary 
%% ------------------------------------------------------------------------------------ %%

%% ------------------------------------------------------------------------------------ %% 
\section{Summary and Conclusions}
    \label{sec:summary}

    Using the high-quality images from the first $\sim100$ deg$^2$ data of the 
    Hyper Suprime-Cam Subaru Strategic Program, we group a large sample of massive 
    central galaxies at $0.3 < z < 0.5$ into two categories based on their dark matter
    halo mass (\mhalo{}$\simgt 10^{14.2} M_{\odot}$ and 
    \mhalo{}$\simlt 10^{14} M_{\odot}$), and study how their stellar mass distributions 
    out to 100 kpc depend on their environments (or \mhalo{}). 
    The main results are:

    \begin{enumerate}

        \item At fixed \mtot{}, we find that massive central galaxies display \mden{} 
            profiles that have large intrinsic scatter in the outer stellar haloes and 
            present subtle but systematic trend with \mhalo{}.
            On average, massive central galaxies in more massive haloes have slightly 
            shallower inner \mden{} profiles and more prominent outer haloes. 
            This suggests that the environment affects the assembly of massive central
            galaxies at fixed ``total'' stellar mass.

        \item At fixed \minn{}, massive central galaxies in more massive haloes 
            possess much more extended stellar haloes than the ones in smaller haloes, 
            despite their median \mden{} profiles at $R < 15$ kpc are very similar. 
            Assuming that \minn{} traces the amount of ``in situ'' stars in massive
            galaxies, this results is consistent with the picture that more massive 
            haloes favor the growth of extended stellar haloes in their central 
            galaxies through more frequent mergers.

        \item The \mtot{}-$R_{\mathrm{50}}$ relation for massive central galaxies 
            also exhibits environmental dependence. 
            Central galaxies in more massive haloes have slightly larger
            $R_{\mathrm{50}}$ than the ones from less massive haloes at fixed \mtot{}.

        \item We also find that massive central galaxies in different environments
            follow different relations between their \mtot{} and \minn{}. 
            Assuming that \mtot{} is a reasonable proxy of ``total'' stellar mass and 
            \minn{} reflects the amount of \textit{in situ} stars, we propose that 
            this relation can help diagnose the assembly of massive central galaxies 
            without using ``size'' with ambiguous definition.

        \item At fixed \mtot{}, the ellipticity profiles of massive central galaxies 
            do not display strong environment dependence at $R < 60$ kpc. 
            At fixed \minn{}, the more extended stellar haloes presented in the \rbcg{}
            galaxies also show higher ellipticity than the \nbcg{} galaxies.

        \item We do not find evidence that suggests the rest-frame color profiles 
            within 60 kpc depend on environment for neither \mtot{}- nor 
            \minn{}--matched samples.

    \end{enumerate}

    These results highlight the importance of deep, high-quality images for studying 
    the assembly of massive haloes and their central galaxies. 
    We are currently working on combining the current results with weak lensing 
    measurements of dark matter haloes around these massive galaxies to gain more physical 
    insights. 
    We will also compare our results with cosmological hydro-simulations 
    such as 
    \textit{Illustris-TNG} (\citealt{Pillepich2017a, Pillepich2017b, Weinberger2017}, 
    \textit{Horizon-AGN} (\citealt{Dubois2014}), or 
    \textit{MassiveBlackII} (\citealt{Khandai2015}).

%% ------------------------------------------------------------------------------------ %%

%% ------------------------------------------------------------------------------------ %% 
%% Acknowledgements 
%% ------------------------------------------------------------------------------------ %% 
\section*{Acknowledgements}

  % Personal 
  The authors thank Frank van~den~Bosch for insightful discussions and thank 
  Shun Saito for helping us estimate the fraction of satellite galaxies in our sample.
  SH thanks Feng-Shan Liu for sharing the \mden{} profile of the $z\sim1$ BCG from 
  his work.

  % HSC part
  The Hyper Suprime-Cam (HSC) collaboration includes the astronomical communities of 
  Japan and Taiwan, and Princeton University.  The HSC instrumentation and software were
  developed by the National Astronomical Observatory of Japan (NAOJ), the Kavli Institute
  for the Physics and Mathematics of the Universe (Kavli IPMU), the University of Tokyo,
  the High Energy Accelerator Research Organization (KEK), the Academia Sinica Institute
  for Astronomy and Astrophysics in Taiwan (ASIAA), and Princeton University.  
  Funding was contributed by the FIRST program from Japanese Cabinet Office, the Ministry 
  of Education, Culture, Sports, Science and Technology (MEXT), the Japan Society for 
  the Promotion of Science (JSPS), Japan Science and Technology Agency (JST), the
  Toray Science Foundation, NAOJ, Kavli IPMU, KEK, ASIAA, and Princeton University.

  % SDSS part
  Funding for SDSS-III has been provided by the Alfred P. Sloan Foundation, the
  Participating Institutions, the National Science Foundation, and the U.S.  Department of
  Energy. The SDSS-III web site is http://www.sdss3.org.  SDSS-III is managed by the
  Astrophysical Research Consortium for the Participating Institutions of the SDSS-III
  Collaboration including the University of Arizona, the Brazilian Participation Group,
  Brookhaven National Laboratory, University of Cambridge, University of Florida, the
  French Participation Group, the German Participation Group, the Instituto de Astrofisica
  de Canarias, the Michigan State/Notre Dame/JINA Participation Group, Johns Hopkins
  University, Lawrence Berkeley National Laboratory, Max Planck Institute for
  Astrophysics, New Mexico State University, New York University, Ohio State University,
  Pennsylvania State University, University of Portsmouth, Princeton University, the
  Spanish Participation Group, University of Tokyo, University of Utah, Vanderbilt
  University, University of Virginia, University of Washington, and Yale University.

  % Pan-STARRS1 part
  The Pan-STARRS1 Surveys (PS1) have been made possible through contributions of the 
  Institute for Astronomy, the University of Hawaii, the Pan-STARRS Project Office, 
  the Max-Planck Society and its participating institutes, the Max Planck Institute 
  for Astronomy, Heidelberg and the Max Planck Institute for Extraterrestrial Physics, 
  Garching, The Johns Hopkins University, Durham University, the University of Edinburgh, 
  Queen's University Belfast, the Harvard-Smithsonian Center for Astrophysics, the Las 
  Cumbres Observatory Global Telescope Network Incorporated, the National Central 
  University of Taiwan, the Space Telescope Science Institute, the National Aeronautics 
  and Space Administration under Grant No. NNX08AR22G issued through the Planetary 
  Science Division of the NASA Science Mission Directorate, the National Science 
  Foundation under Grant No. AST-1238877, the University of Maryland, and Eotvos 
  Lorand University (ELTE).

  % LSST software
  This paper makes use of software developed for the Large Synoptic Survey 
  Telescope. We thank the LSST Project for making their code available as free 
  software at http://dm.lsstcorp.org.

  % KITP
  This research was supported in part by the National Science Foundation under Grant 
  No. NSF PHY11-25915.

  % Software
  This research made use of:
  \href{http://www.stsci.edu/institute/software_hardware/pyraf/stsci\_python}{\texttt{STSCI\_PYTHON}},
      a general astronomical data analysis infrastructure in Python. 
      \texttt{STSCI\_PYTHON} is a product of the Space Telescope Science Institute, 
      which is operated by AURA for NASA;
  \href{http://www.scipy.org/}{\texttt{SciPy}},
      an open source scientific tools for Python (\citealt{SciPy});
  \href{http://www.numpy.org/}{\texttt{NumPy}}, 
      a fundamental package for scientific computing with Python (\citealt{NumPy});
  \href{http://matplotlib.org/}{\texttt{Matplotlib}}, 
      a 2-D plotting library for Python (\citealt{Matplotlib});
  \href{http://www.astropy.org/}{\texttt{Astropy}}, a community-developed 
      core Python package for Astronomy (\citealt{AstroPy}); 
  \href{http://scikit-learn.org/stable/index.html}{\texttt{scikit-learn}},
      a machine-learning library in Python (\citealt{scikit-learn}); 
  \href{http://www.astroml.org/}{\texttt{astroML}}, 
      a machine learning library for astrophysics (\citealt{astroML});
  \href{https://ipython.org}{\texttt{IPython}}, 
      an interactive computing system for Python (\citealt{IPython});
  \href{https://github.com/kbarbary/sep}{\texttt{sep}} 
      Source Extraction and Photometry in Python (\citealt{PythonSEP});
  \href{https://jiffyclub.github.io/palettable/}{\texttt{palettable}},
      color palettes for Python;
  \href{http://dan.iel.fm/emcee/current/}{\texttt{emcee}}, 
      Seriously Kick-Ass MCMC in Python;
  \href{http://bdiemer.bitbucket.org/}{\texttt{Colossus}}, 
      COsmology, haLO and large-Scale StrUcture toolS (\citealt{Colossus}).

%% ------------------------------------------------------------------------------------ %% 
%% References  
%% ------------------------------------------------------------------------------------ %% 
\bibliographystyle{mnras}
\bibliography{redbcg}

%% ------------------------------------------------------------------------------------ %% 
%% Table.1 
%% ------------------------------------------------------------------------------------ %% 
\clearpage
\include{table1}
\clearpage
%% ------------------------------------------------------------------------------------ %%

%% ------------------------------------------------------------------------------------ %% 
%% Appendix Section
%% ------------------------------------------------------------------------------------ %%

\appendix

%% ------------------------------------------------------------------------------------ %% 
  \begin{figure}
      \centering 
      \includegraphics[width=8.5cm]{fig/redbcg_sample_stats_old}
      \caption{
          \textbf{Top-left}: The \logms{}-$g-r$ color relation of the \rbcg{} 
          (red circle) and \nbcg{} (grey dots) samples.
          We apply the $k$-corrections from \texttt{iSEDFit} fitting to the colors.~~          
          \textbf{Top-right}: the histograms of the redshift for the \rbcg{} and 
          \nbcg{} galaxies in both \logmtot$>11.2$ and $11.6<$\logmtot{}$<11.9$
          mass bins.
          The vertical lines highlights the $0.3\leq z \leq 0.5$ redshift range.~~
          \textbf{Bottom-left}: the histograms of \mtot{} for the \rbcg{} (orange-red) 
          and \nbcg{} (grey-black) samples at both $z>0.2$ (step-filled histogram) and 
          $0.3 \leq z \leq 0.5$ (stepped histogram). 
          The vertical lines in both top-left and bottom-left figures highlight the 
          $11.6<$\logmtot{}$<11.9$ mass range that will be used in the comparison of 
          the \mtot{}-matched samples.~~
          \textbf{Bottom-right}: the histograms of \minn{} in similar format. 
          Here the vertical lines highlight the 
          $11.2<$\logminn{}$<11.6$ mass range that is used for comparison.
      }
      \label{fig:sample_stats}
  \end{figure}
%% ------------------------------------------------------------------------------------ %%

%% ------------------------------------------------------------------------------------ %% 
\section{A. Basic Statistical Properties of the Sample} 
	\label{app:basic}

    On the top-left panel of Fig~\ref{fig:sample_stats}, we show the \mtot{}-color 
    relations using the $k$-corrected $g-r$ color. 
    Both samples follow the same tight ``red-sequence'' with little contamination 
    from the ``blue cloud''.
    At fixed \mtot{}, we see little offset in color distributions of the two 
    samples, suggesting that both samples consist of quiescent galaxies with 
    similar average stellar population properties.  
    This is consistent with previous result that suggests the average stellar 
    population of massive central galaxy does not depend on \mhalo{} 
    (e.g.\ \citealt{Park2007}).  
    In this work, we focus on the \mstar{} range of $11.6 \le$\logmtot{}$\le 11.9$, 
    where both samples have acceptable completeness, and their \mtot{} distributions 
    greatly overlap (see the normalized distributions of \mtot{} in the bottom-left 
    panel of Fig~\ref{fig:sample_stats}). 
    As for the \minn{} distributions, the two samples overlap the most within 
    $11.2 \le$\logmtot{}$\le 11.6$, but now they show quite different
    distributions (bottom-right figure).

    The redshift distributions also show small difference
    (upper-right panel) even in the high-\mtot{} bin, where the redshift distribution 
    of the \nbcg{} sample skews toward higher-$z$ end due to the contribution of BOSS 
    spec-$z$.
    Since this could bias the comparison of \mden{} profiles and other properties 
    (please see Appendix\ref{app:redshift} for more details), we address this via matching 
    the two samples in both mass and redshift distributions carefully
    (see Appendix~\ref{app:match}).

%% ------------------------------------------------------------------------------------ %% 
\begin{figure}
    \centering 
    \includegraphics[width=8.2cm]{fig/redbcg_avg_prof_z}
    \caption{
        Comparison of \mden{} profiles of \rbcg{} (orange-red) and \nbcg{} 
        (grey-black) at $11.6 \le$\logmtot$< 11.9$ in redshift bins of 
        $0.3\leq z<0.4$ (solid lines) and $0.4\leq z<0.5$ (dash lines). 
        We show the individual profile in the background using much thinner line, 
        and highlight the median profiles using thicker line and darker color.
        }
    \label{fig:avg_prof_z}
\end{figure}    
%% ------------------------------------------------------------------------------------ %%

%% ------------------------------------------------------------------------------------ %% 
\section{B. Comparisons of \mden{} profiles in different redshift bins}
    \label{app:redshift}

    Given the redshift range for our samples, it is important to evaluate 
    the impacts from the physical extend of seeing and the imaging depth on the \mden{} 
    profiles at different redshift. 
    Under the same seeing, the \mden{} profile of galaxy at higher-$z$ is more 
    vulnerable to the PSF smearing effect at the center. 
    It is also harder to reach to the same \mden{} level under the same imaging depth 
    due to cosmological dimming and background noise.

    In Fig~\ref{fig:avg_prof_z}, we group the \rbcg{} and \nbcg{} galaxies within 
    $11.6 \le$\logmtot$< 11.9$ into two $z$ bins ($0.3\leq z<0.4$ and $0.4\leq z<0.5$),
    and compare their \mden{} profiles. 
    In two redshift bins, the median \mden{} profiles from the same sample follow each 
    other very well outside 10 kpc, but become visibly different in the central 3-4 kpc,
    where the effect from seeing kicks in. 
    Meanwhile, the median \mden{} profiles of \rbcg{} and \nbcg{} in the same $z$ bin 
    are identical in the central region, which indicates similar average seeing 
    conditions.       
    This confirms that \mden{} profile at $> 5$ kpc is safe from the impacts of seeing 
    and difference in redshift.
    More importantly, it also suggests that, once the redshift distributions are 
    carefully matched, the difference of \mden{} profile is likely to be physical 
    even in the central region.  
%% ------------------------------------------------------------------------------------ %%

%% ------------------------------------------------------------------------------------ %% 
\begin{figure*}
    \centering 
    \includegraphics[width=16cm]{fig/redbcg_match}
    \caption{
        \textbf{Left figure} shows the details of the \mtot{}-matching process, 
        corresponding to the results shown in the left figure of   
        On the top panel, we show the overall distributions of \rbcg{} (light orange boxes) 
        and \nbcg{} (light grey dots) galaxies on the \mtot{}-$z$ plane.  
        And, we match the two sample in the \mtot{}-$z$ space outlined by the shaded region.
        We highlight the \rbcg{} galaxies in this region using bigger boxes in red frames, 
        whose size reflects the $P_{\mathrm{Cen}}$ value.  
        We also color-code them using the richness ($\lambda$) of the host cluster. 
        The matched \nbcg{} galaxies are highlighted using darker color and bigger dots. 
        To further evaluate the matching results, we show the distributions of \mtot{} 
        (middle panel) and redshift (bottom panel) separately. 
        On both panesl, we show the histograms along with their kernel density 
        distributions.  
        And, on the top of each panel, two sets of short vertical lines highlight the median 
        value (solid) and the inter-quartile (dash) of each distribution.~~~
        \textbf{Right figure} shows the similar matching results for the \minn{}-matched
        samples used for the right figure of Fig~\ref{fig:prof_1}.
        The format is exactly the same as the left one, except the \minn{} replaces the 
        \mtot{} in the top and middle panels.}
    \label{fig:match}
\end{figure*}
%% ------------------------------------------------------------------------------------ %%

%% ------------------------------------------------------------------------------------ %% 
  \begin{figure*}
      \centering 
      \includegraphics[width=15.5cm]{fig/redbcg_prof_2}
      \caption{
          Comparisons of the \mden{} profiles for \mtot{}-matched \rbcg{} 
          (orange-red) and \nbcg{} (grey-black) galaxies in lower (left; [11.5,11.70]) 
          and higher (right; [11.7, 11.9]) \mtot{} bins. 
          Other formats are in consistent with the right figure of Fig~\ref{fig:prof_1}.
          The difference in median profiles is more significant in higher \mtot{} bin.
          }
      \label{fig:prof_2}
  \end{figure*}
%% ------------------------------------------------------------------------------------ %%

%% ------------------------------------------------------------------------------------ %% 
  \begin{figure*}
      \centering 
      \includegraphics[width=\textwidth]{fig/redbcg_prof_3}
      \caption{
        Comparisons of the \mden{} profiles for \rbcg{} (orange-red) and \nbcg{} 
      	(grey-black) galaxies that are matched using proxies of total \mstar{}. 
        The formats are in consistent with the right figure of Fig~\ref{fig:prof_1}.
        The differences are, here, the samples are matched in slightly differnt ways. 
        From left to right: a) using samples at lower redshift ($0.3 < z < 0.4$); 
        b) using \rbcg{} sample with $\lambda > 20$ instead of 30; 
        c) using \mstar{} within 150 kpc instead of 100 kpc.
        The results are broadly consistent with the one in Fig~\ref{fig:prof_1}.
        }
      \label{fig:prof_3} 
  \end{figure*}
%% ------------------------------------------------------------------------------------ %%

%% ------------------------------------------------------------------------------------ %% 
  \begin{figure*}
      \centering 
      \includegraphics[width=15.5cm]{fig/redbcg_prof_4}
      \caption{
          Comparisons of the \mden{} profiles for \rbcg{} (orange-red) and \nbcg{} 
          (grey-black) galaxies that are matched using the \mstar{} enclosed in the 
          inner region. 
          Left panel shows the results after matching the \minn{} and \meff{} together, 
          and the right panel shows the results when only the \logmtot{}$\ge 11.5$
          \rbcg{} and \nbcg{} galaxies are included.
          Other formats are in consistent with the right figure of Fig~\ref{fig:prof_1}.
          }
      \label{fig:prof_4} 
  \end{figure*}
%% ------------------------------------------------------------------------------------ %%

%% ------------------------------------------------------------------------------------ %% 
\begin{figure*}
    \centering
    \includegraphics[width=16cm]{fig/redbcg_prof_gama_new}
    \caption{
        \textbf{Left:} comparison of \mstar{} estimated by the GAMA survey and 
        the \mtot{} using HSC images in this work. 
        We plot the \logmgama{} against the difference between \logmtot{} and \logmgama{}. 
        The two vertical lines highlight the mass range $11.4 \leq$\logmgama{}$<11.8$ 
        that is used for the comparison.~~
        \textbf{Right:} we compare the \mden{} profiles of \rbcg{} (orange-red) and 
        \nbcg{} (grey-black) galaxies using the samples matched on the 
        \mgama{}-$z$ plane at $11.4 \leq$\logmgama{}$<11.9$ and $0.28 \leq z < 0.4$. 
        The format is very similar to the ones in Fig~\ref{fig:prof_1}.}
    \label{fig:gama}
\end{figure*}
%% ------------------------------------------------------------------------------------ %%

%% ------------------------------------------------------------------------------------ %% 
\section{C. Match the \rbcg{} and \nbcg{} samples in \mstar{} and redshift distributions}
    \label{app:match}

    As explained earlier, it is important to make sure the two samples have similar 
    distributions in both \mstar{} and redshift before comparing their median \mden{} 
    profiles.  
    Here we briefly describe the procedure used in this work. 
    Since the \rbcg{} sample is smaller in size, we always match the \nbcg{} sample to 
    it by searching for the $N$-nearest neighbours on the $M_{\star}$-redshift plane 
    using the KDTree algorithm in the \texttt{scikit-learn} Python library 
    (\citealt{scikit-learn}), and evaluate the quality of the match using the 
    distributions of both parameters (as shown in Fig~\ref{fig:match}).

    As we only keep the unique \nbcg{} galaxies in the matched sample, we manually 
    adjust the value of $N$ to achieve the best match. 
    When the redshift distribution of the \rbcg{} sample becomes bi-model, we also try 
    to split the sample into two redshift bins and match them separately. 
    Typically $N$ is between 3 to 8.
    In Fig~\ref{fig:match}, we demonstrate this procedure using the results for 
    the \mtot{}-matched (Left) and the \minn{}-matched samples in Fig~\ref{fig:prof_1}
    (Right), and the two samples are well matched in the distributions of \mtot{}
    (or \minn{}) and redshift.  
    For all the comparisons of \mden{} profiles in this work, we match the samples 
    in the same way, and make sure the match has the same quality. 
%% ------------------------------------------------------------------------------------ %%

%% ------------------------------------------------------------------------------------ %% 
\section{D. Robustness of the \mden{} Differences} 
	\label{app:robust}

    In Fig~\ref{fig:prof_1}, we compare the \mden{} profiles of \mtot{}- and 
    \minn{}-matched samples of \rbcg{} and \nbcg{} galaxies, and here we test the 
    robustness of the results using a few extra tests that are illustrated in
    Fig~\ref{fig:prof_2}, Fig~\ref{fig:prof_3}, and Fig~\ref{fig:prof_4}, and 
    are briefly described here:

    \begin{enumerate}

        \item In Fig~\ref{fig:prof_2}, we group the samples into two \mtot{} bins. 
            Given the small sample size, we extend slightly toward lower \mtot{} range 
            ($11.5 \leq \log (M_{\star,\ 10\mathrm{kpc}}/M_{\odot}) < 11.7$ and 
             $11.7 \leq \log (M_{\star,\ 10\mathrm{kpc}}/M_{\odot}) < 11.9$). 
            Although the smaller sample leads to larger statistical uncertainties, 
            we can still see similar structural differences in both \mtot{} bins, 
            and the difference becomes more significant in the higher \mtot{} bin.  
            For the lower \mtot{} bin, the difference in the inner region becomes 
            quite uncertain, while the difference in the outskirt is still visible. 
            This potentially suggests that the environmental dependence of structure 
            also varies with \mstar{}, an important implication deserves more 
            investigations in the future.

        \item On the left panel of Fig~\ref{fig:prof_3}, we match the \rbcg{} and 
            \nbcg{} samples in a lower redshift bins ($0.30 < z < 0.42$).
            Despite the larger uncertainties due to smaller samples, we find the 
            results are the same.

        \item On the middle panel of Fig~\ref{fig:prof_3}, we includes \rbcg{} 
            galaxies in poorer clusters ($20 < \lambda < 30$), which should result 
            in overlapped \mhalo{} distributions with the \nbcg{} samples 
            considering the typical uncertainty of $\lambda$.
            This makes the difference in the inner region slightly less significant, 
            but the overall results are the same.

        \item On the right panel of Fig~\ref{fig:prof_3}, in stead of using \mtot{}, 
            we use the \mmax{}--the maximum \mstar{} by integrating the \mden{} 
            profiles to the largest radius allowed.  
            The \mmax{} values are less reliable than \mtot{} due to the 
            uncertainty of background subtraction and contamination from nearby 
            bright objects, but they can serve as different estimates of the ``total''
            \mstar{} of these galaxies.
            As shown in Section~\ref{sec:discussion}, they on average increase
            the \mstar{} by a little bit and affect the \rbcg{} more.
            The differences in the \mden{} profiles still remain very similar.

    \end{enumerate}

    We also test the robustness of the \mtot{}-matched results using the samples with 
    only spectroscopic redshift, the samples in the three GAMA fields, and the \rbcg{} 
    samples without the ones in very massive haloes ($\lambda > 40$).  
    Limited by space, we do not show these results here, but they all verify the 
    robustness of the results.

    For the results from the \minn{}-matched samples:

    \begin{enumerate}

        \item
            We match the two samples using both \minn{} and the \mstar{} within 15 kpc 
            at the same time.  
            This makes the two median \mden{} profiles very similar inside 10-15 
            kpc, while the result in the outskirt remains the same (left panel of 
            Fig~\ref{fig:prof_4}).
            Use \mstar{} within 5 or 20 kpc leads to the same conclusion.

        \item 
            To make sure the two samples are comparable in their overall assembly history,
            we also try to only include the very massive galaxies (\logmtot{}$>11.5$)
            in both samples. 
            This excludes the \nbcg{} galaxies that are much less massive and 
            more ``compact'' in structure. 
            Yet, the results regarding the structural differences remain the same.

    \end{enumerate}
%% ------------------------------------------------------------------------------------ %%

%% ------------------------------------------------------------------------------------ %% 
\section{E. Comparison of \mden{} profiles using \mstar{} from the GAMA survey}
    \label{app:gama}

    The GAMA survey greatly overlaps with the HSC survey, and it provides carefully 
    measured \mstar{} for large sample of galaxies (\citealt{Taylor2011}) that help 
    produce many interesting results (e.g. \citealt{Bauer2013, Ferreras2017}).
    They use 2-D single-\ser{} model to correct the total luminosity of the galaxy 
    (\citealt{Kelvin2012}), and derive the \m2l{} through optical-SED fitting 
    (BC03 model; Chabrier IMF) based on the PSF-matched aperture photometry. 
    Since the \ser{} model is generally more flexible than the \texttt{cModel} one, 
    it is therefore interesting to compare with the \rbcg{} and \nbcg{} galaxies 
    that also have spec-$z$ (at $z < 0.40$) and \mstar{} in GAMA DR2 
    (\citealt{Liske2015}) and see the impact of deep photometry again.

    We summarize the results in Fig~\ref{fig:gama}.  
    On the left panel, we compare the differences between \mtot{} and \mgama{}. 
    HSC survey on average recovers more \mstar{} at high-\mstar{} end, which is 
    consistent with the expectation from deeper photometry, although the 
    systematic differences in the estimates of \m2l{} could play a role here. 
    Meanwhile, it is interesting see that, above \logmtot{}$> 11.8$, \mgama{} 
    becomes increasingly larger than \mtot{}, and most of these massive 
    galaxies have very high \ser{} index from the 2-D fitting. 
    This suggests that the single-\ser{} model is no longer an appropriate one to 
    describe very massive galaxies as it tends to over-estimate the \mstar{} the 
    inner and/or outer regions.

    To verify the cause of the difference in \mstar{}, we further select samples 
    of \rbcg{} and \nbcg{} galaxies with matched \mgama{} and redshift 
    distributions (at $11.4 <$\logmgama{}$<11.8$; see Appendix \ref{app:match}), 
    and compare their \mden{} profiles (right panel). 
    Although these two subsamples are equally massive according to results from 
    GAMA survey, it is clear that the \rbcg{} galaxy has much more extended 
    outer envelope, even though its median \mden{} profile is very similar 
    to the \nbcg{} sample at $< 10$ kpc. 
    We can reproduce very the same trend with the luminosity density profiles 
    (with or without $k$-correction), suggesting that the inaccurate \ser{} 
    model definitely leads to under-estimate of \mstar{}.

%% ------------------------------------------------------------------------------------ %%

%% ------------------------------------------------------------------------------------ %%  
\bsp
\label{lastpage}
\end{document}
%% ------------------------------------------------------------------------------------ %% 
%%%%%%%%%%%%: End of the File %%%%%%%%%%%%
%% ------------------------------------------------------------------------------------ %%