### 2017-11-15 draft/paper1_v15.tex
Final edits

% Version 0; preprint format; Written by SH
% Version 1; emulateapj format; Edited by SH and AL
% Version 2; emulateapj format; Edited by SH and AL
% Version 3; emulateapj format; Edited AL
% Version 4; emulateapj format; Edited SH
% Version 5; MNRAS format; Edited SH
% Version 6; MNRAS format; Paper I edited by JG
% Version 7; MNRAS format; Edited by SH
% Version 10; MNRAS format; Edited by AL and JG
% Version 11; MNRAS format; Edited by SH
% Version 12; MNRAS format; Edited by SH
% Version 13; MNRAS format; Edited by KB
% Version 14; MNRAS format; Edited by SH, response to referee report

\documentclass[fleqn,usenatbib]{mnras}

% Packages
\usepackage{deluxetable}
\usepackage{newtxtext,newtxmath}
\usepackage[T1]{fontenc}
\usepackage{ae,aecompl}
\usepackage{amssymb, amsmath}
\usepackage{graphicx}
\usepackage{natbib}
\usepackage{url}
\usepackage{hyperref}
\usepackage{float}
\usepackage[usenames, dvipsnames]{color}

% Package Settings
\hypersetup{colorlinks=true,
            citecolor=MidnightBlue,
            linkcolor=MidnightBlue,
            filecolor=magenta,      
            urlcolor=cyan}
\urlstyle{same}

% Figure extention
\DeclareGraphicsExtensions{.pdf,.png,.jpg}

%%%%%%%%%%%%: User Defined Commands %%%%%%%%%%%%

% Song Huang's definition 
\def\arcsec{{\prime\prime}}
\def\arcmin{{\prime}}
\def\degree{{\circ}}
\def\h{\hskip -3 mm}
\def\aa{{A\&A}}
\def\aas{{ A\&AS}}
\def\aj{{AJ}}
\def\al{$\alpha$}
\def\bet{$\beta$}
\def\amin{$^\prime$}
\def\annrev{{ARA\&A}}
\def\apj{{ApJ}}
\def\apjs{{ApJS}}
\def\asec{$^{\prime\prime}$}
\def\deg{$^{\circ}$}
\def\ddeg{{\rlap.}$^{\circ}$}
\def\dsec{{\rlap.}$^{\prime\prime}$}
\def\cc{cm$^{-3}$}
\def\flamb{erg s$^{-1}$ cm$^{-2}$ \AA$^{-1}$}
\def\flux{erg s$^{-1}$ cm$^{-2}$}
\def\fnu{erg s$^{-1}$ cm$^{-2}$ Hz$^{-1}$}
\def\hst{{\textit{HST}}}
\def\kms{km s$^{-1}$}
\def\lamb{$\lambda$}
\def\lax{{$\mathrel{\hbox{\rlap{\hbox{\lower4pt\hbox{${\sim}$}}}\hbox{$<$}}}$}}
\def\gax{{$\mathrel{\hbox{\rlap{\hbox{\lower4pt\hbox{${\sim}$}}}\hbox{$>$}}}$}}
\def\simlt{\lower.5ex\hbox{$\; \buildrel < \over {\sim} \;$}}
\def\simgt{\lower.5ex\hbox{$\; \buildrel > \over {\sim} \;$}}
\def\micron{{$\mu$m}}
\def\mnras{{MNRAS}}
\def\nat{{Nature}}
\def\pasp{{PASP}}
\def\perang{\AA$^{-1}$}
\def\peryr{yr$^{-1}$}
\def\reference{\noindent\pp}
\def\refindent{\par\noindent\parskip=2pt\hangindent=3pc\hangafter=1 }
\def\sb{mag~arcsec$^{-2}$}
\def\lsun{$L_\odot$} 
\def\msun{$M_\odot$}
\def\sigs{$\sigma_*$}
\newcommand{\lt}{<}
\newcommand{\gt}{>}

\def\etal{{\ et al.~}}
\def\galfit{{\tt GALFIT}}
\def\ser{{S\'{e}rsic\ }}
\def\redm{\texttt{redMaPPer}}
\def\cmodel{\texttt{cModel}}
% Samples
\def\rbcg{\texttt{cenHighMh}}
\def\nbcg{\texttt{cenLowMh}}
\def\redbcg{{$\lambda \ge 30$}}
\def\nonbcg{{$\lambda < 20$}}
% Mass related 
\def\mstar{{$M_{\star}$}}
\def\mhalo{{$M_{\mathrm{200b}}$}}
\def\logms{{$\log (M_{\star}/M_{\odot})$}}
\def\logmh{{$\log (M_{\mathrm{200b}}/M_{\odot})$}}

\def\minn{{$M_{\star,10\mathrm{kpc}}$}}
\def\meff{{$M_{\star,15\mathrm{kpc}}$}} 
\def\mtot{{$M_{\star,100\mathrm{kpc}}$}}
\def\mout{{$M_{\star,150\mathrm{kpc}}$}}
\def\mmax{{$M_{\star,\mathrm{Max}}$}}
\def\mgama{{$M_{\star,\mathrm{GAMA}}$}}
\def\mcmodel{{$M_{\star,\mathrm{cModel}}$}}

\def\logminn{{$\log (M_{\star,10\mathrm{kpc}}/M_{\odot})$}}
\def\logmtot{{$\log (M_{\star,100\mathrm{kpc}}/M_{\odot})$}}
\def\logmout{{$\log (M_{\star,150\mathrm{kpc}}/M_{\odot})$}}
\def\logmmax{{$\log (M_{\star,\mathrm{Max}}/M_{\odot})$}}
\def\logmgama{{$\log (M_{\star,\mathrm{GAMA}}/M_{\odot})$}}
\def\logmcmodel{{$\log (M_{\star,\mathrm{cModel}}/M_{\odot})$}}

\def\m2l{{$M_{\star}/L$}}
\def\s2n{{$\mathrm{S}/\mathrm{N}$}}
\def\mden{{$\mu_{\star}$}}

% Commenting:
\newcommand{\todo}[1]{\textcolor{red}{\textbf{TODO:~#1}}}
\newcommand{\plan}[1]{\textcolor{cyan}{#1}}
\newcommand{\addref}{{\textcolor{red}{REF}}}
\newcommand{\note}[2]{\textcolor{blue}{\textbf{[Comment (#1): #2]}}}
\newcommand{\song}[1]{\textcolor{cyan}{\textbf{#1}}}
\newcommand{\susan}[1]{\textcolor{Bittersweet}{\textbf{#1}}}
\newcommand{\alexie}[1]{\textcolor{blue}{\textbf{[Alexie: #1]}}}
\newcommand{\jenny}[1]{\textcolor{Bittersweet}{\textbf{[Jenny: #1]}}}
\newcommand{\kevin}[1]{\textcolor{green}{\textbf{[Kevin: #1]}}}
\newcommand{\rachel}[1]{\textcolor{cyan}{\textbf{[Rachel: #1]}}}
\newcommand{\update}[1]{\textcolor{PineGreen}{#1}}
%\newcommand{\update}[1]{#1}

%%%%%%%%%%%%: Title and Affiliations %%%%%%%%%%%%

%% ------------------------------------------------------------------------------------ %% 
\title[Mass Dependent Stellar Haloes in Massive Galaxies]{Individual Stellar Haloes of 
	   Massive Galaxies Measured to 100 kpc at $0.3<z<0.5$ using Hyper Suprime-Cam}

\author[S. Huang et al.]{
        Song Huang,$^{1,2}$\thanks{E-mail: song.huang@ipmu.jp (SH)}
        Alexie Leauthaud,$^{2,1}$
        Jenny E. Greene,$^{4}$
        Kevin Bundy,$^{3,1}$
        \newauthor
        Yen-Ting Lin,$^{5}$
        Masayuki Tanaka,$^{5}$
        Satoshi Miyazaki,$^{5,7}$
        \newauthor
        Yutaka Komiyama$^{5,7}$
        \\
        $^{1}$Kavli-IPMU, The University of Tokyo Institutes for Advanced Study, 
              the University of Tokyo, Kashiwa 277--8583, Japan\\
        $^{2}$Department of Astronomy and Astrophysics, University of California 
              Santa Cruz, 1156 High St., Santa Cruz, CA 95064, U.S.A\\
        $^{3}$UCO/Lick Observatory, University of California, Santa Cruz,
              1156 High Street, Santa Cruz, CA 95064, USA\\
        $^{4}$Department of Astrophysical Sciences, Peyton Hall,
              Princeton University, Princeton, NJ 08540, USA \\
        $^{5}$National Astronomical Observatory of Japan, 2--21--1 Osawa, Mitaka, 
              Tokyo 181--8588, Japan\\
        $^{6}$Academia Sinica Institute of Astronomy and Astrophysics, 
              P.O. Box 23--141, Taipei 10617, Taiwan\\
        $^{7}$SOKENDAI (The Graduate University for Advanced Studies), Mitaka,
              Tokyo, 181--8588, Japan
        }   
%% ------------------------------------------------------------------------------------ %% 
\date{Accepted XXX. Received YYY; in original form ZZZ}        
\pubyear{2017}

%%%%%%%%%%%%: Header and Version %%%%%%%%%%%%
\begin{document}

\label{firstpage}
\pagerange{\pageref{firstpage}--\pageref{lastpage}}

\maketitle

%%%%%%%%%%%%: Abstract and Keywords %%%%%%%%%%%%

\begin{abstract}

    \song{
    Massive galaxies display extended light profiles that can reach several 
    hundreds of kiloparsecs. 
    We use data from the Hyper Suprime-Cam (HSC) survey that is simultaneously wide 
    (${\sim}100$ deg$^2$) and deep ($>28.5$ \sb{} in $i$-band) to study the stellar 
    haloes of a sample of $7000$ massive galaxies at $z=0.4$. 
    The depth of the HSC data enables us to measure surface mass density profiles 
    to 100 kpc for individual galaxies without stacking. 
    As in previous work, we find that more massive galaxies exhibit more extended outer 
    profiles than smaller galaxies.  
    When this extended light is not properly accounted for (because of shallow imaging 
    and/or inadequate profile modeling), the derived stellar mass function can be 
    significantly underestimated at the high mass end.  
    Across our sample, the ellipticity of outer light profiles increases substantially 
    as we probe larger radii.  
    }
    We show for the first time that these ellipticity gradients steepen dramatically as 
    a function of galaxy mass, but we detect no mass-dependence in outer color gradients. 
    Our results support the two-phase formation scenario for massive galaxies in which 
    outer envelopes are built up at a later time from a series of merging events. 
    We provide surface mass density profiles in a convenient tabulated format 
    to facilitate comparisons with predictions from numerical simulations of 
    galaxy formation.

\end{abstract}

\begin{keywords}
    galaxies: elliptical and lenticular, cD --
    galaxies: formation --
    galaxies: photometry -- 
    galaxies: structure -- 
    survey 
\end{keywords}

%%%%%%%%%%%%: Main Text %%%%%%%%%%%%

%% ------------------------------------------------------------------------------------ %% 
    \begin{figure*}
        \centering 
        \includegraphics[width=\textwidth]{fig/redbcg_sdss_compare}
        \caption{
            A comparison between the depth and imaging quality of SDSS and the HSC wide 
            layer for a sample of nearby massive elliptical galaxies at $0.2 < z < 0.5$.  
            These images are generated using $gri$ band images with an arcsinh stretch 
            \citep{Lupton2004}. 
            The HSC \texttt{WIDE} layer is $3.0$-$4.0$ magnitudes deeper than SDSS.
            %This added depth is critical in order map the outskirts of ETGs out to 
            %large radii.
            }
        \label{fig:sdss_compare}
    \end{figure*}
%% ------------------------------------------------------------------------------------ %%

%% ------------------------------------------------------------------------------------ %% 
\section{Introduction}
    \label{sec:intro}

    \song{
    Massive early-type galaxies (ETGs) are predicted to assemble according to a 
    `two-phase' formation scenario 
    (e.g. \citealt{Oser2010, Oser2012}): a rapid growth phase at high-redshift that 
    is dominated by intense dissipative \textit{in situ} star formation (e.g.
    \citealt{Hopkins2008, Dekel2009}), and a second phase that 
    is driven by non-dissipative processes such as dry mergers 
    (e.g. \citealt{Naab2006, Khochfar2006}), with an important role played by 
    minor mergers 
    (e.g. \citealt{Hilz2012, Hilz2013, Oogi2013, Bedorf2013, Laporte2013}).
    Both numerical simulations (e.g. \citealt{Oser2010}) and semi-analytic models 
    (SAM; e.g. \citealt{LeeYi2013, LeeYi2017}) agree that the fraction of stellar 
    mass accreted in the second phase (the \textit{ex situ} component) should increase 
    with total galaxy stellar mass (e.g. \citealt{Lackner2012, Cooper2013, Qu2017}).
    }
    For instance, recent results from the 
    Illustris\footnote{\texttt{http://www.illustris-project.org/}} simulation 
    (\citealt{Vogelsberger2014}, \citealt{Genel2014}) predict that the fraction of  
    accreted stars increases significantly with galaxy mass, reaching
    $f_{\mathrm{accreted}}>0.5$ at \logms{}$>11.5$ 
    (\citealt{RodriguezGomez2016})\footnote{The Illustris simulation
    does not reproduce the observed stellar mass function at the high mass end.}. 
%% ------------------------------------------------------------------------------------ %% 
    \song{
    Given the success of the `two-phase' scenario in explaining the properties of 
    high-$z$ massive quiescent galaxies (e.g. \citealt{vanDokkum2010, vanderWel2011,
    vandeSande2011, Belli2014}) and the dramatic increase of their effective radii 
    ($R_{\mathrm{e}}$; e.g. \citealt{Newman2012, vdWel2014}), it is time to confront 
    this model with additional observations, in particular, the detailed surface 
    mass density profiles of low-redshift massive galaxies. 
    }
    Early studies based on one-dimensional light profiles found that the surface 
    mass density profiles of nearby ETGs are well described by single-\ser{} profiles 
    (e.g. \citealt{Kormendy2009}; except for the most central regions) and that the 
    \ser{} index increases with total luminosity 
    (e.g. \citealt{Graham2013}). 
    A more recent study of the surface brightness profiles of ETGs 
    revealed that ETGs belong to two families: 
    those that follow single-\ser{} law, versus those that significantly deviate from 
    the single-\ser{} profile \citep{Schombert2015}.  
    Two-dimensional analyses have also found that the stellar distributions of massive 
    ETGs are often better described by multiple-component models 
    (e.g. \citealt{Huang2013a, Oh2017}).  
    \citet{Huang2013b}, further suggesting a connection between the multi-component 
    nature of massive galaxies and their two-phase assembly histories.

    To further confront the two-phase scenario requires very deep observations of 
    large samples of massive ETGs to correctly estimate their total stellar masses
    (e.g. \citealt{Bernardi2013, DSouza2014}), as well as to quantify the amplitude and 
    scatter among outer envelopes (e.g. \citealt{Capaccioli2015, Iodice2016, 
    Iodice2017}).
    \song{
    To date, large samples of massive galaxies with deep imaging have been lacking. 
    Even in the nearby universe, it is not trivial to map the low surface
    brightness outskirts of massive galaxies (e.g., 
    \citealt{Capaccioli2015, Iodice2016, Iodice2017, Spavone2017, Mihos2017}). 
    Some of these measurements are based on image stacking methods (e.g. 
    \citealt{Tal2011, dSouza2015}). 
    The number of very massive galaxies is also very limited in the local universe.
    For example, according to the \texttt{MASSIVE} survey (\citealt{Ma2014}), there
    are only $\sim 60$-70 massive galaxies with \logms$> 11.6$ (based on K-band 
    luminosity) within 108 Mpc.
    }

    \song{
    In addition to mass density profiles, the radial profile of ellipticity also 
    contains information about the 3-D geometry (e.g. \citealt{Tremblay1995, 
    Tremblay1996, Chang2013, RodriguezPadilla2013, Mitsuda2017}) and kinematics 
    (e.g. \citealt{Cappellari2012, Weijmans2014}) of stars in massive galaxies.
    If the stellar haloes of massive galaxies are indeed dominated by accreted 
    stars, their ellipticities could be systematically different with the inner 
    regions and may contain clues about the assembly history of massive galaxies 
    (e.g. average time since last merger, average merger mass ratio).
    These issues have not been fully explored in both simulation and observation.
    Certain simulation predicts that a more massive ETG should have rounder shape
    (e.g. \citealt{Wu2014}), while others generate massive galaxies with very 
    elongated haloes (e.g. \citealt{Li2017}).  
    On the observational side, constraints are only available from a few works 
    that use either an image stacking method (e.g. \citealt{Tal2011, dSouza2015}) 
    or a small sample of nearby galaxies (e.g. \citealt{Spavone2017}).
    }

    \song{
    In this paper, we take advantage of the high-quality deep images from the 
    Hyper Suprime-Cam (HSC) Subaru Strategic Program (SSP) \citep[SSP,][]{HSC-DR1} 
    (see \S~\ref{ssec:hsc} for details) to characterize  the light profiles of 
    massive galaxies out to 100 kpc. 
    }
    We select a large sample (${\sim} 7000$) of massive central galaxies 
    at $0.3 < z < 0.5$ using ${\sim} 100$ deg$^2$ of data from the HSC wide layer. 
    %\susan{Note that MNRAS requirement is to use British spelling; however, 
    %they also prefer the -ize variant of word such as recognize, criticize, 
    %minimize, emphasize, organize, ionize, polarize, parametrize (etc.).}

    We use this sample to 
    (1) reliably estimate individual surface mass density (\mden{}) profiles of 
    massive galaxies out to 100 kpc, 
    (2) investigate the dependence of their outer stellar haloes on total stellar 
    mass, and 
    (3) examine the implications in terms of evaluating the high mass end of the 
    galaxy stellar mass function (SMF). 
    In a second paper in this series (Huang \etal in prep.), we will investigate 
    the environmental (dark matter halo mass) dependence of the sizes of massive 
    ETGs (Huang\etal in prep.).

    This paper is organized as follows. 
    \S~\ref{sec:data} presents our data and initial sample selection. 
    \S~\ref{sec:ellipse} describes our procedure for extracting 1-D surface 
    brightness profiles. 
    \S~\ref{sec:mstar} describes how we estimate stellar mass.
    \S~\ref{sec:final} summarizes the final sample selection procedure. 
    Our main results are presented in \S~\ref{sec:result} and discussed in 
    \S~\ref{sec:discussion}. 
    \S~\ref{sec:summary} presents our summary and conclusions.

    Magnitudes use the AB system (\citealt{Oke1983}), and are corrected for galactic 
    extinction using calibrations from \citet{Schlafly11}.
    In this work, we assume $H_0$ = 70~km~s$^{-1}$ Mpc$^{-1}$, ${\Omega}_{\rm m}=0.3$, 
    and ${\Omega}_{\rm \Lambda}=0.7$.
    Stellar mass is denoted \mstar{} and has been derived using a Chabrier Initial Mass 
    Function (IMF; \citealt{Chabrier2003}).

    Halo mass is defined as 
    $M_{\rm 200b}\equiv M(<r_{\rm 200b})=200\bar{\rho} 
    \frac{4}{3}\pi r_{\rm 200b}^3$ where $r_{\rm 200b}$
    is the radius at which the mean interior density is equal to 200 times
    the mean matter density ($\bar{\rho}$).

    We emphasize that in this work we do not attempt to disentangle the 
    galaxy light from any `intra-cluster' light component (ICL; e.g. 
    \citealt{Carlberg1997, Lin2004, Gonzalez2005, Mihos2005}). 
    Although the rising stellar velocity dispersion in the outskirts of massive 
    brightest cluster galaxy (BCG)
    hints at a kinematically separated intra--cluster light (ICL) component (e.g. 
    \citealt{Dressler1979, Carter1999, Kelson2002, Bender2015, Longobardi2015}),
    it is extremely difficult to reliably isolate it photometrically.
    Moreover, both the stellar halo of the main galaxy and the ICL component carry 
    important information regarding the assembly history of the central galaxy and its 
    dark matter halo. 
    Therefore, we adopt the view that the light of the main galaxy and the 
    ICL component trace different scales of a single, smooth, and continuous 
    distribution.

%% ------------------------------------------------------------------------------------ %%

%% ------------------------------------------------------------------------------------ %% 
    \begin{figure*}
        \centering 
        \includegraphics[width=\textwidth]{fig/redbcg_ellipse_example}
        \caption{
            \textbf{Left}: Example of the 1-D surface brightness and ellipticity 
            profile of a massive galaxy at $z=0.23$ in the $i$-band extracted using 
            \texttt{Ellipse}. 
            In this work, we always show the radial profile using a $R^{1/4}$ scaling 
            on the x-axis.
            By using this scale, the de Vaucouleurs profile will appear as a straight 
            line on this figure.  
            We also plot the relative brightness profile of the PSF model normalized 
            at the central surface brightness of the galaxy to highlight the region 
            most strongly affected by seeing.
            The grey shading highlights the region ($r<6$ kpc) that is equivalent 
            to twice the size of the half-width of a 1\asec{} seeing at $z\sim 0.5$.
            Because it is a very conservative estimate of the region, we cannot reliably 
            extract a 1-D profile due to the smearing effect of seeing.
            On the top panel, the dashed line shows the mean ellipticity used for the 
            final isophote. 
            \textbf{Right}: The 3-colour image of this galaxy with isophotes 
            extracted by \texttt{Ellipse}. 
            The thick dotted line highlights the isophote with 
            $\mu_{i}{\sim} 28.5$~\sb.
            }
        \label{fig:ellipse}
    \end{figure*}
%% ------------------------------------------------------------------------------------ %%

%% ------------------------------------------------------------------------------------ %% 
\section{Data And Sample Selection}
    \label{sec:data}

\subsection{The Hyper Suprime-Cam Survey}
    \label{ssec:hsc}

    The Subaru Strategic Program (SSP, \citealt{HSC-SSP, HSC-DR1}) makes use of the 
    new prime-focus camera, the Hyper Suprime-Cam (HSC;~\citealt{Miyazaki2012}, 
    Miyazaki in~prep.), on the 
    8.2-m Subaru telescope at Mauna Kea. 
    The ambitious multi-layer HSC survey takes advantage of the large field of 
    view (FoV;~1.5 deg in diameter) of this camera and will cover $>1000$ deg$^2$ 
    of sky in five broad bands ($grizy$) to a limiting depth of $r {\sim} 26$ mag 
    in the \texttt{WIDE} layer. 
    This work is based on the internal data release \texttt{S15B}, which covers 
    ${\sim} 110$ deg$^2$ in all five bands to full \texttt{WIDE} depth.  
    The regions covered by this release overlap with a number of spectroscopic 
    surveys 
    [e.g.\ Sloan Digital Sky Survey (SDSS)/Baryon Oscillation Spectroscopic Survey 
    (BOSS): \citealt{Eisenstein2011}, \citealt{SDSS-DR12}; 
    Galaxy And Mass Assembly (GAMA): \citealt{Driver2011}, \citealt{Liske2015}].
    \texttt{S15B} release has similar sky coverage with the Public Data Release 1
    (Please see Table 3 in \citealt{HSC-DR1} for detailed comparison).

    The HSC \texttt{WIDE} survey is about $3.0$-$4.0$ magnitudes deeper in terms of 
    the $i$-band surface brightness limit than SDSS. 
    Combined with the excellent imaging resolution (the median $i$-band seeing is 
    0.6\asec) and the wide area, the HSC survey represents an ideal dataset to perform 
    statistical studies of the surface brightness profiles of massive galaxies out to 
    their distant outskirts.  
    Fig~\ref{fig:sdss_compare} illustrates the quality of HSC imaging compared to SDSS 
    for three low-redshift ETGs, and shows that HSC survey data are well suited for 
    mapping the stellar distribution of massive galaxies out to large radii.

	HSC $i$-band images typically have the best seeing compared to other bands because 
	of strict requirements driven by weak lensing science. 
    We therefore use $i$-band images to measure the stellar distributions of 
    massive galaxies.

%% ------------------------------------------------------------------------------------ %% 
\subsection{HSC data processing} 
    \label{sec:pipeline}

    The full details of the HSC data processing can be found in \citet{HSC-PIPE}
    and are briefly summarized here. 
    The HSC SSP data are processed with \texttt{hscPipe 4.0.2}, a derivative of the 
    Large Synoptic Survey Telescope (LSST) pipeline (e.g.\ \citealt{Juric2015}; 
    \citealt{Axelrod2010}), modified for HSC. 
    \texttt{hscPipe} first performs a number of tasks at the single exposure level 
    (bias subtraction, flat fielding, background modelling, object detection and 
    measurements). 
    Astrometric and photometric calibrations are performed at the single exposure 
    level. 
    \texttt{hscPipe} then warps different exposures on to a common World Coordinate 
    System (WCS) and combines them into coadded images. 
    At this stage, \texttt{hscPipe} updates the images with a better astrometric and 
    photometric calibration using stars that are common among exposures.

    The pixel scale of the combined images is $0.168$\asec{}. 
    Photometric calibration is based on data from the Panoramic Survey Telescope 
    and Rapid Response System (Pan-STARRS) 1 imaging survey 
    (\citealt{Schlafly2012}, \citealt{Tonry2012}, \citealt{Magnier2013}). 
    To achieve consistent deblending and photometry across all bands, \texttt{hscPipe} 
    performs multi-band post-processing at the \texttt{coadd} level. 
    First, \texttt{hscPipe} performs object detection on \texttt{coadd} images in 
    each band independently and records the flux peak and the above-threshold region 
    (referred as a \texttt{footprint}) for each source. 
    Next, \texttt{footprints} and peaks from different bands are merged before     
    performing deblending and measurements. 
    Finally, \texttt{hscPipe} selects a reference band for each object based on the 
    $S/N$ in different bands. (For most galaxies in this work, the reference band is 
    the $i$-band.) 
    After fixing the centroids, shape, and other non-amplitude parameters of each 
    object in this reference catalog, \texttt{hscPipe} performs forced photometry 
    on the \texttt{coadd} image in each band.
    This forced photometry approach is optimized to yield accurate galaxy colours
    at $i_{\mathrm{CModel}\leq25.0}$ mag (see \citealt{SynPipe}).

    For each galaxy, \texttt{hscPipe} measures a \cmodel{} magnitude using an approach 
    that is similar to SDSS (\citealt{HSC-PIPE}). 
    However, as opposed to SDSS, the HSC \cmodel{} is based on forced multi-band 
    photometry, which means that it can accurately measure both the 
    \textit{fluxes and colours of galaxies}. 
    The HSC \cmodel{} algorithm fits the flux distribution of each object using a 
    combination of a de~Vaucouleur and an exponential component and accounts for the 
    point spread function (PSF). 
    The performance of this algorithm has been tested using synthetic objects 
    (\citealt{SynPipe}), and the results indicate that, generally speaking, 
    the HSC \cmodel{} photometry is accurate down to $i >25.0$ mag.  
    However, \cmodel{} currently systematically underestimates the total fluxes of 
    massive ETGs with extended stellar distributions. 
    This is caused by an intrinsic limitation of \cmodel{}, as it is incapable of
    modelling profiles with extremely extended outskirts, a problem that is exacerbated 
    at the depth of the HSC survey. 
    In addition, at the depth of the HSC survey, it is challenging to accurately 
    deblend in the vicinity of large ETGs, where satellites and background galaxies 
    often blend with the low surface brightness stellar envelope. 
    The deblending method currently implemented in \texttt{hscPipe} tends to 
    `over-deblend' the outskirts of bright galaxies and leads to an 
    underestimation of the total flux of massive ETGs. (This is discussed further in 
    \citealt{HSC-PIPE}.)   
    For these reasons, our results are based on custom-developed code to measure 
    the luminosities and stellar masses of massive galaxies. 
    We use the HSC \texttt{hscPipe} photometry for two purposes: 
    (1) to perform a first broad sample selection, and (2) to estimate the average 
    colour of massive galaxies.

%% ------------------------------------------------------------------------------------ %% 
\subsection{Initial massive galaxy sample}
    \label{ssec:initial}

    We begin by using a broad flux cut to select an initial sample of massive 
    galaxies at $z < 0.5$ from the HSC photometric catalog. 
    Based on \citet{Leauthaud2016}, $i_{\mathrm{SDSS, cModel}} \leq 21.0$ mag can 
    define a sample that includes almost all \logms{}$\geq 11.5$ galaxies.    
    We therefore perform an initial conservative selection of massive galaxies
    with $i_{\mathrm{HSC, cModel}} \leq 21.5$\footnote{We neglect small differences
    between the response curves of the SDSS-$i$ and HSC-$i$ filters.}. 
    We also limit our sample to regions that have reached the required depth of 
    the \texttt{WIDE} survey in $i$-band as defined in \citet{HSC-DR1}.
    %% The SQL search is saved as "sample.sql" 
    %% The master catalog used here is "dr1_wide_galaxy_icmodel_21.5.fits"
    %%   The original catalogs contain 2275477 objects

    We further select extended objects with no deblending errors, with well-defined 
    centroids, and with useful \texttt{cModel} magnitudes in all five bands. 
    After removing objects that have pixels affected by saturation, cosmic-rays, or 
    other optical artefacts\footnote{each criterion removes less than 8\% of the 
    entire sample.}, this sample corresponds to 1760845 galaxies and is  referred to
    as \texttt{hscPho}. 
    %% Saved as "dr1_wide_galaxy_hscPho.fits"

    Here we limit our study to the very high-mass end where the majority of galaxies 
    have either a spectroscopic redshift or a robust red-sequence photo-$z$ from the 
    \redm{} galaxy cluster catalog\footnote{See: http://risa.stanford.edu/redmapper/} 
    (e.g.\ \citealt{Rykoff2014}; \citealt{Rozo2015b}).

    We match the \texttt{hscPho} sample with a spec-$z$ catalogue compiled by the HSC 
    team.
    The catalogue is created by matching HSC objects with a series of publicly 
    available spectroscopic redshifts (e.g.\ SDSS Data Release (DR) 12  
    \citealt{SDSS-DR12}; GAMA DR2 \citealt{Liske2015}). 
    The spec-$z$ quality flags from different catalogues are homogenized into a single 
    flag that indicates secure redshifts.
    Please see \S~4.4.2 of \citet{HSC-DR1} for details of this catalogue.  
    To ensure reasonable \mstar{}-completeness at the high-\mstar{} end, we focus on
    the redshift range $0.3 \leq z \leq 0.5$. 
    %% This is catalog: "dr1_specz_use.fits"

    Objects without a spectroscopic redshift are matched with central 
    galaxies from the \redm{} SDSS DR8 (\citealt{Rykoff2014}) catalogue using a 
    $2.0^{\arcsec}$ matching radius. 
    Matched objects with a red-sequence photo-$z$ ($0.3 \leq z_{\lambda} < 0.5$) are 
    included in our sample. 
    The accuracy of the red-sequence photo-$z$ is sufficient (median 
    $|z_{\lambda} - z_{\mathrm{Spec}}| {\sim} 0.01$) for our purpose.
    The \redm{} catalogue provides an additional 133 unique redshifts for massive 
    galaxies in our sample.
    %% Estimated from "dr1_redbcg_hsc_sdss_gama_1arcsec.fits"

    In total, at $0.3 \leq z \leq 0.5$, our sample consists of 25286 galaxies with 
    reliable redshift information (referred as \texttt{hscZ}).
    %% Saved as "dr1_wide_galaxy_icmodel_21.5.fits"
    The majority of our redshifts comes from the BOSS and SDSS `legacy' luminous 
    red galaxy (LRG) samples. 
    The GAMA survey provides an additional $14$\% of all spectroscopic redshifts.
    Although the GAMA survey only covers parts of the \texttt{S15B} data release, 
    and hence affects the homogeneity of our sample, it does not affect the results 
    of this work.
    We discuss this more in \S \ref{sec:final}.

    We choose the redshift range $0.3 \leq z \leq 0.5$ to make sure that  
    (1) the inner region of massive galaxies can be resolved, and \mstar{} within 
    10 kpc can be reliably measured; 
    (2) the background noise and cosmological dimming are not major issues so that 
    the \mden{} profile can be measured out to $>100$ kpc; and
    (3) redshift evolution in the stellar population properties can be largely 
    ignored.  
    Also, at higher redshift, the completeness of the spec-$z$ sample starts to 
    decline; at lower redshifts, the over-subtraction of the background level 
    becomes a more serious issue.

    We now describe our 1-D photometric analysis 
    (\S \ref{sec:ellipse}) and our stellar mass estimates (\S \ref{sec:mstar}). 
    We define the final sample in \S \ref{sec:final}.

%%  ------------------------------------------------------------------------------------ %% 
\section{Measurements of 1-D Surface Brightness Profiles}
    \label{sec:ellipse}

    The surface brightness profiles of massive ETG are not well modelled by the 
    de~Vaucouleurs or single-\ser law, especially at the imaging depth of HSC.
    These models fail to simultaneously describe the profile in both the inner 
    and the outer regions and also cannot account for any radial variations in 
    ellipticity and position angle. 
    \song{
    Although they can still be described by more complex models 
    (e.g \citealt{Huang2013a, Huang2013b, Oh2017}), the results are 
    sensitive to the choice of model, the number of components, and internal 
    degeneracies among parameters. 
    2-D modelling is also very sensitive to background subtraction method, 
    especially for massive ETGs (e.g. \citealt{Huang2013a}).
    }

    We therefore perform elliptical isophote fitting using the Image Reduction 
    and Analysis Facility (IRAF) \texttt{Ellipse} algorithm (Jedrzejewski 1987) 
    to estimate the total luminosities of massive galaxies and to measure their 
    1-D stellar mass surface density profiles (\mden{}). 
    This 1-D method is less affected by the issues mentioned above. 
    Also, we only study galaxies in the radial range where our results are less 
    sensitive to either the PSF or the background subtraction.
    We ignore the inner $\sim6$ kpc, which is twice the size of 1\asec{} seeing 
    at $z=0.5$.
    Using this conservative choice, we can safely ignore the smearing effect of 
    seeing outside this radius.
    As we discuss below, we confirm this by comparing our HSC profiles with 
    observations with higher spatial resolution. 
    As for the impact from background subtraction, we focus on the profiles 
    within 100 kpc. 
    This is an empirical but also conservative choice based on the tests we 
    conducted on background-corrected postage-stamps. 
    Once the surrounding objects are appropriately masked out, the extracted 1-D 
    surface brightness profiles rarely see unphysical truncation or fluctuation 
    within 100 kpc, especially for the \logmtot{}$>11.6$ galaxies. 
    Please see Appendix~\ref{app:ellipse} for more details on these tests.

    \song{
    We generate a postage stamp of each galaxy that extends to 750 kpc in radius, 
    along with the bad pixel mask and the PSF model. 
    The postage stamps are large enough to evaluate the local background. 
    We choose to use $i$-band images since they trace the stellar mass distributions 
    of $0.3 \leq z \leq 0.5$ massive galaxies reasonably well 
    (corresponds to rest-frame $g$ or $r$ band). $i$-band images also enable better 
    seeing and lower background levels than the $z$ and $y$ band images, although 
    these bands are better tracers of \mden{}. 
    }

    To overcome the \texttt{hscPipe} `over-deblending' issue, we use a customized 
    procedure on each postage stamp to detect and aggressively mask out 
    neighbouring objects. 
    Furthermore, \texttt{hscPipe} tends to over-subtract the background around 
    bright objects. 
    To improve the background subtraction, we first aggressively mask 
    out all objects (including the central massive galaxy), and derive an 
    empirical background correction using \texttt{SExtractor}.
    These procedures are described in detail in Appendix~\ref{app:ellipse}. 
    We should point out that we do not use the photometric results from our 
    customized process, but simply rely on them for improved local background model
    and appropriate object mask.

    Then, we run \texttt{Ellipse} on the background-corrected, masked postage stamp 
    following the methodology of \citet{Li2012}. 
    In short, we first fit each isophote using a free centroid and shape 
    (ellipticity and position angle). 
    We then fix the centroid (using the mean flux-weighted centroid) and estimate
    the mean ellipticity and position angles of all isophotes. 
    Finally, we extract a 1-D surface brightness profile along the major axis using 
    the mean ellipticity and position angle. 
    We correct these profiles for galactic extinction and cosmological dimming, 
    and we integrate them to various radii to get the luminosity 
    within different physical (elliptical) apertures. 
    Fig~\ref{fig:ellipse} shows an example of the 1-D surface brightness and 
    ellipticity profile for a massive galaxy at $z{\sim}0.2$ and also highlights 
    a few isophotes.

    We test our procedure using different mask sizes and different \texttt{Ellipse} 
    parameters; we also test the procedure with and without our background correction. 
    Based on these tests, we find that our 1-D surface brightness profiles are 
    reliable up to surface brightness levels of $i{\sim}28.5$ \sb. 
    Beyond that, some of our profiles show signs of truncation and/or large 
    fluctuations, which are due to either the uncertainty in the background 
    subtraction or the unmasked flux from other objects.
    We choose to limit our study to surface brightness levels up to ${\sim} 28.5$ \sb. 
    This is a conservative choice but is sufficient to enable us to measure light 
    profiles out to $100$ kpc on a galaxy-by-galaxy basis (no stacking). 
    \song{
    The 1-D method fails to extract profiles for ${\sim}10$\% of the sample due 
    to severe contamination of other objects; these profiles are excluded from the
    analysis.
    }
    For additional technical details on the \texttt{Ellipse} procedure, please see 
    Appendix~\ref{app:ellipse}.

%% ------------------------------------------------------------------------------------ %% 
  \begin{figure*}
      \centering 
      \includegraphics[width=\textwidth]{fig/redbcg_mass_diff_new}
      \caption{              
          \textbf{Left:} Difference between \mcmodel{} and \mtot{} for massive
      	  galaxies (grey dots).
      	  The running-median of the mass difference is shown by large red hexagons. 
      	  On average, \mcmodel{} underestimates the total stellar mass of massive 
          galaxies by 0.1 dex, while in some cases, the difference can exceed 0.2 dex.
          Vertical histograms indicate the mass difference for all galaxies (shaded 
          histogram) and for the ones with \logmtot{}$>11.6$ (empty histogram).
          \textbf{Right:} Difference between \mmax{} and \mtot{} in the same format. 
          The average difference is small (0.02 dex) and has no clear mass--dependence. 
          \textbf{Please note that the scales of the vertical axes are different 
            for these two figures.}
          }
      \label{fig:mass_diff}
  \end{figure*}

%% ------------------------------------------------------------------------------------ %%

%% ------------------------------------------------------------------------------------ %% 
\section{Stellar Masses and Mass Density Profiles}
    \label{sec:mstar}

%% ------------------------------------------------------------------------------------ %% 
\subsection{Stellar masses from SED fitting}
    \label{ssec:isedfit}

    To convert luminosities into \mstar{}, we assume that these massive galaxies 
    can be well described by an average \m2l{}. 
    This is a reasonable assumption considering that they are mostly dominated by 
    old stellar populations and are known to have only shallow colour gradients
    (e.g.\ \citealt{Carollo1993, Davies1993, LaBarbera2012, DSouza2014}). 
    \song{
    We discuss more about this point in Appendix~\ref{app:sed}, and our own 
    measurements of colour profiles (see \S~\ref{ssec:ell_color}) also demonstrate 
    this point. 
    }

    We use the broadband spectral energy distribution (SEDs) fitting 
    (see \citealt{Walcher2011} for a recent review) code 
    \texttt{iSEDFit}\footnote{http://www.sos.siena.edu/~jmoustakas/isedfit/} 
    (\citealt{Moustakas13}) to estimate the average \m2l{} and $k$-corrections using
    five-band HSC \cmodel{} fluxes.
    \song{
    \texttt{iSEDFit} uses a simplified Bayesian approach to estimate the 
    posterior probability distribution functions (PDF) of key stellar population
    parameters.
    }
    Although \cmodel{} tends to underestimate the total fluxes of bright, extended 
    objects, it can still yield accurate \emph{average} colours thanks to the 
    forced-photometry method that takes the PSF convolution into account
    (e.g. \citealt{SynPipe}).

    \song{
    Here we estimate average \m2l{} using the Flexible Stellar Population 
    Synthesis\footnote{http://scholar.harvard.edu/cconroy/sps-models}
    (FSPS; \texttt{v2.4}; \citealt{FSPS}, \citealt{Conroy2010}) model based on the 
    MILES\footnote{http://www.iac.es/proyecto/miles/pages/stellar-libraries}
    (\citealt{MILES1}, \citealt{MILES2}) stellar library, with a
    \citet{Chabrier2003} IMF between 0.1 to 100 \msun and metallicity 
    ($[\mathrm{M}/\mathrm{H}]=\log\ (\mathrm{Z}/\mathrm{Z}_{\odot})$) between 
    0.004 to 0.03. 
    We assume a delayed-$\tau$ model with stochastic star bursts 
    for the star formation history (SFH; see Appendix~\ref{app:sed}) of low-$z$ 
    massive galaxies (e.g.\ \citealt{Kauffmann2003}). 
    The \citet{Calzetti2000} extinction law is adopted in this work.
    The massive ETGs in our sample are not very sensitive to the SFH shape or the 
    internal dust extinction. 
    }

    We construct five-band SEDs using the forced-photometry \cmodel{} magnitudes 
    corrected for Galactic extinction. 
    Presently, \cmodel{} only accounts for the statistical error on the flux 
    measurement and it certainly underestimates the true flux errors of bright 
    galaxies.  
    For this work, we supply \texttt{iSEDFit} with simplified flux errors assuming 
    $S/N = 100$ for the $riz$ bands, and $S/N = 80$ for the $g$ and $y$ band 
    (on average, images in $gy$ bands are shallower in depth and/or have higher 
    background noise).  
    These empirical \s2n{} choices still only provide lower-limits of the true 
    systematic uncertainties from the model-fitting process. 
    In \citealt{SynPipe}, we evaluate the accuracy of HSC \cmodel{} photometry 
    using synthetic galaxies, and show that \cmodel{} provides excellent measurements 
    of 5-band colours, which are crucial for reliable \m2l{} estimates.
    The typical uncertainty of \logms{} is around 0.06-0.08 dex at \logms${\sim} 11.5$.

    In Appendix~\ref{app:sed}, we briefly summarize the basic statistics of 
    the sample by showing the relationships between \mtot{} and stellar age, 
    metallicity, and internal dust extinction. 
    All these properties behave reasonably for massive galaxies in this sample. 
    Using the $k$-corrected optical colour, we can also confirm that the sample 
    follows a tight `red-sequence'.

%% ------------------------------------------------------------------------------------ %%

%% ------------------------------------------------------------------------------------ %% 
\subsection{Definitions of different aperture stellar masses}
    \label{ssec:mtotal}

    \song{
    \texttt{iSEDFit} helps us estimate the best-fitting \mstar{} based on the \cmodel{} 
    photometry (noted as \mcmodel{}) and the average \m2l{} in the $i$-band. 
    Then, we can convert the 1-D luminosity density profiles into stellar mass 
    surface density (\mden{}) profiles with the average \m2l{} and estimate \mstar{} 
    within different radius by integrating the \mden{} profiles. 
    Given exquisite \mden{} profiles extending to $>100$ kpc, the definition and meaning 
    of `total' \mstar{} becomes nuanced.
    At the same time, motivated by the two-phase scenario, \mstar{} within different 
    radius may help us trace different physical components. 
    Considering this, here we define a few benchmark physical apertures throughout 
    this work:   
    }

%% ------------------------------------------------------------------------------------ %%          
    \begin{itemize}

        \item \textbf{The \mstar{} within the inner 10 kpc} 
            \song{
            (hereafter noted \minn{}). 
            Suggested by recent observation (e.g.~\citealt{vanDokkum2010}) and 
            simulation (e.g. \citealt{RodriguezGomez2016}), the \textit{in situ} 
            component dominates the \mstar{} within one effective radius 
            ($R_{\mathrm{e}}$, or 5-10 kpc) of $z{\sim}0$ massive ETGs.
            We therefore use \minn{} as the \mstar{} of the inner `core' and as 
            a proxy for the \textit{in situ} \mstar{}. 
            The high-quality HSC data enable us to reliably measure \minn{} at 
            $0.3 < z < 0.5$ ($1.0^{\arcsec}$ in radii equal 4.4 and 6.1 kpc 
            at redshifts 0.3 and 0.5, respectively).
            We should point out that in simulation an
            (e.g. \citealt{RodriguezGomez2016}), \textit{in situ} component can 
            extend outside the inner 10 kpc, while an \textit{ex situ} component may 
            contribute to \minn{} at the same time.  
            We further discuss this assumption in \ref{ssec:twophase}.
            }

        \item \textbf{The \mstar{} within 100 kpc} 
            (hereafter noted \mtot{}). 
            \song{
            For massive galaxies in our sample, 100 kpc aperture corresponds to 5--10 
            $\times R_{\mathrm{e}}$ and should contain the majority of the \mstar{}. 
            Here we use \mtot{} as a measure of the `total' \mstar{}. 
            We show that, although not perfect, \mtot{} is a better tracer 
            of total \mstar{} than model-dependent results from shallower images 
            that rely on extrapolating the light profiles out to large radii. 
            We should point out that the \s2n{} for surface brightness measurement 
            at \s2n{} is still above the limit set by the intrinsic fluctuation 
            of the background for our massive galaxies 
            (e.g. see \citealt{Pohlen2006}). 
            }

        \item \textbf{The \mstar{} within the largest available aperture} 
            (hereafter noted \mmax{}).
            \song{
            We know that the \mden{} profiles of massive galaxies extend way beyond 
            100 kpc with no clear sign of truncation
            (e.g. \citealt{Gonzalez2005, Tal2011, DSouza2014}).
            Therefore, \mtot{} should be only considered as the lower limit of the
            `total' \mstar{}. 
            Here, we also integrate the \mden{} profile to the edge of the 
            postage stamp, and we select the isophote that gives us the highest  
            \mstar{} and define the \mmax{}. 
            These procedures help us quantify how much extra \mstar{} may be kept 
            at $>100$ kpc.
            }

    \end{itemize}

    \song{
    All aperture \mstar{} are measured after adopting an isophote with fixed 
    ellipticity and position angle, and instances of 10 kpc and 100 kpc refer to 
    the radius along the major axis of the elliptical isophote.
    }

%% ------------------------------------------------------------------------------------ %%

    \song{
    In Fig~\ref{fig:mass_diff}, we compare the different definitions of \mstar{}. 
    As expected, directly measured light out to 100 kpc helps us recover more \mstar{} 
    compared to \mcmodel{}.
    At high-\mstar{} end (e.g. \logmtot{}$>11.6$), the average difference is larger 
    than 0.1 dex and can be as large as 0.2--0.3 dex.  
    For the \cmodel{} photometry in the current \texttt{hscPipe}, the average 
    difference relates to both the intrinsic limitations \cmodel{} algorithm and the 
    over-subtracted background. 
    In \S~\ref{ssec:sbp_compare}, we use the \mden{} profiles to show that a large 
    fraction of these galaxies have \mden{} profiles shallower than the de~Vaucouleurs 
    profile; it is therefore not surprising that \cmodel{} systematically 
    underestimates the luminosity.
    More importantly, the differences clearly depend on total stellar mass, as 
    \mcmodel{} tends to miss more \mstar{} in more massive galaxies.  
    This limitation in \mcmodel{} relates to the mass-dependent nature of the 
    stellar haloes of massive galaxies (see \S~\ref{ssec:sbp_compare} too).
    These differences have important implications for estimates of the stellar 
    mass function and for studies of the environment-dependence of galaxy structure. 
    }

    \song{
    The right panel of Fig~\ref{fig:mass_diff} compares \mmax{} and \mtot{}.  
    Uncertainties in the background subtraction and the impact of neighbouring 
    objects make \mmax{} more uncertain than \mtot{}. 
    None the less, we still see that \mmax{} becomes larger than \mtot{}. 
    The differences are on average very small ($\sim0.02-0.03$ dex) and do not show 
    strong mass dependence.  
    This confirms that, at the current depth of HSC images, \mtot{} can be used as 
    a good proxy of `total' stellar mass. 
    }

%% ------------------------------------------------------------------------------------ %%

%% ------------------------------------------------------------------------------------ %% 
  \begin{figure}
      \centering 
      \includegraphics[width=\columnwidth]{fig/redbcg_completeness}
      \caption{
          Evaluation of the \mstar{} completeness of the HSC massive galaxy sample.
          We compare the volume number density function of the massive galaxies 
          for this work (black line) with the one of a much more complete sample
          from the S82-MGC catalogue (green line). 
          The grey dashed line shows the number density function of HSC massive 
          galaxies in the three GAMA fields for comparison.
          The associated uncertainties derived from bootstrap resampling are shown in 
          shaded regions. 
          The vertical grey line highlights the \logms{}$=11.6$ limit.  
          Below the limit, the HSC massive galaxy sample becomes significantly incomplete in 
          stellar mass. 
          }
      \label{fig:mass_complete}
  \end{figure}     
%% ------------------------------------------------------------------------------------ %%

%% ------------------------------------------------------------------------------------ %%        
\subsection{Stellar mass completeness}
    \label{ssec:complete}

    With the help of the Stripe82 Massive Galaxy Catalogue (\texttt{S82-MGC},
    \citealt{Bundy2015}
    )\footnote{http://www.ucolick.org/\~{}kbundy/massivegalaxies/s82-mgc.html}, 
    we investigate the \mstar{} completeness of our samples. 
    \song{
    The \texttt{S82-MGC} sample matches the deeper SDSS photometric data in the 
    Stripe 82 region (\citealt{Annis2014}) with the near-infrared data from the United 
    Kingdom Infrared Telescope Infrared Deep Sky Survey (UKIDSS; 
    \citealt{Lawrence2007}), and is complete to \logms{}$\geq 11.2$ at $z<0.7$, 
    which makes it sufficient to evaluate the completeness of our HSC sample. 
    \citet{Leauthaud2016} uses this sample to show that the BOSS spec-$z$ sample, 
    which is our main source of redshifts, is about 80\% complete at 
    \logms{}$\geq 11.6$ at $0.3 < z < 0.5$.
    }

    \song{
    Fig~\ref{fig:mass_complete} compares the number density distributions of 
    galaxies from \texttt{S82-MGC} with the 20453 galaxies that are also in 
    our sample\footnote{We do not apply any statistical corrections for completeness 
    and hence, to avoid confusion, we do not use the term `stellar mass function';
    errors on the distributions are estimated via bootstrap resampling.}.
    To be consistent with the \texttt{S82-MGC} catalogue, we estimate the \mcmodel{}
    of these galaxies using \texttt{iSEDFit}.
    We find excellent agreement between HSC \mcmodel{} and the ones from the
    \texttt{S82-MGC} catalogue.
    }

    % The GAMA survey is 80\% complete down to 
    % $10^{10.8}$\msun{} at $z{\sim} 0.3$, but is only 80\% complete to 
    % $10^{12.0}$\msun{} at $z{\sim} 0.5$ according to \citet{Taylor2011} (e.g.\ 
    % their Fig.~6).

    \song{
    We conclude that our sample of massive galaxies is reasonably complete down 
    to \logmcmodel{}${\sim} 11.5$ at $0.3 \leq z \leq 0.5$.    
    Given the average difference between \mtot{} and \mcmodel{}, we chose to focus on 
    galaxies with \logmtot{}$> 11.6$. 
    In \S~\ref{sec:discussion}, we also show results for massive galaxies 
    with $11.4 \le$\logmtot{}$<11.6$, but we caution that our sample is incomplete 
    in this lower mass bin mainly due to the intrinsic incompleteness of the 
    SDSS/BOSS spec-$z$ (see \citealt{Leauthaud2016}).
    }

%% ------------------------------------------------------------------------------------ %%

%% ------------------------------------------------------------------------------------ %% 
  \begin{figure*}
      \centering 
      \includegraphics[width=\textwidth]{fig/redbcg_mass_smf}
      \caption{
          \textbf{(a)} Impact of using \mtot{} on the galaxy stellar mass function. 
          Dashed lines correspond to the observed volume density distribution 
          computed using \mcmodel{}, whereas solid lines correspond to the distribution
          computed using \mtot{}. 
          We do not apply any completeness correction to the distributions here.
          We separate our HSC sample into centrals in haloes more massive than
          \logmh{}$\sim14.2$ (red lines) and centrals in haloes with \logmh{}$<14.0$ 
          (black lines).
          The impact on the SMF can exceed 0.2 dex for massive central galaxies in 
          very massive haloes.
          \textbf{(b)}: The \mstar{} volume density distributions of massive 
          HSC galaxies, using both \mtot{} (black solid line) and \mmax{} (black dotted 
          line).  
          Vertical lines on both plots highlight the \logmtot{}$=11.6$ mass limit. 
          The grey shaded region shows the resampling error on the HSC SMF plus an
          additional 20\% uncertainty to account for the fact that we do not include 
          satellite galaxies and that we fail to extract a 1-D profile for $\sim 10$\% 
          of our galaxies. 
          These issues will be addressed in forthcoming work (Huang\etal in prep.). 
          We compare our results with previous studies: 
          (1): SDSS galaxies at $z{\sim} 0.1$ from \citet{Bernardi2017} with \mstar{} 
          values based on photometry from 2-D \ser{}$+$Exponential model fitting 
          (purple); 
          (2): SDSS galaxies at $z{\sim} 0.1$ from \citet{Moustakas13} based on 
          improved SDSS \texttt{cModel} photometry (blue); 
          (3): \texttt{S82-MGC} galaxies at $0.15 < z< 0.43$ from 
          \citet{Leauthaud2016} based on PSF-matched SDSS--UKIDSS photometry (green).
          }
      \label{fig:smf}
  \end{figure*}
%% ------------------------------------------------------------------------------------ %%

%% ------------------------------------------------------------------------------------ %%  
\section{The Final Sample}
    \label{sec:final}

%% ------------------------------------------------------------------------------------ %%  
\subsection{Candidate massive central galaxies}
    \label{ssec:redmapper}

    \song{
    Typically, a `central' galaxy is defined as a galaxy located in the center of 
    its own dark matter halo, while a galaxy in a sub-halo orbiting within 
    the virial radius of a more massive halo is referred to as a `satellite' 
    (e.g.\citealt{Yang2007}).
    We wish to focus on massive central galaxies here as they are essential
    to the study of galaxy--halo connection.
    Although the satellite fraction is expected to be very low ($<10$\%; e.g. 
    \citealt{Reid2014, Hoshino2015, Saito2016}) at \logmtot{}$> 11.6$, we further use 
    the \redm{} cluster catalogue (\texttt{v5.10}; e.g.\ \citealt{Rykoff2014}; 
    \citealt{Rozo2015b}) based on SDSS DR8 (\citealt{SDSS-DR8}) to help us 
    identify centrals of cluster-level dark matter haloes and reduce satellite
    contamination.
    }

    \song{    
    After matching the \texttt{hscZ} sample with the central galaxies of \redm{} 
    clusters with richness $\lambda \geq 20$\footnote{Due to the depth and 
    resolution of SDSS images, the \redm{} catalogue is not complete down to 
    $\lambda=20$ at $0.3 < z < 0.5$.  
    At $z \geq 0.33$, it starts to miss a small fraction of clusters with 
    $\lambda < 30$, but it does not affect the results in this work.}
    and central probability $P_{\mathrm{Cen}} \geq 0.7$, we find 164 matched galaxies 
    at $0.3 \leq z \leq 0.5$.
    According to available calibration (e.g.\ \citealt{Saro2015, Farahi2016, 
    Simet2016, Melchior2016}), they represent the central galaxies in dark matter 
    haloes with $\log (M_{\mathrm{halo}}/M_{\odot}) \geq 14.0$; we refer to these 
    galaxies as the \rbcg{} sample. 
    }

    \song{
    As the next step, we identify and remove all galaxies within a cylindrical 
    region around each \redm{} cluster. 
    We use a radius equal to $R_{\mathrm{200b}}$ and set the length of the cylinder 
    to twice the value of the photometric redshift ($z_{\lambda}$) uncertainty 
    of each cluster\footnote{We convert $\lambda$ of each cluster to 
    $M_{\mathrm{200b}}$ using the calibration of \citet{Simet2016} and use the 
    mass-concentration relation from \citet{Diemer2015} to compute 
    $R_{\mathrm{200b}}$.}.
    After we remove galaxies associated with \redm{} clusters from our sample, 
    the remaining galaxies are dominated by central galaxies living in haloes with 
    \logmh{}$< 14.0$; we refer to these galaxies as the \nbcg{} sample.
    Using the model presented in \citet{Saito2016}, we estimate that in dark matter 
    haloes with \logmh$<11.4$, $\sim 7$\% of galaxies with \logmcmodel{}$>11.5$ are 
    satellites.
    }
    %% Saved as "dr1_redbcg_use_sed5b.fits"  
    %% Saved as "dr1_nonbcg_use_sed5b.fits"
%% ------------------------------------------------------------------------------------ %%

%% ------------------------------------------------------------------------------------ %% 
\subsection{Summary of sample construction}
    \label{ssec:sample}

    Using ${\sim} 100$ deg$^2$ of HSC data, we select a large sample of massive central 
    galaxies with reliable redshift information, and broadly separate them into two 
    categories based on $M_{\mathrm{halo}}$.

    The following is a summary of our sample construction.

    \begin{itemize}

        \item \texttt{hscPho} sample. This parent sample consists of bright galaxies 
            with $i_{\mathrm{cModel}} \leq 21.0$, good quality imaging, and reliable 
            \texttt{cModel} photometry in all five HSC bands in the \texttt{S15B} 
            data release. 
            This sample is described in \S \ref{ssec:initial}, and it contains 
            1760845 galaxies.

        \item \texttt{hscZ} sample. We limit this \texttt{hscZ} sample to galaxies 
            with reliable redshift information.
            This sample is described in \S \ref{ssec:initial}. 
            It provides us 25286 useful galaxies at $0.3<z<0.5$.

        \item With the help of the \redm{} cluster catalogue, we further select 
            candidates of massive central galaxies.
            We broadly divide the \texttt{hscZ} sample into central galaxies living 
            in haloes with \logmh{}$\geq 14.0$ (\rbcg{}) and central galaxies from 
            the haloes with \logmh{}$<14.0$ (\nbcg{}).
            To ensure the sample is \mstar{}-complete and has minimal satellite 
            contamination, we further focus on the 950 massive galaxies with 
            \logmtot{}$>11.6$ in this work\footnote{As reference, there are 
            2613 massive galaxies with \logmtot{}$>11.5$ in our sample.}.

    \end{itemize}

    The division of our sample into two halo mass bins is mainly relevant for the 
    second paper in this series (Huang \etal in prep.).  
    For the present paper, we consider only the halo mass dependence on our sample when 
    we evaluate impact of mass estimates on the SMF in \S \ref{ssec:smf}. 
    We show the distributions of redshift, \mtot{}, and \minn{} of the massive galaxy 
    sample in Appendix \ref{app:basic}, along with its \mtot{}-$(g-r)$ rest-frame 
    colour relation.

%% ------------------------------------------------------------------------------------ %%

%% ------------------------------------------------------------------------------------ %% 
\section{Results}
    \label{sec:result}

%% ------------------------------------------------------------------------------------ %%     
\subsection{Impact of missing light on the galaxy stellar mass function}
    \label{ssec:smf}

    \song{
    The stellar mass function (SMF) is critical to our understanding of 
    galaxy evolution.
    Photometric method and definitions of \mstar{} can affect the 
    high-\mstar{} end of SMF (e.g. \citealt{Bernardi2013, dSouza2015, Bernardi2017}). 
    The cited works show that, despite being widely adopted in the literature, 
    \cmodel{} photometry can significantly underestimate the \mstar{} of massive 
    galaxies. 
    Still, these works are based on more complex 2-D modelling or stacking of 
    shallow images that suffer from various systematics.
    Now, we characterize the impacts of different photometric measurements on the 
    high-\mstar{} end of the SMF using \mstar{} directly measured out to large 
    radius on deep images (see Fig~\ref{fig:smf}).
    }

    \song{
    The left panel of Fig~\ref{fig:smf} show the volume density distributions 
    (referred to as SMF for simplicity) of massive galaxies using \mcmodel{} 
    and \mtot{} for both the \rbcg{} and \nbcg{} samples.  
    As shown in Fig~\ref{fig:mass_diff}, the mass-dependent differences between these
    two measurements lead to noticeable differences in SMF at high-\mstar{} end.
    \cmodel{} photometry leads to underestimation of volume density at the 
    high-\mstar{} end of SMF. 
    More importantly, it shows that the impact of missing \mstar{} becomes more 
    significant for central galaxies in very massive haloes. 
    We demonstrate that this occurs because galaxies in more massive haloes 
    have more extended stellar envelopes than those in lower mass haloes at fixed 
    \mtot{} in Huang\etal (in prep.).
    On the right panel of Fig~\ref{fig:smf}, we also compare the SMFs using \mtot{} 
    and \mmax{}. 
    As mentioned in \S~\ref{ssec:mtotal}, \mtot{} is still a lower limit of the total 
    \mstar{} of these massive galaxies, and \mmax{} helps us check how much \mstar{} 
    could be left out.  
    Although there is still visible difference at the high-\mstar{} end, the impact
    of going from \mtot{} to \mmax{} is relatively small. 
    It suggests that \mtot{} captures the majority of the total \mstar{}.
    Here we do not attempt to apply any completeness correction, so the SMFs turn 
    over at the low-\mstar{} end. 
    For now, we add a constant ${\sim}20$\% uncertainty to the SMF to account for 
    the satellite galaxies and the galaxies for which we fail to extract 1-D profiles 
    (see Appendix~\ref{app:ellipse}), although these uncertainties should be smaller 
    at high-\mstar{} end. 
    }

    \song{
    On the right panel of Fig~\ref{fig:smf}, we also compare our SMFs with the 
    following works:
    }

    \begin{itemize}

        \item \song{
            The SMF of $0.15 < z < 0.30$ galaxies from the \texttt{S82-MGC} sample
            \citep{Leauthaud2016}, where \mstar{} is based on PSF-matched aperture
            photometry and \texttt{iSEDfit} fitting using \texttt{BC03} model.
            We account for the 0.08 dex average difference with the \texttt{FSPS} 
            ones seen in the Appendix~\ref{app:sed}
            }

        \item \song{
            The SMF for the SDSS-\textit{GALEX} sample at $z{\sim} 0.1$ using 
            SDSS \cmodel{} photometry and \texttt{iSEDfit} stellar mass based on 
            similar assumptions (see \citet{Moustakas13} for details). 
            }

        \item \song{
            The SMF for $z{\sim} 0.1$ SDSS galaxies using 2-D \texttt{SerExp} 
            models\footnote{\ser{}$+$ Exponential disc model that can recover more 
            lights than SDSS \cmodel{}; integrated to infinity.}
            (\citealt{Bernardi2013, Meert2015}) and stellar mass by 
            \citet{Mendel2014}\footnote{We use \texttt{FSPS} model, 
            \citet{Chabrier2003} IMF, and \citet{Calzetti2000} extinction law}.
            (see \citet{Bernardi2017} for details).
            }

    \end{itemize}

    \song{ 
    Due to several systematics (see e.g. \citealt{Bernardi2013, Bernardi2017}), we 
    \emph{do not} attempt to perform detailed comparisons among these SMFs.
    We are currently working to address these issues in Huang \etal (in prep.). 
    Here we simply note that the HSC \mtot{} SMF is closer to the one 
    derived by \citet{Bernardi2017} using SDSS data at $z{\sim}0.1$ and the 
    \texttt{SerExp} model.
    Meanwhile, the differences between the HSC \mtot{} SMF with the others are 
    likely caused by the photometric methods: SDSS \cmodel{} and small-aperture 
    photometry underestimate \mstar{} of massive galaxies.  
    Even before a more in-depth study, it already illustrates an important issue:
    \textbf{It is crucial to understand the impacts from photometric data and 
    methods on the estimates of SMF} before using HSC \mtot{} SMF to study 
    galaxy evolution or 
    comparing it with predictions from models and simulation. 
    This is particularly relevant since a method like \cmodel{} will still be 
    widely adopted in on-going and future imaging surveys.
    }

%% ------------------------------------------------------------------------------------ %%

%% ------------------------------------------------------------------------------------ %%
\subsection{Surface mass density profiles}
    \label{ssec:sbp_compare}

%% ------------------------------------------------------------------------------------ %% 
  \begin{figure*}
      \centering 
      \includegraphics[width=\textwidth]{fig/average_mass_profiles_fsps1_A}
      \caption{
          \textbf{Left}: Median \mden{} profiles in three total stellar mass bins. 
              Thin grey lines in the background show a random subset of individual
              profiles. 
              The scatter between the thin grey lines reflects the true scatter in 
              the profiles of massive galaxies (not measurement error). 
              The shaded region highlights the region that is most strongly affected 
              by the seeing. 
              Two vertical lines indicate 10 kpc (thin, dotted line) and
              100 kpc (thick, dashed line). ~~~ 
          \textbf{Right}: comparison between our \mden{} profiles, previous observations, 
              and simulations. 
              The solid cyan line shows the median profile of massive elliptical 
              galaxies at $z{\sim} 0$ from \citet[][]{Huang2013a}. 
              The red long-dashed line shows the median profile of massive galaxies at 
              $0.25 \leq z < 0.50$ observed by \textit{HST} from \citet[][]{Patel2013}. 
              The purple short-dashed line shows the median radial stellar distributions 
              in massive haloes from simulation using the particle tagging method
              (\citealt{Cooper2013}).}
      \label{fig:avg_prof}
  \end{figure*}
%% ------------------------------------------------------------------------------------ %%

%% ------------------------------------------------------------------------------------ %%
\subsubsection{General trends and comparison with previous work}
    \label{sssec:sbp_inter}

    Previous work on the structural evolution of massive galaxies has often focused on 
    scaling relations such as the `\mstar{}-size' relation.  
    We argue that by comparing \mden{} profiles directly, we can capture more 
    information than afforded by the \mstar{}-size relation. T
    he comparison also has the advantage that it bypasses difficult questions about 
    how to accurately define and measure galaxy `sizes' and `masses'.

    Fig.~\ref{fig:avg_prof} shows the median \mden{} profiles of massive central 
    galaxies at $0.3 < z < 0.5$ in three \mtot{} bins.
    These median profiles along with their uncertainties are derived using the 
    bootstrap resampling method. 
    Note that our sample is not complete in the lowest \mtot{} bin, although the 
    median \mden{} profile may not be significantly affected. 
    As shown in the left panel of Fig~\ref{fig:avg_prof}, we can confidently trace
    the \mden{} profiles of these massive galaxies out to 100 kpc 
    \textbf{individually}.  
    At large scales, some of our \mden{} profiles show signs of unphysical truncation 
    and fluctuation related to inaccurate sky subtraction. 
    In this paper, we do not use profiles beyond 100 kpc, even though the median 
    \mden{} profiles for the two most massive bins behave reasonably well out to 
    ${\sim} 200$ kpc.

    From Fig~\ref{fig:avg_prof} we can see the galaxies in our sample have 
    homogeneous profiles on small radial scales. 
    The amplitude of \mden{} increases with galaxy mass on 10 kpc scales but the 
    slope of \mden{} remains similar.  
    From previous work on this topic, we already know that the inner regions of 
    massive elliptical galaxies display relatively uniform structural (e.g.
    \mden{} profile, isophotal shape: e.g. \citealt{Lauer07, Kormendy2009, 
    Schombert2015}; and kinematic: e.g. \citealt{Cappellari13b}) properties.  
    However, Fig~\ref{fig:avg_prof} reveals a significant diversity in the outer 
    envelopes of massive galaxies.  
    Given the \s2n{} of HSC images at these surface brightness levels, the scatter 
    shown in Fig~\ref{fig:avg_prof} corresponds to \textbf{intrinsic scatter in 
    the stellar envelopes of massive galaxies}. 
    Importantly, Fig~\ref{fig:avg_prof} shows that the global \mden{} profiles 
    of galaxies at these masses are \textbf{clearly not self-similar} out to 100 
    kpc and have outskirts with larger scatter.

    In the right-hand side of Fig.~\ref{fig:avg_prof}, we compare our \mden{} 
    profiles with results from previous work. 
    Most previous studies have focused on surface brightness profiles instead 
    of mass density profiles.  
    Results can also depend on the stacking technique or the model used to extract 
    the profile (e.g., \citealt{Tal2011, DSouza2014}). 
    \citet{Huang2013a} derived \mden{} profiles for a small sample of very nearby 
    ellipticals (within 100 Mpc; median \logms{} ${\sim} 11.3$) based on relatively 
    shallow images from the Carnegie--Irvine Galaxy Survey 
    (CGS, \citealt{CGS1})\footnote{https://cgs.obs.carnegiescience.edu/CGS/Home.html}.  
    This sample is at very low redshift ($z<0.02$), and so the \mden{} profiles from 
    \citet{Huang2013a} galaxies are accurate to smaller scales (down to $r=1$ kpc) 
    than our HSC profiles.  
    Our \mden{} profiles show good agreement with the \citet{Huang2013a} sample in 
    the radial range of overlap (out to 50 kpc). 
    \song{
    The median profiles from \citet{Huang2013a} only reach to ${\sim} 50$ kpc for 
    $z<0.02$ massive galaxies, while our deep HSC images can reliably deliver 
    individual \mden{} profiles for $z{\sim} 0.4$ galaxies out to at least 100 kpc.
    }

    \citet{Patel2013} extracted a median \mden{} profile for massive ETGs at 
    $0.25 < z < 0.50$ using stacked \textit{HST}/ACS images. 
    These galaxies are selected at a constant cumulative number density and are 
    thought to be the progenitors of $z=0$ massive ETGs (e.g. \citealt{Leja2013}).  
    The median \mstar{} of the \citet{Patel2013} sample is 
    ${\sim} 10^{11.2} M_{\odot}$, which is lower than our lowest mass bin. 
    However, \citet{Patel2013} uses the BC03 stellar population model, which leads to 
    \mstar{} that are roughly 0.1 dex lower than our FSPS estimates 
    (see Appendix~\ref{app:sed}). 
    Furthermore, the \citet{Patel2013} images are shallower than ours which means 
    that their \mstar{} could still be underestimated due to missing light in the 
    outskirts. 
    Given these two considerations, it is reasonable to roughly compare the 
    \citet{Patel2013} profile with the one in our lowest \mtot{} bin. 
    The superb resolution of the \textit{HST}/ACS images allows \citet{Patel2013} to 
    accurately measure the \mden{} profile down to 1 kpc without worrying about the
    smearing effect of seeing. 
    The good agreement between our profiles and the ones derived from \textit{HST} 
    imaging demonstrates that our profiles are robust at $r\geq 3$ kpc; therefore, 
    we can accurately measure \minn{}.

    Finally, we also compare our HSC profiles with the predicted median \mden{} 
    profile of central galaxies in massive haloes ($13.5 < \log M_{200,c} < 14.0$) 
    from a cosmological simulation where the \mden{} profiles of galaxies are 
    calculated using the particle tagging technique (e.g., \citealt{Cooper2010}). 
    The simulated \mden{} profile is affected by the resolution limit of the 
    simulation in the inner region but is in good agreement with our median 
    \mden{} profile for the $11.6 <$ \logmtot{} $< 11.8$ bin within 40 kpc. 
    However, when compared to our data for the $11.6 <$ \logmtot{} $< 11.8$ bin 
    outside 40 kpc, the particle tagging method seems to predict an overly  
    prominent stellar halo that has a much shallower outer slope.

    Table~1 provides tabulated values for the median profiles that are 
    displayed in Fig.~\ref{fig:avg_prof}. 
    These profiles are also available  
    \href{http://www.ucolick.org/\~kbundy/massivegalaxies}{here}:
    {\url{http://www.ucolick.org/\~kbundy/massivegalaxies}}. 
    (The files will be made available after the paper is accepted)

%% ------------------------------------------------------------------------------------ %%

%% ------------------------------------------------------------------------------------ %% 
\subsection{Ellipticity and colour profiles}
    \label{ssec:ell_color}

%% ------------------------------------------------------------------------------------ %% 
  \begin{figure*}
      \centering 
      \includegraphics[width=\textwidth]{fig/redbcg_ell_color.pdf}
      \caption{
          Radial profiles of the ellipticty and $k$-corrected rest-frame optical colours 
          of massive galaxies in our sample. 
          The general format of this figure is similar to Fig~\ref{fig:avg_prof}.  
          The \textbf{left panel} displays ellipticity profiles, 
          the \textbf{Upper-right panel} shows $g-r$ colour profiles, and  the
          \textbf{lower-right panel} is for $g-i$ colour profiles. 
          We compare our results with those from (1) \citet{Tal2011} based on stacking 
          large samples of luminous red galaxies in SDSS at $z{\sim} 0.4$ 
          (solid red line on the left panel), 
          (2) the results from a stacking analysis of nearby massive 
          galaxies with high concentration index ($C>2.6$) in 
          \citet[][blue dash lines on the left and upper-right panels]{DSouza2014}, 
          (3) and the average $g-r$ and $g-i$ colour profiles 
          from a large sample of nearby elliptical galaxies in \citet[][blue, solid 
          lines on both right panels]{LaBarbera2010}.
          For colour profiles, we also compare with deep observations of a few nearby 
          massive ETGs: the $g-r$ colour profiles of NGC 4472 (\citealt{Mihos2013}) 
          and NGC 4365 (\citealt{Mihos2017}), and the $g-i$ profile of NGC 1399 
          (\citealt{Iodice2016}).
          }
      \label{fig:ell_color}
  \end{figure*}
%% ------------------------------------------------------------------------------------ %%

%% ------------------------------------------------------------------------------------ %%

    \song{
    Now we look into the ellipticity profile derived by \texttt{Ellipse} in $i$-band 
    and the $g-r$ and $g-i$ colour profiles (galactic extinction and $k$-corrected). 
    }
	The smearing effect of seeing makes the central isophotal shape rounder than the 
	real value, while seeing differences between filters bias the central 
	colour\footnote{Since the HSC $i$-band always has better seeing, the central 
	colour becomes redder if seeing effects are not accounted for}. 
	On large scales, it is more difficult to extract reliable ellipticity and 
	colour profiles out to 100 kpc because at low surface brightness levels, the 
	isophotal shape becomes unstable and is easily affected by contamination. 
	Colour profiles are also more difficult to extract out to large radial scales
	because getting the colour accurately depends on the background subtraction and 
	the noise levels in both bands. 
	In this paper, we focus on the median ellipticity and colour profiles between 
	8 and 60 kpc, where we can safely ignore the issues described above.

	The left panel of Fig \ref{fig:ell_color} shows the ellipticity profiles of 
	massive galaxies and highlights the median profiles for the same three \mtot{} 
	bins shown in Fig \ref{fig:avg_prof}.  
	Our results are also compared with previous work based on image stacking 
	techniques (the PSF-removed $i$-band results from \citealt{Tal2011}, and
    concentrated galaxies with $11.0<$\logms{}$<11.4$ from \citealt{dSouza2015}).
	As expected, ellipticity profiles from image stacking methods yield results that 
	are more shallow than ellipticity profiles measured on a galaxy-by-galaxy basis.     
	Uncertainties in how to align galaxies and the intrinsic isophotal twist can 
	lead to this effect.

	In general, we find that the ellipticities of massive galaxies slowly increase 
	with radius. 
	This trend can even be seen directly "by eye" in HSC images (e.g. 
	Fig \ref{fig:sdss_compare}). 
	More interestingly, the ellipticity profiles vary with \mtot{}: 
	at ${\sim} 10$ kpc, the median ellipticity ($< 0.2$) is similar for all three 
	redshift bins, but the ellipticity of the outer stellar halo increases with 
	\mtot{}. 
	Galaxies with \logmtot{}$>11.8$ have median ellipticity profiles that become 
	steeper at $>10$ kpc. 
	The ellipticity of the outer profile steadily increases from 
	$e\le 0.2$ to $e{\sim} 0.4$ at 50-60 kpc.

	This is consistent with studies of nearby massive galaxies using different
	methods (e.g. \citep{Porter1991, Gonzalez2005, Zibetti2005, Huang2013a, Oh2017}). 
	\song{
	For instance, \citet{Spavone2017} uses very deep images of 6 nearby massive 
	galaxies to show that their average ellipticity in $g$-band also increases with 
	radius in a similar fashion.
	}
	However, to the best of our knowledge, our HSC results are the first to show 
	clear evidence that:  
	(1) the ellipticity of stellar halo in massive ETGs depends strongly on 
	\mtot{}, and 
	(2) the ellipticity of stellar halo also relates to the slope of the \mden{}
	profile [see Fig~\ref{fig:avg_prof} and Fig~\ref{fig:ell_color}; 
	we will discuss this more in Huang\etal (in prep.)]. 
	\song{
	As mentioned in \S~\ref{sec:intro}, such a mass-dependent ellipticity profile
	could reveal more insights about the assembly of stellar haloes of massive
	galaxies. 
	}

    \song{
    Regarding the colour profiles, the right panels of Fig \ref{fig:ell_color} show 
    the $k$-corrected $(g-r)$ and $(g-i)$ colour profiles for all galaxies in our 
    sample, together with the median profiles in three \mtot{} bins. 
    We find that the median rest-frame $(g-r)$ and $(g-i)$ colour profiles present 
    smooth and shallow negative colour gradients out to 50--60 kpc, as observed by 
    many previous works (e.g. \citealt{Carollo1993, LaBarbera2012}) and predicted 
    by some simulations (e.g. \citealt{Hirschmann2015}). 
    Meanwhile, there does not appear to be a significant \mstar{} dependence 
    in the gradient of rest-frame optical colours. 
    }

    We also compare our results with the stacked colour profiles from  
    \citealt{LaBarbera2010}\footnote{We use the median colour profiles of high-mass 
    ETGs; the original profile is in units of $R_{\mathrm{e}}$, and we use a typical 
    $R_{\mathrm{e}}=8.0$ kpc to convert it into physical kpc.} and 
    \citealt{DSouza2014}.
    The median colour profiles from HSC images are systematically steeper than the 
    stacked SDSS ones. 
    Considering differences in the response curves between HSC and SDSS filters, 
    together with the uncertainties of colour measurements, the HSC $(g-r)$ colour 
    profiles are in fairly good agreement with those from SDSS. 
    However, the $(g-i)$ profiles of HSC galaxies are steeper compared to SDSS. 
    The SDSS $i$-band suffers from the so-called `red-halo' effect
    (e.g. \citealt{Wu2005}, \citealt{Tal2011}).  
    This is likely due to the fact that the SDSS $i$-band PSF has a more prominent 
    wing than other bands. 
    Because the PSF model does not capture these wings, more flux is artificially 
    distributed to the outskirts, which leads to seemingly redder colours in the low 
    surface brightness outskirts of galaxies. 
    Because HSC uses thick CCDs, HSC $i$-band images do not suffer from this effect 
    and can be used to determine galaxy colours with higher accuracy. 
    Fairly steep colour profiles have been observed in several very nearby massive 
    ETGs.  
    Fig \ref{fig:ell_color} shows the $g-r$ colour profiles of NGC 4472 
    (\citealt{Mihos2013}) and NGC 4365 (\citealt{Mihos2017})\footnote{
    Both are converted from ($B-V$) colours.}, and the $g-i$ profile of NGC 1399 
    (\citealt{Iodice2016}). 
    These individual profiles display colour gradients similar to our HSC sample.

    \song{
    It is worth mentioning that the individual colour profiles still show large 
    scatter and unphysical trends in the outskirt.  
    The colour profile in the low--surface brightness region is very sensitive to 
    the noise level, contamination of other objects, and uncertainties in background 
    subtraction. 
    The HSC $g$-band images are relatively shallower and the backgrounds from the
    two bands are not jointly constrained. 
    Therefore, we only show the average colour profiles here, and will conduct 
    more careful study of the colour profiles in Huang\etal (in prep.).
    }

    % TODO: Move to discussion
    %Ellipticity profiles contain additional information about the assembly histories 
    %and in future work, it would be of interest to compare our ellipticity profiles 
    %with predictions from hydrodynamic simulations such as Illustris 
    %\citealt{Vogelsberger2014}, \citealt{Genel2014}) or \textit{Horizon-AGN}

%% ------------------------------------------------------------------------------------ %%

%% ------------------------------------------------------------------------------------ %% 
  \begin{figure*}
      \centering 
      \includegraphics[width=\textwidth]{fig/redbcg_insitu_accretion}
      \caption{
          \textbf{Left:} Ratio of the fraction of stars between 10 to 100 
          kpc to the total galaxy mass \mtot{}. 
          We adopt this ratio as a proxy for the fraction of  \textit{ex situ} stars 
          in our sample. 
          Typical observational uncertainties are shown in the upper left hand corner. 
          The solid grey line shows the predicted relation  derived from the Illustris
          simulation at $z=0$ (Fig 4 in \citealt{RodriguezGomez2016}). 
          Regions between the grey dashed lines correspond to the range between the 
          16 and 84 percentile of the distribution. 
          \textbf{Right:} comparison between our median \mden{} profiles 
          with the inner component of the structural decomposition of massive 
          elliptical galaxies at $z<0.02$ from \citet[][Cyan, solid]{Huang2013a}.
          At higher redshifts, the \mden{} profiles of massive galaxies should be 
          dominated by the  \textit{in situ}  component. 
          We compare our profiles with the median \mden{} profile of massive galaxies 
          at $1.0 < z < 1.5$ from \textit{HST} observations 
          \citet[][Red, dashed]{Patel2013}. 
          Both these comparisons suggest that the \mden{} profile within 10 kpc is 
          dominated by \textit{in situ} stars, but there are already contributions
          from the accreted stars at very high \mstar{} end. 
          We also compare with the \mden{} profile of a very massive cD galaxy at
          $z{\sim} 1.1$ discovered by \citet[][Yellow, dashed]{Liu2013} in the Hubble 
          Ultra-Deep Field. 
          It is likely that this object will grow into one of the very massive central
          galaxies in our sample.
          It is interesting to see that its \mden{} profile is very similar to the 
          HSC profile of the most massive \mtot{} bin in the inner $\sim 20$ kpc, so 
          the resulting growth should mostly happen in the outskirt. 
          }
      \label{fig:discussion_1}
  \end{figure*}
%% ------------------------------------------------------------------------------------ %%

%% ------------------------------------------------------------------------------------ %% 
\section{Discussion}
    \label{sec:discussion}

    We have used data from the HSC survey that is both simultaneously 
    deep and wide to trace the stellar mass distributions of $0.3 < z < 0.5$ massive 
    galaxies out to $>100$ kpc and to reveal the mass-dependent nature of their 
    stellar haloes.
    Here we briefly discuss the scientific implications of our results.

%% ------------------------------------------------------------------------------------ %% 
\subsection{The formation of massive galaxies and the assembly of their outer haloes}
    \label{ssec:twophase}

    We find that the outer haloes of massive elliptical galaxies grow more prominent 
    and more elliptical with increasing stellar mass.
    According to the two-phase formation scenario, the inner 5--10 kpc of these 
    massive central galaxies are formed at $z>1$ during an intense period of 
    \textit{in situ} star formation. 
    The outskirts of massive galaxies are then built up through a more gradual second 
    phase of evolution (the \textit{ex situ} phase) that is dominated by mass assembly 
    via accretions.
    Non-dissipative mergers, especially minor mergers\footnote{Typically, minor merger
    means the one with stellar mass ratio smaller than 1:3 or 1:4},  deposit stars
    mostly in the outskirts of centrals and do not have a large impact on the central
    \mden{} profile (e.g. \citealt{BoylanKolchin2008, Oogi2013, Bedorf2013}).
    Given the stochastic nature of the merging process, it is easy to understand why 
    the \mden{} profiles of massive galaxies are similar in the inner region but show
    a large scatter in the outer region.

    State of the art hydrodynamic simulations of massive galaxy formation predict 
    that the fraction of accreted stars should strongly increase with stellar 
    mass and that for very massive galaxies, the \textit{ex situ} may reach up to 
    50--90\% of the total galaxy mass (e.g.
    \citealt{Oser2010, Cooper2013, Dubois2013, LeeYi2013, Hirschmann2015,
    RodriguezGomez2016}).

    This picture is supported by our observations that massive galaxies display more 
    prominent stellar haloes as well as by their negative colour gradients. 
    The fact that the outskirts of these galaxies are slightly bluer than the inner
    regions is consistent with the picture that stellar haloes are built up by a
    series of minor mergers (average merger mass-ratio between 1:3 and 1:10; e.g. 
    \citealt{Huang2016}) because less massive ETGs are typically bluer.

    Also according to this picture, the shape of the stellar halo should preserve
    information about the merging history and possibly even about the shape of the 
    dark matter halo. 
    Simulations show the shape of the stellar and dark matter haloes are closely 
    correlated (e.g. \citealt{Wu2014}) for slowly rotating massive ETGs having 
    undergone multiple minor mergers.
    The more elongated outer halo and the trend between the ellipticity profile and 
    stellar mass may reflect the orbital properties of an accreted satellite. 
    In simulations, satellite orbits become more radial for more massive haloes 
    (e.g. \citealt{Murante2007, Wetzel2011, Jiang2015}. 
    Satellites infalling along radial orbits may help to form elongated stellar 
    haloes.
    In principle, these results may explain the trend that we see. 
    However, in contrast, simulation from \citet{Wu2014} predict that more massive 
    galaxies should have \emph{rounder} outskirts, which is the opposite trend 
    compared to HSC. 
    This difference warrants further investigation.

    At larger scales, the distribution of satellite galaxies in massive haloes is 
    found to be aligned with the major axis of the central galaxy 
    (e.g. \citealt{Brainerd2005, Yang2006b, NiedersteOstholt2010, 
    HuangMandelbaum2016}). 
    This alignment signal is found to be stronger for more luminous galaxies 
    living in more massive haloes (e.g. \citealt{Hirata2007}).
    The fact that more massive central galaxies have steeper ellipticity profiles 
    and become increasingly more elongated in the outskirt may be due to 
    such alignment. 
    Moreover, the shape of the dark matter halo can be estimated by analyzing 
    satellite distributions and weak lensing profiles (\citealt{ClampittJain2016}). 
    It is interesting to point out that the most recent measurement by 
    \citet{Shin2017} around SDSS clusters shows a halo axis-ratio of ${\sim}0.55$, 
    which is only slightly more elongated than the stellar halo of galaxies in our 
    highest \mtot{} bin.

%% ------------------------------------------------------------------------------------ %% 
\subsection{Aperture masses as proxies of the \textit{in situ} and accreted stars}
    \label{ssec:insitu}

    \song{
    As mentioned in \S~\ref{sec:intro}, hydrodynamic simulations often predict that 
    the faction of \textit{ex situ} stars increase with total stellar mass.
    However, it is still extremely difficult to separate the \textit{in situ} and
    \textit{ex situ} components observationally, despite the fact that it is the most 
    important theoretical prediction to be verified. 
    }

    \song{
    Recent deep surveys of nearby disc galaxies have started to provide constraints 
    on their stellar haloes (e.g. \citealt{Courteau2011, Merritt2016, 
    Harmsen2017})\footnote{Although it is still not certain that all stellar haloes
    around disc galaxies are made out of \textit{ex situ} stars.}, as it is
    relatively easier to decompose the stellar halo component. 
    As for massive ETGs, limited observational constraints still mainly depend on 
    multi-component model fitting and/or image stacking analyses 
    (e.g. \citealt{Gonzalez2007, Huang2013a, DSouza2014, Spavone2017}).
    }
    In this work, we propose that \mstar{} values computed within different fixed 
    physical elliptical apertures are worth exploring as proxies of the 
    \textit{ex situ} fraction. 
    We propose to use the mass within 10 kpc (\minn{}) and 100 kpc (\mtot{}) as 
    proxies for the \textit{in situ} component and for the total \mstar{}.

    On the left panel of Fig \ref{fig:discussion_1} we show 
    the relation between \mtot{} and the fraction of mass between 10 and 100 kpc 
    (as a proxy of the mass of the accreted component).
    We compare this with the fraction of \textit{ex situ} stars predicted by the 
    Illustris simulation (\citealt{RodriguezGomez2016}).  
    We find that our proxy for the accreted mass component correlates strongly with  
    \mtot{} and that this relation is surprisingly consistent with predictions from
    \citet{RodriguezGomez2016}.  
    Given the limitations of the Illustris simulation and the imperfect nature of 
    our \textit{ex situ} fraction proxy, the almost perfect agreement seen in 
    Fig~\ref{fig:discussion_1} may well be a coincidence.  
    However, it is encouraging to see that our simple proxy for the fraction of 
    \textit{ex situ} stars seems to match both the slope and the scatter of the 
    predicted relation.  
    There are other hydrodynamic simulations that predict significantly different 
    \textit{ex situ} fractions compared to Illustris (e.g. \citealt{Lackner2012, 
    Qu2017}). 
    \song{
    Furthermore, a few known issues in the original Illustris simulation have been 
    resolved in the new Illustris--TNG version (e.g. \citealt{Genel2017, 
    Pillepich2017}.
    In future works [including Huang\etal (in prep.)], we will explore more 
    comparisons with detailed predictions from new simulations to see how well we 
    can trace \textit{ex situ} fractions in massive galaxies using simple 
    elliptical aperture masses.
    }

    \song{
    Here, we further compare our \mden{} profiles with 
    (1) the profiles of the \textit{in situ} component estimated from different ways,
    and (2) profiles of high-redshift massive galaxies that should be dominated by 
    the \textit{in situ} component.  
    }
    In particular, we compare with the following:

    \begin{itemize}

    \item The median \mden{} profiles of massive ETGs at $1.0 < z < 1.5$ from
            \citealt{Patel2013}.
            These are considered to be the progenitors of 
            ${\sim} 10^{11.5} M_{\odot}$ ETGs at $z=0$ and their inner region 
            should be dominated by \textit{in situ} stars.

        \item The inner component of $z{\sim} 0$ ellipticals from the 2-D 
            decomposition of \citet{Huang2013a}. 
            \citet{Huang2013b} show that this inner component is structurally 
            similar to the compact `red nuggets' at high-$z$.

        \item The \textit{in situ} components of simulated central galaxies in 
            massive haloes from \citet{Cooper2013} (the inner ${\sim} 5$ kpc is 
            quite uncertain due to the resolution).  
            These \mden{} profiles are generated using the particle tagging 
            method (see \citealt{Cooper2010})

    \end{itemize}

    \song{
    Although these simple comparisons support the idea that \minn{} mainly consists 
    of \textit{in situ} stars whereas mass at $\mathrm{R} > 15$--20 kpc is dominated 
    by \textit{ex situ} component, they also show that the \texttt{in situ} component 
    could extend beyond 10 kpc\footnote{We convert these \mden{} profiles to the same 
    \citealt{Chabrier2003} IMF; but there are still differences in median \mstar{} 
    and details in the \m2l{} estimates}, and a fraction of mass within 10 kpc could 
    also come from accretion (see also \citealt{RodriguezGomez2016}). 
    This is certainly the limitation for using aperture masses as proxies of 
    \textit{in situ} and \textit{ex situ} masses. 
    In other works (e.g. \citealt{Huang2013a, Spavone2017}), 
    the outer component from image decomposition is treated as proxy of the 
    \textit{ex situ} component. 
    Since such a photometric component extends to the center of the galaxy and can 
    reach a radius much larger than 100 kpc, it can lead to a higher \textit{ex situ}
    fraction than ours (e.g. \citealt{Spavone2017}). 
    It will be very interesting to apply a similar decomposition method to both the 
    HSC images and the simulated massive galaxies to help us determine the 
    best tracer of the \textit{in situ} and \textit{ex situ} mass.
    }

    \song{
    We also compare our results with a uniquely massive galaxy at high redshift: 
    a \logms{}${\sim} 10^{11.4} M_{\odot}$ galaxy with a distinctive `cD'-like 
    envelope at $z{\sim} 1.1$ (\citealt{Liu2013}).  
    This high-redshift galaxy has a \mden{} profile that follows the median \mden{} 
    profile of our $11.6\leq$\logmtot{}$<11.8$ sample nicely at $\mathrm{R} < 40$ 
    kpc, but the profile becomes much steeper in the outskirt.
    The steeper profile suggests that, while the very outskirt of the stellar halo 
    is still being assembled, some massive high-$z$ galaxies may already have a
    significant fraction of \textit{ex situ} stars. 
    In future works [including Huang\etal (in prep.)], we will also look into the 
    redshift evolution of the outer \mden{} profiles up to $z\sim 1.0$ using both HSC 
    and other observations. 
    Comparisons with relevant simulations  can also tell us how well we can trace
    the evolution of the \textit{ex situ} fraction of stars using aperture masses.
    }
%% ------------------------------------------------------------------------------------ %%

%% ------------------------------------------------------------------------------------ %% 
\section{Summary and Conclusions}
    \label{sec:summary}

    In this work, we use deep images from the Subaru HSC survey to study how the 
    projected stellar mass density profiles and other structural properties of 
    massive central galaxies depend on their total stellar mass. 
    With the help of this high-quality, wide-area data set, we directly map the 
    stellar mass distributions of ${\sim}7000$ massive central galaxies at 
    $0.3 < z < 0.5$ out to $>100$ kpc without resorting to stacking techniques. 
    We group massive central galaxies into two categories based on their host halo 
    mass (\mhalo{}$\simgt 10^{14.0} M_{\odot}$ and \mhalo{}$\simlt 10^{14} M_{\odot}$)
    and three bins of \mtot{}.  
    Our main results are:

    \begin{itemize}

        \item We show that the \mden{} profiles of massive galaxies are relatively
            homogenous within 10--20 kpc. 
            However, there is large scatter in outer profiles of massive galaxies. 
            Galaxies with higher \mtot{} show more prominent stellar haloes and have 
            shallower outer \mden{} slopes. 
            Assuming that stellar haloes are dominated by accreted stars, this result 
            is consistent with the two-phase formation picture of massive galaxies.

        \item We show that, on average, massive galaxies have positive ellipticity 
            gradients out to at least 60 kpc. 
            The average ellipticity profile also depends on \mtot{}: More massive 
            galaxies tend to have steeper ellipticity gradients and become more 
            elongated in stellar haloes.

        \item \song{
            On the other hand, although massive galaxies in all three \mtot{} bins 
            show clearly negative average colour gradients out to 60 kpc, the 
            average $(g-r)$ and $(g-i)$ colour gradients do not show clear 
            dependence on \mtot{}. 
            }

        \item \song{
            We propose that, under reasonable assumptions, stellar masses within 
            different physical apertures could be used to trace the 
            \textit{in situ} and \textit{ex situ} components. 
            Here we show that \minn{} and the \mstar{} between 10 and 100 kpc could 
            be promising proxies for the two physical components, and can recover the 
            trend between stellar mass and \textit{ex situ} fraction from simulation.
            }

        \item We find that the `total' \mstar{} of these massive galaxies can be 
            significantly underestimated with shallow imaging data such as SDSS and/or 
            oversimplified model assumptions (e.g. the \texttt{cModel} or single-\ser). 
            In contrast to previous work, our results do not depend on stacking or any 
            parametric models. 
            Moreover, the degree to which stellar mass is underestimated depends on 
            \mtot{}. 
            A simple model misses more light for massive galaxies because they have 
            more extended envelopes. 
            There is also an \mhalo{}-dependence of this effect, and simple 
            luminosity estimates will miss more light for BCGs in more massive 
            haloes compared to centrals in less massive haloes.
            These effects need to be carefully taken into account when discussing 
            topics such as the evolution of the galaxy stellar mass function.

    \end{itemize}

    These results highlight the advantages of wide-area, deep, and high-quality imaging 
    for studying the evolution of massive galaxies. 
    At present, the HSC survey has already doubled its sky coverage to 
    ${\sim} 200$ deg$^2$ and provides a much larger sample of massive central galaxies. 
    In the near future, we will extend this work to lower \mtot{} by using photometric 
    redshifts, and we will also apply 2-D photometric methods 
    (e.g.\ \citealt{Huang2013a}) to take advantage of the multi-wavelength nature of 
    the HSC survey (e.g. \citealt{Huang2016}). 
    Our current work can also be combined with weak lensing measurements of the dark 
    matter haloes of massive galaxies, and physical insights into the assembly histories 
    of these galaxies can be gained by comparing the measurements with cosmological 
    hydro-simulations such as Illustris (\citealt{Vogelsberger2014}, 
    \citealt{Genel2014}), EAGLE (\citealt{Schaye2015}, \citealt{Crain2015}), 
    or \textit{Horizon-AGN} (\citealt{Dubois2014}).

%% ------------------------------------------------------------------------------------ %%

\section*{Acknowledgements}

  % Personal 
  The authors thank Rachel Mandelbaum and Frank van~den~Bosch for insightful 
  discussions and comments,
  Shun Saito for helping us estimate the fraction of satellite 
  galaxies in our sample, and
  Feng-Shan Liu for sharing the \mden{} profile of the $z\sim1$ BCG from 
  his work.

  % HSC part
  The Hyper Suprime-Cam (HSC) collaboration includes the astronomical communities of 
  Japan and Taiwan, and Princeton University.  The HSC instrumentation and software 
  were developed by  National Astronomical Observatory of Japan (NAOJ), Kavli 
  Institute for the Physics and Mathematics of the Universe (Kavli IPMU), University 
  of Tokyo, High Energy Accelerator Research Organization (KEK), Academia Sinica 
  Institute for Astronomy and Astrophysics in Taiwan (ASIAA), and Princeton 
  University.  
  Funding was contributed by the FIRST program from the Japanese Cabinet Office;  
  Ministry of Education, Culture, Sports, Science and Technology (MEXT);  
  Japan Society for the Promotion of Science (JSPS); 
  Japan Science and Technology Agency (JST); 
  Toray Science Foundation; NAOJ; Kavli IPMU; KEK; ASIAA; and Princeton University.

  % SDSS part
  Funding for SDSS-III has been provided by Alfred P. Sloan Foundation, 
  Participating Institutions, National Science Foundation, and  U.S.  Department of
  Energy. The SDSS-III web site is http://www.sdss3.org.  SDSS-III is managed by 
  Astrophysical Research Consortium for the Participating Institutions of the SDSS-III
  Collaboration, which includes  University of Arizona,  Brazilian Participation Group,
  Brookhaven National Laboratory, University of Cambridge, University of Florida, the
  French Participation Group,  German Participation Group, Instituto de Astrofisica
  de Canarias,  Michigan State/Notre Dame/JINA Participation Group, Johns Hopkins
  University, Lawrence Berkeley National Laboratory, Max Planck Institute for
  Astrophysics, New Mexico State University, New York University, Ohio State University,
  Pennsylvania State University, University of Portsmouth, Princeton University, the
  Spanish Participation Group, University of Tokyo, University of Utah, Vanderbilt
  University, University of Virginia, University of Washington, and Yale University.

  % Pan-STARRS1 part
  The Pan-STARRS1 Surveys (PS1) have been made possible through contributions of  
  Institute for Astronomy;  University of Hawaii;  Pan-STARRS Project Office; 
   Max-Planck Society and its participating institutes:  Max Planck Institute 
  for Astronomy, Heidelberg, and the Max Planck Institute for Extraterrestrial Physics, 
  Garching;  Johns Hopkins University; Durham University;  University of Edinburgh; 
  Queen's University Belfast;  Harvard--Smithsonian Center for Astrophysics;  Las 
  Cumbres Observatory Global Telescope Network Incorporated;  National Central 
  University of Taiwan;  Space Telescope Science Institute; National Aeronautics 
  and Space Administration under Grant No. NNX08AR22G issued through the Planetary 
  Science Division of the NASA Science Mission Directorate; National Science 
  Foundation under Grant No. AST-1238877; University of Maryland; and Eötvös Loránd 
  University (ELTE).

  % LSST software
  This paper makes use of software developed for the Large Synoptic Survey 
  Telescope. We thank the LSST Project for making their code available as free 
  software at http://dm.lsstcorp.org.

  % KITP
  This research was supported in part by National Science Foundation under Grant 
  No. NSF PHY11-25915.

  % Software
  This research made use of:
  \href{http://www.stsci.edu/institute/software_hardware/pyraf/stsci\_python}
  {\texttt{\small{STSCI\_PYTHON}}},
      a general astronomical data analysis infrastructure in Python. 
      \texttt{\small{STSCI\_PYTHON}} is a product of the Space Telescope Science
      Institute, which is operated by Association of Universities for Research 
      in Astronomy (AURA) for NASA;
  \href{http://www.scipy.org/}{\texttt{SciPy}},
      an open source scientific tools for Python (\citealt{SciPy});
  \href{http://www.numpy.org/}{\texttt{NumPy}}, 
      a fundamental package for scientific computing with Python (\citealt{NumPy});
  \href{http://matplotlib.org/}{\texttt{Matplotlib}}, 
      a 2-D plotting library for Python (\citealt{Matplotlib});
  \href{http://www.astropy.org/}{\texttt{Astropy}}, a community-developed 
      core Python package for astronomy (\citealt{AstroPy}); 
  \href{http://scikit-learn.org/stable/index.html}{\texttt{scikit-learn}},
      a machine-learning library in Python (\citealt{scikit-learn}); 
  \href{http://www.astroml.org/}{\texttt{astroML}}, 
      a machine-learning library for astrophysics (\citealt{astroML});
  \href{https://ipython.org}{\texttt{IPython}}, 
      an interactive computing system for Python (\citealt{IPython});
  \href{https://github.com/kbarbary/sep}{\texttt{sep}} 
      Source Extraction and Photometry in Python (\citealt{PythonSEP});
  \href{https://jiffyclub.github.io/palettable/}{\texttt{palettable}},
      color palettes for Python;
  \href{http://dan.iel.fm/emcee/current/}{\texttt{emcee}}, 
      Seriously Kick-Ass MCMC in Python;
  \href{http://bdiemer.bitbucket.org/}{\texttt{Colossus}}, 
      COsmology, haLO and large-Scale StrUcture toolS (\citealt{Colossus}).

%%%%%%%%%%: Bibliographic Section %%%%%%%%%%

\bibliographystyle{mnras}
\bibliography{redbcg}

%%%%%%%%%%: Appendix Section %%%%%%%%%%%%

\appendix

%% ------------------------------------------------------------------------------------ %% 
  \begin{figure}
      \centering 
      \includegraphics[width=\columnwidth]{fig/redbcg_sample_stats}
      \caption{
          \textbf{Top-left}: The \logmtot{}-$(g-r)$ rest-frame colour relation for the 
          HSC massive galaxies.
          We $k$-correct the colour using the \texttt{iSEDFit} fitting results.
          Massive galaxies form a `red-sequence' on this figure, and there is 
          little contamination from blue object at high-mass end.~~          
          \textbf{Top-right}: The redshift distribution of the massive galaxies. 
          The filled and empty histograms are for the \logmtot{}$>11.6$ and 
          \logmtot{}$>11.2$ galaxies.
          The vertical lines highlight the $0.3\leq z \leq 0.5$ redshift range.~~
          \textbf{Bottom-left}: The distributions of \mtot{} of massive galaxies in 
          this sample. 
          Filled histogram shows the distribution for $0.3 < z < 0.5$ galaxies 
          used in this work. 
          The empty histogram shows the distribution for the whole $z>0.2$ sample
          as comparison.~~
          \textbf{Bottom-right}: The distributions of \minn{} in similar format. 
      }
      \label{fig:sample_stats}
  \end{figure}
%% ------------------------------------------------------------------------------------ %%

%% ------------------------------------------------------------------------------------ %% 
    \begin{figure*}
        \centering 
        \includegraphics[width=\textwidth]{fig/redbcg_ellipse_tech}
        \caption{
            \textbf{Left:} Example of the object-mask built for the \texttt{Ellipse}
            run for a typical massive galaxy in the sample. 
            All the shaded regions are masked out. 
            The three dashed lines (one red inner line and two blue lines) around 
            the target at the center outlines the three radii we defined using the 
            flux radius of the target.  
            We increase the mask size for objects detected in different regions 
            separated by these apertures (which are outlined by solid, elliptical 
            apertures with different colours) using slightly different criteria.~~
            \textbf{Middle:}: The zoom-in intensity profile around very low intensity 
            value (top panel), and the curve-of-growth of the enclosed magnitude 
            (bottom panel) of the example galaxy.  
            To highlight the importance of background correction, we show the profiles 
            using both images with (red solid line) and without (black dotted line) 
            background correction. 
            On the top panel, besides the horizontal line that highlights the zero flux 
            level, we also show the uncertainty of the sky background estimate using 
            the grey-shaded region.  
            On the bottom panel, two horizontal lines indicate the magnitudes 
            corresponding to total flux (solid) and flux within 100 kpc (dashed).~~
            \textbf{Right:} Compares the 1-D surface brightness profiles for the same 
            example galaxy using different masks 
            (smaller masking region: red dashed line; larger masks: blue dashed line), 
            or different \texttt{Ellipse} configurations
            (more aggressive pixel-clipping: cyan dashed line; 
            larger step in radius: green dashed line; 
            using mean flux along the isophote instead of median: purple dashed line)
            with the default one (black solid line).
            }
        \label{fig:ell_tech}
    \end{figure*}
%% ------------------------------------------------------------------------------------ %%

%% ------------------------------------------------------------------------------------ %%  
\section{Basic Statistical Properties of the Sample} 
	\label{app:basic}

    Here we show the basic statistics of the massive galaxies used in this work.
    On the top-left panel of Fig~\ref{fig:sample_stats}, we show the \mtot{}-colour 
    relation using the $k$-corrected rest-frame $(g-r)$ colour. 
    These massive galaxies form a clear `red sequence' with little contamination 
    from the `blue cloud' at the very high-mass end.

    In the rest of Fig~\ref{fig:sample_stats}, we also show the distributions of 
    redshift, \mtot{}, and \minn{}.
    In this work, we focus on the massive galaxies with \logmtot{}$>11.6$ at 
    $0.3 < z < 0.5$ where the sample is fairly complete in \mtot{}.

%% ------------------------------------------------------------------------------------ %% 
\section{Extraction of 1-D surface brightness profile} 
    \label{app:ellipse}

    Here we briefly discuss a few technical issues related to the measurements of the 
    1-D surface brightness profiles around massive galaxies.

    To derive a reliable 1-D profile, it is important to mask out all the irrelevant 
    objects around the target.
    At the depth of the HSC images, this becomes a challenging task, especially 
    for massive galaxies with extended outer profiles and many satellites. 
    At this point, the \texttt{hscPipe} tends to over-subtract the background around 
    bright objects.  
    In addition, the \texttt{hscPipe} deblending process performance is not optimized 
    for extended objects. 
    For these reasons, we perform \texttt{SExtractor}-like background subtraction and 
    object detection using the \texttt{SEP} Python library to generate the necessary 
    masks.
    Combining two different local background models and $S/N$ thresholds, we obtain 
    the centroid, shape, and radius that together encloses 90\% of flux for each 
    object, including the one that is very close to the center of bright galaxy 
    (left panel of Fig~\ref{fig:ell_tech}). 
    Based on this information, we then create the mask that covers all contaminating 
    objects around the target after adaptively increasing the sizes of their masks 
    according to their brightness and distance to the central target. 
    Generally speaking, we more aggressively mask out bright objects or objects in 
    the outskirt of the image to reduce their impact on the surface brightness 
    profiles in the outskirt. 
    We also create masks that are less and more aggressive than the default mask to 
    test their impacts on the surface brightness profiles.

    Next, we aggressively mask out all objects on the cut-out image.  
    We then evaluate the background level using the unmasked pixels after median 
    smoothing the masked image using box of $6x6$ pixels.
    This provides estimate of global background level along with its uncertainty. 
    Given the typical background uncertainty, the HSC \texttt{WIDE} image should be 
    able to reach down to $> 29$ \sb surface brightness level in the $i$-band.  
    However, as mentioned, we often find evidence of slightly over-subtracted 
    backgrounds for massive galaxies in our sample. 
    In the current \texttt{hscPipe}, the background on each CCD is modelled with a 
    Chebyshev-polynomial that is fit to the smoothed image after excluding pixels 
    with $S/N >5$.
    This algorithm performs much better than the SDSS version 
    (e.g.\ see \citealt{Blanton2011}), yet still over-subtracts background around 
    bright objects and results in unphysical truncation in their surface brightness 
    profiles.
    We empirically correct this issue using the background model generated by 
    the \texttt{SExtractor} algorithm on the masked image 
    ($200x200$ pixels background box size, and 6 pixels median filtering size of 
    sky boxes).
    This model can account for the slightly over-subtracted background at large 
    scale and reduce the impact from the low--surface brightness `wings' of bright 
    neighbors. 
    We clearly see improvement in both the distributions of background pixels 
    (more symmetric distribution; median value is closer to 0) and the surface 
    brightness profile (middle panel of Fig~\ref{app:ellipse}; the negative 
    intensity and the turn-over of the curve-of-growth in the outskirt of the 
    `Original' profile are successfully corrected) after this correction.
    Also, it is worth mentioning that such correction does not often affect the 
    surface brightness profile within 100 kpc.

    The procedure used to derive 1-D surface brightness profile from the 
    background-corrected, contamination-masked images is  described briefly in 
    \S~\ref{sec:ellipse}. 
    \song{
    The uncertainties of the surface brightness profiles include the error in 
    isophotal intensity derived by the \texttt{ellipse} procedure and the 
    uncertainty from background fluctuation estimated by the distribution of the 
    rebinned background pixels using the above method. 
    }
    In practice, the profile at very low surface brightness level is sensitive to 
    several \texttt{Ellipse} configurations.
    After some tests, we choose to use 0.1 dex in logarithm as the step in semi-major 
    axis length between successive ellipses, and we use the median pixel value over 
    the elliptical annulus after rejecting outlying pixels via 3-$\sigma$ clipping 
    three times.
    We make these  choices to make the final profile less affected by any nearby
    object and also to test the differences between the profiles derived using larger
    step, or mean value on the annulus, or fewer times of $\sigma$-clipping. 
    Generally speaking, the surface brightness profile is very robust against these
    changes, especially within 100 kpc. 
    On the right panel of Fig~\ref{fig:ell_tech}, we compare the surface brightness
    profiles for an example massive galaxy using different masks and \texttt{Ellipse}
    parameters. 
    The profile within 100 kpc is very stable, and the only noticable difference 
    is caused by the less aggressive object-mask in the outskirt.

    We should also mention that we run \texttt{Ellipse} to allow for more 
    sophisticated shapes than simple ellipse (4th Fourier modes that can make 
    isophote more `disky' or `boxy', e.g.\ \citealt{Kormendy2009}) to fit the 
    isophote better.
    We also apply the isophotes from $i$-band images to other bands in a
    `force-photometry' mode \texttt{Ellipse} run to get initial estimates of 
    colour profiles.

    \song{
    We cannot extract reliable 1-D profiles for a small fraction of massive
    galaxies because they are heavily masked out for either physical 
    (e.g.\ late-stage major merger) or nuisance (e.g.\ nearby foreground 
    galaxy or bright star) reasons. 
    This is an intrinsic limitation of the 1-D method, and it removes ${\sim}10$\% 
    of the sample.
    We visually examine the 3-colour images of randomly selected galaxies with 
    failed 1-D profiles.  
    Most of them are relatively small galaxies that are severely contaminated by 
    nearby objects and do not affect the results of this work.  
    Meanwhile, it is worth noting that this work does exclude most major merging 
    systems among massive galaxies. 
    }

%% ------------------------------------------------------------------------------------ %%

%% ------------------------------------------------------------------------------------ %% 
\section{Estimate average {$M_{\star}/L_{\star}$} using \texttt{iSEDFit}} 
    \label{app:sed}

%% ------------------------------------------------------------------------------------ %% 
    \begin{figure*}
        \begin{center}
        \includegraphics[width=\textwidth]{fig/redbcg_isedfit_example.pdf}
        \caption{
            \textbf{Left:} Example of output figure from \texttt{iSEDFit} that shows 
            the SED fitting results. 
            The open boxes show the observed fluxes in the five bands, and the solid 
            blue dots show the best-fitting results, along with the high-resolution
            spectrum for this model reconstructed using the synthetic spectra from
            \texttt{FSPS}. 
            The top-left corner shows the best-fitting stellar population parameters, 
            and the bottom-right corner shows the ID, redshift of this object, and 
            reduced $\chi^2$ of the best-fitting model.~~~
            \textbf{Right:} the posterior distributions of a few key
            parameters.
            From top-left to bottom right are: 
            (1) stellar mass (\logms{}); 
            (2) age of the population (mass and star-formation rate weighted) in Gyr; 
            (3) star formation rate ($\log\ \mathrm{SFR}\ (M_{\odot}/\mathrm{yr})$; 
            instant one and the one averaged over the previous 100 Myr; 
            (4) stellar metallicity ($\mathrm{Z}/\mathrm{Z}_{\odot}$); 
            (5) dust extinction ($\mathrm{A}_V$ in mag);
            (6) birthrate parameter ($\log\ b$; averaged over previous 100 and 
            1000 Myr).
            }
        \label{fig:ised_1}
        \end{center}
    \end{figure*}
%% ------------------------------------------------------------------------------------ %%

%% ------------------------------------------------------------------------------------ %% 
    \begin{figure*}
        \begin{center}
        \includegraphics[width=12cm]{fig/redbcg_isedfit_2.pdf}
        \caption{
            Relationships between \mstar{} and key stellar population parameters from 
            \texttt{iSEDfit}. 
            The four stellar population properties are: 
            \textbf{Top-left}: \mstar{}-weighted stellar population age 
            in Gyr. 
            \textbf{Bottom-left}: SFR-weighted age in Gyr. 
            \textbf{Top-right}: \mstar{}-weighted stellar metallicity in 
            unit of solar value.
            \textbf{bottom-right}: Dust extinction value in $V$-band.
            As expected, most of the HSC massive galaxies are old, metal-rich, and 
            dust-free. 
            }
        \label{fig:ised_2}
        \end{center}
    \end{figure*}
%% ------------------------------------------------------------------------------------ %%

%% ------------------------------------------------------------------------------------ %% 
    \begin{figure}
        \begin{center}
        \includegraphics[width=\columnwidth]{fig/redbcg_isedfit_3.pdf}
        \caption{
            Comparisons of \mstar{} estimated by \texttt{iSEDFit} using different
            model assumptions. 
            In each panel, we plot the \mstar{} from the default model against the
            differences with four other models:
            (A): Default model; 
            (B): Using the Salpeter IMF instead of the Chabrier IMF 
                (\textbf{Top panel}) ;
            (C): Using the BC03 synthetic population model instead of the FSPS one
                (\textbf{Middle panel});
            (D): No dust extinction (\textbf{Bottom panel}). 
            On each panel, the grey-shaded region highlights the typical uncertainty 
            of the $\log($\logms{}$)$.            
            For each pair of models, we highlight the median $\chi^{2}$ values and 
            the fraction of galaxies with $\chi^{2} > 10.0$ at the top. 
            On each panel, we also show the histograms of the \mstar{}-differences on 
            the right side.
            }
        \label{fig:ised_3}
        \end{center}
    \end{figure}
%% ------------------------------------------------------------------------------------ %%

%% ------------------------------------------------------------------------------------ %%

    In \S~\ref{ssec:isedfit}, we briefly explain the SED fitting procedure and 
    the priors used.   
    \song{
    \texttt{iSEDFit} takes a simplified Bayesian approach. 
    In short, it first generates a large grid of SEDs from synthetic stellar 
    population models by drawing randomly from the prior distributions of relevant
    parameters (e.g. age, metallicity, dust extinction, and star formation history).
    Based on these models, it uses the observed photometry and redshift to compute 
    the statistical likelihood and generates the posterior PDF for each parameter.  
    To get the best estimate of a given parameter, \texttt{iSEDFit} integrates the 
    full PDF over all the other nuisance parameters.
    Then, the median value and the 1-$\sigma$ uncertainty are derived based on the 
    marginalized PDF. 
    Please refer to \citet{Moustakas13} for technical details. 
    }
    In Fig~\ref{fig:ised_1}, we show an example of the \texttt{iSEDFit} output by 
    visualizing the five-band HSC SED on top of the best-fitting model along with the PDF 
    of the key parameters.

    Although we only use the best-fitting \m2l{} in this work, it is necessary to make 
    sure the model is reasonable. 
    We show the relationship between \mstar{} and a few key stellar population parameters 
    derived by \texttt{iSEDFit} in Fig~\ref{fig:ised_2}. 
    Degeneracies among these parameters are inevitable based on only five broad-band
    photometry; but as expected, most massive galaxies show old stellar age, high
    stellar metallicity ($1.5 \times Z_{\odot}$ is the highest metallicity allowed by 
    the adopted \texttt{FSPS} SSP models), and low dust extinction.

    Meanwhile, \mstar{} measurement based on SED fitting heavily depends on the 
    adopted SSP model, the form of IMF, dust extinction law, and details in 
    the assumption of SFH (e.g.\ \citealt{Bernardi2017}). 
    For massive galaxies in this sample, the form of the SFH\footnote{We choose 
    to use the delayed-$\tau$ model for SFH; we adopt a flat distribution between 
    0.5 and 14.0 Gyrs as the prior for the look-back time when the star formation 
    turned on. 
    The exponential delayed time-scale ($\tau$) is allowed to change between 
    0.1 and 3.0 with equal probability}, and the contribution from random star 
    burst\footnote{The chance of random star burst is set at 0.2 for every 2 Gyrs. 
    The duration of the star burst is drawn from a logarithmic distribution 
    between 0.03 and 0.3 Gyr; and the mass fraction formed in the burst is from 
    a logarithmic distribution between 0.01 and 1.0.} rarely affect the \mstar{}. 
    But the choices of SSP model, IMF, and dust extinction do systematically impact
    the estimates of \mstar{}; therefore, we look into this with a few 
    additional tests (see Fig~\ref{fig:ised_3}):

    \begin{itemize}

        \item Choosing the \citet{Salpeter1955} IMF results in systematically 
            higher \mstar{} (on average $+0.25$ dex of \logms{}) for these massive 
            galaxies (top panel).
            Although there are multiple lines of evidence that favor Salpeter 
            or an even more `bottom-heavy' IMF in massive galaxies 
            (e.g.\ \citealt{Conroy2012}; \citealt{Cappellari2012}), we still 
            present the main results using Chabrier IMF to accommodate galaxies 
            with lower \mstar{} in the sample, and to be as consistent as possible 
            with previous work.  
            This choice of IMF does not change the main results qualitatively.

        \item \mstar{} based on the \texttt{BC03} models are systematically lower 
            than the ones based on \texttt{FSPS+MILES} models (middle panel)
            The difference shows a large scatter, and can be as large as 0.4 dex,
            although it is not \mstar{}-dependent. 
            The \texttt{BC03} results show a better average ${\chi}^2$ than the 
            \texttt{FSPS} ones. 
            This relates to the higher upper-limit of stellar metallicity 
            ($2 \times \mathrm{Z}_{\odot}$) allowed by the \texttt{BC03} model, 
            which helps fit the shape of the SED in the red-end slightly better.  
            However, the \texttt{BC03} results also show puzzlingly low stellar 
            ages ($< 3$--4 Gyrs) for these massive, red galaxies. 
            This could also lead to underestimated \m2l{} values.
            It is worth noting that both \texttt{FSPS} and \texttt{BC03} 
            models still have difficulties recovering SED at the very red-end 
            (between $z$ and $y$-band) and reproducing the optical colour--colour 
            relations for red-sequence galaxies (e.g.\ \citealt{MIUSCAT2}).
            In this work, we decide to keep using the \texttt{FSPS+MILES} model as 
            the fiducial one.  
            Using results based on the \texttt{BC03} model does not change any of 
            our conclusions here.

        \item On the bottom panel of Fig~\ref{fig:ised_3}, we compare the default 
            model with the SED fitting results without considering the dust 
            extinction. 
            This choice leads to slightly smaller \mstar{} values as expected. 
            Its impact becomes slightly larger at lower \mstar{} end. 
            It does not change any of our conclusions here.

    \end{itemize}

    \song{
    We should also point out that our methodology ignores radial variations in \m2l{}. 
    It is well known that massive ETGs have negative optical colour gradients, 
    indicating gradients in \m2l{} (e.g.\ \citealt{Carollo1993, Davies1993, 
    LaBarbera2012, dSouza2015}). 
    Assuming all massive galaxies in our sample have negative colour gradients and 
    there is a simple monotonic relationship between optical colour and \m2l{}, the
    average \m2l{} we use should in principle underestimate the \mstar{} in the 
    center while overestimating the \mstar{} in the outskirt.
    However, these colour gradients are shallow and they smooth out to a few times the 
    effective radius (e.g. \citealt{LaBarbera2010, Tal2011, DSouza2014}, 
    colour gradients at larger radii are not yet well quantified). 
    Because the gradients are shallow, using an average \m2l{} is unlikely to bias 
    our results on \mstar{} measurements.  
    In \citet{Huang2016}, the authors conduct multi-band decomposition for a sample 
    of very nearby elliptical galaxies and estimate the \m2l{} of each component 
    separately. 
    The sum of all components suggests a slightly higher \mstar{} ($0.05-0.10$ dex 
    when typical uncertainty of \mstar{} is 0.12--0.15 dex) and the mass differences 
    show no dependence on \mstar{}. 
    Colour gradients are discussed in \S~\ref{ssec:ell_color}. 
    In summary, our results about the mass dependence of \mden{} profiles should not 
    be affected by the assumption of a constant \m2l{} ratio because optical colour 
    gradients in our sample do not show a dependence on stellar mass.
    }

%% ------------------------------------------------------------------------------------ %%

%% ------------------------------------------------------------------------------------ %% 
%% Table.1 
%% ------------------------------------------------------------------------------------ %% 
\clearpage
\include{table1}
\clearpage
%% ------------------------------------------------------------------------------------ %%

%% ------------------------------------------------------------------------------------ %%  
\bsp
\label{lastpage}
\end{document}
%% ------------------------------------------------------------------------------------ %% 
%%%%%%%%%%%%: End of the File %%%%%%%%%%%%
%% ------------------------------------------------------------------------------------ %%