### 2017-11-28 draft/response_paper_II.md
Updated response to referee report of Paper II

# Response to Referee Report of Paper II

We are very grateful for the positive referee report and many useful comments from the referee.  We have modified the draft accordingly and have highlighted all the modifications in cyan.

Here we will also address each of the referee's questions in details.

## MODERATE COMMENTS

> 1. First off, a difference of ~20\% is not very large, although a larger difference might be hidden by observational errors in stellar masses, halo masses, and sizes. Some Monte Carlo analysis+discussion on this point should be included (see also point 3 below). Would the Authors expect an intrinsically larger size difference? This is important for models. Although a 20\% systematic difference is predicted by basic hierarchical models (e.g., Shankar et al. 2015), a larger difference of up to a factor of 3 in sizes may be expected if models include processes such as gas dissipation and significant disc instabilities (e.g., Shankar et al. 2014).

The referee is correct that the difference in effective radius is not **very** significant. For the M_100kpc matched samples, we use bootstrap resampling method to see the uncertainties of the average size.  As for the mass-size relation, we use MCMC-Bayesian to fit the relation after considering the uncertainties of both mass and size.  As we show on the figure, the 1-sigma uncertainty regions of the mass-size relations for two samples slightly overlap.

At the same time, we want to emphasize that we consider the direct comparison of mass density profiles (which is the more important results for this work as we can directly map the mass distribution out to >100 kpc) a more reliable and straightforward way to reveal environmental difference.  We also use resampling method and statistical tests to prove that the subtle differences from the comparisons of mass density profiles are robust and significant.  The stellar mass distributions of massive galaxies do depend on halo mass based on our results, although it does not guarantee a very significant difference on the mass-size plane. This is likely because the environment-dependence of mass density profile becomes more significant in the low-density outer halo, but the value of effective radius is not very sensitive to such region.

We do not have any expectation for the size difference at fixed stellar mass.  The comparison of mass density profile seems to show that, although the environment-dependence is significant, environment (or halo mass) still does not play the most crucial role in shaping these massive galaxies.  As for mass-size relation, within the 1-sigma uncertainties allowed by our fitting results, the maximum size difference is still not as significant as predicted by some of the models.

We do appreciate the referee to draw our attention to the Monte-Carlo tests in Huertas-Company+2013. We now mention this critical point in the text.

> 2. When plotting R50 vs Mstar one should also keep in mind that because of the steepness of this relation, any secondary binning (e.g., per Mhalo) may artificially induce spurious differences. A good way to handle this is to follow the procedure by Newman+12 and Huertas-Company+13 who proposed to weigh each size by its median value for its stellar mass ("gamma"). By simply plotting data in this way, gamma vs Mstar or gamma/<gamma> vs Mstar/Mhalo, may erase part of the apparent discrepancy in the R50 vs Mstar plane. It would be nice to see a plot of gamma vs Mhalo, at say fixed M100kpc, if possible...

We thank the referee for the useful suggestion and expand the relevant section to address these issues.  
1) We compare the average differences for the stellar mass and redshift matched samples, and confirm the ~20% size difference.  The mass-matched samples can help handle this issue. 
2) We adopt the definition of gamma in Huertas-Company+13 and estimate the normalized sizes for the two samples using their best-fit slopes of the mass-size relations.  The median gamma for $>10^{14.2}M_{\odot}$ and $<10^{14.0}M_{\odot}$ samples are $4.2\pm0.4$ and $3.8\pm0.3$.  Although the uncertainties are still quite large, it also confirms the result. 
3) Also, we would like to mention that the direct comparison of mass density profiles also confirms the same conclusion.  We do not need to worry about the uncertainty of size there, and only need to consider the uncertainty of the mass-to-light ratio from SED fitting.

Unfortunately, we can only broadly group our sample into two halo mass bins, and we do not have the accurate estimate of the average halo mass for the sample of less massive haloes.  Therefore, it will not be very useful (and reliable) to put them on the figure proposed by the referee.  In the near future, we will use better group catalog and weak lensing measurements from HSC to further look into this relation. Furthermore, we are in the process of building a full forward model of the mass - size - halo mass relation which we are fitting to lensing results. This forward modeling approach has the advantage that it naturally accounts for the secondary binning issues brought up by the referee. Preliminary results from this forward fitting (including lensing) suggest that our results are robust to these effects,

> 3. I find the comparison in Figure 3 (left) with Huertas-Company et al. (2013; HC13) quite intriguing. The median relation by HC13 fits in well with the median relation found by the Authors in higher-mass haloes. If HC13 were missing a large portion of the extra light, I would expect, if anything, a lower median R50-Mstar relation. Besides, as the Authors also point out, HC13 is in good agreement with their results below Mstar~3e11 Msun. Also to note that the background should not be an issue here as Re in centrals and satellites appear similar in HC13 (see their Figure 2 and also Bernardi+17). I am thus not sure why there should be such a systematic difference only in galaxies above 3e11 Msun and in lower mass haloes. More discussion and analysis is needed here.

The referee is right that this is indeed an important point left out in our discussion.  We think that it relates to 1) the redshift difference (HC13 is at z<0.09; our sample has average redshift around 0.4); 2) the uncertainties in both mass and size measurements.  As suggested by the referee, now we include the mass-size relation at similar redshift from Huertas-Company+2013b to the figure, which clearly lies below our mass-size relation.  We would also like to point out that both the stellar mass and effective radius in Huertas-Company+2013 are based on single-Sersic model fitting on shallower SDSS images, which could lead to systematic differences or even bias (as discussed in our Appendix F using the GAMA sample).  It could mean the single-Sersic model on SDSS image could systematically underestimate the mass or overestimate the size (especially for galaxies with high Sersic indices).  This demonstrates the importance of systematics like the imaging depth and model assumption when it comes to the comparisons of mass-size relation.  It is possible that the HC13 result also slightly underestimates the size of z~0 galaxies.  It would be interesting to confirm this using HSC images.  This effort is complicated by issues like saturation on HSC images and over-subtracted background, and is beyond the scope of this work.

In tests which we did not include in the paper, we also fit single-Sersic models to the HSC images.  These models show larger size than the ones measured using 1-D profile for the same massive galaxies, and they often have a very high Sersic index. 
We do not include these results because, at the imaging depth of HSC, the single-Serisc model performs poorly for these galaxies.  We agree with the referee that this point deserves further investigation.  Currently, we are working with a new student to test the mass and size measurements for the same galaxies using different methods and different images.

> 4. HC13b also present results for higher-z galaxies, would it be possible to show some of these in the same Figure?!?  Finally, note that in section 4.2 HC13 state from results of a number of Monte Carlo tests that "...we should be able to detect any size difference greater than a factor of 1.4 between galaxies in extreme environments." This implies that errors could easily wash out a difference of 20\% in their data... All these points must be included in the discussion.

We agree with the referee that comparisons of mass-size relations at different redshifts can greatly help us look into the origin of any environment dependence. 
It is particularly intriguing since several past papers have found a stronger environment dependence of size at high redshift.  We discussed this in Section 4.3, and now we expand the discussion following the referee's suggestion.  We point out that, if these results are robust, it could mean that denser environment at high-redshift accelerate galaxy evolution.

As suggested by the referee, we now try to include the results from Huertas-Company+2013b into the comparison.  On the updated Fig 3, we show the mass-size relation of ETGs in group at 0.2<z<0.5 from Huertas-Company+2013b.  We also correct the systematic difference in stellar population model, and also convert the circulized effective radius into the measured ones after assuming an average b/a of 0.75.  And, regardless of these technical details, we find the Huertas-Company+2013b relation has a similar slope with our results, but shows systematic offset (smaller sizes). This may relate to the fact that the HST/ACS F814W images are much shallower than the HSC ones.  We did not include results from higher redshift, and we think the current figure already demonstrates our concerns for such comparisons:

1) Crucial systematics such as the imaging depth and modeling method must be carefully addressed before we can perform reliable comparison and look into the redshift evolution.  This is beyond the scope of this work. 
2) High redshift observations often use small sample or very special environment (e.g. single massive proto-cluster).
3) It is hard to know the evolutionary connections between the high-z samples and our ones.  Progenitors of our massive galaxies could be disc or star-forming galaxy. 
4) At high redshift, it is even more challenging to measure mass and size reliably.  
5) At the same time, there are also works that find no environment dependent on size for high redshift ETGs.

We think it is probably more straightforward to directly compare the surface brightness or mass density profiles between low- and high-redshift massive galaxies with an evolutionary connection (e.g., using the cumulative number density).  We did compare with stacked profiles of high-z galaxies from Patel+2013 in companion paper (Huang et al. 2016a), and we confirmed that the inner profile does not evolve a lot while the higher redshift galaxy has a steeper outer profile.  Also, we are performing careful comparisons with the massive galaxies in hydrodynamic simulations to see whether they can correctly reproduce the stellar mass distribution and the halo mass dependence.  If so, we will further look into the redshift evolution of the environment dependence using simulations and compare them with results in literature.

The important point based on the Monte Carlo tests now is also emphasized in our discussion.

> 5. I do not think that comparing exactly the same exact M(<10kpc) distributions is necessarily the best way to go to look for possible differences in evolutionary pathways induced by environment. For example, if on average massive galaxies at the centre of clusters are older and/or were born in deeper potential wells, I would expect their inner density distributions to be somewhat steeper at early stages during the cooling phase, or in fact maybe even flatter if at later epochs if and when additional effects such as (stronger?) AGN feedback and/or (more frequent?) SMBH mergers kick in. The situation is complex enough to warrant more investigation in this respect. One way to go could be to fit some Sersic profiles to the mu_star profiles and check whether the Sersic index distributions are different in different environments at fixed Mstar or Mhalo (see e.g. Figure 6 in HC13). Other ways may of course be possible and I leave it to the Authors to device additional routes for more test comparisons between galaxies in different environments. Of course similar comments apply to the comparisons in colour profiles.

We understand and agree on the point made by the referee that the relation between inner profile and the environment is not exactly clear.  We adjust several parts of the text to reflect this point.  And during the discussion in Section 4.2, we now admit that "it is still too early to conclude that the stellar mass distributions at the center of massive ETGs also depend on the environment."  We also briefly discuss the potential impacts of AGN feedback and SMBH merger of the inner mass distributions, although it is not entirely clear that they can affect the stellar mass within ten kpc.  For instance, the largest depleted-core in massive ETGs is still smaller than five kpc.

The reason we emphasize the usage of 10 kpc mass is that we propose to use it as a proxy of the mass formed in the in situ phase.  In our earlier work, we found this assumption is promising when compared with results with recent simulations.  Under the two-phase scenario, it is critical to discuss the assembly of the massive galaxy and the galaxy-halo connection via the angle of the in situ and ex situ components.  It is not an easy task as observational separation of these components is almost impossible.  We wish to use this simple assumption in this work as a starting point, which we are in the process of refining further.  We do see intriguing evidence of halo-mass dependence on the M_100kpc-M_10kpc plane and we are now investigating this using a weak lensing analysis.

As for the Sersic modeling method, we agree with the referee that it could be a useful approach.  As a matter of fact, we did try Serisc fitting to the HSC images of our massive galaxies.  However, at the imaging depth of HSC, we find that single Sersic is a terrible model for these galaxies.  It can not describe the changes of isophotal shape and slope of surface brightness profiles of our massive galaxies.  And the extended low surface brightness stellar haloes seen on the HSC images always drive the Sersic index to very high value, which tends to overestimate the mass in the inner region.  We think the more appropriate approach would be using more complicated or more physically motivated models for these galaxies.  Right now, we are working on this using both real and simulated (from the Illustris simulation) massive galaxies.

## MINOR COMMENTS

> 1. Per se, having more extended sizes/stellar haloes may not necessarily favour the merge hypothesis. One could also argue that galaxies in more massive haloes are simply born more extended (e.g., Buchan & Shankar 2016).

We thank the referee for drawing our attention to the Buchan & Shankar (2016) work.  It is an intriguing idea that the central galaxies of massive haloes could form in a more extreme fashion that leads to their large size.  We expand the discussion a little to include this theory and cite the recommended work.

We think that better observations of massive galaxies at high redshift could help constrain this picture better.  Right now, although there are works that find the significant environmental dependence of mass-size relation at z>1, we think it is still too early to draw any solid conclusion.  Massive central galaxy of high-z proto-cluster often also has very compact size (e.g., Kubo et al. 2017).  In the previous work of this series (Huang et al. 2017; arXiv:1707.01904), we compare our sample with a very massive cD galaxy at z~1 (Liu et al. 2011).  We find its outer profile is still too steep compared to our z~0.4 ones, suggesting latter assembly in the outskirt by minor mergers.

> 2. While reading the introduction I was unclear of how and why the two-phase scenario should be necessarily strictly related to the host halo masses. We know for a fact, for example, that the (unevolved) subhalo mass function is invariant with respect to the host halo mass, and the number of moderate-major mergers is also not a strong function of main progenitor halo mass (e.g., Shankar+15).

The referee is right that in many models and simulations, the moderate-major merger rate does not show a strong correlation with halo mass.  And it is probably too early to say that the frequency of mergers causes the environment dependence on size and structure.  We mentioned the "two-phase" scenario mostly because, in some state-of-art hydrodynamic simulations, the fraction of "ex-situ" or accreted stars around central galaxy increases with halo mass (e.g., see Fig 12 of Pillepich et al. 2017b using the most recent IllustrisTNG simulation).  It seems to suggest that mergers are more important to the mass assembly of central galaxies in more massive halos.  However, it is not entirely clear that mergers can lead to differences in stellar mass distribution at the same total stellar mass.  From simulations, there is evidence that suggests the frequency of minor merger depends on halo mass and could cause the environment dependence of size (e.g., Yoon et al. 2017, use the SAM by Guo et al. 2011).  We already briefly discuss this in Section 4.1 of our work, and acknowledge that this trend still relies on the assumption of the typical time scale for the minor merger).

We rewrite this section of introduction to be more objective and to reflect the point raised by the referee.  We also include Shankar et al. (2015) as citation now.

> 3. A few more words should be included on the methods adopted by Simet+16 and Diemer & Kravtsov15 without forcing the reader to go back to the original papers to get an idea of what these groups are doing...

We have now added more detailed explanations of both works as suggested by the referee.

In short, Simet et al. (2017) calibrate the mass-richness relation for the redMaPPer clusters using the SDSS weak lensing data after carefully addressing all relevant systematic uncertainties.  Diemer & Kravstov (2015) systematically investigates the mass density profiles of dark matter halos under Lambda-CDM cosmology using a suite of dark matter-only numerical simulations.  They provide empirical fitting functions for the relations among halo mass, radius, concentration, and redshift.  We use these functions to estimate the R200 for the haloes of the massive central galaxies in our sample.

> 4. The sentence in Section 2 "...and we can safely ignore structural evolution over our redshift range (~1.5 Gyr of time span)." sounds a little unclear and also out of place in this section presenting the data. Some rewording is required. It is true that models confirm that in this time span not much evolution is expected (e.g., Shankar+15). However, the authors should better emphasize that they do not expect any difference in size induced by environment/time effects in their sample, but this should be provided later on when presenting the results.

Our original intention is to suggest that there should be little size growth or structural evolution during the ~1.5 Gyr between redshift 0.3 and 0.5.  We think it would not be appropriate to look into the environment dependence of structure if we can not ignore the size growth of individual galaxy within the span of the redshift range.

The referee is right that this sentence is not very clear. We changed it into: "...And massive galaxies should experience little structural evolution and size growth between z=0.5 and 0.3 (~1.5 Gyr of time span) based on model prediction (e.g., Shankar et al. 2015)."

> 5. There are some typos around, e.g., "..this results is consistent...", "they also shows that...", etc...

We thank the referee for the careful reading and correct both of these typos.  We have also went through a more careful proof-reading process.